### 2013-06-18 background.tex
first commit

\section{Background}
\label{sec:background}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

### 2013-06-18 guide14-initial-2e.tex
first commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
% 	Template for Producing ASP-DAC 2014 Proceedings
% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% History
% ??/??/?? Designed by Hiroaki Kunieda (ASP-DAC '97 Publication Chair)
% 09/22/97 Modified and small bug fixed by Masaharu Imai 
% 	   (ASP-DAC '98 Publication Chair)
% 11/02/98 Modified by Tsuyoshi Isshiki
% 	   (ASP-DAC 2000 TPS Secretary)
% 7/24/00 Modified by Kiyoharu Hamaguchi
% 	   (ASP-DAC 2001 Publication chair)
% 6/18/02 Modified by Kazutoshi Kobayashi
% 	   (ASP-DAC 2003 Publication Co-Chair)
% 5/27/03 Modified by Kiyoharu Hamaguchi
% 	   (ASP-DAC 2004 TPC secretary)
% 6/10/03 Modified by Kazutoshi Kobayashi for Latex2e
% 	   (ASP-DAC 2004 Publication Co-Chair)
% 6/01/05 Modified by Nozomu Togawa
% 	   (ASP-DAC 2006 Publication Chair)
% 6/01/06 Modified by Hiroyuki Ochi
% 	   (ASP-DAC 2007 Publication Chair)
% 5/30/08 Modified by Nozomu Togawa
% 	   (ASP-DAC 2009 Publication Co-Chair)
% 4/30/10 Modified by Masashi Imai
% 	   (ASP-DAC 2011 Publication Chair)
% 3/20/12 Modified by Masashi Imai
% 	   (ASP-DAC 2014 Publication Chair)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% If you have any problem, please contact ASP-DAC 2014 Publication
% Co-Chairs by E-mail at ``aspdac2014-tpc@mls.aspdac.com.''
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[twocolumn]{article}
%% If you use dvips and ps2pdf, please use Postscript font 
%% and uncomment the line below.
%%\usepackage{times}
\pagestyle{empty}
%set paper size
%for A4 paper
\topmargin      29mm    %bottom margin 30mm
\oddsidemargin  15mm    %left & right margin 15mm

%for 8 1/2" x 11" paper paper, use the following definition
%\topmargin     17mm    %bottom margin 24mm
%\oddsidemargin 18mm    %left margin 18mm & right margin 17mm

%text sizes
\textwidth  180mm
\textheight 238mm
\columnsep  5.0mm
\parindent  3.5mm

%misc parameters
\headsep 0mm  \headheight 0mm
\footskip 18mm
%\footheight 6mm

%conversion to values for LaTeX
\advance\topmargin-1in\advance\oddsidemargin-1in
\evensidemargin\oddsidemargin

\makeatletter
%as Latex considers descenders in its calculation of interline spacing,
%to get 12 point spacing for normalsize text, must set it to 10 points
\def\@normalsize{\@setsize\normalsize{12pt}\xpt\@xpt
\abovedisplayskip 10pt plus2pt minus5pt\belowdisplayskip \abovedisplayskip
\abovedisplayshortskip \z@ plus3pt\belowdisplayshortskip 6pt plus3pt
minus3pt\let\@listi\@listI}

%interline spaceing and title font for section
\def\section{\@startsection {section}{1}{\z@}{20pt plus 2pt minus 2pt}
{8pt plus 2pt minus 2pt}{\centering\normalsize\sc
\edef\@svsec{\thesection.\ }}}
\def\thesection{\Roman{section}}

%interline spacing and title font for subsection
\def\subsection{\@startsection {subsection}{2}{\z@}{16pt plus 2pt minus 2pt}
{6pt plus 2pt minus 2pt}{\normalsize\sl
\edef\@svsec{\thesubsection.\ }}}
\def\thesubsection{\Alph{subsection}}

%figures/tables captions
\long\def\@makecaption#1#2{
\vskip10pt\begin{center} #1 #2 \end{center}\par\vskip 1pt}
\def\fnum@figure{\raggedright{\footnotesize Fig. \thefigure }.%
\footnotesize}
\def\fnum@table{\footnotesize TABLE \thetable\\\footnotesize\sc}
\def\thetable{\Roman{table}}

\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
%date not printed
\date{}

%make title
\title{\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\
\large\textbf{
Preparation of Papers in Two-Column Format\\
for the ASP-DAC 2014 (\LaTeX2e version)}}	% Modified by K. Kobayashi 18/06/02

%for single author
%\author{Center the Authors Names Here \\
%Center the Affiliations Here\\
%Center the City, Stats and Country Here\\
%{\small (it is your option if you want your entire address listed)}}

%for two authors
\author{
You must not show authors' names and affiliations in the\\
initial-submission paper. But you had better to make \\
room for the final submission.\\
\\
\\
\\ % To make more spaces just add \\!
}
\maketitle
\thispagestyle{empty}

{\small\textbf{Abstract---
 Abstract is a brief (50-80 word) synopsis of your paper.
The purpose is to provide a quick outline of your
presentation, giving the reader an overview of the
research. It must be fit within the size allowed, which is
about 3 inches or 7.5 centimeters.}}

\section{Introduction}

These introductions give you basic guidelines for preparing 
initial-submission papers for the ASP-DAC 2014.

The instructions assume that you have computer that can generate a PDF
file.

These instructions have been prepared in the preferred format. For items
not addressed here, please refer to recent issues of IEEE Transactions
and simulate, as closely as possible.

\section{How to Format the Page}

\subsection{Full-Size Camera-Ready Copy}

Prepare Initial-Submission paper in full size format, on A4 size or 8 1/2''
$\times$ 11''(21.5cm $\times$ 27.9 cm) paper.

\subsection{Fonts}
The best results will be obtained if your computer word-processor has
several font sizes. Try to follow the font sizes specified in Table I as
best as you can. As an aid to gauging font size, 1 point is about
0.35mm. Use a proportional, serif font such as Times of Dutch Roman.

\begin{table}[tb]
\caption{Fonts for Initial-Submission and Camera-Ready Papers}
\begin{minipage}{8cm}
\def\arraystretch{1.5}\tabcolsep 2pt
\def\thefootnote{a}\footnotesize
\begin{tabular}{l@{~}l@{~~~}l}
\hline
\parbox[c]{7mm}{Font\newline Size} & Style & Text\\
\hline
 14pt&bold     &Paper title\\
 12pt&         &Authors' names\\
 10pt&         &Authors' affiliations, main text, equations,\\[-5pt]
     &         &first letters in section titles\footnotemark[1]\\
 10pt&italic   &Subheddings\\
 ~9pt&bold     &Abstract\\
 ~8pt&         &Section titles\footnotemark[1], table
                names\footnotemark[1], first letters in table\\[-5pt]
     &         &captions\footnotemark[1],
                tables, figure captions, references,\\[-5pt]
     &         &footnotes, text subscripts and superscripts\\
 ~6pt&         &Table captions\footnotemark[1], table superscripts\\
\hline
\end{tabular}
\footnotetext[1]{\scriptsize Uppercase}
\end{minipage}
\end{table}

\subsection{Formats}

In formatting your A4-size paper, set top margin to 29mm (1.14 inches),
bottom margin to 30mm (1.18 inches), left and right margins to 15mm
(0.59 inches) If you are using paper 8 1/2'' $\times$ 11'', set top margin
to 17mm (0.67 inches), bottom margin to 24mm (0.94 inches), left
margin to 18mm (0.71 inches) and right margins to 17mm (0.67 inches). The column
width is 88mm (3.46 inches) with 5mm (0.2 inches) space between the two
columns.

You should left- and right-justify your columns. On the last page of
your paper, try to adjust the lengths of the two columns so that they
are the same. Use automatic hyphenation if you have it and check
spelling. Either digitize or paste down your figures.

%{\bf Number each of you submitted pages at the top, right corner, in
%non-photographic light blue pencil.}

\section{Figures and Tables}
Position figures and tables at the tops and bottoms of columns. Avoid
placing them in the middle of columns. Large figures and tables may
span across both columns. Figure captions should be below the figures;
table captions should be above the tables. Avoid placing figures and
tables before their first mention in the text. Use the abbreviation
``Fig.1'', even at the beginning of a sentence.

\begin{figure}[tb]
\begin{center}
\begin{minipage}{5cm}
$.$\hrulefill $.$\\$|$\hfill $|$\\$|$\hfill $|$\\$|$\hfill $|$\\
$|$\hfill this is \hfill $|$\\
$|$\hfill a sample \hfill $|$\\
$|$\hfill  figure  \hfill $|$\\
$|$\hfill $|$\\$|$\hfill $|$\\$|$\hfill $|$\\$.$\hrulefill $.$\\
\end{minipage}
\caption{This is a sample figure. Captions exceeding
one line are arranged like this.}
\end{center}
\end{figure}

\textbf{
The paper's graphics should have resolutions of 600dpi for monochrome,
300 dpi for grayscale, and 300 dpi for color. But, please do not include
too much high-resolution images or photos in your paper.
%%% Images or photos over 300dpi do not enhance the quality of your
%%% paper, but increase the size of the PDF file!
%%% Reduce the dpis of them as small as possible. 
} % Modified by N. Togawa 06/01/2005

\section{Helpful Hints}

\subsection{References}
List and number all references at the end of the paper. When reffering
to them in the text, type the corresponding reference number in the
parentheses as shown at the end of this sentence \cite{key}. Number
the citations consecutively. The sentence punctuation follows the
parentheses. Do not use ``Ref.\ \cite{baz}'' or
``reference \cite{baz}'' except at the beginning of a sentence.

\subsection{Footnotes}
Number the footnotes separately in superscripts. Place the actual
footnote at the bottom of the column in which it is cited. Do not put
footnotes in the reference list.

\subsection{Authors names}

Give all authors' names; do not use ``et al'' unless there are six
authors or more. Papers that have not been published, even if they have
been submitted for publication, should be cited as
``unpublished'' \cite{unpub}.  Papers that have been accepted for
publication should be cited as ``in press'' \cite{inpress}.
Capitalize only the first word in a paper title, except for proper nouns
and element symbols.

For papers published in translation journals, please give the English
citation first, followed by the original foreign language
citations \cite{trans}.

\subsection{Notice for \LaTeX\ users}

If you use \LaTeX\ to create your initial-submission paper, we recommend you
to use \texttt{dvipdfm} to produce PDF files from dvi files. If you cannot
use it, please use Type1 fonts instead of ugly Type3 fonts!

\section{Summary and Conclusions}

This template can be downloaded through the ASP-DAC 2014 web site
(http://www.aspdac.com/). If you have any problem, please contact ASP-DAC
2014 publication chair (aspdac2014-pub@mls.aspdac.com).

%this is how to do an unnumbered subsection 
\section*{\sc Acknowledgements}
This article was written by referring to {\em ``Author's guide --
Preparation of Papers in Two-Column Format for VLSI Symposia on
Technology and Circuits''}, the {\em ``Preparation of Papers in
Two-Column Format for the Proceedings of the 32nd ACM/IEEE Design
Automation Conference''} written by Ann Burgmeyer, IEEE and {\em ``the
template for producing IEEE-format articles using \LaTeX''}, written by
Matthew Ward, Worcester Polytechnic Institute.

\begin{thebibliography}{9}
\footnotesize
\bibitem{key}
I. M. Author,
``Some related article I wrote,''
{\em Some Fine Journal}, vol. 17, pp. 1--100, 1987.

\bibitem{baz}
A. N. Expert,
{\em A Book He Wrote,}
His Publisher, 1989.

\bibitem{unpub}
M. Smith,
``Title of paper optional here,''
unpublished.

\bibitem{inpress}
K. Rose,
``Title of paper with only first word capitalized,''	% bug fixed by M. Imai
in press.

\bibitem{trans}
T. Murayama,
``Title of paper published in translation journals,''	% bug fixed by M. Imai
{\em Some English Journal}, vol. 17, pp. 1--100, 1995.	% bug fixed by M. Imai
({\em Original Foreign Journal, vol. 1, pp. 100-200, 1993}.)	% ditto

\end{thebibliography}
\end{document}

### 2013-06-19 introduction.tex
first commit
Added the introduction and motivation

--- 2013-06-18 background.tex
@@
-\section{Background}
-\label{sec:background}
+\section{Introduction and motivation}
+\label{sec:intr-motiv}
+
+Reactive languages~\cite{gber05,amal10} are a class of programming
+languages used for designing and implementing \textit{reactive} systems,
+which continuously respond to input from their environment. These
+languages have been successfully used in programming a plethora of
+systems such as fly by wire in Airbus~\cite{eairbus}, security
+surveillance systems~\cite{amal121}, etc. Usually the working
+environment of these reactive systems is also real-time. Yet, these
+languages do not support describing real-time as first class language
+constructs. For example, one cannot describe a simple real-time delay
+\texttt{delay 0.2 ms} in these languages. These languages are based on
+formal semantics, essential for the formal verification of functional
+correctness of the developed programs, but leave non-functional
+properties such as timing behavior as an implementation
+detail~\cite{ringler00,boldt07}.
+
+These languages only support discrete \textit{logical} clock rather than
+a continuous clock. The period of the discrete logical clock (also
+referred to as a \textit{logical tick}) is determined by the
+responsiveness of the program and as opposed to a discrete
+\textit{physical} clock is elastic. Although this timing model has
+worked well for these languages in designing discrete systems there is
+obviously a need to introduce continuous time since often implementation
+models are in the continuous time
+domain~\cite{DBLP:journals/pieee/SifakisTY03}.
+
+\input{motivation}
+
+The fundamental contribution of this paper is the inclusion of real-time
+constructs in GALS (and its subset synchronous) reactive
+languages. Moreover, our contributions can be further refined as
+follows:
+
+\begin{enumerate}
+\item The real-time is modeled in the $\mathbb{Q}^{>0}$ domain of
+  numbers, i.e., non-negative rational numbers.
+\item We consider the property of \textit{non maximal-parallelism},
+  i.e., we consider scheduling on constrained resources to be an
+  integral part of the solution. To our knowledge we are the very first
+  to do so.
+\item Our solution does not change the functional or timing semantics of
+  GALS and its subset synchronous programs, in fact, our solution does
+  not require one to change the compiler at all.
+\item Our solution gives complete control of under and over
+  approximation of real-time to the system designer.
+\end{enumerate}
+
+The rest of the paper is arranged as follows...
 

### 2013-06-19 motivation.tex
separated introduction and motivation
Added the introduction and motivation

--- 2013-06-18 background.tex
@@
-\section{Background}
-\label{sec:background}
+\subsection{Motivating Example}
+\label{sec:motivating-example}
+
+% system{
+%  interface{
+%   input signal TOUCH;
+%   output signal GREEN_LIGHT;
+%   output char channel S;
+%   input char channel R;
+%  }
+\begin{figure}[t!]
+% \begin{center}
+\begin{minipage}{5cm}
+\begin{verbatim}
+{ // Clock-domain 1
+while(1) {
+  trap(TRAP){
+   // Abort if user touches the screen
+   abort(TOUCH){
+    {sustain GREEN_LIGHT;}
+    || // synchronous parallel operator
+    {
+    //exit after any 
+    // where from 0.5 to 1 ms.
+    /*delay [0.5..1]ms;*/
+    exit(TRAP); 
+    }
+   } do {send S(1);} // Send 1 if touched 
+  } do {send S(0);} // Else Send 0
+  pause; // only construct to consume time
+ }
+}
+>< // asynchronous operator
+{ // Clock-domain 2
+  while(1){ receive S;}
+  // do something with S
+}
+\end{verbatim}
+\end{minipage}
+\caption{An example human reaction time collection system (HRTCS)}
+% }
+% \end{center}
+\label{fig:1}
+\end{figure}
+
+Consider a simple machine used to test human reaction time and collect
+this data for future research purposes. The machine switches on a green
+light on a touch screen for anywhere between \texttt{0.5ms} to
+\texttt{1ms}, if the user can touch this green light then she is
+successful and an integer \texttt{1} is sent to a data-base server, if
+unsuccessful a \texttt{0} is sent instead, this way a database of human
+responsiveness can easily be generated.
+
+The SystemJ~\cite{amal10} code for such a system is shown in
+Figure~\ref{fig:1}. We use the SystemJ language in this paper first of
+all, because being a \textit{Globally Asynchronous Locally Synchronous}
+(GALS) language it helps us easily describe a larger class of systems
+(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
+languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
+allows us to explore the marriage of real-time and logical time, which
+is equally applicable to the synchronous sub-set.
+
+The SystemJ program is divided into two clock-domains, the first
+clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
+signal) on the touch screen sensor and waits for the user to respond. If
+the user is able to touch the screen within the specified time of
+\texttt{0.5ms} to \texttt{1ms} a positive response is sent to the second
+clock-domain, which is interacting with data-base (via CSP style channel
+communication via \texttt{S}) server in Java. The synchronous parallel
+operator \texttt{||} along with the \texttt{trap-exit} (exception
+mechanism) is used to make sure that the green light is switched on for
+only a specified delay. The 2 clock-domains are implemented using the
+asynchronous parallel \texttt{><} operator. All these constructs are
+described in~\cite{amal10}.
+
+This SystemJ example is \textit{incomplete}, the \texttt{delay}
+specification, which models the non-deterministic delay between
+\texttt{0.5ms} and \texttt{1ms} does not work with the SystemJ compiler!
+In fact, there is no way to include such real-time specifications in the
+language itself.
 

### 2013-06-19 background.tex
Added the background section

@@
 \label{sec:background}
+
+\subsection{Brief introduction to SystemJ syntax, semantics and model of
+  computation}
+\label{sec:brief-intr-syst}
+
+\begin{table}[tb]
+\centering
+\caption{SystemJ kernel statements and their meaning}
+\begin{minipage}{8cm}
+  \begin{small}
+   \begin{tabular}{|c|p{80pt}|}
+     \hline                                                                                     
+     \textbf{Kernel Statements} & \textbf{Meaning}\\                                            
+     \hline                                                                                     
+     \hline                                                                                     
+     [\textbf{\texttt{input}}] [\textbf{\texttt{output}}]
+     [\textbf{\texttt{type}}] \textbf{\texttt{signal}} S & declare signal S\\                                      
+     \hline                                                                                     
+     \textbf{\texttt{emit}} S [(value)] & broadcast signal S\\                                                    
+     \hline                                                                                     
+     \textbf{\texttt{present}} (S) \{p\} else \{q\}& do p if S is present, else do q\\                            
+     \hline                                                                                     
+     \textbf{\texttt{abort}} (S) \{p\} & preempt program p if S is present\\                                      
+     \hline                                                                                     
+     \textbf{\texttt{suspend}} (S)\{p\} & suspend p if S is present\\                                             
+     \hline                                                                                     
+     \textbf{\texttt{trap}} (T)\{p\ldots \textbf{\texttt{exit}} T\ldots\} & preempt p if exit is executed\\                         
+     \hline                                                                                     
+     p\textbf{\texttt{$||$}}q & run p and q in lock-step\\                                                        
+     \hline                                                                                     
+     p$><$q & run p and q asynchronously\\                                                      
+     \hline                                                                                     
+     \textbf{\texttt{send}} C([value]) & send a value through C, blocking
+     send\\                                                 
+     \hline                                                                                     
+     \textbf{\texttt{receive}} C() & receive a value through C, blocking
+     receive\\
+     \hline                                                                                     
+     \textbf{\texttt{pause}} & finish a tick and communicate
+     with environment\\
+     \hline                                                                                     
+   \end{tabular}
+  \end{small}
+   % \footnotetext[1]{\scriptsize Uppercase}
+ \end{minipage}
+ \label{tab:1}
+\end{table}
+
+Table~\ref{tab:1} presents the SystemJ kernel statements used to program
+the control flow. The data-flow is programmed in Java. A number of
+derived statements also exist (e.g., \texttt{sustain}, \texttt{await},
+etc) that make programming reactive systems easier. The \textit{Model of
+  Computation} (MoC) of a simple SystemJ is shown in Figure~\ref{fig:2}.
+
+\begin{figure*}[bth]
+\centering
+\begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
+\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
+\begin{verbatim}
+while(true){
+ abort(A){
+  while(true)
+   pause;
+  emit O;
+  // do java computation
+}
+\end{verbatim}%
+\end{minipage}%
+\end{SubFloat}
+\hspace{1.5cm}%
+\begin{SubFloat}[Black box]{\label{fig:2b}Ticks in SystemJ}%
+\includegraphics[scale=0.5]{moc}
+\end{SubFloat}%
+\caption{Simple SystemJ example and corresponding MoC}
+\label{fig:2}
+\end{figure*}
+
+SystemJ consists of entities called clock-domains (CD)s, each running at
+their individual pace. Each CD adheres to the \textit{perfect synchrony}
+hypothesis, i.e., all statements execute instantaneously in zero
+time. Only the \texttt{pause} statement consumes time, just like in
+Esterel~\cite{gber931}. Consider the SystemJ program in
+Figure~\ref{fig:2a}, it is waiting for an input signal producing logical
+ticks (see Figure~\ref{fig:2b}). Once \texttt{A} is received at tick 4,
+output \texttt{O} is emitted to the environment instantaneously.
+
+\subsection{Mapping logical time to physical time}
+\label{sec:mapping-logical-time}
+
+The perfect synchrony hypothesis is ideal for programming the
+synchronous sub-set of SystemJ. But, the real-world is not that
+forgiving, execution of every \textit{micro-step} in the logical zero
+time, requires $\delta$ physical time. Time for a complete reaction to
+one or more input signal can thus be summarized as $\Delta = \Sigma
+(\delta)$. We call $\Delta$ the reaction time. Obviously depending upon
+the amount of computation required, $\Delta$ can vary, hence, in order
+to satisfy the implicit restriction posed by the synchrony hypothesis --
+no input event can be missed, one needs to calculate the \textit{Worst
+  Case Reaction Time} (WCRT) and the resultant WCRT needs to be smaller
+than the speed of the fastest input event, else the synchrony hypothesis
+is violated. A substantial amount of research for calculating the WCRT
+of synchronous programs~\cite{proop10,boldt07,wilhelm08} exists. We
+refer the reader to any of these research articles, since WCRT analysis
+is not the focus of this paper. The opposite of the WCRT is the
+\textit{Best Case Reaction Time} (BCRT), we denote the WCRT and BCRT in
+Figure~\ref{fig:2b} using \texttt{W} and \texttt{B}, respectively.
+
+We now present two lemmas with proof sketch (due to lack of space) that
+we require for the rest of the paper.
+
+\begin{lemma}
+  WCRT and BCRT analysis are invariant to values of conditional
+  expressions.
+\end{lemma}
+
+\begin{proof}
+  The physical time ($\delta$) taken by a conditional expression is at
+  best affected by the type rather than the value of the
+  conditional. Example, comparing the value of 2 floats might take
+  longer than comparing the value of 2 integer types on some given
+  platform. But, the time taken by the comparison instruction is
+  invariant to the value itself.
+\end{proof}
+
+\begin{lemma}
+  WCRT and BCRT analysis are invariant to channel communication.
+\end{lemma}
+\begin{proof}
+  LATER ....
+\end{proof}
 

### 2013-06-19 introduction.tex
Added the background section

@@
 \texttt{delay 0.2 ms} in these languages. These languages are based on
-formal semantics, essential for the formal verification of functional
-correctness of the developed programs, but leave non-functional
-properties such as timing behavior as an implementation
-detail~\cite{ringler00,boldt07}.
-
-These languages only support discrete \textit{logical} clock rather than
-a continuous clock. The period of the discrete logical clock (also
-referred to as a \textit{logical tick}) is determined by the
-responsiveness of the program and as opposed to a discrete
-\textit{physical} clock is elastic. Although this timing model has
-worked well for these languages in designing discrete systems there is
-obviously a need to introduce continuous time since often implementation
-models are in the continuous time
+formal semantics, essential for the verification of functional
+correctness properties of the developed programs, but leave
+non-functional properties such as timing behavior as an implementation
+detail~\cite{ringler00,boldt07}. Arguably rightly so, because physical
+time cannot be incorporated without information about the underlying
+execution platform. But, one can still reason in terms of
+\textit{logical} time. These languages support discrete \textit{logical}
+clock rather than a continuous clock. The period of the discrete logical
+clock (also referred to as a \textit{logical tick}) is determined by the
+responsiveness of the program to the input signals and as opposed to a
+discrete \textit{physical} clock is elastic. Although this timing model
+has worked well for these languages in designing discrete systems there
+is obviously a need to introduce continuous time since often
+implementation models are in the continuous time
 domain~\cite{DBLP:journals/pieee/SifakisTY03}.
@@
   approximation of real-time to the system designer.
+\item We are the first to allow specification of both deterministic and
+  non-deterministic real-time.
 \end{enumerate}

### 2013-06-24 background.tex
First version of the paper

@@
 \begin{minipage}{8cm}
-  \begin{small}
+  \begin{scriptsize}
    \begin{tabular}{|c|p{80pt}|}
@@
    \end{tabular}
-  \end{small}
+  \end{scriptsize}
    % \footnotetext[1]{\scriptsize Uppercase}
@@
 Table~\ref{tab:1} presents the SystemJ kernel statements used to program
-the control flow. The data-flow is programmed in Java. A number of
+the control flow. The data flow is programmed in Java. A number of
 derived statements also exist (e.g., \texttt{sustain}, \texttt{await},
 etc) that make programming reactive systems easier. The \textit{Model of
-  Computation} (MoC) of a simple SystemJ is shown in Figure~\ref{fig:2}.
+  Computation} (MoC) of a simple SystemJ program is shown in
+Figure~\ref{fig:2}.
 
@@
 \hspace{1.5cm}%
-\begin{SubFloat}[Black box]{\label{fig:2b}Ticks in SystemJ}%
-\includegraphics[scale=0.5]{moc}
+\begin{SubFloat}[Black box]{\label{fig:2b}Logical time in SystemJ}%
+\includegraphics[scale=0.4]{moc}
+\end{SubFloat}%
+\begin{SubFloat}{\label{fig:2c}Mapping to physical time}%
+% \includegraphics[scale=0.4]{phy.pdf_t}
+\scalebox{0.4}{\input{phy.pdf_t}}
 \end{SubFloat}%
@@
 SystemJ consists of entities called clock-domains (CD)s, each running at
-their individual pace. Each CD adheres to the \textit{perfect synchrony}
+their individual pace, interacting with the environment via signals and
+with each other via channels and CSP~\cite{pwel05} style
+rendezvous. Each CD adheres to the \textit{perfect synchrony}
 hypothesis, i.e., all statements execute instantaneously in zero
@@
 one or more input signal can thus be summarized as $\Delta = \Sigma
-(\delta)$. We call $\Delta$ the reaction time. Obviously depending upon
+(\delta)$. We call $\Delta$ the reaction time. Obviously, depending upon
 the amount of computation required, $\Delta$ can vary, hence, in order
-to satisfy the implicit restriction posed by the synchrony hypothesis --
+to satisfy the implicit restriction imposed by the synchrony hypothesis,
 no input event can be missed, one needs to calculate the \textit{Worst
@@
 \textit{Best Case Reaction Time} (BCRT), we denote the WCRT and BCRT in
-Figure~\ref{fig:2b} using \texttt{W} and \texttt{B}, respectively.
+Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
+respectively. Figure~\ref{fig:2c} shows a continuously running physical
+clock (analogous to a clock in hardware), the numerical annotations on
+the rising edge of the clock mark the logical ticks from
+Figure~\ref{fig:2b}. The subscripts \texttt{s} and \texttt{e} for the
+numbers show the start and end of the logical ticks, respectively. The
+end of a logical tick and the start of the next logical tick happen
+together. Figure~\ref{fig:2c} clearly shows the mapping of the logical
+time to the physical time. As we can see due to Java computations
+(Figure~\ref{fig:2a}) the physical time between logical ticks 4 and 5 is
+2 clock cycles, while for the rest of the ticks it is just 1 clock
+cycle. This elasticity is inherent (and elegant) to both Esterel and
+SystemJ.
 
-We now present two lemmas with proof sketch (due to lack of space) that
-we require for the rest of the paper.
+% We now present a lemma with proof sketch (due to lack of space) that we
+% require for the rest of the paper and is special to SystemJ, being a
+% GALS language.
 
-\begin{lemma}
-  WCRT and BCRT analysis are invariant to values of conditional
-  expressions.
-\end{lemma}
+% \begin{lemma}
+%   WCRT and BCRT analysis are invariant to values of conditional
+%   expressions.
+% \end{lemma}
 
-\begin{proof}
-  The physical time ($\delta$) taken by a conditional expression is at
-  best affected by the type rather than the value of the
-  conditional. Example, comparing the value of 2 floats might take
-  longer than comparing the value of 2 integer types on some given
-  platform. But, the time taken by the comparison instruction is
-  invariant to the value itself.
-\end{proof}
+% \begin{proof}
+%   The physical time ($\delta$) taken by a conditional expression is at
+%   best affected by the type rather than the value of the
+%   conditional. Example, comparing the value of 2 floats might take
+%   longer than comparing the value of 2 integer types on some given
+%   platform. But, the time taken by the comparison instruction is
+%   invariant to the value itself.
+% \end{proof}
 
-\begin{lemma}
-  WCRT and BCRT analysis are invariant to channel communication.
-\end{lemma}
-\begin{proof}
-  LATER ....
-\end{proof}
+% \begin{lemma}
+%   WCRT and BCRT analysis are invariant to channel communication.
+% \end{lemma}
+% \begin{proof}
+%   LATER ....
+% \end{proof}
 

### 2013-06-24 delays.tex
first commit
First version of the paper

--- 2013-06-18 background.tex
@@
-\section{Background}
-\label{sec:background}
+\section{Introducing real time in SystemJ}
+\label{sec:intr-real-time}
+
+We introduce a single \textit{derived} construct called
+\mbox{\texttt{delay [M..N]}} in the SystemJ language for real time
+control. The resolution of the delay statement is of secondary concern
+and is dependent upon the execution platform, without loss of generality
+through out this paper we consider the resolution to be always in milli
+seconds.
+
+\subsection{Semantics of the delay statement}
+\label{sec:semant-delay-stat}
+
+Given a SystemJ program: \texttt{delay [M..N]; p}, where $M \in
+\mathbb{Q}^{>0}$ and $N \in \mathbb{Q}^{>0}$, statement $p$ is executed
+after real-time $\tau$, such that, $M \leq \tau \leq N$.
+
+We also introduce two variants of the derived statement \texttt{delay}.
+\begin{enumerate*}
+\item Given a SystemJ program \texttt{delay M; p}, where $M \in
+  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time $\tau$,
+  such that, $M \leq \tau \leq \infty$. It is important to note that the
+  lower bound of the delay construct $M$ is \textit{not} an exact delay,
+  but rather the control is allowed to proceed to the next statement
+  anytime after the delay of time $M$.
+\item Give a SystemJ program \texttt{delay [M..M]; p}, where $M \in
+  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time $\tau$,
+  such that, $M \leq \tau \leq M$. In this variant, the delay is exact.
+\end{enumerate*}
+
+The aforementioned variants are specializations of the general case
+\texttt{delay [M..N]}.
+
+\subsection{Rewriting the \texttt{delay} statement}
+\label{sec:rewr-delay-stat}
+
+The introduced \texttt{delay} construct is not a kernel construct, but a
+\textit{derived} construct built from the kernel constructs
+(Table~\ref{tab:1}) in SystemJ. Figure~\ref{fig:3} gives the rewrite of
+the \texttt{delay} construct to kernel statements.
+
+\begin{figure}[tb]
+    \begin{minipage}{\textwidth}
+      \begin{scriptsize}
+\begin{verbatim}
+trap(T){
+ int x = 0;
+ while(true){
+  x = x + 1;
+  if(x == d) exit (T); //wait for "d" ticks
+  pause;
+ }
+}
+\end{verbatim}
+      \end{scriptsize}
+    \end{minipage}
+    \caption{The rewrite of \texttt{delay} construct}
+    \label{fig:3}
+\end{figure}
+
+The fundamental observation is that real-time is converted into logical
+time via the \texttt{pause} construct. The rewrite basically maps the
+physical notion of time back to the elegant logical notion of time. The
+rewrite \textit{delays} a certain number of logical ticks, before
+proceeding to the next statement. The number of logical ticks
+\texttt{``d''} to \textit{delay} (Figure~\ref{fig:3}) is determined by
+the compiler statically at compile time. The value of \texttt{d} is
+intricately tied with the WCRT and BCRT of the program and hence the
+execution platform.
+
+\subsection{Finding the logical delay \texttt{d}}
+\label{sec:find-logic-delay}
+
+\begin{algorithm}[t!]
+  \begin{minipage}{1.0\linewidth}
+    \SetAlgoLined
+    \KwData{WCRT, BCRT, $M \in \mathbb{Q}^{>0}$, $N \in \mathbb{Q}^{>0}$}
+    \KwResult{d}
+    let $l_1 \leftarrow \lceil \frac{M}{WCRT} \rceil$\;
+    let $l_2 \leftarrow \lceil \frac{M}{BCRT} \rceil$\;
+    let $u_1 \leftarrow \lfloor \frac{N}{WCRT} \rfloor$\;
+    let $u_2 \leftarrow \lfloor \frac{N}{BCRT} \rfloor$\;
+    let $F:(l_1,u_1) \rightarrow S_1$\;
+    let $F:(l_2,u_2) \rightarrow S_2$\;
+    let $D \leftarrow S_1 \cap S_2$\;
+    \Return (some $d \in D$)\;
+    \caption{Finding the value of \texttt{d}}
+    \label{alg:1}
+  \end{minipage}
+\end{algorithm}
+
+The computation of \texttt{d} is shown in Algorithm~\ref{alg:1}. This
+algorithm is carried out for each CD (in SystemJ) or a synchronous
+program individually. \textit{The fundamental observation is -- the
+  reaction time for each logical tick is elastic -- varying only between
+  the BCRT and the WCRT, thus any logical number of ticks \texttt{d}
+  that map to the required real-time \texttt{delay} should be chosen in
+  such a way that they are invariant to this elasticity.}
+
+Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
+$M$ and $N$, respectively of the \texttt{delay} construct. We first
+divide $M$ and $N$ with the BCRT and WCRT, respectively. This division
+gives us the number of individual logical ticks required to delay the CD
+(or synchronous program) by the real-time specification. We always
+$ceil$ when dividing $M$ and $floor$ when dividing $N$ to make sure that
+the resultant values are integers (in domain $\mathbb{N}^{>0}$) and
+these functions guarantee that the resultant logical ticks result in
+real-time delays between the required range $M-N$. Next, a linear
+function $F$ maps these calculated values to a set of equidistant
+integer points (values) separated by a unit value -- these points
+represent all the logical ticks running at the WCRT and the BCRT,
+respectively that satisfy the real-time delay requirements. The
+intersection of these two sets gives all the logical ticks that satisfy
+the real-time requirements invariant of the reaction time and its
+elasticity.
+
+Let us revisit our motivating example to elucidate the algorithm. From
+Figure~\ref{fig:1} we know that $M$ is 50ms and $N$ is 100ms,
+respectively. Let the WCRT and BCRT be: 13.332ms and 10ms,
+respectively. Thus, the algorithm proceeds as follows:
+
+\begin{enumerate*}
+\item $l_1 \leftarrow \lceil 50/13.332 \rceil$ and $u_1 \leftarrow
+  \lfloor 100/13.332 \rfloor$. $l_1 = 4$ and $u_1 = 7$. We first
+  calculate the logical ticks that are always running at the WCRT
+  satisfy the required real-time delay.
+\item $l_2 \leftarrow \lceil 50/10 \rceil$ and $u_2 \leftarrow
+  \lfloor 100/10 \rfloor$. $l_1 = 5$ and $u_1 = 10$. We do the same for
+  the BCRT case.
+\item $S_1 = \{4,5,6,7\}$ and $S_2 =\{5,6,7,8,9,10\}$. We then map the
+  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
+  logical ticks that running at the WCRT and BCRT, respectively always
+  satisfy the required real-time constraints.
+\item $D = S_1 \cap S_2$, $D = \{5,6,7\}$. Finally, the intersection of
+  the two sets gives the set $D$ from which we can pick up any value for
+  \texttt{d}.
+\end{enumerate*}
+
+The resultant value for \texttt{d} gives the number of logical ticks,
+which can run at any physical clock-speed, bounded by the BCRT and the
+WCRT of the program and still result in the required real-time delay. We
+think this is an elegant solution, because the technique provides
+hard-real time guarantee while preserving the essence of synchronous and
+GALS programming prescribed by SystemJ and Esterel style
+languages. Moreover, this technique considers non maximal-parallelism,
+i.e., the delays in logical ticks is calculated after scheduling has
+been performed. To our knowledge we are the first to do so.
+
+\subsection{Extending the technique to variants of \texttt{delay}}
+\label{sec:extend-tehcn-vari}
+
+\paragraph{The \texttt{delay(M)} construct}
+\label{sec:extend-techn-vari}
+
+The \texttt{delay(M)} variant is easily accommodated in the
+technique. All one needs to do is find the set $S_1$ and choose a value
+from this set.
+
+\paragraph{The \texttt{delay(M..M)} construct}
+\label{sec:extend-techn-vari}
+
+The \texttt{delay(M..M)} variant is a little more interesting. Like
+before we find sets $S_1$ and $S_2$, and find the intersection of the
+two sets to get the value of \texttt{d}. It is possible (and often
+likely, as suggested by our experiments in
+Section~\ref{sec:experimental-results}) that the resultant set $D$ is
+empty (also possible in the case of \texttt{delay(M..N)}, but never
+possible in case of \texttt{delay(M)}). In such a case, we automatically
+\textit{relax} the upper bound of the \texttt{delay} statement.
+
+\paragraph{Relaxation of the upper real-time bound}
+\label{sec:over-appr-relax}
+
+The relaxation algorithm is shown in Algorithm~\ref{alg:2}.
+
+\begin{algorithm}[t!]
+  \begin{minipage}{1.0\linewidth}
+    \SetAlgoLined
+    \KwData{$S_2$, $D$, WCRT}
+    \KwResult{d}
+    \If {$D = \emptyset$} {
+      let $j_{0}$ be the first element of set $S_2$, s.t., $|S_2|=Q$\;
+      \ShowLn let $N \leftarrow WCRT \times j_0$\;
+      let $d \leftarrow j_0$\;
+    }
+    \Return d\;
+    \caption{Calculating the minimum relaxation of the upper real-time
+      bound}
+    \label{alg:2}
+  \end{minipage}
+\end{algorithm}
+
+This algorithm results in the smallest relaxation required for the
+real-time delay to be satisfied. The algorithm takes as input the WCRT
+and set $S_2$, recall that set $S_2$ represents the logical ticks
+required to satisfy the real-time requirement at the BCRT. We take the
+very first value from set $S_2$ and multiply it with the WCRT to get the
+relaxation $N$. The first element of set $S_2$ is returned as the
+logical tick delay \texttt{d}.
+
+The fundamental observation is that we have to delay for a minimum of
+$M$ units of real-time, hence, under-approximation is out of
+question. We can still over-approximate, but to reduce the resultant
+error, we should over-approximate by least possible value, which is the
+lower bound of set $S_2$. Thus, the lower bound is considered to be the
+only element shared between the two sets $S_1$ and $S_2$ and
+accordingly, the upper real-time bound is relaxed by the multiplication
+of WCRT and the first element of set $S_2$.
+
+\subsection{Programming using the delay construct}
+\label{sec:progr-using-delay}
+
+In this section we provide a number of examples to show the different
+types of real-time programming paradigms that can be incorporated into
+the GALS (and its sub-set synchronous) programming model.
+
+\subsubsection{Non-deterministic time}
+\label{sec:non-determ-time}
+
+A number of programs require non-deterministic timing constructs. One of
+them is our motivating example -- the human response time
+system. Another is a printer-spooler example borrowed from timed
+CSP~\cite{Schneider:1999:CRT:555233}. The spooler and printer need to
+synchronize using channels. The printer might be unable to print
+depending upon the paper tray, similarly, spooler might take sometime to
+send the job depending upon its size. Such real-time constraints can be
+modeled in SystemJ as below:
+
+\begin{scriptsize}
+
+\begin{verbatim}
+{await (job); delay([2..10]ms);  //SPOOLER CD
+ send mid(job) delay (1ms);}
+><
+{receive mid; delay([1..30]ms); //PRINTER CD
+ emit print(#mid);}
+\end{verbatim}
+\end{scriptsize}
+
+\subsubsection{Timeout}
+\label{sec:timeout}
+
+There are many instances when one would like to wait on an input from
+environment for only a specified amount of time. This can be programmed
+as:
+    \begin{scriptsize}
+\begin{verbatim}
+// timeout after 1ms.
+trap(T){
+ {await(A);}||{delay(1ms);exit(T); }
+}
+\end{verbatim}
+    \end{scriptsize}
+
+\subsubsection{Periodic reactions}
+\label{sec:periodic-reactions}
+
+A reaction, or a whole CD, can be programmed to run periodically like
+so:
+
+\begin{scriptsize}
+
+\begin{verbatim}
+{
+delay (1ms); emit S; //do something
+}
+\end{verbatim}
+\end{scriptsize}
+A periodic reaction (or CD) requires special consideration. Since the
+delay statement is converted into \texttt{pause} constructs. One should
+not introduce extra \texttt{pause} constructs when building periodic
+reactions (or CDs). This is essential since, introducing pauses would
+introduce more logical ticks.
+
+\subsubsection{Interaction of preemption and delays}
+\label{sec:inter-preempt-delays}
+
+Preemption plays an important role in reactive languages. One needs to
+carefully consider the interplay of \texttt{delay} semantics with the
+preemption semantics of reactive languages. Other attempts at
+incorporating delays (using external timers) have only been partially
+successful, because of the complex interplay between real-time and
+preemption. Consider the simple example below, which models real-time
+using external timers as in~\cite{rsh94}.
+
+\begin{scriptsize}
+
+\begin{verbatim}
+suspend{S} {
+ emit START_TIMER(10); await (TIMER); 
+ emit O1;
+}
+\end{verbatim}
+\end{scriptsize}
+
+As identified in~\cite{Bourke2009a} the \texttt{suspend} does not play
+well with the external timer. The above program sends a signal to an
+external timer and waits for 10ms to pass by. Consider what happens when
+signals \texttt{S} and \texttt{TIMER} occur together, the \texttt{await}
+statement is never executed and hence, we enter a deadlock. Such
+problems are completely avoided in our technique, because we convert the
+real-time delays into logical delays (\texttt{pause} constructs), which
+bode well with preemption.
+
+\subsubsection{Interaction of channel communication and delays}
+\label{sec:inter-chann-comm}
+
+Channels are an addition in the SystemJ language. Like interaction of
+preemption and delays, conversion of rela-time delays to logical ticks
+also bodes well with channel rendezvous, because the semantics of
+interaction are well defined~\cite{amal10}. More importantly, we need to
+consider the interplay of channel communication and WCRT/BCRT
+analysis. Since, channel communication does not stop logical time
+(see~\cite{amal10}) WCRT/BCRT are unaffected by channel
+communication. The response time to input signals \textit{is} though!
+But, we are unconcerned with the response time, analysis in this paper
+and it remains a future research avenue.
 

### 2013-06-24 introduction.tex
First version of the paper

@@
 non-functional properties such as timing behavior as an implementation
-detail~\cite{ringler00,boldt07}. Arguably rightly so, because physical
+detail~\cite{ringler00,boldt07}. Arguably, rightly so, because physical
 time cannot be incorporated without information about the underlying
@@
 responsiveness of the program to the input signals and as opposed to a
-discrete \textit{physical} clock is elastic. Although this timing model
-has worked well for these languages in designing discrete systems there
+discrete \textit{physical} clock, is elastic. Although this timing model
+has worked well for these languages in designing discrete systems, there
 is obviously a need to introduce continuous time since often
@@
 
-\begin{enumerate}
-\item The real-time is modeled in the $\mathbb{Q}^{>0}$ domain of
-  numbers, i.e., non-negative rational numbers.
+\begin{enumerate*}
+\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
+  i.e., non-negative rational numbers.
 \item We consider the property of \textit{non maximal-parallelism},
@@
   GALS and its subset synchronous programs, in fact, our solution does
-  not require one to change the compiler at all.
-\item Our solution gives complete control of under and over
-  approximation of real-time to the system designer.
-\item We are the first to allow specification of both deterministic and
-  non-deterministic real-time.
-\end{enumerate}
+  not require one to change the mid-end, the back-end or the related
+  optimization phases of the compiler at all.
+\item We are the first, to our knowledge, to allow specification of both
+  deterministic and non-deterministic real-time in GALS and synchronous
+  languages.
+\end{enumerate*}
 

### 2013-06-24 motivation.tex
First version of the paper

@@
 \begin{minipage}{5cm}
+  \begin{scriptsize}
+
 \begin{verbatim}
@@
     //exit after any 
-    // where from 0.5 to 1 ms.
-    /*delay [0.5..1]ms;*/
+    // where from 50 to 100 ms.
+    /*delay [50..100]ms;*/
     exit(TRAP); 
@@
 \end{verbatim}
+  \end{scriptsize}
 \end{minipage}
-\caption{An example human reaction time collection system (HRTCS)}
+\caption{An example human response time collection system (HRTCS)}
 % }
@@
 
-Consider a simple machine used to test human reaction time and collect
+Consider a simple machine used to test human response time and collects
 this data for future research purposes. The machine switches on a green
-light on a touch screen for anywhere between \texttt{0.5ms} to
-\texttt{1ms}, if the user can touch this green light then she is
-successful and an integer \texttt{1} is sent to a data-base server, if
+light on a touch screen for anywhere between \texttt{50ms} to
+\texttt{100ms}, if the user can touch this green light then she is
+successful and an integer \texttt{1} is sent to a database server, if
 unsuccessful a \texttt{0} is sent instead, this way a database of human
@@
 the user is able to touch the screen within the specified time of
-\texttt{0.5ms} to \texttt{1ms} a positive response is sent to the second
-clock-domain, which is interacting with data-base (via CSP style channel
-communication via \texttt{S}) server in Java. The synchronous parallel
-operator \texttt{||} along with the \texttt{trap-exit} (exception
-mechanism) is used to make sure that the green light is switched on for
-only a specified delay. The 2 clock-domains are implemented using the
-asynchronous parallel \texttt{><} operator. All these constructs are
-described in~\cite{amal10}.
+\texttt{50ms} to \texttt{100ms} a positive response is sent to the
+second clock-domain, which is interacting with a database (via CSP style
+channel communication via \texttt{S}) server in Java. The synchronous
+parallel operator \texttt{||} along with the \texttt{trap-exit}
+(exception mechanism) is used to make sure that the green light is
+switched on for only a specified delay. The 2 clock-domains are
+implemented using the asynchronous parallel \texttt{><} operator. All
+these constructs are described in~\cite{amal10} and
+Section~\ref{sec:background}.
 
@@
 specification, which models the non-deterministic delay between
-\texttt{0.5ms} and \texttt{1ms} does not work with the SystemJ compiler!
+\texttt{0.5ms} and \texttt{100ms} does not work with the SystemJ compiler!
 In fact, there is no way to include such real-time specifications in the

### 2013-06-24 related_work.tex
first commit
First version of the paper

--- 2013-06-18 background.tex
@@
-\section{Background}
-\label{sec:background}
+\section{Related work}
+\label{sec:related-work}
+
+A small number of studies have been performed when it comes to
+integrating real-time directly as a programming construct in GALS (and
+its subset synchronous) reactive languages. The most prominent works in
+this area are by Shyamsundar~\cite{rsh94} and Bourke et
+al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
+external timers in \textit{Communicating Reactive Processes} (CRP),
+which like SystemJ is an extension of synchronous language Esterel to
+asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
+external timers do not interact well with preemption constructs in these
+languages. Bourke et al. do a much better job of introducing real-time
+as first class constructs in the Esterel language. They like us provide
+real-time \texttt{delay} as first class Esterel programming constructs,
+and translate them into Esterel kernel constructs. But, unlike us they
+do not translate delays into \texttt{pause} constructs
+directly. Instead, logical ticks are generated by using abstract notion
+of \texttt{event} and \texttt{sample} platform dependent timers. This
+notion makes it a complex and inflexible solution since the number of
+platform timers with certain resolutions need to be determined and
+present on the system for the solution to be realizable. Moreover,
+non-deterministic real-time delays and integration with non-maximal
+parallelism is not studied at all by them. Other works such as
+Quartz~\cite{glog02} also incorporate delay statements in Esterel using
+\texttt{pause} constructs, but again, non-maximal parallelism,
+non-deterministic delays are not studied. Moreover, Quartz is targeted
+at studying timing properties using model checkers rather than
+implementation, like us.
 

### 2013-06-25 exps.tex
first commit
table

--- 2013-06-18 background.tex
@@
-\section{Background}
-\label{sec:background}
+\section{Experimental results}
+\label{sec:experimental-results}
+
+\begin{center}
+\scalebox{0.9}{
+\begin{tabular}{|c| c | c | c | c | c | c | c |}
+	\cline{4-8}
+	\multicolumn{3}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
+	\multirow{12}{*}{JOP} 		& \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{2-8}   
+								& \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{2-8} 
+								& \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{3-8} 
+								&			   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{2-8}
+								& \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{3-8} 
+								&			   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{2-8}
+								& \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
+								&			   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
+								& \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
+								&			   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
+								& \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
+								&			   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
+	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}   
+								& \multicolumn{2}{|c|}{WCRT} 		&0.0329 ms 		& 0.0393 ms		  & 0.0361 ms		  & 0.5072 ms		   & 0.8388 ms\\ \cline{2-8} 
+								& \multirow{2}{*}{Delay 1}	 & M..N &50.3-600.3 ms 	& 2.4-15.6265 ms  & 1230-42317.8276 ms& 10000-95358.8578 ms& 10000-287757.9834 ms \\ \cline{3-8} 
+								&			   & d			 		&18209-18232 	& 398 			  & 1171429			  & 188015		       &343054\\ \cline{2-8}
+								& \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-10.8757 ms& N/A				  & N/A 			   &10000-287757.9834 ms\\ \cline{3-8} 
+								&			   & d			 		& 	N/A	 		& 277			  & N/A				  & N/A       		   &343054\\ \cline{2-8}
+								& \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.3534 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
+								&			   & d			 		& 	N/A	 		& 9		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
+								& \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.9631 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
+								&			   & d				 	& 	N/A	 		& 50			  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
+								& \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-4.79 ms   & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
+								&			   & d			 		& 	N/A	 		& 122 			  &	N/A				  & N/A      		   &N/A\\ \cline{2-8}
+	\hline
+\end{tabular}
+}
+\end{center}
+
+aa
 

### 2013-06-26 exps.tex
table2

@@
 \label{sec:experimental-results}
+\begin{table}[t]
+\centering
+\twocolumn[
 
-\begin{center}
-\scalebox{0.9}{
+\begin{@twocolumnfalse}
+\caption{Benchmark programs using \texttt{delay} constructs} \label{table:benchmark}
+\scalebox{0.85}{
 \begin{tabular}{|c| c | c | c | c | c | c | c |}
@@
 }
-\end{center}
+\end{@twocolumnfalse}
+]
 
-aa
+\end{table}
+In this section, we present a set of SystemJ benchmark programs in which real-time requirements should be met.
+We have carried out a set of experiments to obtain BCRT and WCRT of the benchmark programs on two execution platforms called Java Optimized Processor(JOP) 
+\cite{jop:jnl:jsa2007} and TP-JOP \cite{6119095}. JOP is a hardware implementation of the JVM which enables real-time execution of Java programs by translating
+Java bytecodes into a sequence of JOP's native instructions called \emph{microcode} which is time-predictable. As SystemJ's default compilation
+target is Java source code, JOP is an excellent platform which enables us to analyze timing properties of SystemJ program. On the other hand, there is also
+an option where compiled code could be more tightly coupled to specific platforms such as TP-JOP, which execute SystemJ's kernel statements more
+efficiently. By utilizing our internal tools in conjunction with Worst-Case-Execution-Time (WCET) analyzer \cite{jop:jnl:jsa2007} 
+provided by JOP tools, we were able to estimate BCRT and WCRT of our benchmark programs: (full-name?) HRTCS, Stepper motor controller (Motor) \cite{}, Robot 
+and Access Environment Control System (AECS) \cite{}.
+
+%HRTCS, as already illustrated in Section \ref{sec:intr-motiv}, tests for an ability of one's responsiveness by generating green
+%lights between the real-time range \emph{M - N}. For AECS \ref{} we have replaced the external timer with our delay constructs. Stepper motor controller
+%is originally introduced in \cite{Bourke2009a} which converted into SystemJ.
+
+As one can see in Table \ref{table:benchmark}, BCRT and WCRT of overall programs are smaller when they run on TP-JOP. For example,
+WCRT and BCRT of HRTCS are \(\times\)11.9 and \(\times\)3.4 smaller, respectively, for TP-JOP (0.0028, 0.0329 ms) compared to JOP (0.0334, 0.1120 ms). On the other
+hand, required logical delays \emph{d} are generally bigger on TP-JOP e.g. \(\times\)4.5 - \(\times\)5.9 and \(\times\)30.9 in 
+Motor and Robot respectively. It is expected as their BCRT and WCRT differences are bigger on TP-JOP. 
+This also led to greater increase in minimum upper real-time bounds \emph{N} for TP-JOP in every example.
+In AECS, there are two clock-domains using delay constructs that each has their own BCRT and WCRT. Again, individual BCRT and WCRT is smaller 
+whereas \emph{d} is bigger on TP-JOP. One important property to note here, is that BCRT and WCRT of any clock-domains are invariant of \emph{d} as explained 
+in Section \ref{sec:intr-real-time}. It is our fundamental assumption to find \emph{d} in our benchmark programs.
 

### 2013-06-26 exps.tex
graph added need to write something

--- 2013-06-18 background.tex
@@
-\section{Background}
-\label{sec:background}
+\section{Experimental results}
+\label{sec:experimental-results}
+% \begin{table}[t]
+% \centering
+%\twocolumn[
+%\begin{@twocolumnfalse}
+%\caption{Benchmark programs using \texttt{delay} constructs} \label{table:benchmark}
+%\scalebox{0.85}{
+%\begin{tabular}{|c| c | c | c | c | c | c | c |}
+%	\cline{4-8}
+% 	\multicolumn{3}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
+% 	\multirow{12}{*}{JOP} 		& \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{2-8}   
+% 								& \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{2-8} 
+% 								& \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{3-8} 
+% 								&			   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{2-8}
+% 								& \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{3-8} 
+% 								&			   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{2-8}
+% 								& \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
+% 								&			   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
+% 								& \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
+% 								&			   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
+% 								& \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
+% 								&			   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
+% 	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}   
+% 								& \multicolumn{2}{|c|}{WCRT} 		&0.0329 ms 		& 0.0393 ms		  & 0.0361 ms		  & 0.5072 ms		   & 0.8388 ms\\ \cline{2-8} 
+% 								& \multirow{2}{*}{Delay 1}	 & M..N &50.3-600.3 ms 	& 2.4-15.6265 ms  & 1230-42317.8276 ms& 10000-95358.8578 ms& 10000-287757.9834 ms \\ \cline{3-8} 
+% 								&			   & d			 		&18209-18232 	& 398 			  & 1171429			  & 188015		       &343054\\ \cline{2-8}
+% 								& \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-10.8757 ms& N/A				  & N/A 			   &10000-287757.9834 ms\\ \cline{3-8} 
+% 								&			   & d			 		& 	N/A	 		& 277			  & N/A				  & N/A       		   &343054\\ \cline{2-8}
+% 								& \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.3534 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
+% 								&			   & d			 		& 	N/A	 		& 9		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
+% 								& \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.9631 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
+% 								&			   & d				 	& 	N/A	 		& 50			  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
+% 								& \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-4.79 ms   & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
+% 								&			   & d			 		& 	N/A	 		& 122 			  &	N/A				  & N/A      		   &N/A\\ \cline{2-8}
+% 	\hline
+% \end{tabular}
+% }
+% \end{@twocolumnfalse}
+% ]
+% \end{table}
+\begin{figure}
+\centering
+	\includegraphics[trim=2cm 2cm 2cm 2cm,scale=0.5]{table2}
+\caption{Evaluation of logical delays \emph{d} for on various benchmark programs}\label{fig:comparison}
+\end{figure}
+%
+In this section, we present a set of SystemJ benchmark programs in which
+real-time requirements should be met.
+We have carried out a set of experiments to obtain BCRT and WCRT of the
+benchmark programs on the execution platform called Java Optimized
+Processor(JOP) \cite{jop:jnl:jsa2007}. JOP is a hardware implementation of the
+JVM written in VHDL which enables real-time execution of Java programs by
+translating Java bytecodes into a sequence of JOP's native instructions called
+\emph{microcode} which is time-predictable. As SystemJ's default compilation
+target is Java source code, JOP is an excellent platform which enables us to
+analyze timing properties of SystemJ program.
+
+% On the other hand, there is also an option where compiled code could be more
+% tightly coupled to specific platforms such as TP-JOP, which execute SystemJ's
+% kernel statements more efficiently.
+
+By utilizing in-house tools in conjunction with Worst-Case-Execution-Time (WCET)
+analyzer \cite{jop:jnl:jsa2007} provided by JOP tools, we were able to estimate
+BCRT and WCRT of the following benchmark programs: (full-name?) HRTCS \cite{},
+Stepper motor controller (Motor) \cite{}, Robot and Access Environment Control
+System (AECS) \cite{}. We have also added a set of customized hardware counters
+to the original JOP design in order to collect clock cycles required for each
+logical tick of given SystemJ program. All examples except Robot are existing
+applications translated into pure SystemJ programs. The Robot example models a
+combination of robotic arms and conveyors which splits goods based on their
+volumetric sizes.
+
+As one can see in Figure \ref{fig:comparison},
+
+% HRTCS, as already illustrated in Section \ref{sec:intr-motiv}, tests for an
+% ability of one's responsiveness by generating green lights between the
+% real-time range \emph{M - N}. For AECS \ref{} we have replaced the external
+% timer with our delay constructs. Stepper motor controller is originally
+% introduced in \cite{Bourke2009a} which converted into SystemJ.
+
+% As one can see in Figure \ref{fig:comparison}, BCRT and WCRT of overall programs
+% are smaller when they run on TP-JOP. For example, WCRT and BCRT of HRTCS are
+% \(\times\)11.9 and \(\times\)3.4 smaller, respectively, for TP-JOP (0.0028,
+% 0.0329 ms) compared to JOP (0.0334, 0.1120 ms). On the other hand, required
+% logical delays \emph{d} are generally bigger on TP-JOP e.g. \(\times\)4.5 -
+% \(\times\)5.9 and \(\times\)30.9 in Motor and Robot respectively. It is expected
+% as their BCRT and WCRT differences are bigger on TP-JOP.
+% This also led to greater increase in minimum upper real-time bounds \emph{N} for
+% TP-JOP in every example.
+% In AECS, there are two clock-domains using delay constructs that each has their
+% own BCRT and WCRT. Again, individual BCRT and WCRT is smaller whereas \emph{d}
+% is bigger on TP-JOP. One important property to note here, is that BCRT and WCRT
+% of any clock-domains are invariant of \emph{d} as explained in Section
+% \ref{sec:intr-real-time}. It is our fundamental assumption to find \emph{d} in
+% our benchmark programs.
 

### 2013-06-26 exps.tex
little change

@@
 System (AECS) \cite{}. We have also added a set of customized hardware counters
-to the original JOP design in order to collect clock cycles required for each
-logical tick of given SystemJ program. All examples except Robot are existing
+to the original JOP design in order to collect numbers of clock cycles required for 
+logical ticks of given SystemJ program. All examples except Robot are existing
 applications translated into pure SystemJ programs. The Robot example models a

### 2013-06-26 exps.tex
texts added

@@
 System (AECS) \cite{}. We have also added a set of customized hardware counters
-to the original JOP design in order to collect numbers of clock cycles required for 
-logical ticks of given SystemJ program. All examples except Robot are existing
-applications translated into pure SystemJ programs. The Robot example models a
-combination of robotic arms and conveyors which splits goods based on their
-volumetric sizes.
+to the original JOP design in order to collect numbers of clock cycles required
+for logical ticks of given SystemJ program. All examples except Robot are
+existing applications translated into pure SystemJ programs. The Robot example
+models a combination of robotic arms and conveyors which splits goods based on
+their volumetric sizes. Finally, garbage collector (GC) was disabled in all
+experiments to avoid unwanted side-effects such as lengthening clock-domain
+logical tick times. Whenever the GC triggers, experiment terminates. BCRT and
+WCRT were then calculated based on data collected until that point.
 
-As one can see in Figure \ref{fig:comparison},
+As one can see in Figure \ref{fig:comparison}, BCRT and WCRT were calculated for
+all examples. Given the real-time ranges \mbox{\texttt{delay [M..N]}} provided
+by programmers, required logical delays \emph{d} were determined by applying the
+Algorithm \ref{alg:1}. In HRTCS, real-time delay of the range 50.3..200.3 (ms)
+is only satisfied when \emph{d} is between 1506 and 1787 (not shown in Figure).
+It is then compiler's job to statically choose one of the values between this
+range.
+
+However, it is not the case that given real-time delays in a particular
+application are always realizable on JOP. For instance, all other examples other
+than HRTCS are required to relax the upper bound for a given real-time delay in
+order to obtain a feasible number of \emph{d}. Therefore, Algorithm \ref{alg:2}
+is applied to those examples to determine the minimum upper real-time bound
+\emph{N} such that there exist a solution for \emph{D} in Algorithm \ref{alg:1}.
+Unlike in HRTCS, however, this algorithm gives a unique number for logical tick
+delay \emph{d} such that \emph{d} = Min(\emph{$S_2$}) = Max(\emph{$S_1$}).
+The compiler, therefore, will choose \emph{d} based on the newly computed \emph{N} for
+the examples Robot, Motor and AECS.
 

### 2013-06-26 exps.tex
updating citation

@@
 BCRT and WCRT of the following benchmark programs: (full-name?) HRTCS \cite{},
-Stepper motor controller (Motor) \cite{}, Robot and Access Environment Control
-System (AECS) \cite{}. We have also added a set of customized hardware counters
+Stepper motor controller (Motor) \cite{Bourke2009a}, Robot and Access Environment Control
+System (AECS) \cite{aecs_ispa}. We have also added a set of customized hardware counters
 to the original JOP design in order to collect numbers of clock cycles required
@@
 delay \emph{d} such that \emph{d} = Min(\emph{$S_2$}) = Max(\emph{$S_1$}).
-The compiler, therefore, will choose \emph{d} based on the newly computed \emph{N} for
-the examples Robot, Motor and AECS.
+The compiler, therefore, will choose \emph{d} based on the newly computed
+\emph{N} for the examples Robot, Motor and AECS.
 

### 2013-06-27 exps.tex
My changes

--- 2013-06-26 exps.tex
@@
 \label{sec:experimental-results}
-% \begin{table}[t]
-% \centering
-%\twocolumn[
-%\begin{@twocolumnfalse}
-%\caption{Benchmark programs using \texttt{delay} constructs} \label{table:benchmark}
-%\scalebox{0.85}{
-%\begin{tabular}{|c| c | c | c | c | c | c | c |}
-%	\cline{4-8}
-% 	\multicolumn{3}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
-% 	\multirow{12}{*}{JOP} 		& \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{2-8}   
-% 								& \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{2-8} 
-% 								& \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{3-8} 
-% 								&			   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{3-8} 
-% 								&			   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
-% 								&			   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
-% 								&			   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
-% 								&			   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
-% 	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}   
-% 								& \multicolumn{2}{|c|}{WCRT} 		&0.0329 ms 		& 0.0393 ms		  & 0.0361 ms		  & 0.5072 ms		   & 0.8388 ms\\ \cline{2-8} 
-% 								& \multirow{2}{*}{Delay 1}	 & M..N &50.3-600.3 ms 	& 2.4-15.6265 ms  & 1230-42317.8276 ms& 10000-95358.8578 ms& 10000-287757.9834 ms \\ \cline{3-8} 
-% 								&			   & d			 		&18209-18232 	& 398 			  & 1171429			  & 188015		       &343054\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-10.8757 ms& N/A				  & N/A 			   &10000-287757.9834 ms\\ \cline{3-8} 
-% 								&			   & d			 		& 	N/A	 		& 277			  & N/A				  & N/A       		   &343054\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.3534 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
-% 								&			   & d			 		& 	N/A	 		& 9		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.9631 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
-% 								&			   & d				 	& 	N/A	 		& 50			  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-4.79 ms   & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
-% 								&			   & d			 		& 	N/A	 		& 122 			  &	N/A				  & N/A      		   &N/A\\ \cline{2-8}
-% 	\hline
-% \end{tabular}
-% }
-% \end{@twocolumnfalse}
-% ]
-% \end{table}
-\begin{figure}
+\begin{table}[t]
   \centering
-	\includegraphics[trim=2cm 2cm 2cm 2cm,scale=0.5]{table2}
-\caption{Evaluation of logical delays \emph{d} for on various benchmark programs}\label{fig:comparison}
-\end{figure}
-%
+  \twocolumn[
+
+  \begin{@twocolumnfalse}
+    \caption{Benchmark programs using \texttt{delay} constructs} \label{table:benchmark}
+    \scalebox{0.85}{
+      \begin{tabular}{|c| c | c | c | c | c | c | c |}
+	\cline{4-8}
+	\multicolumn{3}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
+	\multirow{12}{*}{JOP} 		& \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{2-8}   
+        & \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{2-8} 
+        & \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{3-8} 
+        &			   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{2-8}
+        & \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{3-8} 
+        &			   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{2-8}
+        & \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
+        &			   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
+        & \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
+        &			   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
+        & \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
+        &			   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
+	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}   
+        & \multicolumn{2}{|c|}{WCRT} 		&0.0329 ms 		& 0.0393 ms		  & 0.0361 ms		  & 0.5072 ms		   & 0.8388 ms\\ \cline{2-8} 
+        & \multirow{2}{*}{Delay 1}	 & M..N &50.3-600.3 ms 	& 2.4-15.6265 ms  & 1230-42317.8276 ms& 10000-95358.8578 ms& 10000-287757.9834 ms \\ \cline{3-8} 
+        &			   & d			 		&18209-18232 	& 398 			  & 1171429			  & 188015		       &343054\\ \cline{2-8}
+        & \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-10.8757 ms& N/A				  & N/A 			   &10000-287757.9834 ms\\ \cline{3-8} 
+        &			   & d			 		& 	N/A	 		& 277			  & N/A				  & N/A       		   &343054\\ \cline{2-8}
+        & \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.3534 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
+        &			   & d			 		& 	N/A	 		& 9		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
+        & \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.9631 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
+        &			   & d				 	& 	N/A	 		& 50			  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
+        & \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-4.79 ms   & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
+        &			   & d			 		& 	N/A	 		& 122 			  &	N/A				  & N/A      		   &N/A\\ \cline{2-8}
+	\hline
+      \end{tabular}
+    }
+  \end{@twocolumnfalse}
+  ]
+
+\end{table}
 In this section, we present a set of SystemJ benchmark programs in which
-real-time requirements should be met.
-We have carried out a set of experiments to obtain BCRT and WCRT of the
-benchmark programs on the execution platform called Java Optimized
-Processor(JOP) \cite{jop:jnl:jsa2007}. JOP is a hardware implementation of the
-JVM written in VHDL which enables real-time execution of Java programs by
-translating Java bytecodes into a sequence of JOP's native instructions called
-\emph{microcode} which is time-predictable. As SystemJ's default compilation
-target is Java source code, JOP is an excellent platform which enables us to
-analyze timing properties of SystemJ program.
+real-time requirements should be met.  We have carried out a set of
+experiments to obtain BCRT and WCRT of the benchmark programs on two
+execution platforms called Java Optimized Processor(JOP)
+\cite{jop:jnl:jsa2007} and TP-JOP \cite{6119095}. JOP is a hardware
+implementation of the JVM which enables real-time execution of Java
+programs by translating Java bytecodes into a sequence of JOP's native
+instructions called \emph{microcode} which is time-predictable. As
+SystemJ's default compilation target is Java source code, JOP is an
+excellent platform which enables us to analyze timing properties of
+SystemJ program. On the other hand, there is also an option where
+compiled code could be more tightly coupled to specific platforms such
+as TP-JOP, which execute SystemJ's kernel statements more
+efficiently. By utilizing our internal tools in conjunction with
+Worst-Case-Execution-Time (WCET) analyzer \cite{jop:jnl:jsa2007}
+provided by JOP tools, we were able to estimate BCRT and WCRT of our
+benchmark programs: (full-name?) HRTCS, Stepper motor controller (Motor)
+\cite{}, Robot and Access Environment Control System (AECS) \cite{}.
 
-% On the other hand, there is also an option where compiled code could be more
-% tightly coupled to specific platforms such as TP-JOP, which execute SystemJ's
-% kernel statements more efficiently.
+% HRTCS, as already illustrated in Section \ref{sec:intr-motiv}, tests
+% for an ability of one's responsiveness by generating green lights
+% between the real-time range \emph{M - N}. For AECS \ref{} we have
+% replaced the external timer with our delay constructs. Stepper motor
+% controller is originally introduced in \cite{Bourke2009a} which
+% converted into SystemJ.
 
-By utilizing in-house tools in conjunction with Worst-Case-Execution-Time (WCET)
-analyzer \cite{jop:jnl:jsa2007} provided by JOP tools, we were able to estimate
-BCRT and WCRT of the following benchmark programs: (full-name?) HRTCS \cite{},
-Stepper motor controller (Motor) \cite{}, Robot and Access Environment Control
-System (AECS) \cite{}. We have also added a set of customized hardware counters
-to the original JOP design in order to collect numbers of clock cycles required for 
-logical ticks of given SystemJ program. All examples except Robot are existing
-applications translated into pure SystemJ programs. The Robot example models a
-combination of robotic arms and conveyors which splits goods based on their
-volumetric sizes.
-
-As one can see in Figure \ref{fig:comparison},
-
-% HRTCS, as already illustrated in Section \ref{sec:intr-motiv}, tests for an
-% ability of one's responsiveness by generating green lights between the
-% real-time range \emph{M - N}. For AECS \ref{} we have replaced the external
-% timer with our delay constructs. Stepper motor controller is originally
-% introduced in \cite{Bourke2009a} which converted into SystemJ.
-
-% As one can see in Figure \ref{fig:comparison}, BCRT and WCRT of overall programs
-% are smaller when they run on TP-JOP. For example, WCRT and BCRT of HRTCS are
-% \(\times\)11.9 and \(\times\)3.4 smaller, respectively, for TP-JOP (0.0028,
-% 0.0329 ms) compared to JOP (0.0334, 0.1120 ms). On the other hand, required
-% logical delays \emph{d} are generally bigger on TP-JOP e.g. \(\times\)4.5 -
-% \(\times\)5.9 and \(\times\)30.9 in Motor and Robot respectively. It is expected
-% as their BCRT and WCRT differences are bigger on TP-JOP.
-% This also led to greater increase in minimum upper real-time bounds \emph{N} for
-% TP-JOP in every example.
-% In AECS, there are two clock-domains using delay constructs that each has their
-% own BCRT and WCRT. Again, individual BCRT and WCRT is smaller whereas \emph{d}
-% is bigger on TP-JOP. One important property to note here, is that BCRT and WCRT
-% of any clock-domains are invariant of \emph{d} as explained in Section
-% \ref{sec:intr-real-time}. It is our fundamental assumption to find \emph{d} in
-% our benchmark programs.
+As one can see in Table \ref{table:benchmark}, BCRT and WCRT of overall
+programs are smaller when they run on TP-JOP. For example, WCRT and BCRT
+of HRTCS are \(\times\)11.9 and \(\times\)3.4 smaller, respectively, for
+TP-JOP (0.0028, 0.0329 ms) compared to JOP (0.0334, 0.1120 ms). On the
+other hand, required logical delays \emph{d} are generally bigger on
+TP-JOP e.g. \(\times\)4.5 - \(\times\)5.9 and \(\times\)30.9 in Motor
+and Robot respectively. It is expected as their BCRT and WCRT
+differences are bigger on TP-JOP.  This also led to greater increase in
+minimum upper real-time bounds \emph{N} for TP-JOP in every example.  In
+AECS, there are two clock-domains using delay constructs that each has
+their own BCRT and WCRT. Again, individual BCRT and WCRT is smaller
+whereas \emph{d} is bigger on TP-JOP. One important property to note
+here, is that BCRT and WCRT of any clock-domains are invariant of
+\emph{d} as explained in Section \ref{sec:intr-real-time}. It is our
+fundamental assumption to find \emph{d} in our benchmark programs.
 

### 2013-07-01 background.tex
My changes
Included all changes by Prof. Salcic

@@
 
-\begin{figure*}[bth]
-\centering
-\begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
-\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
-\begin{verbatim}
-while(true){
- abort(A){
-  while(true)
-   pause;
-  emit O;
-  // do java computation
-}
-\end{verbatim}%
-\end{minipage}%
-\end{SubFloat}
-\hspace{1.5cm}%
-\begin{SubFloat}[Black box]{\label{fig:2b}Logical time in SystemJ}%
-\includegraphics[scale=0.4]{moc}
-\end{SubFloat}%
-\begin{SubFloat}{\label{fig:2c}Mapping to physical time}%
-% \includegraphics[scale=0.4]{phy.pdf_t}
-\scalebox{0.4}{\input{phy.pdf_t}}
-\end{SubFloat}%
-\caption{Simple SystemJ example and corresponding MoC}
-\label{fig:2}
-\end{figure*}
-
 SystemJ consists of entities called clock-domains (CD)s, each running at
 their individual pace, interacting with the environment via signals and
-with each other via channels and CSP~\cite{pwel05} style
-rendezvous. Each CD adheres to the \textit{perfect synchrony}
-hypothesis, i.e., all statements execute instantaneously in zero
-time. Only the \texttt{pause} statement consumes time, just like in
-Esterel~\cite{gber931}. Consider the SystemJ program in
-Figure~\ref{fig:2a}, it is waiting for an input signal producing logical
-ticks (see Figure~\ref{fig:2b}). Once \texttt{A} is received at tick 4,
-output \texttt{O} is emitted to the environment instantaneously.
+with each other via channels and CSP~\cite{choa85} style
+rendezvous. Each CD has its own notion of logical time. CDs adheres to
+the \textit{perfect synchrony} hypothesis, i.e., all statements execute
+instantaneously in zero time. Only the \texttt{pause} statement consumes
+time, just like in Esterel~\cite{gber931}. Consider the SystemJ program
+in Figure~\ref{fig:2a}, it is waiting for an input signal producing
+logical ticks (see Figure~\ref{fig:2c}). Once \texttt{A} is received at
+tick 1 or 4, output \texttt{O} is emitted to the environment
+instantaneously.
 
@@
 
+\begin{figure}[t!]
+% \centering
+\begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
+\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
+  \begin{scriptsize}
+\begin{verbatim}
+{
+int counter = 0;
+while(true){
+ abort(A){
+  while(true) pause; 
+  emit O;
+  if(counter != 0) { // do java computation leading to WCRT}
+  ++counter;
+  } // abort end
+ } // loop end
+} // SystemJ CD end
+\end{verbatim}%
+  \end{scriptsize}
+\end{minipage}%
+\end{SubFloat}
+\hspace{1.5cm}%
+% \begin{SubFloat}[Black box]{\label{fig:2b}Logical time in SystemJ}%
+% \includegraphics[scale=0.4]{moc}
+% \end{SubFloat}%
+% \vspace{-10pt}
+
+\begin{SubFloat}{\label{fig:2c}Mapping to physical time}%
+% \includegraphics[scale=0.4]{phy.pdf_t}
+\scalebox{0.68}{\input{phy.pdf_t}}
+\end{SubFloat}%
+\caption{Simple SystemJ example and corresponding MoC}
+\label{fig:2}
+\end{figure}
+
 The perfect synchrony hypothesis is ideal for programming the
-synchronous sub-set of SystemJ. But, the real-world is not that
-forgiving, execution of every \textit{micro-step} in the logical zero
-time, requires $\delta$ physical time. Time for a complete reaction to
-one or more input signal can thus be summarized as $\Delta = \Sigma
-(\delta)$. We call $\Delta$ the reaction time. Obviously, depending upon
-the amount of computation required, $\Delta$ can vary, hence, in order
-to satisfy the implicit restriction imposed by the synchrony hypothesis,
-no input event can be missed, one needs to calculate the \textit{Worst
-  Case Reaction Time} (WCRT) and the resultant WCRT needs to be smaller
-than the speed of the fastest input event, else the synchrony hypothesis
-is violated. A substantial amount of research for calculating the WCRT
-of synchronous programs~\cite{proop10,boldt07,wilhelm08} exists. We
-refer the reader to any of these research articles, since WCRT analysis
-is not the focus of this paper. The opposite of the WCRT is the
-\textit{Best Case Reaction Time} (BCRT), we denote the WCRT and BCRT in
-Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
-respectively. Figure~\ref{fig:2c} shows a continuously running physical
-clock (analogous to a clock in hardware), the numerical annotations on
-the rising edge of the clock mark the logical ticks from
-Figure~\ref{fig:2b}. The subscripts \texttt{s} and \texttt{e} for the
-numbers show the start and end of the logical ticks, respectively. The
-end of a logical tick and the start of the next logical tick happen
-together. Figure~\ref{fig:2c} clearly shows the mapping of the logical
-time to the physical time. As we can see due to Java computations
-(Figure~\ref{fig:2a}) the physical time between logical ticks 4 and 5 is
-2 clock cycles, while for the rest of the ticks it is just 1 clock
-cycle. This elasticity is inherent (and elegant) to both Esterel and
-SystemJ.
+synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
+requires real-time to compute. Formally, execution of every micro-step
+in the logical zero time, requires $\delta$ physical time. Time for a
+logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
+where $\mathcal{F}$ is some function dependent upon the schedule. We
+call $\Delta$ the reaction time. Obviously, depending upon the amount of
+computation required and schedule, $\Delta$ can vary, hence, in order to
+satisfy the implicit restriction imposed by the synchrony hypothesis --
+no input event \textit{from the environment} can be missed -- one needs
+to calculate the \textit{Worst Case Reaction Time} (WCRT) and the
+resultant WCRT needs to be smaller than the time between any two
+consecutive events on any input, else the synchrony hypothesis is
+violated. Techniques for calculating the WCRT of synchronous
+programs~\cite{boldt07} exist and this analysis is not the focus of this
+paper. The opposite of the WCRT is the \textit{Best Case Reaction Time}
+(BCRT), we denote the WCRT and BCRT in Figure~\ref{fig:2c} using
+\texttt{W} and \texttt{B}, respectively. Figure~\ref{fig:2c} shows a
+continuously running physical clock (analogous to a clock in digital
+hardware), the numerical annotations on the rising edge of the clock
+mark the logical ticks from Figure~\ref{fig:2c}. The subscripts
+\texttt{s} and \texttt{e} for the numbers show the start and end of the
+logical ticks, respectively. The end of a logical tick and the start of
+the next logical tick happen together. Figure~\ref{fig:2c} shows the
+mapping of the logical time to the physical time. For example, the
+logical tick \texttt{1} starts at the first rising edge and completes at
+the second rising edge of the physical clock. Where as logical tick
+\texttt{4} starts at the $4^{th}$ rising edge, but finishes at the
+$6^{th}$ rising edge of the physical clock. This difference can be
+accounted for by the java computations required when signal \texttt{A}
+is input by the environment (see Figure~\ref{fig:2a}). This elasticity
+is inherent (and elegant) to both Esterel and SystemJ.
 

### 2013-07-01 conclusion.tex
first commit
Included all changes by Prof. Salcic

--- 2013-06-18 background.tex
@@
-\section{Background}
-\label{sec:background}
+\section{Conclusion}
+\label{sec:concl-future-work}
+
+In this paper we have described a novel way to introduce real-time in
+\textit{Globally Asynchronous Locally Synchronous} (GALS) languages, in
+particular SystemJ, and their subset the synchronous languages like
+Esterel. The fundamental idea is to convert real-time delay into logical
+delays (\texttt{pause} construct) that interact well with rest of the
+constructs in these languages, especially, preemption and channel
+communication. We do \textit{not} need or use external timers to
+introduce these real-time delays, thereby resolving the problems of
+interaction between external timers and preemption. Moreover, we also do
+not require specific timer resolutions, thereby making our solution
+elegant (in the spirit of GALS/synchronous programming) and flexible. We
+rewrite the delay construct into reactive kernel constructs, calculate
+the \textit{Worst} and \textit{Base} case reaction times and then
+determine \textit{delay} \texttt{d}, number of logical ticks, which is
+invariant to the \textit{elasticity} of logical time. To our knowledge
+we are the first to introduce non-deterministic delays in such languages
+and show their usefulness in designing real-world systems.
 

### 2013-07-01 delays.tex
Included all changes by Prof. Salcic

@@
-\section{Introducing real time in SystemJ}
+\section{Introducing real-time in SystemJ}
 \label{sec:intr-real-time}
 
-We introduce a single \textit{derived} construct called
-\mbox{\texttt{delay [M..N]}} in the SystemJ language for real time
-control. The resolution of the delay statement is of secondary concern
-and is dependent upon the execution platform, without loss of generality
-through out this paper we consider the resolution to be always in milli
-seconds.
+We introduce a single \textit{derived} (consisting of kernel constructs)
+statement called \mbox{\texttt{delay (M..N)}} in the SystemJ language
+for real-time control. The resolution of the delay statement is of
+secondary concern and is dependent upon the execution platform.
 
@@
 
-Given a SystemJ program: \texttt{delay [M..N]; p}, where $M \in
+Given a SystemJ program: \texttt{delay (M..N); p}, where $M \in
 \mathbb{Q}^{>0}$ and $N \in \mathbb{Q}^{>0}$, statement $p$ is executed
-after real-time $\tau$, such that, $M \leq \tau \leq N$.
+after real-time delay of $\tau$ units, such that, $M \leq \tau \leq N$.
 
@@
 \begin{enumerate*}
-\item Given a SystemJ program \texttt{delay M; p}, where $M \in
-  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time $\tau$,
-  such that, $M \leq \tau \leq \infty$. It is important to note that the
-  lower bound of the delay construct $M$ is \textit{not} an exact delay,
-  but rather the control is allowed to proceed to the next statement
-  anytime after the delay of time $M$.
-\item Give a SystemJ program \texttt{delay [M..M]; p}, where $M \in
-  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time $\tau$,
-  such that, $M \leq \tau \leq M$. In this variant, the delay is exact.
+\item Given a SystemJ program \texttt{delay (M); p}, where $M \in
+  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay of
+  $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
+  note that the lower bound of the delay construct $M$ is \textit{not}
+  an exact delay, but rather the control is allowed to proceed to the
+  next statement anytime after the delay of time $M$.
+\item Given a SystemJ program \texttt{delay (M..M); p}, where $M \in
+  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay of
+  $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
+  delay is exact.
 \end{enumerate*}
 
-The aforementioned variants are specializations of the general case
-\texttt{delay [M..N]}.
+The aforementioned variants are special cases of the general case
+\texttt{delay (M..N)}.
 
@@
   x = x + 1;
+  pause;
   if(x == d) exit (T); //wait for "d" ticks
-  pause;
- }
-}
+ } // d is the number of logical ticks calculated
+} // by Algorithm 1
 \end{verbatim}
@@
   reaction time for each logical tick is elastic -- varying only between
-  the BCRT and the WCRT, thus any logical number of ticks \texttt{d}
+  the BCRT and the WCRT, thus any number of logical ticks \texttt{d}
   that map to the required real-time \texttt{delay} should be chosen in
@@
 these functions guarantee that the resultant logical ticks result in
-real-time delays between the required range $M-N$. Next, a linear
-function $F$ maps these calculated values to a set of equidistant
-integer points (values) separated by a unit value -- these points
-represent all the logical ticks running at the WCRT and the BCRT,
-respectively that satisfy the real-time delay requirements. The
-intersection of these two sets gives all the logical ticks that satisfy
-the real-time requirements invariant of the reaction time and its
-elasticity.
+real-time delays between the required range $M-N$. Next, a function $F$
+maps these calculated values to a set of equidistant integer points
+(values) separated by a unit value -- these points represent all the
+logical ticks running at the WCRT and the BCRT, respectively that
+satisfy the real-time delay requirements. The intersection of these two
+sets gives all the logical ticks that satisfy the real-time requirements
+invariant of the reaction time and its elasticity.
 
-Let us revisit our motivating example to elucidate the algorithm. From
-Figure~\ref{fig:1} we know that $M$ is 50ms and $N$ is 100ms,
-respectively. Let the WCRT and BCRT be: 13.332ms and 10ms,
+Let us revisit our motivating example to illustrate the algorithm. From
+Figure~\ref{fig:1} we know that $M$ is 50.3ms and $N$ is 200.3ms,
+respectively. Let the WCRT and BCRT be: 0.112ms and 0.0334ms,
 respectively. Thus, the algorithm proceeds as follows:
@@
 \begin{enumerate*}
-\item $l_1 \leftarrow \lceil 50/13.332 \rceil$ and $u_1 \leftarrow
-  \lfloor 100/13.332 \rfloor$. $l_1 = 4$ and $u_1 = 7$. We first
+\item $l_1 \leftarrow \lceil 50.3/0.112 \rceil$ and $u_1 \leftarrow
+  \lfloor 200.3/0.112 \rfloor$. $l_1 = 450$ and $u_1 = 1788$. We first
   calculate the logical ticks that are always running at the WCRT
   satisfy the required real-time delay.
-\item $l_2 \leftarrow \lceil 50/10 \rceil$ and $u_2 \leftarrow
-  \lfloor 100/10 \rfloor$. $l_1 = 5$ and $u_1 = 10$. We do the same for
-  the BCRT case.
-\item $S_1 = \{4,5,6,7\}$ and $S_2 =\{5,6,7,8,9,10\}$. We then map the
+\item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
+  \lfloor 200.3/0.0334 \rfloor$. $l_1 = 1506$ and $u_1 = 5997$. We do
+  the same for the BCRT case.
+\item $S_1 = [450..1788]$ and $S_2 =[1506,5997]$. We then map the
   resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
@@
   satisfy the required real-time constraints.
-\item $D = S_1 \cap S_2$, $D = \{5,6,7\}$. Finally, the intersection of
-  the two sets gives the set $D$ from which we can pick up any value for
-  \texttt{d}.
+\item $D = S_1 \cap S_2$, $D = [1506,1788]$. Finally, the intersection
+  of the two sets gives the set $D$ from which we can pick up any value
+  for \texttt{d}.
 \end{enumerate*}
@@
 which can run at any physical clock-speed, bounded by the BCRT and the
-WCRT of the program and still result in the required real-time delay. We
+WCRT of the program and still result in the desired real-time delay. We
 think this is an elegant solution, because the technique provides
@@
 languages. Moreover, this technique considers non maximal-parallelism,
-i.e., the delays in logical ticks is calculated after scheduling has
-been performed. To our knowledge we are the first to do so.
+i.e., the delays in logical ticks is calculated after scheduling of CDs
+or synchronous programs has been performed. To our knowledge we are the
+first to do so.
 
@@
 
-A number of programs require non-deterministic timing constructs. One of
-them is our motivating example -- the human response time
-system. Another is a printer-spooler example borrowed from timed
-CSP~\cite{Schneider:1999:CRT:555233}. The spooler and printer need to
-synchronize using channels. The printer might be unable to print
-depending upon the paper tray, similarly, spooler might take sometime to
-send the job depending upon its size. Such real-time constraints can be
-modeled in SystemJ as below:
+Many real-world systems require non-deterministic timing constructs. One
+of them is our motivating example -- the human response time
+system.
 
-\begin{scriptsize}
+% Another is a printer-spooler example borrowed from timed
+% CSP~\cite{Schneider:1999:CRT:555233}. The spooler and printer need to
+% synchronize using channels. The printer might be unable to print
+% depending upon the paper tray, similarly, spooler might take sometime
+% to send the job depending upon its size. Such real-time constraints
+% can be modeled in SystemJ as below:
 
-\begin{verbatim}
-{await (job); delay([2..10]ms);  //SPOOLER CD
- send mid(job) delay (1ms);}
-><
-{receive mid; delay([1..30]ms); //PRINTER CD
- emit print(#mid);}
-\end{verbatim}
-\end{scriptsize}
+% \begin{scriptsize}
+
+% \begin{verbatim}
+% {await (job); delay(2..10 ms);  //SPOOLER CD
+%  send mid(job) delay (1ms);}
+% ><
+% {receive mid; delay(1..30 ms); //PRINTER CD
+%  emit print(#mid);}
+% \end{verbatim}
+% \end{scriptsize}
 
@@
 as:
-    \begin{scriptsize}
+% \begin{scriptsize}
 \begin{verbatim}
 // timeout after 1ms.
-trap(T){
- {await(A);}||{delay(1ms);exit(T); }
-}
+trap(T){{await(A);}||{delay((1..1)ms);exit(T);}}
 \end{verbatim}
-    \end{scriptsize}
+% \end{scriptsize}
 
@@
 
-\begin{scriptsize}
-
+% \begin{scriptsize}
 \begin{verbatim}
-{
-delay (1ms); emit S; //do something
-}
+while(true) { delay ((1..1)ms); emit S; 
+ //do something }
 \end{verbatim}
-\end{scriptsize}
+% \end{scriptsize}
 A periodic reaction (or CD) requires special consideration. Since the
@@
 not introduce extra \texttt{pause} constructs when building periodic
-reactions (or CDs). This is essential since, introducing pauses would
-introduce more logical ticks.
+reactions (or CDs). This is essential since introducing additional
+pauses would introduce more logical ticks.
 
@@
 carefully consider the interplay of \texttt{delay} semantics with the
-preemption semantics of reactive languages. Other attempts at
+preemption semantics of reactive languages. Previous attempts at
 incorporating delays (using external timers) have only been partially
@@
 
-\begin{scriptsize}
-
+% \begin{scriptsize}
 \begin{verbatim}
-suspend{S} {
- emit START_TIMER(10); await (TIMER); 
- emit O1;
-}
+suspend{S} {emit START_TIMER(10); await (TIMER); 
+ emit O1;}
 \end{verbatim}
-\end{scriptsize}
+% \end{scriptsize}
 
@@
 external timer and waits for 10ms to pass by. Consider what happens when
-signals \texttt{S} and \texttt{TIMER} occur together, the \texttt{await}
-statement is never executed and hence, we enter a deadlock. Such
-problems are completely avoided in our technique, because we convert the
-real-time delays into logical delays (\texttt{pause} constructs), which
-bode well with preemption.
+signals \texttt{S} and \texttt{TIMER} occur in the same logical tick,
+the \texttt{await} statement is never executed and hence, we enter a
+deadlock. Such problems are completely avoided in our technique, because
+we convert the real-time delays into logical delays (\texttt{pause}
+constructs), which interact well with preemption.
 
@@
 
-Channels are an addition in the SystemJ language. Like interaction of
-preemption and delays, conversion of rela-time delays to logical ticks
-also bodes well with channel rendezvous, because the semantics of
-interaction are well defined~\cite{amal10}. More importantly, we need to
-consider the interplay of channel communication and WCRT/BCRT
+Channels, used for communication between reactions in asynchronously
+running CDs, are an addition in the SystemJ language. Like interaction
+of preemption and delays, conversion of rela-time delays to logical
+ticks also interacts well with channel rendezvous, because the semantics
+of interaction are well defined~\cite{amal10}. More importantly, we need
+to consider the interplay of channel communication and WCRT/BCRT
 analysis. Since, channel communication does not stop logical time
 (see~\cite{amal10}) WCRT/BCRT are unaffected by channel
-communication. The response time to input signals \textit{is} though!
-But, we are unconcerned with the response time, analysis in this paper
-and it remains a future research avenue.
+communication. % The response time to input signals \textit{is} though!
+% But, we are unconcerned with the response time, analysis in this paper
+% and it remains a future research avenue.
 

### 2013-07-01 exps.tex
Included all changes by Prof. Salcic

--- 2013-06-26 exps.tex
@@
 \label{sec:experimental-results}
+% <<<<<<< HEAD
+% \begin{table}[t]
+%   \centering
+%   \twocolumn[
+
+%   \begin{@twocolumnfalse}
+%     \caption{Benchmark programs using \texttt{delay} constructs} \label{table:benchmark}
+%     \scalebox{0.85}{
+%       \begin{tabular}{|c| c | c | c | c | c | c | c |}
+% 	\cline{4-8}
+% 	\multicolumn{3}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
+% 	\multirow{12}{*}{JOP} 		& \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{2-8}   
+%         & \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{2-8} 
+%         & \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{3-8} 
+%         &			   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{2-8}
+%         & \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{3-8} 
+%         &			   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{2-8}
+%         & \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
+%         &			   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
+%         & \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
+%         &			   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
+%         & \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
+%         &			   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
+% 	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}   
+%         & \multicolumn{2}{|c|}{WCRT} 		&0.0329 ms 		& 0.0393 ms		  & 0.0361 ms		  & 0.5072 ms		   & 0.8388 ms\\ \cline{2-8} 
+%         & \multirow{2}{*}{Delay 1}	 & M..N &50.3-600.3 ms 	& 2.4-15.6265 ms  & 1230-42317.8276 ms& 10000-95358.8578 ms& 10000-287757.9834 ms \\ \cline{3-8} 
+%         &			   & d			 		&18209-18232 	& 398 			  & 1171429			  & 188015		       &343054\\ \cline{2-8}
+%         & \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-10.8757 ms& N/A				  & N/A 			   &10000-287757.9834 ms\\ \cline{3-8} 
+%         &			   & d			 		& 	N/A	 		& 277			  & N/A				  & N/A       		   &343054\\ \cline{2-8}
+%         & \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.3534 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
+%         &			   & d			 		& 	N/A	 		& 9		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
+%         & \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.9631 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
+%         &			   & d				 	& 	N/A	 		& 50			  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
+%         & \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-4.79 ms   & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
+%         &			   & d			 		& 	N/A	 		& 122 			  &	N/A				  & N/A      		   &N/A\\ \cline{2-8}
+% 	\hline
+%       \end{tabular}
+%     }
+%   \end{@twocolumnfalse}
+%   ]
+
+% \end{table}
+% In this section, we present a set of SystemJ benchmark programs in which
+% real-time requirements should be met.  We have carried out a set of
+% experiments to obtain BCRT and WCRT of the benchmark programs on two
+% execution platforms called Java Optimized Processor(JOP)
+% \cite{jop:jnl:jsa2007} and TP-JOP \cite{6119095}. JOP is a hardware
+% implementation of the JVM which enables real-time execution of Java
+% programs by translating Java bytecodes into a sequence of JOP's native
+% instructions called \emph{microcode} which is time-predictable. As
+% SystemJ's default compilation target is Java source code, JOP is an
+% excellent platform which enables us to analyze timing properties of
+% SystemJ program. On the other hand, there is also an option where
+% compiled code could be more tightly coupled to specific platforms such
+% as TP-JOP, which execute SystemJ's kernel statements more
+% efficiently. By utilizing our internal tools in conjunction with
+% Worst-Case-Execution-Time (WCET) analyzer \cite{jop:jnl:jsa2007}
+% provided by JOP tools, we were able to estimate BCRT and WCRT of our
+% benchmark programs: (full-name?) HRTCS, Stepper motor controller (Motor)
+% \cite{}, Robot and Access Environment Control System (AECS) \cite{}.
+
+% % HRTCS, as already illustrated in Section \ref{sec:intr-motiv}, tests
+% % for an ability of one's responsiveness by generating green lights
+% % between the real-time range \emph{M - N}. For AECS \ref{} we have
+% % replaced the external timer with our delay constructs. Stepper motor
+% % controller is originally introduced in \cite{Bourke2009a} which
+% % converted into SystemJ.
+
+% As one can see in Table \ref{table:benchmark}, BCRT and WCRT of overall
+% programs are smaller when they run on TP-JOP. For example, WCRT and BCRT
+% of HRTCS are \(\times\)11.9 and \(\times\)3.4 smaller, respectively, for
+% TP-JOP (0.0028, 0.0329 ms) compared to JOP (0.0334, 0.1120 ms). On the
+% other hand, required logical delays \emph{d} are generally bigger on
+% TP-JOP e.g. \(\times\)4.5 - \(\times\)5.9 and \(\times\)30.9 in Motor
+% and Robot respectively. It is expected as their BCRT and WCRT
+% differences are bigger on TP-JOP.  This also led to greater increase in
+% minimum upper real-time bounds \emph{N} for TP-JOP in every example.  In
+% AECS, there are two clock-domains using delay constructs that each has
+% their own BCRT and WCRT. Again, individual BCRT and WCRT is smaller
+% whereas \emph{d} is bigger on TP-JOP. One important property to note
+% here, is that BCRT and WCRT of any clock-domains are invariant of
+% \emph{d} as explained in Section \ref{sec:intr-real-time}. It is our
+% fundamental assumption to find \emph{d} in our benchmark programs.
+% =======
 % \begin{table}[t]
@@
 % \end{table}
+
 \begin{figure}
@@
 	\includegraphics[trim=2cm 2cm 2cm 2cm,scale=0.5]{table2}
-\caption{Evaluation of logical delays \emph{d} for on various benchmark programs}\label{fig:comparison}
+        \caption{Evaluation of logical delays \emph{d} for on various
+          benchmark programs}\label{fig:comparison}
 \end{figure}
 %
-In this section, we present a set of SystemJ benchmark programs in which
-real-time requirements should be met.
-We have carried out a set of experiments to obtain BCRT and WCRT of the
-benchmark programs on the execution platform called Java Optimized
-Processor(JOP) \cite{jop:jnl:jsa2007}. JOP is a hardware implementation of the
-JVM written in VHDL which enables real-time execution of Java programs by
-translating Java bytecodes into a sequence of JOP's native instructions called
-\emph{microcode} which is time-predictable. As SystemJ's default compilation
-target is Java source code, JOP is an excellent platform which enables us to
-analyze timing properties of SystemJ program.
+We have carried out a number of experiments on a set of applications
+with real-time constraints. The benchmark set is shown in
+Figure~\ref{fig:comparison}. HRCTS is the human response system
+described in Figure~\ref{fig:1}. AECS (Access and Environmental Control
+System) is an application that controls an intelligent room environment
+described in~\cite{aecs_ispa}. Motor is a stepper motor controller
+from~\cite{Bourke2009a}. Robot is an industrial automation application
+that consists of two manufacturing belts and a robotic arm that splits
+goods according to their volumes. HRCTS consists of 2 CDs, but only one
+with real-time delays. AECS also has 2 CDs; CD1 consist of a single
+delay statement, while CD2 has 2 delay statements. Robot and Motor both
+are synchronous programs with 1 and 5 delay statements,
+respectively. All these examples are written in the SystemJ
+language. Finally, except for HRCTS all other examples have exact (and
+deterministic) delay requirements. These delay requirements are shown in
+Figure~\ref{fig:comparison} (\texttt{M..N}).
+
+% In this section, we present a set of SystemJ benchmark programs in which
+% real-time requirements should be met. We have carried out a set of
+% experiments to obtain BCRT and WCRT of the benchmark programs on the
+% execution platform called Java Optimized Processor(JOP)
+% \cite{jop:jnl:jsa2007}. JOP is a hardware implementation of the JVM
+% written in VHDL which enables real-time execution of Java programs by
+% translating Java bytecodes into a sequence of JOP's native instructions
+% called \emph{microcode} which is time-predictable. As SystemJ's default
+% compilation target is Java source code, JOP is an excellent platform
+% which enables us to analyze timing properties of SystemJ program.
 
@@
 
-By utilizing in-house tools in conjunction with Worst-Case-Execution-Time (WCET)
-analyzer \cite{jop:jnl:jsa2007} provided by JOP tools, we were able to estimate
-BCRT and WCRT of the following benchmark programs: (full-name?) HRTCS \cite{},
-Stepper motor controller (Motor) \cite{}, Robot and Access Environment Control
-System (AECS) \cite{}. We have also added a set of customized hardware counters
-to the original JOP design in order to collect numbers of clock cycles required
-for logical ticks of given SystemJ program. All examples except Robot are
-existing applications translated into pure SystemJ programs. The Robot example
-models a combination of robotic arms and conveyors which splits goods based on
-their volumetric sizes. Finally, garbage collector (GC) was disabled in all
-experiments to avoid unwanted side-effects such as lengthening clock-domain
-logical tick times. Whenever the GC triggers, experiment terminates. BCRT and
-WCRT were then calculated based on data collected until that point.
+We use the \textit{Java Optimized Processor}
+(JOP)~\cite{jop:jnl:jsa2007}, to carry out all our experiments. JOP
+implements the JVM specification in hardware directly, and has been
+shown to be real-time amenable. We statically estimate the WCRT and BCRT
+of these benchmark applications using the real-time analysis tools
+provided by the JOP design framework. We disable the \textit{Garbage
+  Collector} (GC) in all experiments to avoid unwanted side-effects such
+as lengthening clock-domain logical tick times.
 
-As one can see in Figure \ref{fig:comparison}, BCRT and WCRT were calculated for
-all examples. Given the real-time ranges \mbox{\texttt{delay [M..N]}} provided
-by programmers, required logical delays \emph{d} were determined by applying the
-Algorithm \ref{alg:1}. In HRTCS, real-time delay of the range 50.3..200.3 (ms)
-is only satisfied when \emph{d} is between 1506 and 1787 (not shown in Figure).
-It is then compiler's job to statically choose one of the values between this
+As one can see in Figure \ref{fig:comparison}, BCRT and WCRT were
+calculated for all examples. Given the real-time \mbox{\texttt{delay
+    (M..N)}} provided by programmers, required logical delay \texttt{d}
+was determined by applying Algorithm~\ref{alg:1}. In HRTCS, real-time
+delay of the range 50.3..200.3 (ms) is only satisfied when \emph{d} is
+between 1506 and 1787 (as shown previously in
+Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}).  It is
+then compiler's job to statically choose one of the values between this
 range.
 
-However, it is not the case that given real-time delays in a particular
-application are always realizable on JOP. For instance, all other examples other
-than HRTCS are required to relax the upper bound for a given real-time delay in
-order to obtain a feasible number of \emph{d}. Therefore, Algorithm \ref{alg:2}
-is applied to those examples to determine the minimum upper real-time bound
-\emph{N} such that there exist a solution for \emph{D} in Algorithm \ref{alg:1}.
-Unlike in HRTCS, however, this algorithm gives a unique number for logical tick
-delay \emph{d} such that \emph{d} = Min(\emph{$S_2$}) = Max(\emph{$S_1$}).
-The compiler, therefore, will choose \emph{d} based on the newly computed \emph{N} for
-the examples Robot, Motor and AECS.
+The upper bound \texttt{N} needed to be relaxed for all examples other
+than HRCTS to obtain a feasible \texttt{d}. Therefore, Algorithm
+\ref{alg:2} is applied to those examples to determine the minimum upper
+real-time bound \emph{N} such that there exist a solution for \emph{D}
+in Algorithm \ref{alg:1}.  Unlike in HRTCS, however, this algorithm
+gives a unique number for logical tick delay \emph{d} such that \emph{d}
+= Min(\emph{$S_2$}) = Max(\emph{$S_1$}).  The compiler, therefore, will
+choose \emph{d} based on the newly computed \emph{N} for the examples:
+Robot, Motor and AECS.
 
@@
 % our benchmark programs.
+% >>>>>>> 4bffc4b1a43388f5ebf6cc737b9d3b246ee77dc7
 

### 2013-07-01 introduction.tex
Included all changes by Prof. Salcic

@@
 
-Reactive languages~\cite{gber05,amal10} are a class of programming
-languages used for designing and implementing \textit{reactive} systems,
-which continuously respond to input from their environment. These
-languages have been successfully used in programming a plethora of
-systems such as fly by wire in Airbus~\cite{eairbus}, security
-surveillance systems~\cite{amal121}, etc. Usually the working
-environment of these reactive systems is also real-time. Yet, these
-languages do not support describing real-time as first class language
-constructs. For example, one cannot describe a simple real-time delay
-\texttt{delay 0.2 ms} in these languages. These languages are based on
-formal semantics, essential for the verification of functional
-correctness properties of the developed programs, but leave
-non-functional properties such as timing behavior as an implementation
-detail~\cite{ringler00,boldt07}. Arguably, rightly so, because physical
-time cannot be incorporated without information about the underlying
-execution platform. But, one can still reason in terms of
-\textit{logical} time. These languages support discrete \textit{logical}
-clock rather than a continuous clock. The period of the discrete logical
-clock (also referred to as a \textit{logical tick}) is determined by the
-responsiveness of the program to the input signals and as opposed to a
-discrete \textit{physical} clock, is elastic. Although this timing model
-has worked well for these languages in designing discrete systems, there
-is obviously a need to introduce continuous time since often
-implementation models are in the continuous time
-domain~\cite{DBLP:journals/pieee/SifakisTY03}.
+Reactive languages~\cite{gber931,amal10} are a class of programming
+languages used for designing and implementing reactive systems, which
+continuously respond to input from their environment. These languages
+have been successfully used in programming a plethora of systems such as
+fly-by-wire in Airbus~\cite{eairbus}, security surveillance
+systems~\cite{amal121}, etc.  Usually these systems also need to meet
+real-time constraints enforced by the environment.  Yet, the languages
+used to program these systems do not support describing real-time
+constructs as first class language constructs.  For example, one cannot
+describe a simple real-time delay (postponement) of an operation for 0.2
+ms. These languages are based on formal semantics, essential for the
+formal reasoning about and verification of correctness of functional
+properties of the developed programs, but leave nonfunctional properties
+such as timing behavior as an implementation
+detail~\cite{boldt07}. Arguably, rightly so, because physical time
+cannot be incorporated without information about the underlying
+execution platform.  But, one can still reason in terms of logical
+time. These languages support discrete logic clock rather than a
+discrete physical (real-time) clock. The real-time period of the
+discrete logic clock (also referred to as a logic tick) is not fixed and
+it is determined by the responsiveness of the program to the input
+signals. Unlike a discrete physical clock, which has fixed real-time
+period, the period of the logic clock is elastic. Although the timing
+model with logic clock has worked well for reactive synchronous
+languages in designing discrete systems, there is a need to introduce
+real-time since, often, implementation models are in the continuous or
+discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
+Moreover, adding timing capabilities as the ﬁrst class language
+constructs create additional opportunity for formally verifying
+functional and non-functional requirements before system deployment.
 
@@
 
-The fundamental contribution of this paper is the inclusion of real-time
-constructs in GALS (and its subset synchronous) reactive
-languages. Moreover, our contributions can be further refined as
-follows:
-
+The major motivations for introducing delay mechanism and at the same
+time contributions of the paper are: (1) delays allow real-time based
+synchronization between concurrent behaviors, (2) delays model relays on
+the use of relative instead of absolute time, i.e., a delay in selected
+time units is counted from the currently executing SystemJ statement and
+(3) delays allow certain behaviours to having real-time features, while
+the other behaviours use only logical time and (4) real-time delay does
+not affect the model of logical time in SystemJ program as long as the
+amount of delay is within boundaries that can be determined statically
+by program analysis. All these as well as semantics of the delay
+construct are discussed and illustrated in the remaining part of the
+paper. Our contributions can be further refined as follows:
 \begin{enumerate*}
@@
   i.e., non-negative rational numbers.
-\item We consider the property of \textit{non maximal-parallelism},
-  i.e., we consider scheduling on constrained resources to be an
-  integral part of the solution. To our knowledge we are the very first
-  to do so.
-\item Our solution does not change the functional or timing semantics of
-  GALS and its subset synchronous programs, in fact, our solution does
-  not require one to change the mid-end, the back-end or the related
-  optimization phases of the compiler at all.
+\item We consider the property of \textit{non
+    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
+  not always sufficient resources for processes to execute, so
+  effectively processes are allocated and scheduled with different
+  optimization criteria -- reducing computation time, reduced power
+  consumption, etc. This allocation and scheduling is an integral part
+  of any language compilation and needs to be considered when
+  introducing real-time delays.
+% \item Our solution does not change the functional or timing semantics of
+%   GALS and its subset synchronous programs, in fact, our solution does
+%   not require one to change the mid-end, the back-end or the related
+%   optimization phases of the compiler at all.
 \item We are the first, to our knowledge, to allow specification of both
   deterministic and non-deterministic real-time in GALS and synchronous
-  languages.
+  languages. \textit{Deterministic delay:} is the ability to specify a
+  single \textit{real-time} delay. \textit{Non-deterministic delay:} is
+  the ability to specify a range, of real-time values for a delay
+  statement, e.g., Figure~\ref{fig:1}.
 \end{enumerate*}
 
-The rest of the paper is arranged as follows...
+The rest of the paper is arranged as follows:
+Section~\ref{sec:background} gives the background on SystemJ and related
+techniques required for the rest of the
+paper. Section~\ref{sec:intr-real-time} introduces real-time delays in
+the GALS paradigm and describes their compilation into logical
+time. Section~\ref{sec:experimental-results} gives the experimental
+results for a set of real-time
+applications. Section~\ref{sec:related-work} positions this work in
+relation to other approaches to introducing real-time. Finally, we
+conclude in Section~\ref{sec:concl-future-work}.
 

### 2013-07-01 motivation.tex
Included all changes by Prof. Salcic

--- 2013-06-18 background.tex
@@
-\section{Background}
-\label{sec:background}
+\subsection{Motivating Example}
+\label{sec:motivating-example}
+
+% system{
+%  interface{
+%   input signal TOUCH;
+%   output signal GREEN_LIGHT;
+%   output char channel S;
+%   input char channel R;
+%  }
+\begin{figure}[t!]
+% \begin{center}
+\begin{minipage}{5cm}
+  \begin{scriptsize}
+
+\begin{verbatim}
+{ // Clock-domain 1
+while(true) {
+  trap(TRAP){
+   // Abort if user touches the screen
+   abort(TOUCH){
+    {sustain GREEN_LIGHT;}
+    || // synchronous parallel operator
+    {
+    //exit after any 
+    // where from 50.3 to 200.3 ms.
+    delay (50.3 .. 200.3 ms);
+    exit(TRAP); 
+    }
+   } do {send S(1);} // Send 1 if touched 
+  } do {send S(0);} // Else Send 0
+  pause; // only construct to consume time
+ }
+}
+>< // asynchronous operator
+{ // Clock-domain 2
+  while(true){ receive S;}
+  // do something with S
+}
+\end{verbatim}
+  \end{scriptsize}
+\end{minipage}
+\caption{An example human responsiveness system (HRTCS)}
+% }
+% \end{center}
+\label{fig:1}
+\end{figure}
+
+Consider a simple machine used to test human response time and collect
+this data for future research purposes. The machine switches on a green
+light on a touch screen for anywhere between \texttt{50.3ms} to
+\texttt{200.3ms}, if the user can touch this green light then she is
+successful and an integer \texttt{1} is sent to a database, if
+unsuccessful a \texttt{0} is sent instead, Different statistics and
+analyses on human responsiveness can be based on collected data.
+
+The SystemJ psuedocode~\cite{amal10} code for such a system is shown in
+Figure~\ref{fig:1}. We use the SystemJ language in this paper first of
+all, because being a \textit{Globally Asynchronous Locally Synchronous}
+(GALS) language it helps us easily describe a larger class of systems
+(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
+languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
+allows us to explore the marriage of real-time and logical time, which
+is equally applicable to the synchronous sub-set.
+
+The SystemJ program is divided into two clock-domains, the first
+clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
+signal) on the touch screen sensor and waits for the user to respond. If
+the user is able to touch the screen within the specified time of
+\texttt{50.3ms} to \texttt{200.3ms} a positive response is sent to the
+second clock-domain. All the SystemJ programming constructs used to
+implement this system are described in~\cite{amal10} and
+Table~\ref{tab:1} in Section~\ref{sec:background}. The delay
+speciﬁcation, which models the non-deterministic delay between 50.3ms
+and \texttt{200.3ms} is not supported in the current language and cannot
+be compiled with the current SystemJ compiler.
 

### 2013-07-01 related_work.tex
Included all changes by Prof. Salcic

@@
 external timers do not interact well with preemption constructs in these
-languages. Bourke et al. do a much better job of introducing real-time
-as first class constructs in the Esterel language. They like us provide
-real-time \texttt{delay} as first class Esterel programming constructs,
-and translate them into Esterel kernel constructs. But, unlike us they
-do not translate delays into \texttt{pause} constructs
-directly. Instead, logical ticks are generated by using abstract notion
-of \texttt{event} and \texttt{sample} platform dependent timers. This
-notion makes it a complex and inflexible solution since the number of
-platform timers with certain resolutions need to be determined and
-present on the system for the solution to be realizable. Moreover,
-non-deterministic real-time delays and integration with non-maximal
-parallelism is not studied at all by them. Other works such as
-Quartz~\cite{glog02} also incorporate delay statements in Esterel using
-\texttt{pause} constructs, but again, non-maximal parallelism,
+languages. Bourke et al. introduce real-time as first class constructs
+in the Esterel language. They like us provide real-time \texttt{delay}
+as first class Esterel programming constructs, and translate them into
+Esterel kernel constructs. But, unlike us they do not translate delays
+into \texttt{pause} constructs directly. Instead, logical ticks are
+generated by using abstract notion of \texttt{event} and \texttt{sample}
+platform dependent timers. This notion makes it a complex and inflexible
+solution since the number of platform timers with certain resolutions
+need to be determined and present on the system for the solution to be
+realizable. Moreover, non-deterministic real-time delays and integration
+with non-maximal parallelism is not studied at all. Our solution does
+not require external timers (like~\cite{rsh94}) or logical notion of
+timers (like~\cite{Bourke2009a}). Since in our solution we calculate the
+delay in logical ticks \textit{after} allocation and scheduling, it
+gives the compiler developers the chance to optimize for many different
+criteria such as computation time, energy, power, etc, without worrying
+about or violating the real-time delay specification. Other works such
+as Quartz~\cite{glog02} also incorporate delay statements in Esterel
+using \texttt{pause} constructs, but again, non-maximal parallelism and
 non-deterministic delays are not studied. Moreover, Quartz is targeted

### 2013-07-02 exps.tex
need to make \centering work for table

@@
 % =======
-% \begin{table}[t]
-% \centering
-%\twocolumn[
-%\begin{@twocolumnfalse}
-%\caption{Benchmark programs using \texttt{delay} constructs} \label{table:benchmark}
-%\scalebox{0.85}{
-%\begin{tabular}{|c| c | c | c | c | c | c | c |}
-%	\cline{4-8}
-% 	\multicolumn{3}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
-% 	\multirow{12}{*}{JOP} 		& \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{2-8}   
-% 								& \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{2-8} 
-% 								& \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{3-8} 
-% 								&			   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{3-8} 
-% 								&			   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
-% 								&			   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
-% 								&			   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
-% 								&			   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
+
+\begin{table}[t]
+ \centering
+  \twocolumn[
+\begin{@twocolumnfalse}
+\caption{Benchmark programs using \texttt{delay} constructs} \label{fig:comparison}
+\scalebox{0.85}{
+\begin{tabular}{| c | c | c | c | c | c | c |}
+	\cline{3-7}
+	\multicolumn{2}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
+								 \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{1-7}
+								 \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{1-7}
+								 \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{2-7}
+											   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{1-7}
+								 \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{2-7}
+											   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{1-7}
+								 \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{2-7}
+											   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{1-7}
+								 \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{2-7}
+											   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{1-7}
+								 \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{2-7}
+											   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
 % 	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}
@@
 % 	\hline
-% \end{tabular}
-% }
-% \end{@twocolumnfalse}
-% ]
-% \end{table}
+		\end{tabular}
+		}
+  \end{@twocolumnfalse}
+  ]
+\end{table}
 
-\begin{figure}
-\centering
-	\includegraphics[trim=2cm 2cm 2cm 2cm,scale=0.5]{table2}
-        \caption{Evaluation of logical delays \emph{d} for on various
-          benchmark programs}\label{fig:comparison}
-\end{figure}
-%
+% \begin{figure}
+% \centering
+% 	\includegraphics[trim=2cm 2cm 2cm 2cm,scale=0.5]{table2}
+%         \caption{Evaluation of logical delays \emph{d} for on various
+%           benchmark programs}\label{fig:comparison}
+% \end{figure}
+
 We have carried out a number of experiments on a set of applications
 with real-time constraints. The benchmark set is shown in
-Figure~\ref{fig:comparison}. HRCTS is the human response system
+Table~\ref{fig:comparison}. HRCTS is the human response system
 described in Figure~\ref{fig:1}. AECS (Access and Environmental Control
@@
 deterministic) delay requirements. These delay requirements are shown in
-Figure~\ref{fig:comparison} (\texttt{M..N}).
+Table~\ref{fig:comparison} (\texttt{M..N}).
 
@@
 
-As one can see in Figure \ref{fig:comparison}, BCRT and WCRT were
+As one can see in Table \ref{fig:comparison}, BCRT and WCRT were
 calculated for all examples. Given the real-time \mbox{\texttt{delay

### 2013-07-06 background.tex
Made the final changes

@@
 
-SystemJ consists of entities called clock-domains (CD)s, each running at
-their individual pace, interacting with the environment via signals and
-with each other via channels and CSP~\cite{choa85} style
+SystemJ consists of entities called \textit{Clock Domains} (CD)s, each
+running at their individual pace, interacting with the environment via
+signals and with each other via channels and CSP~\cite{choa85} style
 rendezvous. Each CD has its own notion of logical time. CDs adheres to
@@
 logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
-where $\mathcal{F}$ is some function dependent upon the schedule. We
-call $\Delta$ the reaction time. Obviously, depending upon the amount of
-computation required and schedule, $\Delta$ can vary, hence, in order to
-satisfy the implicit restriction imposed by the synchrony hypothesis --
-no input event \textit{from the environment} can be missed -- one needs
-to calculate the \textit{Worst Case Reaction Time} (WCRT) and the
-resultant WCRT needs to be smaller than the time between any two
-consecutive events on any input, else the synchrony hypothesis is
-violated. Techniques for calculating the WCRT of synchronous
-programs~\cite{boldt07} exist and this analysis is not the focus of this
-paper. The opposite of the WCRT is the \textit{Best Case Reaction Time}
-(BCRT), we denote the WCRT and BCRT in Figure~\ref{fig:2c} using
-\texttt{W} and \texttt{B}, respectively. Figure~\ref{fig:2c} shows a
-continuously running physical clock (analogous to a clock in digital
-hardware), the numerical annotations on the rising edge of the clock
-mark the logical ticks from Figure~\ref{fig:2c}. The subscripts
-\texttt{s} and \texttt{e} for the numbers show the start and end of the
-logical ticks, respectively. The end of a logical tick and the start of
-the next logical tick happen together. Figure~\ref{fig:2c} shows the
-mapping of the logical time to the physical time. For example, the
-logical tick \texttt{1} starts at the first rising edge and completes at
-the second rising edge of the physical clock. Where as logical tick
-\texttt{4} starts at the $4^{th}$ rising edge, but finishes at the
-$6^{th}$ rising edge of the physical clock. This difference can be
-accounted for by the java computations required when signal \texttt{A}
-is input by the environment (see Figure~\ref{fig:2a}). This elasticity
-is inherent (and elegant) to both Esterel and SystemJ.
+where $\mathcal{F}$ is some function dependent upon the
+\underline{schedule}. We call $\Delta$ the reaction time. Depending upon
+the amount of computation required and schedule, $\Delta$ can vary,
+hence, in order to satisfy the implicit restriction imposed by the
+synchrony hypothesis -- no input event \textit{from the environment} can
+be missed -- one needs to calculate the \textit{Worst Case Reaction
+  Time} (WCRT) and the resultant WCRT needs to be smaller than the time
+between any two consecutive events on any input, else the synchrony
+hypothesis is violated. Techniques for calculating the WCRT of
+synchronous programs~\cite{boldt07} exist and this analysis is not the
+focus of this paper. The opposite of the WCRT is the \textit{Best Case
+  Reaction Time} (BCRT), we denote the WCRT and BCRT in
+Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
+respectively. Figure~\ref{fig:2c} shows a continuously running physical
+clock (analogous to a clock in digital hardware), the numerical
+annotations on the rising edge of the clock mark the logical ticks from
+Figure~\ref{fig:2c}. The subscripts \texttt{s} and \texttt{e} for the
+numbers show the start and end of the logical ticks, respectively. The
+end of a logical tick and the start of the next logical tick happen
+together. Figure~\ref{fig:2c} shows the mapping of the logical time to
+the physical time. For example, the logical tick \texttt{1} starts at
+the first rising edge and completes at the second rising edge of the
+physical clock. Where as logical tick \texttt{4} starts at the $4^{th}$
+rising edge, but finishes at the $6^{th}$ rising edge of the physical
+clock. This difference can be accounted for by the java computations
+required when signal \texttt{A} is input by the environment (see
+Figure~\ref{fig:2a}). This elasticity is inherent (and elegant) to both
+Esterel and SystemJ.
 

### 2013-07-06 delays.tex
Made the final changes

@@
 
-We introduce a single \textit{derived} (consisting of kernel constructs)
+We introduce a single \textit{derived} (built from kernel constructs)
 statement called \mbox{\texttt{delay (M..N)}} in the SystemJ language
@@
 sets gives all the logical ticks that satisfy the real-time requirements
-invariant of the reaction time and its elasticity.
+invariant of the logical time and its elasticity.
 
 Let us revisit our motivating example to illustrate the algorithm. From
-Figure~\ref{fig:1} we know that $M$ is 50.3ms and $N$ is 200.3ms,
-respectively. Let the WCRT and BCRT be: 0.112ms and 0.0334ms,
+Figure~\ref{fig:1} we know that $M$ is 50.3 ms and $N$ is 200.3 ms,
+respectively. Let the WCRT and BCRT be: 0.112 ms and 0.0334 ms,
 respectively. Thus, the algorithm proceeds as follows:
@@
   \lfloor 200.3/0.112 \rfloor$. $l_1 = 450$ and $u_1 = 1788$. We first
-  calculate the logical ticks that are always running at the WCRT
+  calculate the logical ticks that are always running at the WCRT and
   satisfy the required real-time delay.
@@
 \item $D = S_1 \cap S_2$, $D = [1506,1788]$. Finally, the intersection
-  of the two sets gives the set $D$ from which we can pick up any value
+  of the two sets gives the set $D$ from which we can choose any value
   for \texttt{d}.
@@
 think this is an elegant solution, because the technique provides
-hard-real time guarantee while preserving the essence of synchronous and
-GALS programming prescribed by SystemJ and Esterel style
-languages. Moreover, this technique considers non maximal-parallelism,
-i.e., the delays in logical ticks is calculated after scheduling of CDs
-or synchronous programs has been performed. To our knowledge we are the
-first to do so.
+hard-real time guarantees while preserving the essence (elastic logical
+tick) of synchronous and GALS programming prescribed by SystemJ and
+Esterel style languages. Moreover, this technique considers non
+maximal-parallelism, i.e., the delays in logical ticks is calculated
+after scheduling of CDs or synchronous programs has been performed. To
+our knowledge we are the first to do so.
 
@@
 
-The relaxation algorithm is shown in Algorithm~\ref{alg:2}.
+The relaxation technique is shown in Algorithm~\ref{alg:2}.
 
@@
 
-Many real-world systems require non-deterministic timing constructs. One
-of them is our motivating example -- the human response time
-system.
+Many real-world systems require non-deterministic timing
+constructs. Where the exact real-time delay is not known or should not
+be known apriori. One such system is our motivating example -- the human
+response time system.
 
@@
 % \begin{verbatim}
-% {await (job); delay(2..10 ms);  //SPOOLER CD
+% {await (job); delay (2..10 ms);  //SPOOLER CD
 %  send mid(job) delay (1ms);}
@@
 // timeout after 1ms.
-trap(T){{await(A);}||{delay((1..1)ms);exit(T);}}
+trap(T){{await(A);}||{delay(1..1 ms);exit(T);}}
 \end{verbatim}
@@
 well with the external timer. The above program sends a signal to an
-external timer and waits for 10ms to pass by. Consider what happens when
-signals \texttt{S} and \texttt{TIMER} occur in the same logical tick,
-the \texttt{await} statement is never executed and hence, we enter a
-deadlock. Such problems are completely avoided in our technique, because
-we convert the real-time delays into logical delays (\texttt{pause}
-constructs), which interact well with preemption.
+external timer and waits for 10 ms to pass by. Consider what happens
+when signals \texttt{S} and \texttt{TIMER} occur in the same logical
+tick, the \texttt{await} statement is never executed (due to suspend)
+and hence, we enter a deadlock. Such problems are completely avoided in
+our technique, because we convert the real-time delays into logical
+delays (\texttt{pause} constructs), which interact well with preemption.
 
@@
 running CDs, are an addition in the SystemJ language. Like interaction
-of preemption and delays, conversion of rela-time delays to logical
+of preemption and delays, conversion of real-time delays to logical
 ticks also interacts well with channel rendezvous, because the semantics

### 2013-07-06 introduction.tex
Made the final changes

@@
 systems~\cite{amal121}, etc.  Usually these systems also need to meet
-real-time constraints enforced by the environment.  Yet, the languages
-used to program these systems do not support describing real-time
-constructs as first class language constructs.  For example, one cannot
-describe a simple real-time delay (postponement) of an operation for 0.2
-ms. These languages are based on formal semantics, essential for the
-formal reasoning about and verification of correctness of functional
-properties of the developed programs, but leave nonfunctional properties
-such as timing behavior as an implementation
-detail~\cite{boldt07}. Arguably, rightly so, because physical time
-cannot be incorporated without information about the underlying
-execution platform.  But, one can still reason in terms of logical
-time. These languages support discrete logic clock rather than a
-discrete physical (real-time) clock. The real-time period of the
-discrete logic clock (also referred to as a logic tick) is not fixed and
-it is determined by the responsiveness of the program to the input
+real-time constraints enforced by the environment. Yet, these languages
+do not support describing real-time statements as first class language
+constructs.  For example, one cannot describe a simple real-time delay
+(postponement) of an operation for 0.2 ms. These languages are based on
+formal semantics, essential for the formal reasoning about and
+verification of correctness of functional properties of the developed
+programs, but leave nonfunctional properties such as timing behavior as
+an implementation detail~\cite{boldt07}. Arguably, rightly so, because
+physical time cannot be incorporated without information about the
+underlying execution platform.  But, one can still reason in terms of
+logical time. These languages support discrete logical clock rather than
+a discrete physical (real-time) clock. The real-time period of the
+discrete logical clock (also referred to as a logical tick) is not fixed
+and it is determined by the responsiveness of the program to the input
 signals. Unlike a discrete physical clock, which has fixed real-time
-period, the period of the logic clock is elastic. Although the timing
-model with logic clock has worked well for reactive synchronous
+period, the period of the logical clock is elastic. Although the timing
+model with logical clock has worked well for reactive synchronous/GALS
 languages in designing discrete systems, there is a need to introduce
@@
 discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
-Moreover, adding timing capabilities as the ﬁrst class language
-constructs create additional opportunity for formally verifying
-functional and non-functional requirements before system deployment.
+Moreover, adding timing capabilities as first class language constructs
+create additional opportunity for formally verifying functional and
+non-functional requirements before system deployment.
 
@@
 time units is counted from the currently executing SystemJ statement and
-(3) delays allow certain behaviours to having real-time features, while
-the other behaviours use only logical time and (4) real-time delay does
-not affect the model of logical time in SystemJ program as long as the
-amount of delay is within boundaries that can be determined statically
-by program analysis. All these as well as semantics of the delay
-construct are discussed and illustrated in the remaining part of the
-paper. Our contributions can be further refined as follows:
+(3) delays allow mixing behaviors with real-time features and others
+with only logical time and (4) real-time delay does not affect the model
+of logical time in SystemJ program as long as the amount of delay is
+within boundaries that can be determined statically by program
+analysis. All these as well as semantics of the delay construct are
+discussed and illustrated in the remaining part of the paper. Our
+contributions can be further refined as follows:
 \begin{enumerate*}
@@
   effectively processes are allocated and scheduled with different
-  optimization criteria -- reducing computation time, reduced power
+  optimization criteria -- reducing computation time, reducing power
   consumption, etc. This allocation and scheduling is an integral part
@@
   the ability to specify a range, of real-time values for a delay
-  statement, e.g., Figure~\ref{fig:1}.
+  statement, e.g., delay statement in Figure~\ref{fig:1}.
 \end{enumerate*}

### 2013-07-06 motivation.tex
Made the final changes

@@
     {
-    //exit after any 
-    // where from 50.3 to 200.3 ms.
+     //exit after any where from 50.3 to 200.3 ms.
      delay (50.3 .. 200.3 ms);
@@
 { // Clock-domain 2
-  while(true){ receive S;}
-  // do something with S
+  while(true){ receive S; // do something with S}
 }
@@
 this data for future research purposes. The machine switches on a green
-light on a touch screen for anywhere between \texttt{50.3ms} to
-\texttt{200.3ms}, if the user can touch this green light then she is
-successful and an integer \texttt{1} is sent to a database, if
-unsuccessful a \texttt{0} is sent instead, Different statistics and
-analyses on human responsiveness can be based on collected data.
+light on a touch screen for anywhere between 50.3 ms to 200.3 ms, if the
+user can touch this green light then she is successful and an integer
+\texttt{1} is sent to a database, if unsuccessful a \texttt{0} is sent
+instead, Different statistics and analyses on human responsiveness can
+be based on collected data.
 
-The SystemJ psuedocode~\cite{amal10} code for such a system is shown in
-Figure~\ref{fig:1}. We use the SystemJ language in this paper first of
+The SystemJ psuedocode~\cite{amal10} for such a system is shown in
+Figure~\ref{fig:1}. We use the SystemJ language in this paper, first of
 all, because being a \textit{Globally Asynchronous Locally Synchronous}
@@
 signal) on the touch screen sensor and waits for the user to respond. If
-the user is able to touch the screen within the specified time of
-\texttt{50.3ms} to \texttt{200.3ms} a positive response is sent to the
-second clock-domain. All the SystemJ programming constructs used to
-implement this system are described in~\cite{amal10} and
-Table~\ref{tab:1} in Section~\ref{sec:background}. The delay
-speciﬁcation, which models the non-deterministic delay between 50.3ms
-and \texttt{200.3ms} is not supported in the current language and cannot
-be compiled with the current SystemJ compiler.
+the user is able to touch the screen within the specified time of 50.3
+ms to 200.3 ms a positive response is sent to the second
+clock-domain. All the SystemJ programming constructs used to implement
+this system are described in~\cite{amal10} and Table~\ref{tab:1} in
+Section~\ref{sec:background}. The delay speciﬁcation, which models the
+non-deterministic delay between 50.3 ms and 200.3 ms is not supported in
+the current language and cannot be compiled with the current SystemJ
+compiler. We introduce real-time \texttt{delay} statements in SystemJ.
 

### 2013-11-17 background.tex
Updated paper

@@
 Table~\ref{tab:1} presents the SystemJ kernel statements used to program
-the control flow. The data flow is programmed in Java. A number of
-derived statements also exist (e.g., \texttt{sustain}, \texttt{await},
-etc) that make programming reactive systems easier. The \textit{Model of
-  Computation} (MoC) of a simple SystemJ program is shown in
-Figure~\ref{fig:2}.
+the control flow. The data flow is programmed in the safe subset of
+Java~\cite{scj2013} amenable to formal verification and timing analysis,
+while avoiding all the pitfalls of programming in 'C' style languages. A
+number of derived statements also exist (e.g., \texttt{sustain},
+\texttt{await}, etc) that make programming reactive systems easier. The
+\textit{Model of Computation} (MoC) of a simple SystemJ program is shown
+in Figure~\ref{fig:2}.
 
@@
 signals and with each other via channels and CSP~\cite{choa85} style
-rendezvous. Each CD has its own notion of logical time. CDs adheres to
+rendezvous. Each CD has its own notion of logical time. CDs adhere to
 the \textit{perfect synchrony} hypothesis, i.e., all statements execute
@@
 while(true){
- abort(A){
+ abort(immediate A){
   while(true) pause; 
+ } // abort end
   emit O;
-  if(counter != 0) { // do java computation leading to WCRT}
+  if(counter != 0) { // java computation leading to WCRT}
   ++counter;
-  } // abort end
  } // loop end
@@
 the first rising edge and completes at the second rising edge of the
-physical clock. Where as logical tick \texttt{4} starts at the $4^{th}$
+physical clock, whereas logical tick \texttt{4} starts at the $4^{th}$
 rising edge, but finishes at the $6^{th}$ rising edge of the physical
-clock. This difference can be accounted for by the java computations
+clock. This difference can be accounted for by the Java computations
 required when signal \texttt{A} is input by the environment (see

### 2013-11-17 conclusion.tex
Updated paper

@@
 communication. We do \textit{not} need or use external timers to
-introduce these real-time delays, thereby resolving the problems of
+introduce these real-time delays, instead we depend upon an underlying
+platform that is timing analyzable, thereby resolving the problems of
 interaction between external timers and preemption. Moreover, we also do
@@
 rewrite the delay construct into reactive kernel constructs, calculate
-the \textit{Worst} and \textit{Base} case reaction times and then
+the \textit{Worst} and \textit{Best} case reaction times and then
 determine \textit{delay} \texttt{d}, number of logical ticks, which is
 invariant to the \textit{elasticity} of logical time. To our knowledge
-we are the first to introduce non-deterministic delays in such languages
-and show their usefulness in designing real-world systems.
+we are the first to introduce precise and imprecise delays in such
+languages and show their usefulness in designing real-world systems.
 

### 2013-11-17 delays.tex
Updated paper

@@
 rewrite \textit{delays} a certain number of logical ticks, before
-proceeding to the next statement. The number of logical ticks
-\texttt{``d''} to \textit{delay} (Figure~\ref{fig:3}) is determined by
-the compiler statically at compile time. The value of \texttt{d} is
-intricately tied with the WCRT and BCRT of the program and hence the
-execution platform.
+proceeding to the next statement. The number of logical ticks \texttt{d}
+to \textit{delay} (Figure~\ref{fig:3}) is determined by the compiler
+statically at compile time. The value of \texttt{d} is intricately tied
+with the WCRT and BCRT of the program and hence the execution platform.
 
@@
 these functions guarantee that the resultant logical ticks result in
-real-time delays between the required range $M-N$. Next, a function $F$
-maps these calculated values to a set of equidistant integer points
+real-time delays between the required range $[M,N]$. Next, a function
+$F$ maps these calculated values to a set of equidistant integer points
 (values) separated by a unit value -- these points represent all the
@@
   the same for the BCRT case.
-\item $S_1 = [450..1788]$ and $S_2 =[1506,5997]$. We then map the
+\item $S_1 = [450,1788]$ and $S_2 =[1506,5997]$. We then map the
   resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
@@
 think this is an elegant solution, because the technique provides
-hard-real time guarantees while preserving the essence (elastic logical
-tick) of synchronous and GALS programming prescribed by SystemJ and
-Esterel style languages. Moreover, this technique considers non
-maximal-parallelism, i.e., the delays in logical ticks is calculated
+real-time guarantees while preserving the essence (elastic logical tick)
+of synchronous and GALS programming prescribed by SystemJ and Esterel
+style languages. Moreover, this technique considers non
+maximal-parallelism, i.e., the delay in logical ticks is calculated
 after scheduling of CDs or synchronous programs has been performed. To
@@
 empty (also possible in the case of \texttt{delay(M..N)}, but never
-possible in case of \texttt{delay(M)}). In such a case, we automatically
-\textit{relax} the upper bound of the \texttt{delay} statement.
+possible in case of \texttt{delay(M)}). Note that $S_1$ and $S_2$ can
+never be empty sets and will always contain at minimum a single
+element. In such a case, we automatically \textit{relax} the upper bound
+of the \texttt{delay} statement.
 
@@
     \If {$D = \emptyset$} {
-      let $j_{0}$ be the first element of set $S_2$, s.t., $|S_2|=Q$\;
+      let $j_{0}$ be the first element of set $S_2$\;
       \ShowLn let $N \leftarrow WCRT \times j_0$\;
@@
 
-This algorithm results in the smallest relaxation required for the
-real-time delay to be satisfied. The algorithm takes as input the WCRT
-and set $S_2$, recall that set $S_2$ represents the logical ticks
-required to satisfy the real-time requirement at the BCRT. We take the
-very first value from set $S_2$ and multiply it with the WCRT to get the
-relaxation $N$. The first element of set $S_2$ is returned as the
-logical tick delay \texttt{d}.
+This algorithm results in the \underline{smallest} relaxation required
+for the real-time delay to be satisfied. The algorithm takes as input
+the WCRT and set $S_2$, recall that set $S_2$ represents the logical
+ticks required to satisfy the real-time requirement at the BCRT. We take
+the very first value from set $S_2$ and multiply it with the WCRT to get
+the relaxation $N$. The first element of set $S_2$ is returned as the
+logical tick delay \texttt{d}. The fundamental observation is that we
+have to delay for a minimum of $M$ units of real-time, hence,
+under-approximation is out of question. We can still over-approximate,
+but to reduce the resultant error, we should over-approximate by least
+possible value, which is the lower bound of set $S_2$. Thus, the lower
+bound is considered to be the only element shared between the two sets
+$S_1$ and $S_2$ and accordingly, the upper real-time bound is relaxed by
+the multiplication of WCRT and the first element of set $S_2$.
 
-The fundamental observation is that we have to delay for a minimum of
-$M$ units of real-time, hence, under-approximation is out of
-question. We can still over-approximate, but to reduce the resultant
-error, we should over-approximate by least possible value, which is the
-lower bound of set $S_2$. Thus, the lower bound is considered to be the
-only element shared between the two sets $S_1$ and $S_2$ and
-accordingly, the upper real-time bound is relaxed by the multiplication
-of WCRT and the first element of set $S_2$.
+For example, assume $BCRT=5$ and $WCRT=100$, with a required delay of
+$(200..200)$. Algorithm~\ref{alg:1}, results in $D$ being an empty set,
+since $S_1=[2]$ and $S_2=[40]$, respectively. In such a case, the
+resultant relaxation is: $40 \times 100$ using Algorithm~\ref{alg:2},
+which results in a time of $4000$ units. Such large over-approximations
+can be avoided by breaking the critical paths (worst case ticks) in the
+program (by inserting pause constructs), just like in hardware design,
+which is very \textit{unsurprising} considering Esterel has its origins
+in hardware.
+
+\subsection{The compiler flow}
+\label{sec:systemj-tool-chain}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[scale=0.5]{./tool}
+  \caption{The tool-chain flow of SystemJ compiler with delays}
+  \label{fig:4}
+\end{figure}
+
+The interaction of the low-level execution platform and the described
+high-level technique of introducing real-time delays in the SystemJ
+language is shown in Figure~\ref{fig:4}. We use static-analysis
+techniques to find out the WCRT/BCRT of the programs automatically after
+rewrites and standard compiler optimizations. We require timing
+analyzable platforms, in our case \textit{Java Optimized Processor}
+(JOP)~\cite{jop:jnl:jsa2007}. Moreover, since algorithms~\ref{alg:1}
+and~\ref{alg:2} are performed after scheduling for different constraints
+such as faster execution time, or reduced power consumption, we,
+inherently satisfy the criteria of exploiting non-maximal
+parallelism. The low-level analysis guarantees a precise WCRT/BCRT for
+the logical tick, which is then used for calculating $d$ in
+Figure~\ref{fig:3}.
 
@@
 
-% Another is a printer-spooler example borrowed from timed
-% CSP~\cite{Schneider:1999:CRT:555233}. The spooler and printer need to
-% synchronize using channels. The printer might be unable to print
-% depending upon the paper tray, similarly, spooler might take sometime
-% to send the job depending upon its size. Such real-time constraints
-% can be modeled in SystemJ as below:
-
-% \begin{scriptsize}
-
-% \begin{verbatim}
-% {await (job); delay (2..10 ms);  //SPOOLER CD
-%  send mid(job) delay (1ms);}
-% ><
-% {receive mid; delay(1..30 ms); //PRINTER CD
-%  emit print(#mid);}
-% \end{verbatim}
-% \end{scriptsize}
-
 \subsubsection{Timeout}
@@
 // timeout after 1ms.
-trap(T){{await(A);}||{delay(1..1 ms);exit(T);}}
+trap(T){{await(A);}||
+        {delay(1..1 ms);exit(T);}}
 \end{verbatim}
@@
 \begin{verbatim}
-suspend{S} {emit START_TIMER(10); await (TIMER); 
- emit O1;}
+suspend(S) {emit START_TIMER(10); 
+            await (TIMER);emit O1;}
 \end{verbatim}
@@
 
-As identified in~\cite{Bourke2009a} the \texttt{suspend} does not play
-well with the external timer. The above program sends a signal to an
-external timer and waits for 10 ms to pass by. Consider what happens
-when signals \texttt{S} and \texttt{TIMER} occur in the same logical
-tick, the \texttt{await} statement is never executed (due to suspend)
-and hence, we enter a deadlock. Such problems are completely avoided in
-our technique, because we convert the real-time delays into logical
-delays (\texttt{pause} constructs), which interact well with preemption.
+As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
+with the external timer. The above program sends a signal to an external
+timer and waits for 10 ms to pass by. Consider what happens when signals
+\texttt{S} and \texttt{TIMER} occur in the same logical tick, the
+\texttt{await} statement is never executed (due to suspend) and hence,
+we enter a deadlock. Such problems are completely avoided in our
+technique, because we convert the real-time delays into logical delays
+(\texttt{pause} constructs), which interact well with preemption.
 

### 2013-11-17 exps.tex
table
Updated paper

@@
 \begin{table}[t]
+\twocolumn[
 \centering
-  \twocolumn[
 \begin{@twocolumnfalse}
 \caption{Benchmark programs using \texttt{delay} constructs} \label{fig:comparison}
-\scalebox{0.85}{
+\resizebox{18cm}{2.4cm}{
 \begin{tabular}{| c | c | c | c | c | c | c |}
@@
 with real-time constraints. The benchmark set is shown in
-Table~\ref{fig:comparison}. HRCTS is the human response system
-described in Figure~\ref{fig:1}. AECS (Access and Environmental Control
-System) is an application that controls an intelligent room environment
-described in~\cite{aecs_ispa}. Motor is a stepper motor controller
+Table~\ref{fig:comparison}. HRCTS is the human response system described
+in Figure~\ref{fig:1}. AECS (Access and Environmental Control System) is
+an application that controls an intelligent room environment described
+in~\cite{aecs_ispa}. Motor is a stepper motor controller
 from~\cite{Bourke2009a}. Robot is an industrial automation application
@@
 respectively. All these examples are written in the SystemJ
-language. Finally, except for HRCTS all other examples have exact (and
-deterministic) delay requirements. These delay requirements are shown in
-Table~\ref{fig:comparison} (\texttt{M..N}).
+language. The requirement to have deterministic or non-deterministic
+delays varies between each benchmark. These delay requirements are shown
+in Table~\ref{fig:comparison} (\texttt{M..N}).
 
@@
 of these benchmark applications using the real-time analysis tools
-provided by the JOP design framework. We disable the \textit{Garbage
-  Collector} (GC) in all experiments to avoid unwanted side-effects such
-as lengthening clock-domain logical tick times.
+provided by the JOP design framework. The \textit{Safety Critical Java}
+(SCJ) specification we use for real-time analysis avoids garbage
+collection (GC) and hence, GC overheads and unpredictability are
+completely avoided.
+
+% We disable the \textit{Garbage Collector} (GC) in all experiments to
+% avoid unwanted side-effects such as lengthening clock-domain logical
+% tick times.
 
@@
 Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}).  It is
-then compiler's job to statically choose one of the values between this
+then compiler's job to statically choose one of the values in this
 range.
 
-The upper bound \texttt{N} needed to be relaxed for all examples other
-than HRCTS to obtain a feasible \texttt{d}. Therefore, Algorithm
-\ref{alg:2} is applied to those examples to determine the minimum upper
-real-time bound \emph{N} such that there exist a solution for \emph{D}
-in Algorithm \ref{alg:1}.  Unlike in HRTCS, however, this algorithm
-gives a unique number for logical tick delay \emph{d} such that \emph{d}
-= Min(\emph{$S_2$}) = Max(\emph{$S_1$}).  The compiler, therefore, will
-choose \emph{d} based on the newly computed \emph{N} for the examples:
-Robot, Motor and AECS.
+% The upper bound \texttt{N} needed to be relaxed for all examples other
+% than HRCTS to obtain a feasible \texttt{d}. Therefore,
+Algorithm \ref{alg:2} is applied to those examples, where relaxation was
+required, to determine the minimum upper real-time bound \emph{N} such
+that there exist a solution for \emph{D} in Algorithm \ref{alg:1}. This
+algorithm gives a unique number for logical tick delay \emph{d} such
+that \emph{d} = Min(\emph{$S_2$}) = Max(\emph{$S_1$}).  The compiler,
+therefore, will choose \emph{d} based on the newly computed \emph{N} for
+the examples: Robot, Motor and AECS.
 

### 2013-11-17 introduction.tex
Updated paper

@@
 time contributions of the paper are: (1) delays allow real-time based
-synchronization between concurrent behaviors, (2) delays model relays on
-the use of relative instead of absolute time, i.e., a delay in selected
-time units is counted from the currently executing SystemJ statement and
-(3) delays allow mixing behaviors with real-time features and others
-with only logical time and (4) real-time delay does not affect the model
-of logical time in SystemJ program as long as the amount of delay is
-within boundaries that can be determined statically by program
+synchronization between concurrent behaviors, (2) delay model relies on
+the use of relative instead of absolute real-time, i.e., a delay in
+selected time units is counted from the currently executing SystemJ
+statement (3) delays allow mixing behaviors with real-time features and
+others with only logical time and (4) real-time delay does not affect
+the model of logical time in SystemJ program as long as the amount of
+delay is within boundaries that can be determined statically by program
 analysis. All these as well as semantics of the delay construct are
@@
   deterministic and non-deterministic real-time in GALS and synchronous
-  languages. \textit{Deterministic delay:} is the ability to specify a
-  single \textit{real-time} delay. \textit{Non-deterministic delay:} is
+  languages. \textit{Deterministic delay:} is the ability to specify an
+  exact \textit{real-time} delay. \textit{Non-deterministic delay:} is
   the ability to specify a range, of real-time values for a delay

### 2013-11-17 related_work.tex
Updated paper

@@
 
-A small number of studies have been performed when it comes to
-integrating real-time directly as a programming construct in GALS (and
-its subset synchronous) reactive languages. The most prominent works in
-this area are by Shyamsundar~\cite{rsh94} and Bourke et
+A number of studies have been performed when it comes to integrating
+real-time directly as a programming construct in GALS (and its subset
+synchronous) reactive languages. The most prominent works in this area
+are by Shyamsundar~\cite{rsh94} and Bourke et
 al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
@@
 criteria such as computation time, energy, power, etc, without worrying
-about or violating the real-time delay specification. Other works such
-as Quartz~\cite{glog02} also incorporate delay statements in Esterel
-using \texttt{pause} constructs, but again, non-maximal parallelism and
-non-deterministic delays are not studied. Moreover, Quartz is targeted
-at studying timing properties using model checkers rather than
-implementation, like us.
+about or violating the real-time delay specification.
+
+Other works such as Quartz~\cite{glog02} and Bertin
+et.al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate delay
+statements in Esterel using \texttt{pause} and special \textit{pragmas}
+(annotations), respectively.  The Quartz solution is targeted at
+studying timing properties using model checkers rather than
+implementation, like us. The solution introduced by Bertin
+et.al.~\cite{Bertin:2000:TVR:1947412.1947439} translates real-time
+pragmas into Timed automata~\cite{alur94} for model-checking real-time
+deadlines. Our solution on the other hand removes all continuous time
+elements and produces single discrete time model thereby making the
+solution amenable to existing functional and real-time verification
+tools developed for reactive languages and more in the spirit of logical
+time. Finally, a major difference between Bertin et.al and our approach
+is that their approach requires programmers to annotate the time taken
+by 'C' code fragments manually, this is extremely error prone, we use
+static low-level program analysis to find out such times automatically.
 

### 2013-11-20 exps.tex
changin exp section

@@
 an application that controls an intelligent room environment described
-in~\cite{aecs_ispa}. Motor is a stepper motor controller
-from~\cite{Bourke2009a}. Robot is an industrial automation application
-that consists of two manufacturing belts and a robotic arm that splits
-goods according to their volumes. HRCTS consists of 2 CDs, but only one
-with real-time delays. AECS also has 2 CDs; CD1 consist of a single
-delay statement, while CD2 has 2 delay statements. Robot and Motor both
-are synchronous programs with 1 and 5 delay statements,
-respectively. All these examples are written in the SystemJ
-language. The requirement to have deterministic or non-deterministic
-delays varies between each benchmark. These delay requirements are shown
-in Table~\ref{fig:comparison} (\texttt{M..N}).
+in~\cite{aecs_ispa}. The system interacts with two external timers
+through input/output signals in order to measure how long the entrance
+door to the room is being opened and time the illegal presence (e.g.
+intruder) that is being detected. The program has been modified such
+that the delay constructs are used instead of external timers as shown
+in~\ref{delaycode}.
+
+\begin{figure}[h!]
+	\begin{SubFloat}{\label{delaycode:a}sss}
+		\centering
+		\begin{minipage}[b]{0.3\linewidth}
+			\footnotesize
+\begin{verbatim}
+emit TimerTriggered;
+trap(T){
+  abort(DoorOpened){
+    await(TimeOut);
+    exit(T); 
+  }
+}
+\end{verbatim}
+		\end{minipage}
+	\end{SubFloat}
+\hspace{1cm}%
+	\begin{SubFloat}{\label{delaycode:b}sss}
+		\centering
+		\begin{minipage}[b]{0.3\linewidth}
+			\footnotesize
+\begin{verbatim}
+// No emit statement
+trap(T){
+  abort(DoorOpened){
+    delay(2.4 .. 7.4772 ms);
+    exit(T);  
+  }
+}
+\end{verbatim}
+		\end{minipage}
+	\end{SubFloat}
+\caption{www}
+\label{delaycode}
+\end{figure}
+
+Motor is a stepper motor controller from~\cite{Bourke2009a}. Robot is an
+industrial automation application that consists of two manufacturing
+belts and a robotic arm that splits goods according to their volumes.
+HRCTS consists of 2 CDs, but only one with real-time delays. AECS also
+has 2 CDs; CD1 consist of a single delay statement, while CD2 has 2
+delay statements. Robot and Motor both are synchronous programs with 1
+and 5 delay statements, respectively. All these examples are written in
+the SystemJ language. The requirement to have deterministic or
+non-deterministic delays varies between each benchmark. These delay
+requirements are shown in Table~\ref{fig:comparison} (\texttt{M..N}).
 

### 2013-11-21 exps.tex
changing exp

@@
 
+{\renewcommand{\arraystretch}{0.9}
+\begin{table}[t]
+	\small
+	\centering
+	\caption{Program specification}
+	\label{fig:benchlist}
+	\scalebox{0.9}{
+\begin{tabular}{cccc}
+\toprule
+\emph{Benchmark}       & \emph{Required delays} & \emph{\# of CDs} & 
+\emph{Lines of code}        
+\\ \midrule
+HRTCS                  & 50.3-200.3 ms & 2                  & 56
+\\ \midrule
+\multirow{5}{*}{Motor} & 2.4 ms        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
+                       & 1.667 ms      &                    &                      \\ 
+                       & 0.05 ms       &                    &                      \\ 
+                       & 0.3 ms        &                    &                      \\ 
+											 & 0.733 ms      &                    &
+\\ \midrule
+Printer                & TBC           & 2                  & 39
+\\ \midrule
+AECS/CD1               & 10000 ms      & -                  & 144
+\\ \midrule
+AECS/CD2               & 10000 ms      & -                  & 123
+\\ \bottomrule
+\end{tabular}
+}
+\vspace{-20pt}
+\end{table}
+}
+
+{\renewcommand{\arraystretch}{0.6}
 \begin{table}[t]
@@
 \begin{@twocolumnfalse}
-\caption{Benchmark programs using \texttt{delay} constructs} \label{fig:comparison}
-\resizebox{18cm}{2.4cm}{
-\begin{tabular}{| c | c | c | c | c | c | c |}
-	\cline{3-7}
-	\multicolumn{2}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
-								 \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{1-7}
-								 \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{1-7}
-								 \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{2-7}
-											   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{1-7}
-								 \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{2-7}
-											   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{1-7}
-								 \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{2-7}
-											   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{1-7}
-								 \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{2-7}
-											   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{1-7}
-								 \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{2-7}
-											   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
-% 	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}
-% 								& \multicolumn{2}{|c|}{WCRT} 		&0.0329 ms 		& 0.0393 ms		  & 0.0361 ms		  & 0.5072 ms		   & 0.8388 ms\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 1}	 & M..N &50.3-600.3 ms 	& 2.4-15.6265 ms  & 1230-42317.8276 ms& 10000-95358.8578 ms& 10000-287757.9834 ms \\ \cline{3-8}
-% 								&			   & d			 		&18209-18232 	& 398 			  & 1171429			  & 188015		       &343054\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-10.8757 ms& N/A				  & N/A 			   &10000-287757.9834 ms\\ \cline{3-8}
-% 								&			   & d			 		& 	N/A	 		& 277			  & N/A				  & N/A       		   &343054\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.3534 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8}
-% 								&			   & d			 		& 	N/A	 		& 9		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.9631 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8}
-% 								&			   & d				 	& 	N/A	 		& 50			  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
-% 								& \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-4.79 ms   & N/A				  &	N/A      		   &N/A\\ \cline{3-8}
-% 								&			   & d			 		& 	N/A	 		& 122 			  &	N/A				  & N/A      		   &N/A\\ \cline{2-8}
-% 	\hline
+	\caption{Actual delays obtained for \texttt{delay} constructs in the
+benchmark programs based on their BCRT and WCRT}
+\label{fig:comparison}
+\begin{tabular}{ c  c  c  c  c  c  c }
+	\toprule
+	\multicolumn{2}{c}{}											& \emph{HRTCS} 		&
+	\emph{Motor} 		  & \emph{Robot} 			  & \emph{AECS/CD1}   &
+	\emph{AECS/CD2} \\ \midrule
+	\multicolumn{2}{c}{\emph{BCRT}} 		&0.0334 ms 		& 0.0361 ms		  &
+	0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cmidrule(r){1-7}
+	\multicolumn{2}{c}{\emph{WCRT}} 		&0.1120 ms 		& 0.1116 ms		  &
+	0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cmidrule(r){1-7}
+								 \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cmidrule(r){2-7}
+											   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cmidrule(r){1-7}
+								 \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cmidrule(r){2-7}
+											   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cmidrule(r){1-7}
+								 \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cmidrule(r){2-7}
+											   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cmidrule(r){1-7}
+								 \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cmidrule(r){2-7}
+											   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cmidrule(r){1-7}
+								 \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cmidrule(r){2-7}
+											   & d			 		& 	N/A	 		& 21 			  &	N/A
+												 & N/A      		   &N/A\\ \bottomrule
 \end{tabular}
-}
 \end{@twocolumnfalse}
@@
 \end{table}
-
-% \begin{figure}
-% \centering
-% 	\includegraphics[trim=2cm 2cm 2cm 2cm,scale=0.5]{table2}
-%         \caption{Evaluation of logical delays \emph{d} for on various
-%           benchmark programs}\label{fig:comparison}
-% \end{figure}
+}
 
 We have carried out a number of experiments on a set of applications
-with real-time constraints. The benchmark set is shown in
-Table~\ref{fig:comparison}. HRCTS is the human response system described
-in Figure~\ref{fig:1}. AECS (Access and Environmental Control System) is
-an application that controls an intelligent room environment described
-in~\cite{aecs_ispa}. The system interacts with two external timers
-through input/output signals in order to measure how long the entrance
-door to the room is being opened and time the illegal presence (e.g.
-intruder) that is being detected. The program has been modified such
-that the delay constructs are used instead of external timers as shown
-in~\ref{delaycode}.
+with real-time constraints. The characteristics of the benchmark set
+including delay specifications are shown in Table~\ref{fig:benchlist}.
 
+HRTCS is the human response system described in
+Section~\ref{sec:motivating-example}. AECS (Access and Environmental
+Control System) is an application that controls an intelligent room
+environment described in~\cite{aecs_ispa}.  The system interacts with
+two external timers through input/output signals in order to measure how
+long the entrance door to the room is being opened and time the illegal
+presence (e.g. intruder) that is being detected. The program has been
+modified such that the delay constructs are used instead of external
+timers as shown in Figure~\ref{delaycode}. In the original AECS, the
+system needed to trigger external timer by emitting the output signal in
+order to wait for a specific amount of time (e.g.
+\texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
+expires, it notifies back to the system through the input signal
+\texttt{TimeOut} which is polled at the \texttt{await} statement. With
+the new delay constructs, those timers are not necessary and delays can
+be expressed directly in the code as shown in Figure~\ref{delaycode:b}.
 \begin{figure}[h!]
-	\begin{SubFloat}{\label{delaycode:a}sss}
+	\begin{SubFloat}{\label{delaycode:a}}
 		\centering
 		\begin{minipage}[b]{0.3\linewidth}
-			\footnotesize
+			\scriptsize
 \begin{verbatim}
@@
 \hspace{1cm}%
-	\begin{SubFloat}{\label{delaycode:b}sss}
+	\begin{SubFloat}{\label{delaycode:b}}
 		\centering
-		\begin{minipage}[b]{0.3\linewidth}
-			\footnotesize
+		\begin{minipage}[b]{0.5\linewidth}
+			\scriptsize
 \begin{verbatim}
@@
   abort(DoorOpened){
-    delay(2.4 .. 7.4772 ms);
+    delay(2.4 ms);
     exit(T);  
@@
 	\end{SubFloat}
-\caption{www}
+\caption{Replacing external timer with delay construct}
 \label{delaycode}
 \end{figure}
+Motor is a stepper motor controller from~\cite{Bourke2009a}, used for
+producing monochrome images on paper by applying a correct application
+of current to the print head (a row of resistors). In this example,
+duration of current applied to the print head is controlled based on
+feedback logic where delays are expressed in real-time Esterel
+constructs. We have rewritten the program in SystemJ with our delay
+construct and compare between desired and relaxed delay according to
+WCRT and BCRT of the program. Printer is an example from [], which
+models a printer and a spooler in timed CSP. Their operation is rather
+simple: as soon as the spooler receives a print job, passes it to the
+printer through channel for processing. The example demonstrates that by
+specifying delay bounds (e.g. 1 to 5 time units) between two subsequent
+behaviors, the job is passed along the channel as soon as both processes
+become ready.
 
-Motor is a stepper motor controller from~\cite{Bourke2009a}. Robot is an
-industrial automation application that consists of two manufacturing
-belts and a robotic arm that splits goods according to their volumes.
-HRCTS consists of 2 CDs, but only one with real-time delays. AECS also
+%Robot is an industrial automation application that consists of two
+%manufacturing belts and a robotic arm that splits goods according to
+%their volumes.
+
+HRTCS consists of 2 CDs, but only one with real-time delays. AECS also
 has 2 CDs; CD1 consist of a single delay statement, while CD2 has 2
-delay statements. Robot and Motor both are synchronous programs with 1
-and 5 delay statements, respectively. All these examples are written in
-the SystemJ language. The requirement to have deterministic or
-non-deterministic delays varies between each benchmark. These delay
-requirements are shown in Table~\ref{fig:comparison} (\texttt{M..N}).
+delay statements. In this case, as explained previously, delay
+statements replaced the external timers and corresponding interface
+signals. The system triggers an alarm after 10 seconds of detection of
+an intruder or generates a beep sound when an entrance door is being
+opened for the same amount of time. For Printer, each spooler and
+printer process is mapped to a single clock-domain. Spooler CD sends a
+job to the Printer CD between [] to [] ms after receiving it from the
+enviornment (e.g. a user). On the other hand, Printer CD notifies
+environment that it starts printing between [] to [] ms once it receives
+the job from the Spooler CD. Actual print job is started between [] to
+[] after the notification. Lastly, Motor is a synchronous program with 5
+delay statements where they controls amount of current into the stepper
+motor coils.
+
+%All these examples are written in
+%the SystemJ language. The requirement to have deterministic or
+%non-deterministic delays varies between each benchmark.  These delay
+%requirements are shown in Table~\ref{fig:comparison} (\texttt{M..N}).
 
@@
 collection (GC) and hence, GC overheads and unpredictability are
-completely avoided.
+completely avoided. Note that only HRTCS and Printer have
+non-deterministic delays (\texttt{M..N}) as their program specification
+where others use exact delays (\texttt{M}). The actual delay after
+applying relaxation for the upper real-time bound is shown in
+Table~\ref{fig:comparison}.
 
@@
 calculated for all examples. Given the real-time \mbox{\texttt{delay
-    (M..N)}} provided by programmers, required logical delay \texttt{d}
-was determined by applying Algorithm~\ref{alg:1}. In HRTCS, real-time
-delay of the range 50.3..200.3 (ms) is only satisfied when \emph{d} is
-between 1506 and 1787 (as shown previously in
+(M..N)}} provided by programmers, required logical delay \texttt{d} was
+determined by applying Algorithm~\ref{alg:1}. In HRTCS, real-time delay
+of the range 50.3..200.3 (ms) is only satisfied when \emph{d} is between
+1506 and 1787 (as shown previously in
 Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}).  It is
@@
 % The upper bound \texttt{N} needed to be relaxed for all examples other
-% than HRCTS to obtain a feasible \texttt{d}. Therefore,
+% than HRTCS to obtain a feasible \texttt{d}. Therefore,
 Algorithm \ref{alg:2} is applied to those examples, where relaxation was
@@
 therefore, will choose \emph{d} based on the newly computed \emph{N} for
-the examples: Robot, Motor and AECS.
+the examples: Printer, Motor and AECS.
 

### 2013-11-22 exps.tex
exps.tex

@@
 be expressed directly in the code as shown in Figure~\ref{delaycode:b}.
+
 \begin{figure}[h!]
@@
 \end{figure}
+
 Motor is a stepper motor controller from~\cite{Bourke2009a}, used for
@@
 duration of current applied to the print head is controlled based on
-feedback logic where delays are expressed in real-time Esterel
-constructs. We have rewritten the program in SystemJ with our delay
-construct and compare between desired and relaxed delay according to
-WCRT and BCRT of the program. Printer is an example from [], which
-models a printer and a spooler in timed CSP. Their operation is rather
-simple: as soon as the spooler receives a print job, passes it to the
-printer through channel for processing. The example demonstrates that by
-specifying delay bounds (e.g. 1 to 5 time units) between two subsequent
-behaviors, the job is passed along the channel as soon as both processes
-become ready.
+feedback logic using real-time delays which are expressed in Esterel
+constructs. We have rewritten the program in SystemJ and compared
+between desired and relaxed delay according to WCRT and BCRT of the
+program.  Printer is an example from [], which models a printer and a
+spooler in timed CSP. Their operation is rather simple: as soon as the
+spooler receives a print job, passes it to the printer through channel
+for processing. The example demonstrates that by specifying delay bounds
+(e.g. 1 to 5 time units) between two subsequent behaviors, the job is
+passed along the channel as soon as both processes become ready.
 
@@
 delay statements. In this case, as explained previously, delay
-statements replaced the external timers and corresponding interface
-signals. The system triggers an alarm after 10 seconds of detection of
-an intruder or generates a beep sound when an entrance door is being
-opened for the same amount of time. For Printer, each spooler and
-printer process is mapped to a single clock-domain. Spooler CD sends a
-job to the Printer CD between [] to [] ms after receiving it from the
+statements replace the external timers and corresponding interface
+signals. The system triggers an alarm 10 seconds after detection of an
+intruder or generates a beep sound when an entrance door is being opened
+for the same amount of time. In Printer, each spooler and printer
+process is mapped to a single clock-domain. Spooler CD sends a job to
+the Printer CD between [] to [] ms after receiving it from the
 enviornment (e.g. a user). On the other hand, Printer CD notifies
@@
 the job from the Spooler CD. Actual print job is started between [] to
-[] after the notification. Lastly, Motor is a synchronous program with 5
+[] after the notification. In this example, bounded delay ensures that
+both channel communication and print job should occur within the
+specified amount of time. Lastly, Motor is a synchronous program with 5
 delay statements where they controls amount of current into the stepper
@@
 
-As one can see in Table \ref{fig:comparison}, BCRT and WCRT were
-calculated for all examples. Given the real-time \mbox{\texttt{delay
-(M..N)}} provided by programmers, required logical delay \texttt{d} was
-determined by applying Algorithm~\ref{alg:1}. In HRTCS, real-time delay
-of the range 50.3..200.3 (ms) is only satisfied when \emph{d} is between
-1506 and 1787 (as shown previously in
+%As one can see in Table \ref{fig:comparison}, 
+As a first step, BCRT and WCRT were calculated for all examples. Given
+the real-time \mbox{\texttt{delay (M..N)}} provided by programmers,
+required logical delay \texttt{d} was determined by applying
+Algorithm~\ref{alg:1}. In HRTCS, real-time delay of the range
+50.3..200.3 (ms) is only satisfied when \emph{d} is between 1506 and
+1787 (as shown previously in
 Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}).  It is
@@
 % than HRTCS to obtain a feasible \texttt{d}. Therefore,
-Algorithm \ref{alg:2} is applied to those examples, where relaxation was
-required, to determine the minimum upper real-time bound \emph{N} such
-that there exist a solution for \emph{D} in Algorithm \ref{alg:1}. This
-algorithm gives a unique number for logical tick delay \emph{d} such
-that \emph{d} = Min(\emph{$S_2$}) = Max(\emph{$S_1$}).  The compiler,
-therefore, will choose \emph{d} based on the newly computed \emph{N} for
-the examples: Printer, Motor and AECS.
+On the other hand, Algorithm \ref{alg:2} is applied to those examples,
+where relaxation was required, to determine the minimum upper real-time
+bound \emph{N} such that there exist a solution for \emph{D} in
+Algorithm \ref{alg:1}. This algorithm gives a unique number for logical
+tick delay \emph{d} such that \emph{d} = Min(\emph{$S_2$}) =
+Max(\emph{$S_1$}).  The compiler, therefore, will choose \emph{d} based
+on the newly computed \emph{N} for the examples: Printer, Motor and
+AECS.
 

### 2013-11-25 exps.tex
Updated table

@@
 	\centering
-	\caption{Program specification}
+	\caption{An overview of the benchmark programs}
 	\label{fig:benchlist}
 	\scalebox{0.9}{
-\begin{tabular}{cccc}
+		\begin{tabular}{llp{0.5cm}p{1cm}}
 \toprule
-\emph{Benchmark}       & \emph{Required delays} & \emph{\# of CDs} & 
+\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
 \emph{Lines of code}        
 \\ \midrule
-HRTCS                  & 50.3-200.3 ms & 2                  & 56
+HRTCS                  & \texttt{delay(50.3..200.3 ms)} ($CD1$) & 2                  & 56
 \\ \midrule
-\multirow{5}{*}{Motor} & 2.4 ms        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
-                       & 1.667 ms      &                    &                      \\ 
-                       & 0.05 ms       &                    &                      \\ 
-                       & 0.3 ms        &                    &                      \\ 
-											 & 0.733 ms      &                    &
+\multirow{5}{*}{Motor} & \texttt{delay(2.4..2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
+											 & \texttt{delay(1.667..1.667 ms)}      &                    &                      \\ 
+											 & \texttt{delay(0.05..0.05 ms)}       &                    &                      \\ 
+											 & \texttt{delay(0.3..0.3 ms)}        &                    &                      \\ 
+											 & \texttt{delay(0.733..0.733 ms)}      &                    &
 \\ \midrule
-Printer                & TBC           & 2                  & 39
+\multirow{3}{*}{Printer}
+& \texttt{delay(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
+& \texttt{delay(1.2..30.3 ms)} ($CD2$)		  &														&											\\
+& \texttt{delay(5.6..100.2 ms)} ($CD2$)     &														&											
 \\ \midrule
-AECS/CD1               & 10000 ms      & -                  & 144
-\\ \midrule
-AECS/CD2               & 10000 ms      & -                  & 123
+\multirow{2}{*}{AECS} & \texttt{delay(10000 ms)} ($CD1$) 
+& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
+& \texttt{delay(10000..10000 ms)} ($CD2$) & &
 \\ \bottomrule
@@
 \label{fig:comparison}
-\begin{tabular}{ c  c  c  c  c  c  c }
+\begin{tabular}{ c c l l l l l l }
 	\toprule
-	\multicolumn{2}{c}{}											& \emph{HRTCS} 		&
-	\emph{Motor} 		  & \emph{Robot} 			  & \emph{AECS/CD1}   &
-	\emph{AECS/CD2} \\ \midrule
-	\multicolumn{2}{c}{\emph{BCRT}} 		&0.0334 ms 		& 0.0361 ms		  &
-	0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cmidrule(r){1-7}
-	\multicolumn{2}{c}{\emph{WCRT}} 		&0.1120 ms 		& 0.1116 ms		  &
-	0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cmidrule(r){1-7}
-								 \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cmidrule(r){2-7}
-											   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cmidrule(r){1-7}
-								 \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cmidrule(r){2-7}
-											   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cmidrule(r){1-7}
-								 \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cmidrule(r){2-7}
-											   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cmidrule(r){1-7}
-								 \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cmidrule(r){2-7}
-											   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cmidrule(r){1-7}
-								 \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cmidrule(r){2-7}
+	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
+	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
+	\midrule
+	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0334 ms	& 0.0361 ms & 0.009613 ms &
+	0.0263 ms & 0.1683 ms & 0.1881 ms\\ 
+	\cmidrule(r){1-8}
+	\multicolumn{2}{c}{\emph{WCRT}} &0.1120 ms & 0.1116 ms & 0.0603 ms
+	& 0.0491 ms & 0.8148 ms & 1.0115 ms\\ 
+	\cmidrule(r){1-8}
+
+	\multirow{2}{*}{Delay 1} & M..N & 50.3-200.3 ms	& 2.4-8.4882 ms	  &
+	2.7-16.9378 ms & 1.2-30.3 ms & 10000-42483.2198 ms& 10000-53882.0045 ms \\
+	\cmidrule(r){2-8}
+	& d &1506-1888 & 68 & 281 & 46-617 & 59428 &53160\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 2} & M..N & N/A & 1.668-5.2452 ms & N/A &
+	5.6-100.2 ms & N/A &10000-53882.0045 ms\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 48 & N/A & 213-2041 & N/A &53160\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 3} & M..N & N/A & 0.05-0.2232 ms & N/A & N/A & N/A
+	&N/A\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 2 & N/A & N/A & N/A &N/A\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 4} & M..N & N/A & 0.3-1.0044 ms & N/A & N/A & N/A
+	&N/A\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 9 & N/A & N/A & N/A &N/A\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 5} & M..N & N/A & 0.833-2.3436 ms & N/A & N/A & N/A
+	&N/A\\ 
+	\cmidrule(r){2-8}
 	& d & N/A & 21 & N/A
-												 & N/A      		   &N/A\\ \bottomrule
+	& N/A & N/A &N/A\\ \bottomrule
 \end{tabular}
@@
 with real-time constraints. The characteristics of the benchmark set
-including delay specifications are shown in Table~\ref{fig:benchlist}.
+including required real-time delays and program sizes are shown in
+Table~\ref{fig:benchlist}.
 
@@
   abort(DoorOpened){
-    delay(2.4 ms);
+    delay(10000 ms);
     exit(T);  
@@
 producing monochrome images on paper by applying a correct application
-of current to the print head (a row of resistors). In this example,
-duration of current applied to the print head is controlled based on
-feedback logic using real-time delays which are expressed in Esterel
-constructs. We have rewritten the program in SystemJ and compared
-between desired and relaxed delay according to WCRT and BCRT of the
-program.  Printer is an example from [], which models a printer and a
-spooler in timed CSP. Their operation is rather simple: as soon as the
-spooler receives a print job, passes it to the printer through channel
-for processing. The example demonstrates that by specifying delay bounds
-(e.g. 1 to 5 time units) between two subsequent behaviors, the job is
-passed along the channel as soon as both processes become ready.
+of current to a print head (a row of resistors). In this example,
+duration of current applied is controlled by the feedback logic using
+real-time delays which are expressed in Esterel constructs. We have
+rewritten the program in SystemJ and compared between desired and
+relaxed delays according to WCRT and BCRT of the program. Printer is an
+example from [], which models a printer and a spooler in timed CSP.
+Their operation is rather simple: as soon as the spooler receives a
+print job, passes it to the printer through a channel where actual
+printing process take place. The example demonstrates that by specifying
+delay bounds (e.g. 1 to 5 time units) on receiving and sending channel
+ends data exchange will occur as soon as both processes become ready.
 
@@
 delay statements. In this case, as explained previously, delay
-statements replace the external timers and corresponding interface
+statements replace the external timers and corresponding interface (I/O)
 signals. The system triggers an alarm 10 seconds after detection of an
@@
 collection (GC) and hence, GC overheads and unpredictability are
-completely avoided. Note that only HRTCS and Printer have
-non-deterministic delays (\texttt{M..N}) as their program specification
-where others use exact delays (\texttt{M}). The actual delay after
-applying relaxation for the upper real-time bound is shown in
-Table~\ref{fig:comparison}.
+completely avoided.
 
@@
 
-%As one can see in Table \ref{fig:comparison}, 
+Table~\ref{fig:comparison} shows the result of our algorithm applied to
+each benchmark program. Note that we allowed the compiler to find the
+minimum upper real-time bounds for all delay constructs, e.g. bounded
+(\texttt{M..N}) or exact (\texttt{M..M}) delays, when it fails to find a
+feasible solution. However, a user can prohibit this behavior through
+configuration that such case is an error.
+
 As a first step, BCRT and WCRT were calculated for all examples. Given
 the real-time \mbox{\texttt{delay (M..N)}} provided by programmers,
-required logical delay \texttt{d} was determined by applying
+required logical delay \textit{d} was determined by applying
 Algorithm~\ref{alg:1}. In HRTCS, real-time delay of the range
@@
 1787 (as shown previously in
-Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}).  It is
-then compiler's job to statically choose one of the values in this
-range.
+Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}). Similarly,
+Printer/CD2 satisfies real-time ranges of 1.2..30.3 (ms) and 5.6..100.2
+(ms) when \emph{d} is 46..617 and 213..2041 respectively. It is then
+compiler's job to statically choose one of the values in this range.
 
@@
 Max(\emph{$S_1$}).  The compiler, therefore, will choose \emph{d} based
-on the newly computed \emph{N} for the examples: Printer, Motor and
+on the newly computed \emph{N} for the examples: Printer/CD1, Motor and
 AECS.

### 2013-11-25 exps.tex
cont

@@
 Table~\ref{fig:comparison} shows the result of our algorithm applied to
-each benchmark program. Note that we allowed the compiler to find the
-minimum upper real-time bounds for all delay constructs, e.g. bounded
-(\texttt{M..N}) or exact (\texttt{M..M}) delays, when it fails to find a
-feasible solution. However, a user can prohibit this behavior through
-configuration that such case is an error.
+each benchmark program. Note that we allowed the compiler to relax the
+upper real-time bounds even for bounded (\texttt{M..N}) or exact
+(\texttt{M..M}) delays when it fails to find a feasible solution.
+However, a user can prohibit this behavior through configuration that
+such case is an error.
 

### 2013-11-26 exps.tex
fig is groupplot

@@
 relaxed delays according to WCRT and BCRT of the program. Printer is an
-example from [], which models a printer and a spooler in timed CSP.
-Their operation is rather simple: as soon as the spooler receives a
-print job, passes it to the printer through a channel where actual
-printing process take place. The example demonstrates that by specifying
-delay bounds (e.g. 1 to 5 time units) on receiving and sending channel
-ends data exchange will occur as soon as both processes become ready.
+example from~\cite{Schneider:1999:CRT:555233}, which models a printer and
+a spooler in timed CSP.  Their operation is rather simple: as soon as
+the spooler receives a print job, passes it to the printer through a
+channel where actual printing process take place. The example
+demonstrates that by specifying delay bounds (e.g. 1 to 5 time units) on
+receiving and sending channel ends data exchange will occur as soon as
+both processes become ready.
 
@@
 process is mapped to a single clock-domain. Spooler CD sends a job to
-the Printer CD between [] to [] ms after receiving it from the
+the Printer CD between 2.7 to 10.6 ms after receiving it from the
 enviornment (e.g. a user). On the other hand, Printer CD notifies
-environment that it starts printing between [] to [] ms once it receives
-the job from the Spooler CD. Actual print job is started between [] to
-[] after the notification. In this example, bounded delay ensures that
-both channel communication and print job should occur within the
-specified amount of time. Lastly, Motor is a synchronous program with 5
-delay statements where they controls amount of current into the stepper
-motor coils.
+environment that it starts printing between 1.2 to 30.3 ms once it
+receives the job from the Spooler CD. Actual print job is started
+between 5.6 to 100.2 ms after the notification. In this example, bounded
+delay ensures that both channel communication and print job should occur
+within the specified amount of time. Lastly, Motor is a synchronous
+program with 5 delay statements where they controls amount of current
+into the stepper motor coils.
 
@@
 
+\subsection{Improving results}
+
+Previous results shows that the compiler was only able to find
+\textit{d} for HRTCS and Printer/CD2 without relaxing upper bound.
+Figure~\ref{fig:graph} shows the accuracy of our result by comparing
+differences between programmer specified upper bounds and the actual
+bounds obtained from our algorithm.
+
+\begin{figure}[h!]
+	\centering
+	\begin{tikzpicture}[scale=0.95]
+	\begin{groupplot}[group style={group size=2 by 3,
+			xlabels at=edge bottom,
+			ylabels at=edge left,
+		},
+			footnotesize,ybar,xtick=data,
+			symbolic x
+			coords={HRTCS,Motor,Printer/CD1,Printer/CD2,AECS/CD1,AECS/CD2},
+			height=3.3cm,
+			xlabel=Benchmark,
+			ylabel style={text width=1.4cm},
+			ylabel=Difference factor,
+			legend style={
+				anchor=north,
+				at={(-0.15,-3.9)},
+				legend columns=-1,
+			}
+%      nodes near coords
+		]
+			\nextgroupplot
+			\addplot coordinates {(HRTCS,1)};
+			\nextgroupplot
+			\addplot coordinates {(Motor, 3.54)};
+			\addplot coordinates {(Motor, 3.15)};
+			\addplot coordinates {(Motor, 4.46)};
+			\addplot coordinates {(Motor, 3.35)};
+			\addplot coordinates {(Motor, 2.81)};
+			\legend{Delay1,Delay2,Delay3,Delay4,Delay5}
+			\nextgroupplot
+			\addplot coordinates {(Printer/CD1, 1.6)};
+			\nextgroupplot
+			\addplot coordinates {(Printer/CD2, 1)};
+			\addplot coordinates {(Printer/CD2, 1)};
+			\nextgroupplot
+			\addplot coordinates {(AECS/CD1, 4.25)};
+			\nextgroupplot
+			\addplot coordinates {(AECS/CD2, 4.25)};
+			\addplot coordinates {(AECS/CD2, 4.25)};
+	\end{groupplot}
+\end{tikzpicture}
+\caption{Accuracy of the results}
+\label{fig:graph}
+\end{figure}
+
 % HRTCS, as already illustrated in Section \ref{sec:intr-motiv}, tests for an

### 2013-11-26 exps.tex
updated the graph

@@
 
-Previous results shows that the compiler was only able to find
-\textit{d} for HRTCS and Printer/CD2 without relaxing upper bound.
-Figure~\ref{fig:graph} shows the accuracy of our result by comparing
-differences between programmer specified upper bounds and the actual
-bounds obtained from our algorithm.
+The previous results show that HRTCS and Printer/CD2 were the only
+examples that the compiler was able to find \textit{d} without relaxing
+upper bound of \texttt{delay}. Figure~\ref{fig:graph} shows the accuracy
+of our result by comparing differences between programmer specified
+upper bounds and the actual bounds obtained from our algorithm.
 
@@
 	\centering
-	\begin{tikzpicture}[scale=0.95]
-	\begin{groupplot}[group style={group size=2 by 3,
-			xlabels at=edge bottom,
-			ylabels at=edge left,
-		},
-			footnotesize,ybar,xtick=data,
+\begin{tikzpicture}[scale=0.9]
+	\begin{axis}[
+			ybar,xtick=data,
 			symbolic x
-			coords={HRTCS,Motor,Printer/CD1,Printer/CD2,AECS/CD1,AECS/CD2},
-			height=3.3cm,
+			coords={HRTCS,Printer/CD1,Printer/CD2,AECS/CD1,AECS/CD2},
+			height=5cm,
+			width=9.5cm,
+%      xlabel=Benchmark,
+			x tick label style={rotate=45,font=\footnotesize},
+			ylabel=Difference factor,
+		]
+			\addplot coordinates {(HRTCS,1)(Printer/CD1,1.6)(Printer/CD2,1)(AECS/CD1,4.25)(AECS/CD2,4.25)};
+			\addplot coordinates {(Printer/CD2, 1)(AECS/CD2, 4.25)};
+	\end{axis}
+\end{tikzpicture}
+\begin{tikzpicture}[scale=0.9]
+	\begin{axis}[
+			ybar,xtick=data,
+			symbolic x
+			coords={Motor},
+			height=5cm,
+			width=9.5cm,
 			xlabel=Benchmark,
-			ylabel style={text width=1.4cm},
+			xlabel shift={4pt},
 			ylabel=Difference factor,
@@
 				anchor=north,
-				at={(-0.15,-3.9)},
+				at={(0.45,-0.44)},
 				legend columns=-1,
@@
 		]
-			\nextgroupplot
-			\addplot coordinates {(HRTCS,1)};
-			\nextgroupplot
 			\addplot coordinates {(Motor, 3.54)};
@@
 			\legend{Delay1,Delay2,Delay3,Delay4,Delay5}
-			\nextgroupplot
-			\addplot coordinates {(Printer/CD1, 1.6)};
-			\nextgroupplot
-			\addplot coordinates {(Printer/CD2, 1)};
-			\addplot coordinates {(Printer/CD2, 1)};
-			\nextgroupplot
-			\addplot coordinates {(AECS/CD1, 4.25)};
-			\nextgroupplot
-			\addplot coordinates {(AECS/CD2, 4.25)};
-			\addplot coordinates {(AECS/CD2, 4.25)};
-	\end{groupplot}
+	\end{axis}
 \end{tikzpicture}

### 2013-11-26 exps.tex
exprs

@@
 			width=9.5cm,
+			minor y tick num=1,
+			ymajorgrids=true,
+			ytick={0,...,5},
 %      xlabel=Benchmark,
@@
 		]
-			\addplot coordinates {(HRTCS,1)(Printer/CD1,1.6)(Printer/CD2,1)(AECS/CD1,4.25)(AECS/CD2,4.25)};
-			\addplot coordinates {(Printer/CD2, 1)(AECS/CD2, 4.25)};
+			\addplot coordinates
+			{(HRTCS,1)(Printer/CD1,1.6)(Printer/CD2,1)(AECS/CD1,4.25)(AECS/CD2,5.39)};
+			\addplot coordinates {(Printer/CD2, 1)(AECS/CD2, 5.39)};
 	\end{axis}
@@
 			ybar,xtick=data,
-			symbolic x
-			coords={Motor},
+			symbolic x coords={Motor},
+			ytick={0,...,5},
+			ymin={1.5},
+			minor y tick num=1,
+			ymajorgrids=true,
+%      extra y ticks={2.5},
+%      extra y tick labels={asdf},
+%      extra y tick style={grid=major},
 			height=5cm,
@@
 \end{figure}
+
+There are two main reasons that influence size of the relaxation: 1)
+execution speed of a target platform 2) and the nature of the program
+which runs on that specific platform. As one can easily speculate, clock
+frequency of the processor directly impacts on the value of WCRT. On the
+other hand, programmer can write a program such that large data
+computation in a single tick increases WCRT of the program. Assuming we
+do not change the architecture of the execution platform (i.e. JOP), the
+difference factor can only be reduced by modifying the benchmark
+programs by dividing the single longest tick, that corresponds to WCRT,
+into a number of smaller ticks.
 

### 2013-11-29 delays.tex
Updated text and added discussion text

@@
 \label{sec:semant-delay-stat}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[scale=0.6]{./sem}
+  \caption{Pictorial representation of the semantics of the delay
+    statement and its variants}
+  \label{fig:5}
+\end{figure}
 
@@
 
-The aforementioned variants are special cases of the general case
-\texttt{delay (M..N)}.
+Figure~\ref{fig:5} pictorially represents the semantics of these delay
+constructs. The construct $delay (M..N)$ requires that once executed the
+control-flow suspends execution for \textit{at\_least} $M$ units and
+\textit{at\_most} $N$ units, depicted as bounded box in
+Figure~\ref{fig:5}. There is no upper bound for the second delay form
+$delay(M)$, which requires that the control-flow suspends for
+\textit{at\_least} $M$ units, but the following statement can be
+executed after some countable delay, depicted with the unbounded box in
+Figure~\ref{fig:5}. Finally, $delay(M..M)$, requires that there is an
+\textit{exact} delay of $M$ units, i.e., the control-flow suspends
+execution for \textit{at\_least} and \textit{at\_most} $M$ units.
 
@@
 
-\subsection{The compiler flow}
-\label{sec:systemj-tool-chain}
+\subsection{The tool-chain flow}
+\label{sec:tool-chain-flow}
 
-\begin{figure}[t!]
+\begin{figure}[h!]
   \centering
-  \includegraphics[scale=0.5]{./tool}
+  \includegraphics[scale=0.45]{./tool.pdf}
+  \label{fig:4}
   \caption{The tool-chain flow of SystemJ compiler with delays}
-  \label{fig:4}
 \end{figure}
@@
 
-\subsection{Programming using the delay construct}
-\label{sec:progr-using-delay}
-
-In this section we provide a number of examples to show the different
-types of real-time programming paradigms that can be incorporated into
-the GALS (and its sub-set synchronous) programming model.
-
-\subsubsection{Non-deterministic time}
-\label{sec:non-determ-time}
-
-Many real-world systems require non-deterministic timing
-constructs. Where the exact real-time delay is not known or should not
-be known apriori. One such system is our motivating example -- the human
-response time system.
-
-\subsubsection{Timeout}
-\label{sec:timeout}
-
-There are many instances when one would like to wait on an input from
-environment for only a specified amount of time. This can be programmed
-as:
-% \begin{scriptsize}
-\begin{verbatim}
-// timeout after 1ms.
-trap(T){{await(A);}||
-        {delay(1..1 ms);exit(T);}}
-\end{verbatim}
-% \end{scriptsize}
-
-\subsubsection{Periodic reactions}
-\label{sec:periodic-reactions}
-
-A reaction, or a whole CD, can be programmed to run periodically like
-so:
-
-% \begin{scriptsize}
-\begin{verbatim}
-while(true) { delay ((1..1)ms); emit S; 
- //do something }
-\end{verbatim}
-% \end{scriptsize}
-A periodic reaction (or CD) requires special consideration. Since the
-delay statement is converted into \texttt{pause} constructs. One should
-not introduce extra \texttt{pause} constructs when building periodic
-reactions (or CDs). This is essential since introducing additional
-pauses would introduce more logical ticks.
-
-\subsubsection{Interaction of preemption and delays}
-\label{sec:inter-preempt-delays}
-
-Preemption plays an important role in reactive languages. One needs to
-carefully consider the interplay of \texttt{delay} semantics with the
-preemption semantics of reactive languages. Previous attempts at
-incorporating delays (using external timers) have only been partially
-successful, because of the complex interplay between real-time and
-preemption. Consider the simple example below, which models real-time
-using external timers as in~\cite{rsh94}.
-
-% \begin{scriptsize}
-\begin{verbatim}
-suspend(S) {emit START_TIMER(10); 
-            await (TIMER);emit O1;}
-\end{verbatim}
-% \end{scriptsize}
-
-As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
-with the external timer. The above program sends a signal to an external
-timer and waits for 10 ms to pass by. Consider what happens when signals
-\texttt{S} and \texttt{TIMER} occur in the same logical tick, the
-\texttt{await} statement is never executed (due to suspend) and hence,
-we enter a deadlock. Such problems are completely avoided in our
-technique, because we convert the real-time delays into logical delays
-(\texttt{pause} constructs), which interact well with preemption.
-
-\subsubsection{Interaction of channel communication and delays}
-\label{sec:inter-chann-comm}
-
-Channels, used for communication between reactions in asynchronously
-running CDs, are an addition in the SystemJ language. Like interaction
-of preemption and delays, conversion of real-time delays to logical
-ticks also interacts well with channel rendezvous, because the semantics
-of interaction are well defined~\cite{amal10}. More importantly, we need
-to consider the interplay of channel communication and WCRT/BCRT
-analysis. Since, channel communication does not stop logical time
-(see~\cite{amal10}) WCRT/BCRT are unaffected by channel
-communication. % The response time to input signals \textit{is} though!
-% But, we are unconcerned with the response time, analysis in this paper
-% and it remains a future research avenue.
+Our compiler has a switch that enables over-approximation as described
+in Section~\ref{sec:extend-tehcn-vari} if the delay requirements are not
+met. If this compiler switch is not enabled the compiler simply outputs
+an error indicating the relaxation required. The system designer can
+then either set the compiler switch of change the program by shortening
+the critical path by introducing \texttt{pause} constructs. We show how
+introducing \texttt{pause} constructs can help meet timing requirements
+explicitly, using examples, later on in
+Section~\ref{sec:experimental-results}.
 

### 2013-11-29 discussion.tex
Updated text and added discussion text

\section{Discussion -- addressing perceived limitations}
\label{sec:disc-perc-limit}

We dedicate this section to discuss the details of the delay semantics,
the dependence on time analyzable platforms, and the various uses and
programming paradigms that can be built using the $delay(M..N)$
construct.

\subsection{Programming using the delay construct}
\label{sec:progr-using-delay}

All three delay constructs are useful for programming real systems. % The
% first form of the delay construct $delay(M..N)$ is used to program
% non-determinism like in the HRCTS system and a real printer-spooler
% embedded system, etc. The second form is useful for programming low
% priority periodic tasks.

\subsubsection{Non-deterministic time}
\label{sec:non-determ-time}

Many real-world systems require non-deterministic timing
constructs. Where the exact real-time delay is not known or should not
be known apriori. One such system is our motivating example -- the human
response time system (HRCTS) another one we presented was a real
printer-spooler embedded controller in
Section~\ref{sec:experimental-results}.

\subsubsection{Timeout}
\label{sec:timeout}

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. This can be programmed
as:
% \begin{scriptsize}
\begin{verbatim}
// timeout after 1ms.
trap(T){{await(A);}||
        {delay(1 ms);exit(T);}}
\end{verbatim}
% \end{scriptsize}

In the above case, the program waits for signal \texttt{A} from the
environment for \textit{at\_least} \texttt{1 ms}, but it may wait longer
if other reactions/CDs are scheduled for execution. This second form of
delay statement us very useful for low priority tasks.

\subsubsection{Periodic reactions}
\label{sec:periodic-reactions}

The third form, is useful for hard-real time guarantees, e.g., a
reaction, or a whole CD, can be programmed to run periodically like so:

\begin{verbatim}
while(true) { delay ((1..1)ms); emit S; 
 //do something }
\end{verbatim}

The above program will emit signal \texttt{S} every \texttt{1 ms}. A
periodic reaction (or CD) requires special consideration, like any other
real-time periodic task. One should be aware of the interaction delays
and other \texttt{pause} constructs in the body of the reaction (or
CD). For example, consider the (perceived) periodic program and its
rewrite below.

\begin{verbatim}
while(true) {      while(true) {
 delay(1..1 ms);    while(true) { 
 //extra pause
 pause;              trap(T) {
 emit S;              int x = 0;
}                     while(true) {
                       x = x + 1;
                       pause;
                       if (x == d)exit(T);
                      }
                    }
                    //extra pause
                    pause;
                    emit S;
                   }
\end{verbatim}

The system designer needs to be aware that just introducing a delay
construct does not make the reaction (or CD) periodic. The program on
the left does \textit{not} emit signal \texttt{S} every \texttt{1 ms},
because of the extra \texttt{pause} statement. Introducing additional
pauses introduces more logical ticks consequently increasing the
real-time delay.

\subsection{Need for a time analyzable platform -- interaction of
  preemption and delays}
\label{sec:inter-preempt-delays}

As mentioned previously, the proposed solution might need
over-approximation techniques (Algorithm~\ref{alg:2}) to guarantee
timing requirements (cf. Section~\ref{sec:extend-tehcn-vari}). This
might give the readers an impression that timer based solution proposed
for standard real-time operating systems can be utilized for
GALS/synchronous programs as well. This is a fallacy. Bourke et
al~\cite{Bourke2009a} have investigated exactly this technique
unsuccessfully. We elaborate upon the interplay of timers and reactivity
to make the problem of using timers explicit.

Preemption plays an important role in reactive languages. One needs to
carefully consider the interplay of \texttt{delay} semantics with the
preemption semantics of reactive languages. Previous attempts at
incorporating delays (using external timers) have only been partially
successful, because of the complex interplay between real-time and
preemption. Consider the simple example below, which models real-time
using external timers as in~\cite{rsh94}. Furthermore, synchronous/GALS
languages have always been targeted at real-time analyzable platforms
and constrained
environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}, one can use
our technique within such a setting with ease.

% \begin{scriptsize}
\begin{verbatim}
suspend(S) {emit START_TIMER(10); 
            await (TIMER);emit O1;}
\end{verbatim}
% \end{scriptsize}

As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
with the external timer. The above program sends a signal to an external
timer and waits for 10 ms to pass by. Consider what happens when signals
\texttt{S} and \texttt{TIMER} occur in the same logical tick, the
\texttt{await} statement is never executed (due to suspend) and hence,
we enter a deadlock. Such problems are completely avoided in our
technique, because we convert the real-time delays into logical delays
(\texttt{pause} constructs), which interact well with preemption.

\subsection{Other semantic discussion}
\label{sec:other-semant-dissc}

\subsubsection{Interaction of channel communication and delays}
\label{sec:inter-chann-comm}

Channels, used for communication between reactions in asynchronously
running CDs, are an addition in the SystemJ language. Like interaction
of preemption and delays, conversion of real-time delays to logical
ticks also interacts well with channel rendezvous, because the semantics
of interaction are well defined~\cite{amal10}. More importantly, we need
to consider the interplay of channel communication and WCRT/BCRT
analysis. Since, channel communication does not stop logical time
(see~\cite{amal10}) WCRT/BCRT are unaffected by channel
communication. % The response time to input signals \textit{is} though!
% But, we are unconcerned with the response time, analysis in this paper
% and it remains a future research avenue.

\subsubsection{Interaction of synchronous parallelism and delay}
\label{sec:inter-synchr-parall}

{\color{red}Maybe we should talk about how the control flow in other
  reactions and CDs is not affected by the delay statement.}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

### 2013-11-29 motivation.tex
Updated paper
Updated text and added discussion text

@@
    abort(TOUCH){
-    {sustain GREEN_LIGHT;}
+    {sustain GREEN_LIGHT;} //reaction R1
     || // synchronous parallel operator
@@
      exit(TRAP); 
-    }
+    } // reaction R2
    } do {send S(1);} // Send 1 if touched 
   } do {send S(0);} // Else Send 0
-  pause; // only construct to consume time
+  pause; // only construct to consume logical time
  }
@@
 
-The SystemJ psuedocode~\cite{amal10} for such a system is shown in
+The SystemJ pseudo-code~\cite{amal10} for such a system is shown in
 Figure~\ref{fig:1}. We use the SystemJ language in this paper, first of

### 2013-11-29 delays.tex
Updated the paper

@@
 Figure~\ref{fig:5} pictorially represents the semantics of these delay
-constructs. The construct $delay (M..N)$ requires that once executed the
-control-flow suspends execution for \textit{at\_least} $M$ units and
-\textit{at\_most} $N$ units, depicted as bounded box in
-Figure~\ref{fig:5}. There is no upper bound for the second delay form
-$delay(M)$, which requires that the control-flow suspends for
-\textit{at\_least} $M$ units, but the following statement can be
-executed after some countable delay, depicted with the unbounded box in
-Figure~\ref{fig:5}. Finally, $delay(M..M)$, requires that there is an
-\textit{exact} delay of $M$ units, i.e., the control-flow suspends
-execution for \textit{at\_least} and \textit{at\_most} $M$ units.
+constructs. The construct \texttt{delay (M..N)} requires that upon
+execution the control-flow is suspended for \textit{at least} $M$ units
+and \textit{at most} $N$ units, depicted as bounded box in
+Figure~\ref{fig:5}. There is no upper bound for the second delay variant
+\texttt{delay(M)}, which requires that the control-flow suspends for
+\textit{at least} $M$ units, but the following statement can be executed
+after some countable delay, depicted with the unbounded box in
+Figure~\ref{fig:5}. Finally, \texttt{delay (M..M)} requires that there
+is an \textit{exact} delay of $M$ units, i.e., the control-flow suspends
+execution for \textit{at least} and \textit{at most} $M$ units.
 
@@
 logical tick delay \texttt{d}. The fundamental observation is that we
-have to delay for a minimum of $M$ units of real-time, hence,
+have to delay for at least $M$ units of real-time, hence,
 under-approximation is out of question. We can still over-approximate,
@@
 can be avoided by breaking the critical paths (worst case ticks) in the
-program (by inserting pause constructs), just like in hardware design,
-which is very \textit{unsurprising} considering Esterel has its origins
-in hardware.
+program (by inserting \texttt{pause} constructs), just like in hardware
+design, which is very \textit{unsurprising} considering Esterel has its
+origins in hardware.
 
@@
   \includegraphics[scale=0.45]{./tool.pdf}
+  \caption{The tool-chain flow of SystemJ compiler with delays}
   \label{fig:4}
-  \caption{The tool-chain flow of SystemJ compiler with delays}
 \end{figure}
@@
 Our compiler has a switch that enables over-approximation as described
-in Section~\ref{sec:extend-tehcn-vari} if the delay requirements are not
-met. If this compiler switch is not enabled the compiler simply outputs
-an error indicating the relaxation required. The system designer can
-then either set the compiler switch of change the program by shortening
-the critical path by introducing \texttt{pause} constructs. We show how
-introducing \texttt{pause} constructs can help meet timing requirements
-explicitly, using examples, later on in
-Section~\ref{sec:experimental-results}.
+in Section~\ref{sec:extend-tehcn-vari} provided the delay requirements
+are not met. If this compiler switch is not enabled, the compiler simply
+outputs an error indicating the relaxation required. The system designer
+can then either set the compiler switch or change the program by
+shortening the critical path by introducing \texttt{pause}
+constructs. We show how introducing \texttt{pause} constructs can help
+reduce the WCRT in Section~\ref{sec:experimental-results}.
 

### 2013-11-29 discussion.tex
Updated the paper

@@
 the dependence on time analyzable platforms, and the various uses and
-programming paradigms that can be built using the $delay(M..N)$
+programming paradigms that can be built using the \texttt{delay (M..N)}
 construct.
@@
 be known apriori. One such system is our motivating example -- the human
-response time system (HRCTS) another one we presented was a real
+response time system (HRCTS). Another one we presented was a real
 printer-spooler embedded controller in
@@
 if other reactions/CDs are scheduled for execution. This second form of
-delay statement us very useful for low priority tasks.
+delay statement is very useful for programming low priority tasks.
 
@@
 periodic reaction (or CD) requires special consideration, like any other
-real-time periodic task. One should be aware of the interaction delays
-and other \texttt{pause} constructs in the body of the reaction (or
-CD). For example, consider the (perceived) periodic program and its
+real-time periodic task. One should be aware of the interaction of
+delays and other \texttt{pause} constructs in the body of the reaction
+(or CD). For example, consider the (perceived) periodic program and its
 rewrite below.
@@
 because of the extra \texttt{pause} statement. Introducing additional
-pauses introduces more logical ticks consequently increasing the
+pauses introduces more logical ticks, consequently increasing the
 real-time delay.
@@
 might give the readers an impression that timer based solution proposed
-for standard real-time operating systems can be utilized for
-GALS/synchronous programs as well. This is a fallacy. Bourke et
-al~\cite{Bourke2009a} have investigated exactly this technique
+for standard real-time operating systems should be utilized for
+GALS/synchronous programs. This is a fallacy. Bourke et
+al~\cite{Bourke2009a} have investigated this technique
 unsuccessfully. We elaborate upon the interplay of timers and reactivity
@@
 preemption. Consider the simple example below, which models real-time
-using external timers as in~\cite{rsh94}. Furthermore, synchronous/GALS
-languages have always been targeted at real-time analyzable platforms
-and constrained
-environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}, one can use
-our technique within such a setting with ease.
+using external timers as in~\cite{rsh94}.
 
@@
 (\texttt{pause} constructs), which interact well with preemption.
+Furthermore, synchronous/GALS languages have always been targeted at
+real-time analyzable platforms and constrained
+environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}, one can use
+our technique within such a setting with ease.
 
@@
 
-\subsubsection{Interaction of synchronous parallelism and delay}
-\label{sec:inter-synchr-parall}
+% \subsubsection{Interaction of synchronous parallelism and delay}
+% \label{sec:inter-synchr-parall}
 
-{\color{red}Maybe we should talk about how the control flow in other
-  reactions and CDs is not affected by the delay statement.}
+% {\color{red}Maybe we should talk about how the control flow in other
+%   reactions and CDs is not affected by the delay statement.}
 

### 2013-11-29 exps.tex
Updated text and added discussion text
Updated the paper

@@
 
-{\renewcommand{\arraystretch}{0.9}
-\begin{table}[t]
-	\small
-	\centering
-	\caption{An overview of the benchmark programs}
-	\label{fig:benchlist}
-	\scalebox{0.9}{
-		\begin{tabular}{llp{0.5cm}p{1cm}}
-\toprule
-\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
-\emph{Lines of code}        
-\\ \midrule
-HRTCS                  & \texttt{delay(50.3..200.3 ms)} ($CD1$) & 2                  & 56
-\\ \midrule
-\multirow{5}{*}{Motor} & \texttt{delay(2.4..2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
-											 & \texttt{delay(1.667..1.667 ms)}      &                    &                      \\ 
-											 & \texttt{delay(0.05..0.05 ms)}       &                    &                      \\ 
-											 & \texttt{delay(0.3..0.3 ms)}        &                    &                      \\ 
-											 & \texttt{delay(0.733..0.733 ms)}      &                    &
-\\ \midrule
-\multirow{3}{*}{Printer}
-& \texttt{delay(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
-& \texttt{delay(1.2..30.3 ms)} ($CD2$)		  &														&											\\
-& \texttt{delay(5.6..100.2 ms)} ($CD2$)     &														&											
-\\ \midrule
-\multirow{2}{*}{AECS} & \texttt{delay(10000 ms)} ($CD1$) 
-& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
-& \texttt{delay(10000..10000 ms)} ($CD2$) & &
-\\ \bottomrule
-\end{tabular}
-}
-\vspace{-20pt}
-\end{table}
-}
-
-{\renewcommand{\arraystretch}{0.6}
-\begin{table}[t]
-\twocolumn[
-\centering
-\begin{@twocolumnfalse}
-	\caption{Actual delays obtained for \texttt{delay} constructs in the
-benchmark programs based on their BCRT and WCRT}
-\label{fig:comparison}
-\begin{tabular}{ c c l l l l l l }
-	\toprule
-	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
-	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
-	\midrule
-	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0334 ms	& 0.0361 ms & 0.009613 ms &
-	0.0263 ms & 0.1683 ms & 0.1881 ms\\ 
-	\cmidrule(r){1-8}
-	\multicolumn{2}{c}{\emph{WCRT}} &0.1120 ms & 0.1116 ms & 0.0603 ms
-	& 0.0491 ms & 0.8148 ms & 1.0115 ms\\ 
-	\cmidrule(r){1-8}
-
-	\multirow{2}{*}{Delay 1} & M..N & 50.3-200.3 ms	& 2.4-8.4882 ms	  &
-	2.7-16.9378 ms & 1.2-30.3 ms & 10000-42483.2198 ms& 10000-53882.0045 ms \\
-	\cmidrule(r){2-8}
-	& d &1506-1888 & 68 & 281 & 46-617 & 59428 &53160\\ 
-	\cmidrule(r){1-8}
-	\multirow{2}{*}{Delay 2} & M..N & N/A & 1.668-5.2452 ms & N/A &
-	5.6-100.2 ms & N/A &10000-53882.0045 ms\\ 
-	\cmidrule(r){2-8}
-	& d & N/A & 48 & N/A & 213-2041 & N/A &53160\\ 
-	\cmidrule(r){1-8}
-	\multirow{2}{*}{Delay 3} & M..N & N/A & 0.05-0.2232 ms & N/A & N/A & N/A
-	&N/A\\ 
-	\cmidrule(r){2-8}
-	& d & N/A & 2 & N/A & N/A & N/A &N/A\\ 
-	\cmidrule(r){1-8}
-	\multirow{2}{*}{Delay 4} & M..N & N/A & 0.3-1.0044 ms & N/A & N/A & N/A
-	&N/A\\ 
-	\cmidrule(r){2-8}
-	& d & N/A & 9 & N/A & N/A & N/A &N/A\\ 
-	\cmidrule(r){1-8}
-	\multirow{2}{*}{Delay 5} & M..N & N/A & 0.833-2.3436 ms & N/A & N/A & N/A
-	&N/A\\ 
-	\cmidrule(r){2-8}
-	& d & N/A & 21 & N/A
-	& N/A & N/A &N/A\\ \bottomrule
-\end{tabular}
-\end{@twocolumnfalse}
-]
-\end{table}
-}
-
 We have carried out a number of experiments on a set of applications
@@
 Control System) is an application that controls an intelligent room
-environment described in~\cite{aecs_ispa}.  The system interacts with
-two external timers through input/output signals in order to measure how
-long the entrance door to the room is being opened and time the illegal
-presence (e.g. intruder) that is being detected. The program has been
-modified such that the delay constructs are used instead of external
-timers as shown in Figure~\ref{delaycode}. In the original AECS, the
-system needed to trigger external timer by emitting the output signal in
-order to wait for a specific amount of time (e.g.
+environment described in~\cite{aecs_ispa}. The system interacts with two
+external timers through input/output signals in order to measure how
+long the entrance door to the room is being opened to detect intruders.
+The program has been modified such that the delay constructs are used
+instead of external timers as shown in Figure~\ref{delaycode}. In the
+original AECS, the system needed to trigger external timer by emitting
+the output signal in order to wait for a specific amount of time (e.g.
 \texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
-expires, it notifies back to the system through the input signal
-\texttt{TimeOut} which is polled at the \texttt{await} statement. With
-the new delay constructs, those timers are not necessary and delays can
-be expressed directly in the code as shown in Figure~\ref{delaycode:b}.
+expires, it notifies the system through the input signal
+\texttt{TimeOut}, which is polled using the \texttt{await}
+statement. With the new delay constructs, these timers are not necessary
+and delays can be expressed directly in the code as shown in
+Figure~\ref{delaycode:b}.
 
 \begin{figure}[h!]
-	\begin{SubFloat}{\label{delaycode:a}}
+  \begin{SubFloat}{\label{delaycode:a}Original program}
     \centering
@@
   \hspace{1cm}%
-	\begin{SubFloat}{\label{delaycode:b}}
+  \begin{SubFloat}{\label{delaycode:b}Program with delay constructs}
     \centering
@@
 of current to a print head (a row of resistors). In this example,
-duration of current applied is controlled by the feedback logic using
-real-time delays which are expressed in Esterel constructs. We have
-rewritten the program in SystemJ and compared between desired and
-relaxed delays according to WCRT and BCRT of the program. Printer is an
-example from~\cite{Schneider:1999:CRT:555233}, which models a printer and
-a spooler in timed CSP.  Their operation is rather simple: as soon as
-the spooler receives a print job, passes it to the printer through a
-channel where actual printing process take place. The example
-demonstrates that by specifying delay bounds (e.g. 1 to 5 time units) on
-receiving and sending channel ends data exchange will occur as soon as
-both processes become ready.
+duration of the current applied is controlled by the feedback logic
+using real-time delays, which are expressed using reactive
+constructs. Printer is an example from~\cite{Schneider:1999:CRT:555233},
+which models a printer and a spooler in timed CSP.  Their operation is
+rather simple: as soon as the spooler receives a print job, passes it to
+the printer through a channel where actual printing process take
+place. This example along with HRCTS demonstrates the use of
+non-deterministic delays.
 
@@
 intruder or generates a beep sound when an entrance door is being opened
-for the same amount of time. In Printer, each spooler and printer
-process is mapped to a single clock-domain. Spooler CD sends a job to
-the Printer CD between 2.7 to 10.6 ms after receiving it from the
-enviornment (e.g. a user). On the other hand, Printer CD notifies
+for the same amount of time. In the Printer example, spooler and printer
+processes are mapped to separate CDs. Spooler CD sends a job to the
+Printer CD between 2.7 to 10.6 ms after receiving it from the
+environment (e.g. a user). On the other hand, Printer CD notifies
 environment that it starts printing between 1.2 to 30.3 ms once it
@@
 within the specified amount of time. Lastly, Motor is a synchronous
-program with 5 delay statements where they controls amount of current
-into the stepper motor coils.
+program with 5 delay statements that control the amount of current
+applied to the stepper motor coils.
 
@@
 upper real-time bounds even for bounded (\texttt{M..N}) or exact
-(\texttt{M..M}) delays when it fails to find a feasible solution.
-However, a user can prohibit this behavior through configuration that
-such case is an error.
+(\texttt{M..M}) delays when it failed to find a feasible solution.
+However, a user can prohibit such behavior using compiler switch as
+described in Section~\ref{sec:tool-chain-flow}.
 
@@
 1787 (as shown previously in
-Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}). Similarly,
-Printer/CD2 satisfies real-time ranges of 1.2..30.3 (ms) and 5.6..100.2
-(ms) when \emph{d} is 46..617 and 213..2041 respectively. It is then
-compiler's job to statically choose one of the values in this range.
+Section~\ref{sec:find-logic-delay}). Similarly, Printer/CD2 satisfies
+real-time ranges of 1.2..30.3 (ms) and 5.6..100.2 (ms) when \emph{d} is
+46..617 and 213..2041, respectively. It is then the compiler's job to
+statically choose one of the values in this range.
 
@@
 where relaxation was required, to determine the minimum upper real-time
-bound \emph{N} such that there exist a solution for \emph{D} in
+bound \emph{N} such that there exists a solution for \emph{D} in
 Algorithm \ref{alg:1}. This algorithm gives a unique number for logical
@@
 upper bound of \texttt{delay}. Figure~\ref{fig:graph} shows the accuracy
-of our result by comparing differences between programmer specified
-upper bounds and the actual bounds obtained from our algorithm.
+of our results by comparing differences between programmer specified
+upper bounds (Table~\ref{fig:benchlist}) and the actual bounds obtained
+using our techniques (Table~\ref{fig:comparison}).
 
@@
 
-There are two main reasons that influence size of the relaxation: 1)
-execution speed of a target platform 2) and the nature of the program
-which runs on that specific platform. As one can easily speculate, clock
-frequency of the processor directly impacts on the value of WCRT. On the
-other hand, programmer can write a program such that large data
-computation in a single tick increases WCRT of the program. Assuming we
-do not change the architecture of the execution platform (i.e. JOP), the
-difference factor can only be reduced by modifying the benchmark
-programs by dividing the single longest tick, that corresponds to WCRT,
-into a number of smaller ticks.
+There are two influential factors in determining the relaxation: 1)
+execution speed of a target platform and 2) the nature of the program
+that runs on this platform. Clock frequency of the processor directly
+impacts the value of WCRT. On the other hand, programmer can write a
+program such that large data computations in a single logical tick
+increase WCRT of the program. Assuming we do not change the architecture
+of the execution platform (i.e. JOP), the difference factor can only be
+reduced by modifying the benchmark programs such that the critical paths
+are broken up using \texttt{pause} constructs, thereby reducing the
+WCRT.
 
@@
 % >>>>>>> 4bffc4b1a43388f5ebf6cc737b9d3b246ee77dc7
+{\renewcommand{\arraystretch}{0.9}
+\begin{table}[t]
+	\small
+	\centering
+	\caption{An overview of the benchmark programs}
+	\label{fig:benchlist}
+	\scalebox{0.9}{
+		\begin{tabular}{llp{0.5cm}p{1cm}}
+\toprule
+\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
+\emph{Lines of code}        
+\\ \midrule
+HRTCS                  & \texttt{delay(50.3..200.3 ms)} ($CD1$) & 2                  & 56
+\\ \midrule
+\multirow{5}{*}{Motor} & \texttt{delay(2.4..2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
+											 & \texttt{delay(1.667..1.667 ms)}      &                    &                      \\ 
+											 & \texttt{delay(0.05..0.05 ms)}       &                    &                      \\ 
+											 & \texttt{delay(0.3..0.3 ms)}        &                    &                      \\ 
+											 & \texttt{delay(0.733..0.733 ms)}      &                    &
+\\ \midrule
+\multirow{3}{*}{Printer}
+& \texttt{delay(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
+& \texttt{delay(1.2..30.3 ms)} ($CD2$)		  &														&											\\
+& \texttt{delay(5.6..100.2 ms)} ($CD2$)     &														&											
+\\ \midrule
+\multirow{2}{*}{AECS} & \texttt{delay(10000 ms)} ($CD1$) 
+& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
+& \texttt{delay(10000..10000 ms)} ($CD2$) & &
+\\ \bottomrule
+\end{tabular}
+}
+\vspace{-20pt}
+\end{table}
+}
+
+{\renewcommand{\arraystretch}{0.6}
+\begin{table}[t!]
+\twocolumn[
+\centering
+\begin{@twocolumnfalse}
+	\caption{Actual delays obtained for \texttt{delay} constructs in the
+benchmark programs based on their BCRT and WCRT}
+\begin{tabular}{ c c l l l l l l }
+	\toprule
+	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
+	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
+	\midrule
+	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0334 ms	& 0.0361 ms & 0.009613 ms &
+	0.0263 ms & 0.1683 ms & 0.1881 ms\\ 
+	\cmidrule(r){1-8}
+	\multicolumn{2}{c}{\emph{WCRT}} &0.1120 ms & 0.1116 ms & 0.0603 ms
+	& 0.0491 ms & 0.8148 ms & 1.0115 ms\\ 
+	\cmidrule(r){1-8}
+
+	\multirow{2}{*}{Delay 1} & M..N & 50.3-200.3 ms	& 2.4-8.4882 ms	  &
+	2.7-16.9378 ms & 1.2-30.3 ms & 10000-42483.2198 ms& 10000-53882.0045 ms \\
+	\cmidrule(r){2-8}
+	& d &1506-1888 & 68 & 281 & 46-617 & 59428 &53160\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 2} & M..N & N/A & 1.668-5.2452 ms & N/A &
+	5.6-100.2 ms & N/A &10000-53882.0045 ms\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 48 & N/A & 213-2041 & N/A &53160\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 3} & M..N & N/A & 0.05-0.2232 ms & N/A & N/A & N/A
+	&N/A\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 2 & N/A & N/A & N/A &N/A\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 4} & M..N & N/A & 0.3-1.0044 ms & N/A & N/A & N/A
+	&N/A\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 9 & N/A & N/A & N/A &N/A\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 5} & M..N & N/A & 0.833-2.3436 ms & N/A & N/A & N/A
+	&N/A\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 21 & N/A
+	& N/A & N/A &N/A\\ \bottomrule
+\end{tabular}
+\label{fig:comparison}
+\end{@twocolumnfalse}
+]
+\end{table}
+}
 

### 2013-11-29 related_work.tex
Updated the paper

@@
 
-Other works such as Quartz~\cite{glog02} and Bertin
-et.al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate delay
+Other works such as Quartz~\cite{glog02} and Bertin et
+al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate delay
 statements in Esterel using \texttt{pause} and special \textit{pragmas}
@@
 studying timing properties using model checkers rather than
-implementation, like us. The solution introduced by Bertin
-et.al.~\cite{Bertin:2000:TVR:1947412.1947439} translates real-time
-pragmas into Timed automata~\cite{alur94} for model-checking real-time
+implementation, like us. The solution introduced by Bertin et
+al.~\cite{Bertin:2000:TVR:1947412.1947439} translates real-time pragmas
+into Timed automata~\cite{alur94} for model-checking real-time
 deadlines. Our solution on the other hand removes all continuous time
@@
 tools developed for reactive languages and more in the spirit of logical
-time. Finally, a major difference between Bertin et.al and our approach
+time. Finally, a major difference between Bertin et al and our approach
 is that their approach requires programmers to annotate the time taken

### 2013-12-02 exps.tex
Updated exps

@@
 \label{sec:experimental-results}
-% <<<<<<< HEAD
-% \begin{table}[t]
-%   \centering
-%   \twocolumn[
-
-%   \begin{@twocolumnfalse}
-%     \caption{Benchmark programs using \texttt{delay} constructs} \label{table:benchmark}
-%     \scalebox{0.85}{
-%       \begin{tabular}{|c| c | c | c | c | c | c | c |}
-% 	\cline{4-8}
-% 	\multicolumn{3}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
-% 	\multirow{12}{*}{JOP} 		& \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{2-8}   
-%         & \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{2-8} 
-%         & \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{3-8} 
-%         &			   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{2-8}
-%         & \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{3-8} 
-%         &			   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{2-8}
-%         & \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
-%         &			   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
-%         & \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
-%         &			   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
-%         & \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
-%         &			   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
-% 	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}   
-%         & \multicolumn{2}{|c|}{WCRT} 		&0.0329 ms 		& 0.0393 ms		  & 0.0361 ms		  & 0.5072 ms		   & 0.8388 ms\\ \cline{2-8} 
-%         & \multirow{2}{*}{Delay 1}	 & M..N &50.3-600.3 ms 	& 2.4-15.6265 ms  & 1230-42317.8276 ms& 10000-95358.8578 ms& 10000-287757.9834 ms \\ \cline{3-8} 
-%         &			   & d			 		&18209-18232 	& 398 			  & 1171429			  & 188015		       &343054\\ \cline{2-8}
-%         & \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-10.8757 ms& N/A				  & N/A 			   &10000-287757.9834 ms\\ \cline{3-8} 
-%         &			   & d			 		& 	N/A	 		& 277			  & N/A				  & N/A       		   &343054\\ \cline{2-8}
-%         & \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.3534 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
-%         &			   & d			 		& 	N/A	 		& 9		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
-%         & \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.9631 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
-%         &			   & d				 	& 	N/A	 		& 50			  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
-%         & \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-4.79 ms   & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
-%         &			   & d			 		& 	N/A	 		& 122 			  &	N/A				  & N/A      		   &N/A\\ \cline{2-8}
-% 	\hline
-%       \end{tabular}
-%     }
-%   \end{@twocolumnfalse}
-%   ]
 
@@
 Table~\ref{fig:benchlist}.
+{\renewcommand{\arraystretch}{0.9}
+\begin{table}[h!]
+	\small
+	\centering
+	\caption{An overview of the benchmark programs}
+	\label{fig:benchlist}
+	\scalebox{0.9}{
+		\begin{tabular}{llp{0.5cm}p{1cm}}
+\toprule
+\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
+\emph{Lines of code}        
+\\ \midrule
+HRTCS                  & \texttt{delay(50.3..200.3 ms)} ($CD1$) & 2                  & 56
+\\ \midrule
+\multirow{5}{*}{Motor} & \texttt{delay(2.4..2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
+											 & \texttt{delay(1.667..1.667 ms)}      &                    &                      \\ 
+											 & \texttt{delay(0.05..0.05 ms)}       &                    &                      \\ 
+											 & \texttt{delay(0.3..0.3 ms)}        &                    &                      \\ 
+											 & \texttt{delay(0.733..0.733 ms)}      &                    &
+\\ \midrule
+\multirow{3}{*}{Printer}
+& \texttt{delay(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
+& \texttt{delay(1.2..30.3 ms)} ($CD2$)		  &														&											\\
+& \texttt{delay(5.6..100.2 ms)} ($CD2$)     &														&											
+\\ \midrule
+\multirow{2}{*}{AECS} & \texttt{delay(10000 ms)} ($CD1$) 
+& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
+& \texttt{delay(10000..10000 ms)} ($CD2$) & &
+\\ \bottomrule
+\end{tabular}
+}
+\end{table}
+}
+{\renewcommand{\arraystretch}{0.6}
+\begin{table*}[t!]
+\centering
+	\caption{Actual delays obtained for \texttt{delay} constructs in the
+benchmark programs based on their BCRT and WCRT}
+\begin{tabular}{ c c l l l l l l }
+	\toprule
+	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
+	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
+	\midrule
+	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0.0396 ms	& 0.0425 ms & 0.0190 ms &
+	0.0263 ms & 0.1784 ms & 0.1284 ms\\ 
+	\cmidrule(r){1-8}
+	\multicolumn{2}{c}{\emph{WCRT}} &0.1152 ms & 0.1096 ms & 0.0647 ms
+	& 0.0491 ms & 0.3476 ms & 0.2991 ms\\ 
+	\cmidrule(r){1-8}
+
+	\multirow{2}{*}{Delay 1} & M..N & 50.3-200.3 ms	& 2.4-6.2493 ms	  &
+	2.7-10.6 ms & 1.2-30.3 ms & 10000-19484.3481 ms& 10000-23278.2752 ms \\
+	\cmidrule(r){2-8}
+	& d &1272-1739 & 57 & 143-163 & 46-617 & 56062 &77844\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 2} & M..N & N/A & 1.668-4.3855 ms & N/A &
+	5.6-100.2 ms & N/A &10000-23278.2752 ms\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 40 & N/A & 213-2041 & N/A &53160\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 3} & M..N & N/A & 0.05-0.2193 ms & N/A & N/A & N/A
+	&N/A\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 2 & N/A & N/A & N/A &N/A\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 4} & M..N & N/A & 0.3-0.8771 ms & N/A & N/A & N/A
+	&N/A\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 8 & N/A & N/A & N/A &N/A\\ 
+	\cmidrule(r){1-8}
+	\multirow{2}{*}{Delay 5} & M..N & N/A & 0.733-1.9735 ms & N/A & N/A & N/A
+	&N/A\\ 
+	\cmidrule(r){2-8}
+	& d & N/A & 18 & N/A
+	& N/A & N/A &N/A\\ \bottomrule
+\end{tabular}
+\label{fig:comparison}
+\end{table*}
+}
 
@@
 Algorithm~\ref{alg:1}. In HRTCS, real-time delay of the range
-50.3..200.3 (ms) is only satisfied when \emph{d} is between 1506 and
-1787 (as shown previously in
-Section~\ref{sec:find-logic-delay}). Similarly, Printer/CD2 satisfies
-real-time ranges of 1.2..30.3 (ms) and 5.6..100.2 (ms) when \emph{d} is
-46..617 and 213..2041, respectively. It is then the compiler's job to
-statically choose one of the values in this range.
+50.3..200.3 (ms) is only satisfied when \emph{d} is between 1272 and
+1739 (as shown previously in Section~\ref{sec:find-logic-delay}). Delay
+2.7..10.6 (ms) is satisfied in Printer/CD1 with \emph{d} of 143..163.
+Similarly, Printer/CD2 satisfies real-time ranges of 1.2..30.3 (ms) and
+5.6..100.2 (ms) when \emph{d} is 46..617 and 213..2041, respectively. It
+is then the compiler's job to statically choose one of the values in
+this range.
 
@@
 Max(\emph{$S_1$}). The compiler, therefore, will choose \emph{d} based
-on the newly computed \emph{N} for the examples: Printer/CD1, Motor and
-AECS.
+on the newly computed \emph{N} for the examples: Motor and AECS.
 
@@
 
-The previous results show that HRTCS and Printer/CD2 were the only
-examples that the compiler was able to find \textit{d} without relaxing
-upper bound of \texttt{delay}. Figure~\ref{fig:graph} shows the accuracy
-of our results by comparing differences between programmer specified
-upper bounds (Table~\ref{fig:benchlist}) and the actual bounds obtained
-using our techniques (Table~\ref{fig:comparison}).
+The previous results show that HRTCS and Printer were the only examples
+that the compiler was able to find \textit{d} without relaxing upper
+bound of \texttt{delay}. 
+This is because there are two influential factors in
+determining the relaxation: 1) execution speed of a target platform and 
+%Figure~\ref{fig:graph} shows the accuracy
+%of our results by comparing differences between programmer specified
+%upper bounds (Table~\ref{fig:benchlist}) and the actual bounds obtained
+%using our techniques (Table~\ref{fig:comparison}).
+2) the nature of the program that runs on this platform. Clock frequency
+of the processor directly impacts the value of WCRT. On the other hand,
+programmer can write a program such that large data (or control)
+computations in a single logical tick increase WCRT of the program.
+Assuming we do not change the architecture of the execution platform
+(i.e. JOP), the size of relaxation can only be reduced by modifying the
+benchmark programs such that the critical paths are broken up using
+\texttt{pause} constructs, thereby reducing the WCRT.
+Figure~\ref{fig:improv} shows segments of code from AECS whose WCRT is
+reduced by such technique.
 
@@
 	\centering
+	\begin{SubFloat}{\label{fig:improv-a}Original program}
+    \begin{minipage}[b]{0.3\linewidth}
+			\scriptsize
+\begin{verbatim}
+present(EntryRequest)
+  id = #EntryRequest;
+else
+  id = #ExitRequest;
+if(E.isValiedRequest(id)){
+  // code ..
+}
+\end{verbatim}
+\end{minipage}
+\end{SubFloat}
+  \hspace{1cm}%
+  \begin{SubFloat}{\label{fig:improv-b}Program with reduced WCRT}
+    \centering
+    \begin{minipage}[b]{0.35\linewidth}
+      \scriptsize
+\begin{verbatim}
+present(EntryRequest)
+  id = #EntryRequest;
+else
+  id = #ExitRequest;
+pause; // Inserting pause
+if(E.isValiedRequest(id)){
+  // code ..
+}
+\end{verbatim}
+		\end{minipage}
+	\end{SubFloat}
+	\caption{Improving WCRT by shortening a critical path}
+	\label{fig:improv}
+\end{figure}
+
+In Figure~\ref{fig:improv-a}, the program simply checks for the id of a
+person entering or exiting an intelligent room through the signals
+called \texttt{EntryRequest} and \texttt{ExitRequest}, and verifies the
+id using data logic written in Java (\texttt{isValidRequest}). We
+identified this execution path caused lengthening the WCRT of the
+program. In this case, we divide two consequtive statements, i.e.
+\texttt{present} and \texttt{if}, by inserting \texttt{pause} such that
+each of these statement is executed in two different ticks. The same
+technique is applied to Motor example to reduce WCRT of the program.
+Note that we can also optimize algorithms of the data computations used
+in each Java method to reduce WCRT. However it is out of scope of this
+paper, hence not discussed here.
+
+The result of the upper bound relaxation with reduced WCRT is shown in
+Figure~\ref{fig:graph}. As one can see from the Figure, the upper bounds
+for delay statements in AECS/CD2 has been decreased by 11.81\% which is
+higher than all those in Motor which ranges from 7.54 to 8.05\%. It is
+because AECS has relatively a large amount of data computation while
+Motor is a completely control-dominated program. Therefore, the result
+shows that the size of the relaxation depends on how designers write
+their program as well as types of the application (i.e. data or control
+dominated).
+
+%For AECS/CD2, the upper bound has been reduced
+%by 11.81\%, whereas for Motor, the reduced size ranges from 7.54 to
+%8.05\% for each delay statement.
+
+\begin{figure}[ht!]
+	\centering
+%\begin{tikzpicture}[scale=0.9]
+%  \begin{axis}[
+%      ybar,xtick=data,
+%      symbolic x
+%      coords={AECS/CD2},
+%      height=5cm,
+%      width=9.5cm,
+%      minor y tick num=1,
+%      yminorgrids=true,
+%      ymajorgrids=true,
+%      xlabel=Benchmark,
+%      ylabel=Percentage reduction,
+%    ]
+%      \addplot coordinates
+%      {(AECS/CD2, 11.82)};
+%  \end{axis}
+%\end{tikzpicture}
 \begin{tikzpicture}[scale=0.9]
@@
 			ybar,xtick=data,
-			symbolic x
-			coords={HRTCS,Printer/CD1,Printer/CD2,AECS/CD1,AECS/CD2},
-			height=5cm,
-			width=9.5cm,
+			symbolic x coords={Motor,AECS/CD2},
+			ytick={0,1,...,12},
+			ymin={6},
 			minor y tick num=1,
+			enlarge x limits=0.5,
+			enlarge y limits=0.1,
+			yminorgrids=true,
 			ymajorgrids=true,
-			ytick={0,...,5},
-%      xlabel=Benchmark,
-			x tick label style={rotate=45,font=\footnotesize},
-			ylabel=Difference factor,
-		]
-			\addplot coordinates
-			{(HRTCS,1)(Printer/CD1,1.6)(Printer/CD2,1)(AECS/CD1,4.25)(AECS/CD2,5.39)};
-			\addplot coordinates {(Printer/CD2, 1)(AECS/CD2, 5.39)};
-	\end{axis}
-\end{tikzpicture}
-\begin{tikzpicture}[scale=0.9]
-	\begin{axis}[
-			ybar,xtick=data,
-			symbolic x coords={Motor},
-			ytick={0,...,5},
-			ymin={1.5},
-			minor y tick num=1,
-			ymajorgrids=true,
-%      extra y ticks={2.5},
-%      extra y tick labels={asdf},
-%      extra y tick style={grid=major},
-			height=5cm,
+			height=7cm,
 			width=9.5cm,
@@
 			xlabel shift={4pt},
-			ylabel=Difference factor,
+			ylabel=Percentage reduction (\%),
 			legend style={
 				anchor=north,
-				at={(0.45,-0.44)},
+				at={(0.45,-0.29)},
 				legend columns=-1,
 			}
-%      nodes near coords
 		]
-			\addplot coordinates {(Motor, 3.54)};
-			\addplot coordinates {(Motor, 3.15)};
-			\addplot coordinates {(Motor, 4.46)};
-			\addplot coordinates {(Motor, 3.35)};
-			\addplot coordinates {(Motor, 2.81)};
+			\addplot coordinates {(Motor, 8.05) (AECS/CD2, 11.82)};
+			\addplot coordinates {(Motor, 8.00)};
+			\addplot coordinates {(Motor, 6.42)};
+			\addplot coordinates {(Motor, 7.53)};
+			\addplot coordinates {(Motor, 7.89)};
 			\legend{Delay1,Delay2,Delay3,Delay4,Delay5}
@@
 \end{tikzpicture}
-\caption{Accuracy of the results}
+\caption{Reduction in upper bound relaxation with improved WCRT}
 \label{fig:graph}
 \end{figure}
-
-There are two influential factors in determining the relaxation: 1)
-execution speed of a target platform and 2) the nature of the program
-that runs on this platform. Clock frequency of the processor directly
-impacts the value of WCRT. On the other hand, programmer can write a
-program such that large data computations in a single logical tick
-increase WCRT of the program. Assuming we do not change the architecture
-of the execution platform (i.e. JOP), the difference factor can only be
-reduced by modifying the benchmark programs such that the critical paths
-are broken up using \texttt{pause} constructs, thereby reducing the
-WCRT.
 
@@
 % >>>>>>> 4bffc4b1a43388f5ebf6cc737b9d3b246ee77dc7
-{\renewcommand{\arraystretch}{0.9}
-\begin{table}[t]
-	\small
-	\centering
-	\caption{An overview of the benchmark programs}
-	\label{fig:benchlist}
-	\scalebox{0.9}{
-		\begin{tabular}{llp{0.5cm}p{1cm}}
-\toprule
-\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
-\emph{Lines of code}        
-\\ \midrule
-HRTCS                  & \texttt{delay(50.3..200.3 ms)} ($CD1$) & 2                  & 56
-\\ \midrule
-\multirow{5}{*}{Motor} & \texttt{delay(2.4..2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
-											 & \texttt{delay(1.667..1.667 ms)}      &                    &                      \\ 
-											 & \texttt{delay(0.05..0.05 ms)}       &                    &                      \\ 
-											 & \texttt{delay(0.3..0.3 ms)}        &                    &                      \\ 
-											 & \texttt{delay(0.733..0.733 ms)}      &                    &
-\\ \midrule
-\multirow{3}{*}{Printer}
-& \texttt{delay(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
-& \texttt{delay(1.2..30.3 ms)} ($CD2$)		  &														&											\\
-& \texttt{delay(5.6..100.2 ms)} ($CD2$)     &														&											
-\\ \midrule
-\multirow{2}{*}{AECS} & \texttt{delay(10000 ms)} ($CD1$) 
-& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
-& \texttt{delay(10000..10000 ms)} ($CD2$) & &
-\\ \bottomrule
-\end{tabular}
-}
-\vspace{-20pt}
-\end{table}
-}
-
-{\renewcommand{\arraystretch}{0.6}
-\begin{table}[t!]
-\twocolumn[
-\centering
-\begin{@twocolumnfalse}
-	\caption{Actual delays obtained for \texttt{delay} constructs in the
-benchmark programs based on their BCRT and WCRT}
-\begin{tabular}{ c c l l l l l l }
-	\toprule
-	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
-	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
-	\midrule
-	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0334 ms	& 0.0361 ms & 0.009613 ms &
-	0.0263 ms & 0.1683 ms & 0.1881 ms\\ 
-	\cmidrule(r){1-8}
-	\multicolumn{2}{c}{\emph{WCRT}} &0.1120 ms & 0.1116 ms & 0.0603 ms
-	& 0.0491 ms & 0.8148 ms & 1.0115 ms\\ 
-	\cmidrule(r){1-8}
-
-	\multirow{2}{*}{Delay 1} & M..N & 50.3-200.3 ms	& 2.4-8.4882 ms	  &
-	2.7-16.9378 ms & 1.2-30.3 ms & 10000-42483.2198 ms& 10000-53882.0045 ms \\
-	\cmidrule(r){2-8}
-	& d &1506-1888 & 68 & 281 & 46-617 & 59428 &53160\\ 
-	\cmidrule(r){1-8}
-	\multirow{2}{*}{Delay 2} & M..N & N/A & 1.668-5.2452 ms & N/A &
-	5.6-100.2 ms & N/A &10000-53882.0045 ms\\ 
-	\cmidrule(r){2-8}
-	& d & N/A & 48 & N/A & 213-2041 & N/A &53160\\ 
-	\cmidrule(r){1-8}
-	\multirow{2}{*}{Delay 3} & M..N & N/A & 0.05-0.2232 ms & N/A & N/A & N/A
-	&N/A\\ 
-	\cmidrule(r){2-8}
-	& d & N/A & 2 & N/A & N/A & N/A &N/A\\ 
-	\cmidrule(r){1-8}
-	\multirow{2}{*}{Delay 4} & M..N & N/A & 0.3-1.0044 ms & N/A & N/A & N/A
-	&N/A\\ 
-	\cmidrule(r){2-8}
-	& d & N/A & 9 & N/A & N/A & N/A &N/A\\ 
-	\cmidrule(r){1-8}
-	\multirow{2}{*}{Delay 5} & M..N & N/A & 0.833-2.3436 ms & N/A & N/A & N/A
-	&N/A\\ 
-	\cmidrule(r){2-8}
-	& d & N/A & 21 & N/A
-	& N/A & N/A &N/A\\ \bottomrule
-\end{tabular}
-\label{fig:comparison}
-\end{@twocolumnfalse}
-]
-\end{table}
-}
 

### 2013-12-03 exps.tex
Updated exps.tex 2

@@
 that the compiler was able to find \textit{d} without relaxing upper
-bound of \texttt{delay}. 
-This is because there are two influential factors in
-determining the relaxation: 1) execution speed of a target platform and 
-%Figure~\ref{fig:graph} shows the accuracy
-%of our results by comparing differences between programmer specified
-%upper bounds (Table~\ref{fig:benchlist}) and the actual bounds obtained
-%using our techniques (Table~\ref{fig:comparison}).
-2) the nature of the program that runs on this platform. Clock frequency
-of the processor directly impacts the value of WCRT. On the other hand,
-programmer can write a program such that large data (or control)
-computations in a single logical tick increase WCRT of the program.
-Assuming we do not change the architecture of the execution platform
-(i.e. JOP), the size of relaxation can only be reduced by modifying the
-benchmark programs such that the critical paths are broken up using
-\texttt{pause} constructs, thereby reducing the WCRT.
-Figure~\ref{fig:improv} shows segments of code from AECS whose WCRT is
-reduced by such technique.
+bound of \texttt{delay}. Figure~\ref{fig:graph1} shows the accuracy of
+our results by comparing differences between programmer specified upper
+bounds (Table~\ref{fig:benchlist}) and the actual bounds obtained using
+our techniques (Table~\ref{fig:comparison}).
+
+\begin{figure}[h!]
+	\centering
+\begin{tikzpicture}[scale=0.9]
+	\begin{axis}[
+			ybar,xtick=data,
+			symbolic x
+			coords={HRTCS,Printer/CD1,Printer/CD2,AECS/CD2},
+			height=5cm,
+			width=9.5cm,
+			minor y tick num=1,
+			yminorgrids=true,
+			ymajorgrids=true,
+			ytick={0,0.5,...,2},
+			enlarge x limits=0.3,
+%      xlabel=Benchmark,
+			x tick label style={rotate=45,font=\footnotesize},
+			ylabel=Difference factor,
+		]
+			\addplot coordinates
+			{(HRTCS,1)(Printer/CD1,1)(Printer/CD2,1)(AECS/CD2,1.94)};
+			\addplot coordinates {(Printer/CD2, 1)(AECS/CD2, 1.94)};
+	\end{axis}
+\end{tikzpicture}
+\begin{tikzpicture}[scale=0.9]
+	\begin{axis}[
+			ybar,xtick=data,
+			symbolic x coords={Motor},
+			ytick={0,0.5,...,5},
+			ymin={1.5},
+			minor y tick num=1,
+			yminorgrids=true,
+			ymajorgrids=true,
+%      extra y ticks={2.5},
+%      extra y tick labels={asdf},
+%      extra y tick style={grid=major},
+			height=5cm,
+			width=9.5cm,
+			xlabel=Benchmark,
+			xlabel shift={4pt},
+			ylabel=Difference factor,
+			legend style={
+				anchor=north,
+				at={(0.45,-0.44)},
+				legend columns=-1,
+			}
+%      nodes near coords
+		]
+			\addplot coordinates {(Motor, 2.60)};
+			\addplot coordinates {(Motor, 2.63)};
+			\addplot coordinates {(Motor, 4.39)};
+			\addplot coordinates {(Motor, 2.92)};
+			\addplot coordinates {(Motor, 2.69)};
+			\legend{Delay1,Delay2,Delay3,Delay4,Delay5}
+	\end{axis}
+\end{tikzpicture}
+\caption{Accuracy of the results}
+\label{fig:graph1}
+\end{figure}
+
+There are two influential factors in determining the relaxation: 1)
+execution speed of a target platform and 2) the nature of the program
+that runs on this platform. Clock frequency of the processor directly
+impacts the value of WCRT. On the other hand, programmer can write a
+program such that large data (or control) computations in a single
+logical tick increase WCRT of the program.  Assuming we do not change
+the architecture of the execution platform (i.e. JOP), the differece
+factor can can only be reduced by modifying the benchmark programs such
+that the critical paths are broken up using \texttt{pause} constructs,
+thereby reducing the WCRT.  Figure~\ref{fig:improv} shows segments of
+code from AECS whose WCRT is reduced by such technique.
 
@@
 each of these statement is executed in two different ticks. The same
-technique is applied to Motor example to reduce WCRT of the program.
-Note that we can also optimize algorithms of the data computations used
-in each Java method to reduce WCRT. However it is out of scope of this
-paper, hence not discussed here.
+technique is applied to Motor example, which is purely
+control-dominated, to reduce WCRT of the program. Note that we can also
+optimize algorithms of the data computations used in each Java method to
+reduce WCRT. However it is out of scope of this paper, hence not
+discussed here.
 
 The result of the upper bound relaxation with reduced WCRT is shown in
-Figure~\ref{fig:graph}. As one can see from the Figure, the upper bounds
-for delay statements in AECS/CD2 has been decreased by 11.81\% which is
-higher than all those in Motor which ranges from 7.54 to 8.05\%. It is
-because AECS has relatively a large amount of data computation while
-Motor is a completely control-dominated program. Therefore, the result
-shows that the size of the relaxation depends on how designers write
-their program as well as types of the application (i.e. data or control
-dominated).
+Figure~\ref{fig:graph2}. As one can see from the Figure, the upper
+bounds for delay statements in AECS/CD2 have been decreased by 11.81\%
+which is higher than all those in Motor which ranges from 7.54 to
+8.05\%. It was as expected since amount of computation required
+in control flow is mostly less demanding than data such that breaking
+the critical paths in Motor did not reduce WCRT as much as AECS/CD2.
+Therefore, we can see from the result that the size of the relaxation
+depends on how designers write their program as well as typs of the
+application (i.e. data or control dominated).
 
@@
 \caption{Reduction in upper bound relaxation with improved WCRT}
-\label{fig:graph}
+\label{fig:graph2}
 \end{figure}

### 2013-12-04 delays.tex
Updated the paper

@@
 
-The \texttt{delay(M..M)} variant is a little more interesting. Like
-before we find sets $S_1$ and $S_2$, and find the intersection of the
-two sets to get the value of \texttt{d}. It is possible (and often
-likely, as suggested by our experiments in
-Section~\ref{sec:experimental-results}) that the resultant set $D$ is
-empty (also possible in the case of \texttt{delay(M..N)}, but never
-possible in case of \texttt{delay(M)}). Note that $S_1$ and $S_2$ can
-never be empty sets and will always contain at minimum a single
-element. In such a case, we automatically \textit{relax} the upper bound
-of the \texttt{delay} statement.
+The \texttt{delay(M..M)} variant is more interesting. Like before, we
+find sets $S_1$ and $S_2$, and find the intersection of the two sets to
+get the value of \texttt{d}. It is possible (and often likely, as
+suggested by our experiments in Section~\ref{sec:experimental-results})
+that the resultant set $D$ is empty (also possible in the case of
+\texttt{delay(M..N)}, but never possible in case of
+\texttt{delay(M)}). Note that $S_1$ and $S_2$ can never be empty sets
+and will always contain at minimum a single element. In such a case, we
+can \textit{relax} the upper bound of the \texttt{delay} statement.
 

### 2013-12-04 discussion.tex
Updated exps
Updated the paper

@@
 be known apriori. One such system is our motivating example -- the human
-response time system (HRCTS). Another one we presented was a real
+response time system (HRTCS). Another one we presented was a real
 printer-spooler embedded controller in
@@
 as:
-% \begin{scriptsize}
+
+\begin{figure}[h!]
+  \centering
+    \begin{minipage}[b]{\linewidth}
 \begin{verbatim}
@@
 \end{verbatim}
-% \end{scriptsize}
+    \end{minipage}
+  \caption{Programming low priority timeout tasks with delay statements}
+  \label{fig:timeout}
+\end{figure}
 
 In the above case, the program waits for signal \texttt{A} from the
-environment for \textit{at\_least} \texttt{1 ms}, but it may wait longer
-if other reactions/CDs are scheduled for execution. This second form of
-delay statement is very useful for programming low priority tasks.
+environment for \textit{at least} \texttt{1 ms}, but it may wait longer
+if other reactions/CDs are scheduled for execution. This second variant
+of delay statement is very useful for programming low priority tasks.
 
@@
 
-The third form, is useful for hard-real time guarantees, e.g., a
+The third variant is useful for hard-real time guarantees, e.g., a
 reaction, or a whole CD, can be programmed to run periodically like so:
 
+\begin{figure}[h!]
+  \centering
+  \begin{minipage}[b]{\linewidth}
 \begin{verbatim}
@@
 \end{verbatim}
+  \end{minipage}
+  \caption{Programming periodic tasks with delay statements}
+  \label{fig:periodic}
+\end{figure}
 
@@
 
+\begin{figure}[h!]
+  \centering
+  \begin{SubFloat}{\label{pp:a}Original program}
+    \centering
+    \begin{minipage}[b]{0.3\linewidth}
 \begin{verbatim}
-while(true) {      while(true) {
- delay(1..1 ms);    while(true) { 
+while(true) {
+ delay(1..1 ms);
  //extra pause
- pause;              trap(T) {
- emit S;              int x = 0;
-}                     while(true) {
+ pause;
+ emit S;
+}
+\end{verbatim}
+  \end{minipage}
+\end{SubFloat}
+\hspace{9pt}
+  \begin{SubFloat}{\label{pp:b}Rewritten program}
+    \begin{minipage}[h]{0.5\linewidth}
+\begin{verbatim}
+while(true) {
+ while(true) { 
+  trap(T) {
+   int x = 0;
+   while(true) {
     x = x + 1;
     pause;
-                       if (x == d)exit(T);
+    if (x == d)
+     exit(T);
    }
@@
  }
+}
 \end{verbatim}
+    \end{minipage}
+  \end{SubFloat}
+\caption{Extra pause statements in periodic tasks}
+  \label{fig:periodic2}
+\end{figure}
 
@@
 pauses introduces more logical ticks, consequently increasing the
-real-time delay.
+real-time delay. Static analysis techniques from real-time community can
+be used to automatically determine the delay specification in such
+cases, but describing these techniques is out of the scope of this work.
 
@@
 
-% \begin{scriptsize}
+\begin{figure}[h!]
+  \centering
+  \begin{minipage}[b]{\linewidth}
 \begin{verbatim}
-suspend(S) {emit START_TIMER(10); 
-            await (TIMER);emit O1;}
+suspend(S) {
+ emit START_TIMER(10); 
+ await (TIMER);emit O1;
+}
 \end{verbatim}
-% \end{scriptsize}
+  \end{minipage}
+  \caption{Interaction of preemption and external timers}
+  \label{fig:preemp}
+\end{figure}
 
@@
 
-\subsection{Other semantic discussion}
-\label{sec:other-semant-dissc}
+% \subsection{Other semantic discussion}
+% \label{sec:other-semant-dissc}
 
-\subsubsection{Interaction of channel communication and delays}
+\subsection{Interaction of channel communication and delays}
 \label{sec:inter-chann-comm}

### 2013-12-04 exps.tex
Updated the paper

@@
 Motor is a stepper motor controller from~\cite{Bourke2009a}, used for
-producing monochrome images on paper by applying a correct application
-of current to a print head (a row of resistors). In this example,
-duration of the current applied is controlled by the feedback logic
-using real-time delays, which are expressed using reactive
-constructs. Printer is an example from~\cite{Schneider:1999:CRT:555233},
-which models a printer and a spooler in timed CSP.  Their operation is
-rather simple: as soon as the spooler receives a print job, passes it to
-the printer through a channel where actual printing process take
-place. This example along with HRCTS demonstrates the use of
-non-deterministic delays.
+producing monochrome images on paper by correct application of current
+to a print head (a row of resistors). In this example, duration of the
+current applied is controlled by the feedback logic using real-time
+delays, which are expressed using reactive constructs. Printer is an
+example from~\cite{Schneider:1999:CRT:555233}, which models a printer
+and a spooler in timed CSP. Its operation can be described as follows:
+as soon as the spooler receives a print job it passes it to the printer
+through a channel where actual printing process take place. This example
+along with HRTCS demonstrates the use of non-deterministic delays.
 
@@
 Printer CD between 2.7 to 10.6 ms after receiving it from the
-environment (e.g. a user). On the other hand, Printer CD notifies
+environment (e.g. a user). On the other hand, Printer CD notifies the
 environment that it starts printing between 1.2 to 30.3 ms once it
@@
 between 5.6 to 100.2 ms after the notification. In this example, bounded
-delay ensures that both channel communication and print job should occur
-within the specified amount of time. Lastly, Motor is a synchronous
-program with 5 delay statements that control the amount of current
-applied to the stepper motor coils.
+delay ensures that both; channel communication and print job should
+occur within the specified amount of time. Lastly, Motor is a
+synchronous program with 5 delay statements that control the amount of
+current applied to the stepper motor coils.
 
@@
 We use the \textit{Java Optimized Processor}
-(JOP)~\cite{jop:jnl:jsa2007}, to carry out all our experiments. JOP
-implements the JVM specification in hardware directly, and has been
-shown to be real-time amenable. We statically estimate the WCRT and BCRT
-of these benchmark applications using the real-time analysis tools
-provided by the JOP design framework. The \textit{Safety Critical Java}
-(SCJ) specification we use for real-time analysis avoids garbage
-collection (GC) and hence, GC overheads and unpredictability are
-completely avoided.
+(JOP)~\cite{jop:jnl:jsa2007} to carry out all our experiments. JOP
+implements the JVM specification in hardware directly and has been shown
+to be real-time amenable. We statically estimate the WCRT and BCRT of
+these benchmark applications using the real-time analysis tools provided
+by the JOP design framework. The \textit{Safety Critical Java} (SCJ)
+specification we use for real-time analysis avoids garbage collection
+(GC) and hence, GC overheads and unpredictability are completely
+avoided.
 
@@
 
-The previous results show that HRTCS and Printer were the only examples
-that the compiler was able to find \textit{d} without relaxing upper
-bound of \texttt{delay}. Figure~\ref{fig:graph1} shows the accuracy of
-our results by comparing differences between programmer specified upper
+The previous results show that for HRTCS and Printer that the compiler
+was able to find \textit{d} without relaxing upper bound of
+\texttt{delay}. Figure~\ref{fig:graph1} shows the accuracy of our
+results by comparing differences between programmer specified upper
 bounds (Table~\ref{fig:benchlist}) and the actual bounds obtained using
-our techniques (Table~\ref{fig:comparison}).
+our techniques (Table~\ref{fig:comparison}) for the other examples
+requiring relaxation. AECS/CD1 delay is not included in
+Figure~\ref{fig:graph2}, because although it requires relaxation, the
+delay specification (Table~\ref{fig:benchlist}) does not have an upper
+bound.
 
@@
 		]
-			\addplot coordinates
+			\addplot + [postaction={pattern=north east lines}] coordinates
 			{(HRTCS,1)(Printer/CD1,1)(Printer/CD2,1)(AECS/CD2,1.94)};
-			\addplot coordinates {(Printer/CD2, 1)(AECS/CD2, 1.94)};
+			\addplot + [postaction={pattern=north west
+                          lines}] coordinates 
+                        {(Printer/CD2, 1)(AECS/CD2, 1.94)};
 	\end{axis}
@@
 		]
-			\addplot coordinates {(Motor, 2.60)};
-			\addplot coordinates {(Motor, 2.63)};
-			\addplot coordinates {(Motor, 4.39)};
-			\addplot coordinates {(Motor, 2.92)};
-			\addplot coordinates {(Motor, 2.69)};
+			\addplot + [postaction={pattern=north east lines}] coordinates
+                        {(Motor, 2.60)};
+			\addplot + [postaction={pattern=north west lines}] coordinates
+                        {(Motor, 2.63)};
+			\addplot + [postaction={pattern=dots}] coordinates
+                        {(Motor, 4.39)};
+			\addplot + [postaction={pattern=horizontal lines}] coordinates
+                        {(Motor, 2.92)};
+			\addplot + [postaction={pattern=crosshatch}] coordinates
+                        {(Motor, 2.69)};
 			\legend{Delay1,Delay2,Delay3,Delay4,Delay5}
@@
 id using data logic written in Java (\texttt{isValidRequest}). We
-identified this execution path caused lengthening the WCRT of the
-program. In this case, we divide two consequtive statements, i.e.
+identified this as the critical execution path accounting for the WCRT.
+In this case, we divide two consequtive statements, i.e.
 \texttt{present} and \texttt{if}, by inserting \texttt{pause} such that
@@
 The result of the upper bound relaxation with reduced WCRT is shown in
-Figure~\ref{fig:graph2}. As one can see from the Figure, the upper
-bounds for delay statements in AECS/CD2 have been decreased by 11.81\%
-which is higher than all those in Motor which ranges from 7.54 to
-8.05\%. It was as expected since amount of computation required
-in control flow is mostly less demanding than data such that breaking
+Figure~\ref{fig:graph2}.  The upper bounds for delay statements in
+AECS/CD2 decreased by 11.81\%, while that for Motor example reduced from
+7.54 to 8.05\%. This is expected since the amount of computation
+required in control flow is less demanding than data and hence, breaking
 the critical paths in Motor did not reduce WCRT as much as AECS/CD2.
 Therefore, we can see from the result that the size of the relaxation
-depends on how designers write their program as well as typs of the
+depends on how designers write their program as well as type of the
 application (i.e. data or control dominated).
@@
 		]
-			\addplot coordinates {(Motor, 8.05) (AECS/CD2, 11.82)};
-			\addplot coordinates {(Motor, 8.00)};
-			\addplot coordinates {(Motor, 6.42)};
-			\addplot coordinates {(Motor, 7.53)};
-			\addplot coordinates {(Motor, 7.89)};
+			\addplot + [postaction={pattern=north east lines}] coordinates
+                        {(Motor, 8.05) (AECS/CD2, 11.82)};
+			\addplot + [postaction={pattern=north west lines}] coordinates
+                        {(Motor, 8.00)};
+			\addplot + [postaction={pattern=dots}] coordinates
+                        {(Motor, 6.42)};
+			\addplot + [postaction={pattern=horizontal lines}] coordinates
+                        {(Motor, 7.53)};
+			\addplot + [postaction={pattern=crosshatch}] coordinates
+                        {(Motor, 7.89)};
 			\legend{Delay1,Delay2,Delay3,Delay4,Delay5}

### 2013-12-04 motivation.tex
Updated the paper

@@
 
-Consider a simple machine used to test human response time and collect
-this data for future research purposes. The machine switches on a green
-light on a touch screen for anywhere between 50.3 ms to 200.3 ms, if the
-user can touch this green light then she is successful and an integer
+Consider a machine used to test human response time and collect this
+data for future research purposes. The machine switches on a green light
+on a touch screen for anywhere between 50.3 ms to 200.3 ms, if the user
+can touch this green light then she is successful and an integer
 \texttt{1} is sent to a database, if unsuccessful a \texttt{0} is sent
-instead, Different statistics and analyses on human responsiveness can
-be based on collected data.
+instead. Different statistics and analyses on human responsiveness can
+be based on the collected data.
 
@@
 this system are described in~\cite{amal10} and Table~\ref{tab:1} in
-Section~\ref{sec:background}. The delay speciﬁcation, which models the
+Section~\ref{sec:background}. The delay specification, which models the
 non-deterministic delay between 50.3 ms and 200.3 ms is not supported in

### 2013-12-04 background.tex

@@
 Java~\cite{scj2013} amenable to formal verification and timing analysis,
-while avoiding all the pitfalls of programming in 'C' style languages. A
+while avoiding all the pitfalls of programming in `C' style languages. A
 number of derived statements also exist (e.g., \texttt{sustain},
@@
 requires real-time to compute. Formally, execution of every micro-step
-in the logical zero time, requires $\delta$ physical time. Time for a
+in the logical zero time requires $\delta$ physical time. Time for a
 logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,

### 2013-12-04 introduction.tex

@@
   exact \textit{real-time} delay. \textit{Non-deterministic delay:} is
-  the ability to specify a range, of real-time values for a delay
+  the ability to specify a range of real-time values for a delay
   statement, e.g., delay statement in Figure~\ref{fig:1}.

### 2013-12-04 related_work.tex

@@
 not require external timers (like~\cite{rsh94}) or logical notion of
-timers (like~\cite{Bourke2009a}). Since in our solution we calculate the
-delay in logical ticks \textit{after} allocation and scheduling, it
-gives the compiler developers the chance to optimize for many different
-criteria such as computation time, energy, power, etc, without worrying
-about or violating the real-time delay specification.
+timers (like~\cite{Bourke2009a}). In our solution we calculate the delay
+in logical ticks \textit{after} allocation and scheduling, it gives the
+compiler developers the chance to optimize for many different criteria
+such as computation time, energy, power, etc, without worrying about or
+violating the real-time delay specification.
 
@@
 tools developed for reactive languages and more in the spirit of logical
-time. Finally, a major difference between Bertin et al and our approach
+time. Finally, a major difference between Bertin et al. and our approach
 is that their approach requires programmers to annotate the time taken
-by 'C' code fragments manually, this is extremely error prone, we use
+by `C' code fragments manually, this is extremely error prone, we use
 static low-level program analysis to find out such times automatically.

### 2013-12-04 exps.tex
Added a bar for the second delay stmt
updated the number in exps

@@
 AECS/CD2 decreased by 11.81\%, while that for Motor example reduced from
-7.54 to 8.05\%. This is expected since the amount of computation
+6.42 to 8.05\%. This is expected since the amount of computation
 required in control flow is less demanding than data and hence, breaking
@@
 			\addplot + [postaction={pattern=north west lines}] coordinates
-                        {(Motor, 8.00)};
+                        {(Motor, 8.00) (AECS/CD2, 11.82)};
 			\addplot + [postaction={pattern=dots}] coordinates

### 2013-12-04 background.tex
code style changed

@@
 \begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
-\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
-  \begin{scriptsize}
-\begin{verbatim}
+\begin{minipage}[b]{1\linewidth}% a minipage to control the width...
+		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,delay,suspend}]
 {
@@
 } // SystemJ CD end
-\end{verbatim}%
-  \end{scriptsize}
+\end{lstlisting}%
 \end{minipage}%

### 2013-12-04 delays.tex
code style changed

@@
 \begin{figure}[tb]
+	\vspace{-10pt}
     \begin{minipage}{\textwidth}
-      \begin{scriptsize}
-\begin{verbatim}
+		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,delay,suspend}]
 trap(T){
@@
 } // by Algorithm 1
-\end{verbatim}
-      \end{scriptsize}
+\end{lstlisting}
     \end{minipage}

### 2013-12-04 discussion.tex
code style changed

@@
   \centering
-    \begin{minipage}[b]{\linewidth}
-\begin{verbatim}
+			\vspace{-10pt}
+			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={await,emit,trap,pause,exit,delay}]
 // timeout after 1ms.
@@
         {delay(1 ms);exit(T);}}
-\end{verbatim}
-    \end{minipage}
+\end{lstlisting}
   \caption{Programming low priority timeout tasks with delay statements}
   \label{fig:timeout}
+			\vspace{-10pt}
 \end{figure}
@@
   \centering
-  \begin{minipage}[b]{\linewidth}
-\begin{verbatim}
+	\vspace{-10pt}
+		\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay}]
 while(true) { delay ((1..1)ms); emit S; 
  //do something }
-\end{verbatim}
-  \end{minipage}
+ \end{lstlisting}
   \caption{Programming periodic tasks with delay statements}
   \label{fig:periodic}
+	\vspace{-10pt}
 \end{figure}
@@
     \centering
-    \begin{minipage}[b]{0.3\linewidth}
-\begin{verbatim}
+			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay}]
 while(true) {
@@
 }
-\end{verbatim}
-  \end{minipage}
+\end{lstlisting}
 \end{SubFloat}
@@
   \begin{SubFloat}{\label{pp:b}Rewritten program}
-    \begin{minipage}[h]{0.5\linewidth}
-\begin{verbatim}
+			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay}]
 while(true) {
@@
 }
-\end{verbatim}
-    \end{minipage}
+\end{lstlisting}
   \end{SubFloat}
@@
   \centering
-  \begin{minipage}[b]{\linewidth}
-\begin{verbatim}
+	\vspace{-10pt}
+		\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay,suspend}]
 suspend(S) {
@@
 }
-\end{verbatim}
-  \end{minipage}
+		\end{lstlisting}
   \caption{Interaction of preemption and external timers}
   \label{fig:preemp}
+	\vspace{-10pt}
 \end{figure}

### 2013-12-04 exps.tex
code style changed

@@
 {\renewcommand{\arraystretch}{0.9}
+%\renewcommand{\texttt}[1] {\begin{lstlisting}[language=Java,morekeywords={delay}] #1 \end{lstlisting}}
 \begin{table}[h!]
@@
 \begin{figure}[h!]
+	\vspace{-10pt}
   \begin{SubFloat}{\label{delaycode:a}Original program}
     \centering
-    \begin{minipage}[b]{0.3\linewidth}
       \scriptsize
-\begin{verbatim}
+		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,delay,suspend}]
 emit TimerTriggered;
@@
     \centering
-    \begin{minipage}[b]{0.5\linewidth}
-      \scriptsize
-\begin{verbatim}
+		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,delay,suspend}]
 // No emit statement
@@
 }
-\end{verbatim}
-		\end{minipage}
+\end{lstlisting}
 	\end{SubFloat}
@@
 	\centering
+	\vspace{-10pt}
 	\begin{SubFloat}{\label{fig:improv-a}Original program}
-    \begin{minipage}[b]{0.3\linewidth}
-			\scriptsize
-\begin{verbatim}
+    \begin{minipage}[b]{0.45\linewidth}
+		\begin{lstlisting}[style=sysj,morekeywords={present,emit,trap,pause,exit,delay,suspend}]
 present(EntryRequest)
@@
 }
-\end{verbatim}
+\end{lstlisting}
 \end{minipage}
 \end{SubFloat}
-  \hspace{1cm}%
   \begin{SubFloat}{\label{fig:improv-b}Program with reduced WCRT}
     \centering
-    \begin{minipage}[b]{0.35\linewidth}
-      \scriptsize
-\begin{verbatim}
+    \begin{minipage}[b]{0.45\linewidth}
+		\begin{lstlisting}[style=sysj,morekeywords={emit,present,trap,pause,exit,delay,suspend}]
 present(EntryRequest)
@@
 }
-\end{verbatim}
+\end{lstlisting}
 		\end{minipage}
@@
 	\label{fig:improv}
+	\vspace{-10pt}
 \end{figure}

### 2013-12-04 motivation.tex
code style changed

@@
 % \begin{center}
-\begin{minipage}{5cm}
-  \begin{scriptsize}
-
-\begin{verbatim}
+	\vspace{-10pt}
+		\begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,delay,suspend}]
 { // Clock-domain 1
@@
 }
-\end{verbatim}
-  \end{scriptsize}
-\end{minipage}
+\end{lstlisting}
 \caption{An example human responsiveness system (HRTCS)}

### 2013-12-04 background.tex
updated some figs (pos)
Fixed some commenting issues in the figures

@@
 \begin{figure}[t!]
-% \centering
+\centering
 \begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
@@
   emit O;
-  if(counter != 0) { // java computation leading to WCRT}
+  if(counter != 0) {/* java computation leading to WCRT */}
   ++counter;
@@
 \end{SubFloat}
-\hspace{1.5cm}%
+%\hspace{1.5cm}%
 % \begin{SubFloat}[Black box]{\label{fig:2b}Logical time in SystemJ}%

### 2013-12-04 motivation.tex
Fixed some commenting issues in the figures

@@
 { // Clock-domain 2
-  while(true){ receive S; // do something with S}
+  while(true){ receive S; /* do something with S */}
 }

### 2013-12-04 exps.tex
updated some figs (pos)
Made it ugly

@@
 	\vspace{-10pt}
+	\centering
   \begin{SubFloat}{\label{delaycode:a}Original program}
-    \centering
       \scriptsize
@@
   \begin{SubFloat}{\label{delaycode:b}Program with delay constructs}
-    \centering
 		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,delay,suspend}]
@@
 		]
-			\addplot + [postaction={pattern=north east lines}] coordinates
+			\addplot + [fill=white,postaction={pattern=north east lines}] coordinates
                         {(Motor, 8.05) (AECS/CD2, 11.82)};
-			\addplot + [postaction={pattern=north west lines}] coordinates
+			\addplot + [fill=white,postaction={pattern=north west lines}] coordinates
                         {(Motor, 8.00) (AECS/CD2, 11.82)};
-			\addplot + [postaction={pattern=dots}] coordinates
+			\addplot + [fill=white,postaction={pattern=dots}] coordinates
                         {(Motor, 6.42)};
-			\addplot + [postaction={pattern=horizontal lines}] coordinates
+			\addplot + [fill=white,postaction={pattern=horizontal lines}] coordinates
                         {(Motor, 7.53)};
-			\addplot + [postaction={pattern=crosshatch}] coordinates
+			\addplot + [fill=white,postaction={pattern=crosshatch}] coordinates
                         {(Motor, 7.89)};

### 2014-01-27 background.tex
Updated the paper

@@
 synchronous programs~\cite{boldt07} exist and this analysis is not the
-focus of this paper. The opposite of the WCRT is the \textit{Best Case
-  Reaction Time} (BCRT), we denote the WCRT and BCRT in
-Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
-respectively. Figure~\ref{fig:2c} shows a continuously running physical
-clock (analogous to a clock in digital hardware), the numerical
-annotations on the rising edge of the clock mark the logical ticks from
-Figure~\ref{fig:2c}. The subscripts \texttt{s} and \texttt{e} for the
-numbers show the start and end of the logical ticks, respectively. The
-end of a logical tick and the start of the next logical tick happen
-together. Figure~\ref{fig:2c} shows the mapping of the logical time to
-the physical time. For example, the logical tick \texttt{1} starts at
-the first rising edge and completes at the second rising edge of the
-physical clock, whereas logical tick \texttt{4} starts at the $4^{th}$
-rising edge, but finishes at the $6^{th}$ rising edge of the physical
-clock. This difference can be accounted for by the Java computations
-required when signal \texttt{A} is input by the environment (see
-Figure~\ref{fig:2a}). This elasticity is inherent (and elegant) to both
-Esterel and SystemJ.
+focus of this paper. The opposite of the WCRT, the shortest time for a
+reaction, is called the \textit{Best Case Reaction Time} (BCRT), we
+denote the WCRT and BCRT in Figure~\ref{fig:2c} using \texttt{W} and
+\texttt{B}, respectively. Figure~\ref{fig:2c} shows a continuously
+running physical clock (analogous to a clock in digital hardware), the
+numerical annotations on the rising edge of the clock mark the logical
+ticks from Figure~\ref{fig:2c}. The subscripts \texttt{s} and \texttt{e}
+for the numbers show the start and end of the logical ticks,
+respectively. The end of a logical tick and the start of the next
+logical tick happen together. Figure~\ref{fig:2c} shows the mapping of
+the logical time to the physical time. For example, the logical tick
+\texttt{1} starts at the first rising edge and completes at the second
+rising edge of the physical clock, whereas logical tick \texttt{4}
+starts at the $4^{th}$ rising edge, but finishes at the $6^{th}$ rising
+edge of the physical clock. This difference can be accounted for by the
+Java computations required when signal \texttt{A} is input by the
+environment (see Figure~\ref{fig:2a}). This elasticity is inherent (and
+elegant) to both Esterel and SystemJ.
 

### 2014-01-27 conclusion.tex
Updated the paper

--- 2013-06-18 background.tex
@@
-\section{Background}
-\label{sec:background}
+\section{Conclusion}
+\label{sec:concl-future-work}
+
+In this paper we have described a novel way to introduce real-time in
+\textit{Globally Asynchronous Locally Synchronous} (GALS) languages, in
+particular SystemJ, and their subset the synchronous languages like
+Esterel. The fundamental idea is to convert real-time wait constructs
+into logical waits (\texttt{pause} construct) that interact well with
+rest of the constructs in these languages, especially, preemption and
+channel communication. We do \textit{not} need or use external timers to
+introduce these real-time wait constructs, instead we depend upon an
+underlying platform that is timing analyzable, thereby resolving the
+problems of interaction between external timers and
+preemption. Moreover, we also do not require specific timer resolutions,
+thereby making our solution elegant (in the spirit of GALS/synchronous
+programming) and flexible. We rewrite the wait constructs into reactive
+kernel constructs, calculate the \textit{Worst} and \textit{Best} case
+reaction times and then determine wait \texttt{d} in number of logical
+ticks, which is invariant to the \textit{elasticity} of logical time. To
+our knowledge we are the first to introduce exact and inexact wait
+constructs in such languages and show their usefulness in designing
+real-world systems.
 

### 2014-01-27 delays.tex
Updated the paper

@@
 
-We introduce a single \textit{derived} (built from kernel constructs)
-statement called \mbox{\texttt{delay (M..N)}} in the SystemJ language
-for real-time control. The resolution of the delay statement is of
-secondary concern and is dependent upon the execution platform.
+We introduce a single \textit{derived} statement called
+\mbox{\texttt{wait\_inbetween (M..N)}}, built from kernel constructs, in
+the SystemJ language for real-time control. The other two constructs
+\mbox{\texttt{wait\_atleast (M)}} and \mbox{\texttt{wait\_exactly (M)}}
+are variants of this single derived construct.
 
-\subsection{Semantics of the delay statement}
+% The resolution of the delay statement is of secondary concern and is
+% dependent upon the execution platform.
+
+\subsection{Semantics of the \texttt{wait\_inbetween} statement}
 \label{sec:semant-delay-stat}
 
-\begin{figure}[t!]
-  \centering
-  \includegraphics[scale=0.6]{./sem}
-  \caption{Pictorial representation of the semantics of the delay
-    statement and its variants}
-  \label{fig:5}
-\end{figure}
+% \begin{figure}[t!]
+%   \centering
+%   \includegraphics[scale=0.6]{./sem}
+%   \caption{Pictorial representation of the semantics of the delay
+%     statement and its variants}
+%   \label{fig:5}
+% \end{figure}
 
-Given a SystemJ program: \texttt{delay (M..N); p}, where $M \in
-\mathbb{Q}^{>0}$ and $N \in \mathbb{Q}^{>0}$, statement $p$ is executed
-after real-time delay of $\tau$ units, such that, $M \leq \tau \leq N$.
+Given a SystemJ program: \texttt{wait\_inbetween (M..N); p}, where $M
+\in \mathbb{Q}^{>0}$ and $N \in \mathbb{Q}^{>0}$, statement $p$ is
+executed after real-time delay of $\tau$ units, such that, $M \leq \tau
+\leq N$.
 
-We also introduce two variants of the derived statement \texttt{delay}.
+We also introduce two variants of the derived statement
+\texttt{wait\_inbetween}.
 \begin{enumerate*}
-\item Given a SystemJ program \texttt{delay (M); p}, where $M \in
-  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay of
-  $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
-  note that the lower bound of the delay construct $M$ is \textit{not}
+\item Given a SystemJ program \texttt{wait\_atleast (M); p}, where $M
+  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
+  of $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
+  note that the lower bound of the wait construct; $M$ is \textit{not}
   an exact delay, but rather the control is allowed to proceed to the
-  next statement anytime after the delay of time $M$.
-\item Given a SystemJ program \texttt{delay (M..M); p}, where $M \in
-  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay of
-  $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
-  delay is exact.
+  next statement anytime after waiting for \textit{at least} time $M$.
+\item Given a SystemJ program \texttt{wait\_exactly (M); p}, where $M
+  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
+  of $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
+  waiting time is exact.
 \end{enumerate*}
 
-Figure~\ref{fig:5} pictorially represents the semantics of these delay
-constructs. The construct \texttt{delay (M..N)} requires that upon
-execution the control-flow is suspended for \textit{at least} $M$ units
-and \textit{at most} $N$ units, depicted as bounded box in
-Figure~\ref{fig:5}. There is no upper bound for the second delay variant
-\texttt{delay(M)}, which requires that the control-flow suspends for
-\textit{at least} $M$ units, but the following statement can be executed
-after some countable delay, depicted with the unbounded box in
-Figure~\ref{fig:5}. Finally, \texttt{delay (M..M)} requires that there
-is an \textit{exact} delay of $M$ units, i.e., the control-flow suspends
-execution for \textit{at least} and \textit{at most} $M$ units.
+% Figure~\ref{fig:5} pictorially represents the semantics of these delay
+% constructs. The construct \texttt{delay (M..N)} requires that upon
+% execution, the control-flow is suspended for \textit{at least} $M$ units
+% and \textit{at most} $N$ units, depicted as bounded box in
+% Figure~\ref{fig:5}. There is no upper bound for the delay variant
+% \texttt{delay(M)}, which requires that the control-flow suspends for
+% \textit{at least} $M$ units, but the following statement can be executed
+% after some countable delay, depicted with the unbounded box in
+% Figure~\ref{fig:5}. Finally, \texttt{delay (M..M)} requires that there
+% is an \textit{exact} delay of $M$ units, i.e., the control-flow suspends
+% execution for \textit{at least} and \textit{at most} $M$ units.
 
-\subsection{Rewriting the \texttt{delay} statement}
+\subsection{Rewriting the \texttt{wait\_inbetween} statement}
 \label{sec:rewr-delay-stat}
 
-The introduced \texttt{delay} construct is not a kernel construct, but a
-\textit{derived} construct built from the kernel constructs
-(Table~\ref{tab:1}) in SystemJ. Figure~\ref{fig:3} gives the rewrite of
-the \texttt{delay} construct to kernel statements.
+The introduced \texttt{wait\_inbetween} construct is not a kernel
+construct, but a \textit{derived} construct built from the kernel
+constructs (Table~\ref{tab:1}) in SystemJ. Figure~\ref{fig:3} gives the
+rewrite of the \texttt{wait\_inbetween} construct to kernel statements.
 
@@
     \end{minipage}
-    \caption{The rewrite of \texttt{delay} construct}
+    \caption{The rewrite of \texttt{wait\_inbetween} construct}
     \label{fig:3}
@@
 physical notion of time back to the elegant logical notion of time. The
-rewrite \textit{delays} a certain number of logical ticks, before
+rewrite \textit{waits} a certain number of logical ticks, before
 proceeding to the next statement. The number of logical ticks \texttt{d}
-to \textit{delay} (Figure~\ref{fig:3}) is determined by the compiler
+to \textit{wait} (Figure~\ref{fig:3}) is determined by the compiler
 statically at compile time. The value of \texttt{d} is intricately tied
@@
 
-\subsection{Finding the logical delay \texttt{d}}
+\subsection{Finding the logical wait \texttt{d}}
 \label{sec:find-logic-delay}
@@
   the BCRT and the WCRT, thus any number of logical ticks \texttt{d}
-  that map to the required real-time \texttt{delay} should be chosen in
-  such a way that they are invariant to this elasticity.}
+  that map to the required real-time \mbox{\texttt{wait\_inbetween
+      (M..N)}} should be chosen in such a way that they are invariant to
+  this elasticity.}
 
 Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
-$M$ and $N$, respectively of the \texttt{delay} construct. We first
-divide $M$ and $N$ with the BCRT and WCRT, respectively. This division
-gives us the number of individual logical ticks required to delay the CD
-(or synchronous program) by the real-time specification. We always
-$ceil$ when dividing $M$ and $floor$ when dividing $N$ to make sure that
-the resultant values are integers (in domain $\mathbb{N}^{>0}$) and
-these functions guarantee that the resultant logical ticks result in
-real-time delays between the required range $[M,N]$. Next, a function
-$F$ maps these calculated values to a set of equidistant integer points
-(values) separated by a unit value -- these points represent all the
-logical ticks running at the WCRT and the BCRT, respectively that
-satisfy the real-time delay requirements. The intersection of these two
-sets gives all the logical ticks that satisfy the real-time requirements
-invariant of the logical time and its elasticity.
+$M$ and $N$, respectively of the \texttt{wait\_inbetween} construct. We
+first divide $M$ and $N$ with the BCRT and WCRT, respectively. This
+division gives us the number of individual logical ticks required to
+postpone the CD (or synchronous program) by the real-time
+specification. We always $ceil$ when dividing $M$ and $floor$ when
+dividing $N$ to make sure that the resultant values are integers (in
+domain $\mathbb{N}^{>0}$) and these functions guarantee that the
+resultant logical ticks result in real-time postponements between the
+required range $[M,N]$. Next, a function $F$ maps these calculated
+values to a set of equidistant integer points (values) separated by a
+unit value -- these points represent all the logical ticks running at
+the WCRT and the BCRT, respectively that satisfy the real-time wait
+requirements. The intersection of these two sets gives all the logical
+ticks that satisfy the real-time requirements invariant of the logical
+time and its elasticity.
 
@@
   calculate the logical ticks that are always running at the WCRT and
-  satisfy the required real-time delay.
+  satisfy the required real-time wait.
 \item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
@@
 
+\begin{figure}[t!]
+  \centering
+  \includegraphics[scale=0.6]{FF4}
+  \caption{The pictorial representation of the solution for the HRTCS system}
+  \label{fig:hrtcssoln}
+\end{figure}
+
 The resultant value for \texttt{d} gives the number of logical ticks,
 which can run at any physical clock-speed, bounded by the BCRT and the
-WCRT of the program and still result in the desired real-time delay. We
-think this is an elegant solution, because the technique provides
+WCRT of the program and still result in the desired real-time
+wait. Figure~\ref{fig:hrtcssoln} gives the solution space for the
+aforementioned calculation. The patterned space is the only overlapping
+region and hence gives the solution space for the above example.
+
+We think this is an elegant solution, because the technique provides
 real-time guarantees while preserving the essence (elastic logical tick)
@@
 style languages. Moreover, this technique considers non
-maximal-parallelism, i.e., the delay in logical ticks is calculated
-after scheduling of CDs or synchronous programs has been performed. To
-our knowledge we are the first to do so.
+maximal-parallelism, i.e., the postponement in logical ticks is
+calculated after scheduling of CDs or synchronous programs has been
+performed. To our knowledge we are the first to do so.
 
-\subsection{Extending the technique to variants of \texttt{delay}}
+\subsection{Extending the technique to variants of wait}
 \label{sec:extend-tehcn-vari}
 
-\paragraph{The \texttt{delay(M)} construct}
+\paragraph{The \texttt{wait\_atleast(M)} construct}
 \label{sec:extend-techn-vari}
 
-The \texttt{delay(M)} variant is easily accommodated in the
+The \texttt{wait\_atleast(M)} variant is easily accommodated in the
 technique. All one needs to do is find the set $S_1$ and choose a value
@@
 
-\paragraph{The \texttt{delay(M..M)} construct}
+\paragraph{The \texttt{wait\_exactly(M)} construct}
 \label{sec:extend-techn-vari}
 
-The \texttt{delay(M..M)} variant is more interesting. Like before, we
-find sets $S_1$ and $S_2$, and find the intersection of the two sets to
-get the value of \texttt{d}. It is possible (and often likely, as
-suggested by our experiments in Section~\ref{sec:experimental-results})
+The \texttt{wait\_exactly(M)} variant is more interesting. Like before,
+we find sets $S_1$ and $S_2$, and find the intersection of the two sets
+to get the value of \texttt{d}. It is possible % (and often likely, as
+% suggested by our experiments in
+% Section~\ref{sec:experimental-results})
 that the resultant set $D$ is empty (also possible in the case of
-\texttt{delay(M..N)}, but never possible in case of
-\texttt{delay(M)}). Note that $S_1$ and $S_2$ can never be empty sets
-and will always contain at minimum a single element. In such a case, we
-can \textit{relax} the upper bound of the \texttt{delay} statement.
+\texttt{wait\_inbetween(M..N)}, but never possible in case of
+\texttt{wait\_atleast(M)}). Note that $S_1$ and $S_2$ can never be empty
+sets and will always contain at minimum a single element. For such cases
+we provide two solutions listed below.
+
+% In such a case, we can \textit{relax} the upper bound of the
+% \texttt{delay} statement.
 
@@
 This algorithm results in the \underline{smallest} relaxation required
-for the real-time delay to be satisfied. The algorithm takes as input
-the WCRT and set $S_2$, recall that set $S_2$ represents the logical
-ticks required to satisfy the real-time requirement at the BCRT. We take
-the very first value from set $S_2$ and multiply it with the WCRT to get
-the relaxation $N$. The first element of set $S_2$ is returned as the
-logical tick delay \texttt{d}. The fundamental observation is that we
-have to delay for at least $M$ units of real-time, hence,
-under-approximation is out of question. We can still over-approximate,
-but to reduce the resultant error, we should over-approximate by least
-possible value, which is the lower bound of set $S_2$. Thus, the lower
-bound is considered to be the only element shared between the two sets
-$S_1$ and $S_2$ and accordingly, the upper real-time bound is relaxed by
-the multiplication of WCRT and the first element of set $S_2$.
+for the real-time postponement to be satisfied. The algorithm takes as
+input the WCRT and set $S_2$, recall that set $S_2$ represents the
+logical ticks required to satisfy the real-time requirement at the
+BCRT. We take the very first value from set $S_2$ and multiply it with
+the WCRT to get the relaxation $N$. The first element of set $S_2$ is
+returned as the logical tick \texttt{d}. The fundamental observation is
+that we have to postpone the proceeding statement for at least $M$ units
+of real-time, hence, under-approximation is out of question. We can
+still over-approximate, but to reduce the resultant error, we should
+over-approximate by the least possible value, which is the lower bound
+of set $S_2$. Thus, the lower bound is considered to be the only element
+shared between the two sets $S_1$ and $S_2$ and accordingly, the upper
+real-time bound is relaxed by the multiplication of WCRT and the first
+element of set $S_2$.
 
-For example, assume $BCRT=5$ and $WCRT=100$, with a required delay of
-$(200..200)$. Algorithm~\ref{alg:1}, results in $D$ being an empty set,
-since $S_1=[2]$ and $S_2=[40]$, respectively. In such a case, the
-resultant relaxation is: $40 \times 100$ using Algorithm~\ref{alg:2},
-which results in a time of $4000$ units. Such large over-approximations
-can be avoided by breaking the critical paths (worst case ticks) in the
-program (by inserting \texttt{pause} constructs), just like in hardware
-design, which is very \textit{unsurprising} considering Esterel has its
-origins in hardware.
+For example, assume $BCRT=5$ and $WCRT=100$, with a required
+postponement of $(200..200)$. Algorithm~\ref{alg:1}, results in $D$
+being an empty set, since $S_1=[2]$ and $S_2=[40]$, respectively. In
+such a case, the resultant relaxation is: $40 \times 100$ using
+Algorithm~\ref{alg:2}, which results in a time of $4000$ units. Such
+large over-approximations can be avoided by breaking the critical paths
+(worst case ticks) in the program (by inserting \texttt{pause}
+constructs), just like in hardware design, which is very
+\textit{unsurprising} considering Esterel has its origins in hardware.
+
+\paragraph{Period execution of the clock domains}
+\label{sec:peri-exec-clock}
+
+The aforementioned technique maintains the essence of synchronous and
+GALS programs -- the logical tick, but in doing so is unable to provide
+the exact real-time wait of $M$ time units in case of
+\texttt{wait\_exactly(M)} statement, thereby violating its
+semantics. The second option is to execute the CD periodically at the
+WCRT. WCRT determines the execution speed of any CD guaranteeing that no
+input events from the environment will be missed
+(cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
+the CD at periodically at the WCRT, we can choose the value of
+\textit{d} from set $S1$. This technique allows meeting the synchronous
+hypothesis, while at the same time guaranteeing the semantics of the
+\texttt{wait\_exactly(M)} statement. But, compared to the previous
+option, this technique is very inflexible and results in reduced
+response time and increased power consumption.
+
+We believe that the designer should be allowed to choose either of these
+techniques and hence, provide this option via compiler switches.
 
@@
   \includegraphics[scale=0.45]{./tool.pdf}
-  \caption{The tool-chain flow of SystemJ compiler with delays}
+  \caption{The tool-chain flow of SystemJ compiler with waits}
   \label{fig:4}
@@
 The interaction of the low-level execution platform and the described
-high-level technique of introducing real-time delays in the SystemJ
+high-level technique of introducing real-time waits in the SystemJ
 language is shown in Figure~\ref{fig:4}. We use static-analysis
@@
 Our compiler has a switch that enables over-approximation as described
-in Section~\ref{sec:extend-tehcn-vari} provided the delay requirements
+in Section~\ref{sec:extend-tehcn-vari} provided the wait requirements
 are not met. If this compiler switch is not enabled, the compiler simply
-outputs an error indicating the relaxation required. The system designer
-can then either set the compiler switch or change the program by
-shortening the critical path by introducing \texttt{pause}
-constructs. We show how introducing \texttt{pause} constructs can help
-reduce the WCRT in Section~\ref{sec:experimental-results}.
+decides to execute the CD at the periodically, where the period is the
+WCRT. The system designer can then either set the compiler switch or
+change the program by shortening the critical path by introducing
+\texttt{pause} constructs. We show how introducing \texttt{pause}
+constructs can help reduce the WCRT in
+Section~\ref{sec:experimental-results}.
 

### 2014-01-27 discussion.tex
Updated the paper

@@
-\section{Discussion -- addressing perceived limitations}
+\section{Discussion}
 \label{sec:disc-perc-limit}
 
-We dedicate this section to discuss the details of the delay semantics,
-the dependence on time analyzable platforms, and the various uses and
-programming paradigms that can be built using the \texttt{delay (M..N)}
-construct.
+We dedicate this section to discuss the details of the wait construct
+semantics, their usage and the advantages and limitations of our
+proposed technique to introducing these real-time constructs in a
+reactive setting.
 
-\subsection{Programming using the delay construct}
+\subsection{Programming using the wait\_exactly construct -- the
+  periodic reactions}
 \label{sec:progr-using-delay}
 
-All three delay constructs are useful for programming real systems. % The
+% All three delay constructs are useful for programming real systems. % The
 % first form of the delay construct $delay(M..N)$ is used to program
@@
 
-\subsubsection{Non-deterministic time}
-\label{sec:non-determ-time}
+% \subsubsection{Non-deterministic time}
+% \label{sec:non-determ-time}
 
-Many real-world systems require non-deterministic timing
-constructs. Where the exact real-time delay is not known or should not
-be known apriori. One such system is our motivating example -- the human
-response time system (HRTCS). Another one we presented was a real
-printer-spooler embedded controller in
-Section~\ref{sec:experimental-results}.
+% Many real-world systems require non-deterministic timing
+% constructs. Where the exact real-time delay is not known or should not
+% be known apriori. One such system is our motivating example -- the human
+% response time system (HRTCS). Another one we presented was a real
+% printer-spooler embedded controller in
+% Section~\ref{sec:experimental-results}.
 
-\subsubsection{Timeout}
-\label{sec:timeout}
+% \subsubsection{Timeout}
+% \label{sec:timeout}
 
-There are many instances when one would like to wait on an input from
-environment for only a specified amount of time. This can be programmed
-as:
+% There are many instances when one would like to wait on an input from
+% environment for only a specified amount of time. This can be programmed
+% as:
+
+% \begin{figure}[h!]
+%   \centering
+% 			\vspace{-10pt}
+% 			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={await,emit,trap,pause,exit,delay}]
+% // timeout after 1ms.
+% trap(T){{await(A);}||
+%         {delay(1 ms);exit(T);}}
+% \end{lstlisting}
+%   \caption{Programming low priority timeout tasks with delay statements}
+%   \label{fig:timeout}
+% 			\vspace{-10pt}
+% \end{figure}
+
+% In the above case, the program waits for signal \texttt{A} from the
+% environment for \textit{at least} \texttt{1 ms}, but it may wait longer
+% if other reactions/CDs are scheduled for execution. This second variant
+% of delay statement is very useful for programming low priority tasks.
+
+% \subsubsection{Periodic reactions}
+% \label{sec:periodic-reactions}
+
+% The third variant is useful for hard-real time guarantees, e.g., a
+% reaction, or a whole CD, can be programmed to run periodically like so:
 
@@
 	\vspace{-10pt}
-			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={await,emit,trap,pause,exit,delay}]
-// timeout after 1ms.
-trap(T){{await(A);}||
-        {delay(1 ms);exit(T);}}
+        \begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exactly}]
+          while(true){wait_exactly (1 ms); emit S; //do something
+          }
  \end{lstlisting}
-  \caption{Programming low priority timeout tasks with delay statements}
-  \label{fig:timeout}
-			\vspace{-10pt}
-\end{figure}
-
-In the above case, the program waits for signal \texttt{A} from the
-environment for \textit{at least} \texttt{1 ms}, but it may wait longer
-if other reactions/CDs are scheduled for execution. This second variant
-of delay statement is very useful for programming low priority tasks.
-
-\subsubsection{Periodic reactions}
-\label{sec:periodic-reactions}
-
-The third variant is useful for hard-real time guarantees, e.g., a
-reaction, or a whole CD, can be programmed to run periodically like so:
-
-\begin{figure}[h!]
-  \centering
-	\vspace{-10pt}
-		\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay}]
-while(true) { delay ((1..1)ms); emit S; 
- //do something }
- \end{lstlisting}
-  \caption{Programming periodic tasks with delay statements}
+  \caption{Programming periodic tasks with wait\_exactly statement}
   \label{fig:periodic}
@@
 
-The above program will emit signal \texttt{S} every \texttt{1 ms}. A
-periodic reaction (or CD) requires special consideration, like any other
-real-time periodic task. One should be aware of the interaction of
-delays and other \texttt{pause} constructs in the body of the reaction
-(or CD). For example, consider the (perceived) periodic program and its
-rewrite below.
+The program in Figure~\ref{fig:preemp}, reproduced from
+Section~\ref{sec:progr-using-exact}, is supposed to emit signal
+\texttt{S} every \texttt{1 ms}. Such periodic reactions (or CDs) require
+special consideration, like any other real-time periodic task. One
+should be aware of the interaction of wait and other \texttt{pause}
+constructs in the body of the reaction (or CD). For example, consider
+the (perceived) periodic program and its rewrite below.
 
@@
     \centering
-			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay}]
+			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exactly}]
 while(true) {
- delay(1..1 ms);
+ wait_exactly(1 ms);
  //extra pause
@@
 
-The system designer needs to be aware that just introducing a delay
+The system designer needs to be aware that just introducing a wait
 construct does not make the reaction (or CD) periodic. The program on
@@
 pauses introduces more logical ticks, consequently increasing the
-real-time delay. Static analysis techniques from real-time community can
-be used to automatically determine the delay specification in such
-cases, but describing these techniques is out of the scope of this work.
+real-time wait constructs. Static analysis techniques from real-time
+community can be used to automatically determine the wait specification,
+in such cases, but describing these techniques is out of the scope of
+this work.
 
-\subsection{Need for a time analyzable platform -- interaction of
-  preemption and delays}
-\label{sec:inter-preempt-delays}
+\subsection{Need for time analyzable platforms}
+\label{sec:need-time-analyzable}
+
+The proposed solution relies on time analyzable platforms, i.e.,
+platforms where WCRT and BCRT can be calculated using static analysis
+tools. Many might consider this to be too much of a restriction. In this
+section we debunk this \textit{perceived} disadvantage.
+
+\subsubsection{Interaction of timers and reactive constructs}
+\label{sec:inter-timers-react}
 
@@
 Preemption plays an important role in reactive languages. One needs to
-carefully consider the interplay of \texttt{delay} semantics with the
+carefully consider the interplay of wait constructs semantics with the
 preemption semantics of reactive languages. Previous attempts at
-incorporating delays (using external timers) have only been partially
+incorporating delay (using external timers) have only been partially
 successful, because of the complex interplay between real-time and
@@
 we enter a deadlock. Such problems are completely avoided in our
-technique, because we convert the real-time delays into logical delays
-(\texttt{pause} constructs), which interact well with preemption.
+technique, because we convert the real-time wait constructs into logical
+waits (\texttt{pause} constructs), which interact well with preemption.
 Furthermore, synchronous/GALS languages have always been targeted at
 real-time analyzable platforms and constrained
-environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}, one can use
+environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}. One can use
 our technique within such a setting with ease.
+
+\subsubsection{The timer resolution problem}
+\label{sec:resolution-real-time}
+
+\begin{figure}[h!]
+  \centering
+	\vspace{-10pt}
+		\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay,suspend}]
+{
+ //reaction R1
+ emit START_TIMER(1 ms); 
+ await (TIMER);emit O1;
+}
+||
+{
+ //reaction R2
+ // do something
+}
+		\end{lstlisting}
+  \caption{Resolution of external timers}
+  \label{fig:resolution}
+	\vspace{-10pt}
+\end{figure}
+
+There is yet another problem with timer based systems. Let us consider
+the program in Figure~\ref{fig:resolution}. On a cursory look this
+example should work fine. An external timer is started in reaction R1,
+which counts down from 1 ms. Once this time has elapsed a signal
+\texttt{TIMER} is generated from this external timer, which should emit
+O1 in turn. There is no \texttt{suspend} construct encapsulating the
+timer and hence all seems fine. Let us now consider what happens to the
+whole CD, including reaction R2; while reaction R1 is waiting for the
+\texttt{TIMER} signal, reaction R2 is making progress. Input signals can
+only be captured from the environment at the logical tick boundaries
+(cf. Section~\ref{sec:mapping-logical-time}). In
+Figure~\ref{fig:resolution}, reaction R2 determines the length of the
+tick, because it is performing the heavier computation. Consider what
+happens if the external timer generates the \texttt{TIMER} signal, but
+the system is still performing computation and the logical tick has not
+yet finished -- we \textit{miss} this \texttt{TIMER} signal from the
+environment. Thus, even in the absence of the \texttt{suspend}
+construct, we are not guaranteed that the program will capture the
+external timer signal. In fact, the WCRT needs to be smaller than 1 ms
+for the above program to perform as expected. But, this again implies
+need for a time analyzable platform even when using external timers.
+
+In general we observe that the WCRT determines the lowest resolution of
+any real-time construct whether it be external timer independent, like
+the proposed technique, or external timer dependent, like above. This
+discovery is one of the key insights for proposing the solution
+described in this paper.
 
@@
 
-\subsection{Interaction of channel communication and delays}
+\subsection{Interaction of channel communication and wait constructs}
 \label{sec:inter-chann-comm}
@@
 running CDs, are an addition in the SystemJ language. Like interaction
-of preemption and delays, conversion of real-time delays to logical
-ticks also interacts well with channel rendezvous, because the semantics
-of interaction are well defined~\cite{amal10}. More importantly, we need
-to consider the interplay of channel communication and WCRT/BCRT
-analysis. Since, channel communication does not stop logical time
-(see~\cite{amal10}) WCRT/BCRT are unaffected by channel
+of preemption and waits, conversion of real-time wait constructs to
+logical ticks also interacts well with channel rendezvous, because the
+semantics of interaction are well defined~\cite{amal10}. More
+importantly, we need to consider the interplay of channel communication
+and WCRT/BCRT analysis. Since, channel communication does not stop
+logical time (see~\cite{amal10}) WCRT/BCRT are unaffected by channel
 communication. % The response time to input signals \textit{is} though!

### 2014-01-27 exps.tex
Updated the paper

@@
 \begin{table}[h!]
-	\small
+	\scriptsize
 	\centering
@@
 \\ \midrule
-HRTCS                  & \texttt{delay(50.3..200.3 ms)} ($CD1$) & 2                  & 56
+HRTCS                  & \texttt{wait\_inbetween(50.3..200.3 ms)} ($CD1$) & 2                  & 56
 \\ \midrule
-\multirow{5}{*}{Motor} & \texttt{delay(2.4..2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
-											 & \texttt{delay(1.667..1.667 ms)}      &                    &                      \\ 
-											 & \texttt{delay(0.05..0.05 ms)}       &                    &                      \\ 
-											 & \texttt{delay(0.3..0.3 ms)}        &                    &                      \\ 
-											 & \texttt{delay(0.733..0.733 ms)}      &                    &
+\multirow{5}{*}{Motor} & \texttt{wait\_exactly(2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
+											 &
+                                                                                         \texttt{wait\_exactly(1.667
+                                                                                           ms)}      &                    &                      \\ 
+											 &
+                                                                                    \texttt{wait\_exactly(0.05
+                                                                                           ms)}       &                    &                      \\ 
+											 &
+                                                                                    \texttt{wait\_exactly(0.3
+                                                                                           ms)}        &                    &                      \\ 
+											 &
+                                                                                         \texttt{wait\_exactly(0.733
+                                                                                           ms)}      &                    &
 \\ \midrule
 \multirow{3}{*}{Printer}
-& \texttt{delay(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
-& \texttt{delay(1.2..30.3 ms)} ($CD2$)		  &														&											\\
-& \texttt{delay(5.6..100.2 ms)} ($CD2$)     &														&											
+& \texttt{wait\_inbetween(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
+& \texttt{wait\_inbetween(1.2..30.3 ms)} ($CD2$)		  &														&											\\
+& \texttt{wait\_inbetween(5.6..100.2 ms)} ($CD2$)     &														&											
 \\ \midrule
-\multirow{2}{*}{AECS} & \texttt{delay(10000 ms)} ($CD1$) 
+\multirow{2}{*}{AECS} & \texttt{wait\_atleast(10000 ms)} ($CD1$) 
 & \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
-& \texttt{delay(10000..10000 ms)} ($CD2$) & &
+& \texttt{wait\_exactly(10000 ms)} ($CD2$) & &
 \\ \bottomrule
@@
 \centering
-	\caption{Actual delays obtained for \texttt{delay} constructs in the
+	\caption{Actual delays obtained for wait constructs in the
 benchmark programs based on their BCRT and WCRT}
@@
 	\cmidrule(r){2-8}
-	& d &1272-1739 & 57 & 143-163 & 46-617 & 56062 &77844\\ 
+	& d &1506-1788 & 57 & 143-163 & 46-617 & 56062 &77844\\ 
 	\cmidrule(r){1-8}
@@
 HRTCS is the human response system described in
-Section~\ref{sec:motivating-example}. AECS (Access and Environmental
-Control System) is an application that controls an intelligent room
-environment described in~\cite{aecs_ispa}. The system interacts with two
-external timers through input/output signals in order to measure how
-long the entrance door to the room is being opened to detect intruders.
-The program has been modified such that the delay constructs are used
-instead of external timers as shown in Figure~\ref{delaycode}. In the
-original AECS, the system needed to trigger external timer by emitting
-the output signal in order to wait for a specific amount of time (e.g.
+Section~\ref{sec:motivating-example}. Motor is a stepper motor
+controller from~\cite{Bourke2009a}, used for producing monochrome images
+on paper by correct application of current to a print head (a row of
+resistors). In this example, duration of the current applied is
+controlled by the feedback logic using real-time waits, which are
+expressed using reactive constructs. Printer is an example
+from~\cite{Schneider:1999:CRT:555233}, which models a printer and a
+spooler in timed CSP. Its operation can be described as follows: as soon
+as the spooler receives a print job it passes it to the printer through
+a channel where actual printing process take place. This example along
+with HRTCS demonstrates the use of non exact wait statements.
+
+AECS (Access and Environmental Control System) is an application that
+controls an intelligent room environment described
+in~\cite{aecs_ispa}. The system interacts with two external timers
+through input/output signals in order to measure how long the entrance
+door to the room is being opened to detect intruders.  The program has
+been modified such that the wait constructs are used instead of external
+timers as shown in Figure~\ref{delaycode}. In the original AECS, the
+system needed to trigger external timer by emitting the output signal in
+order to wait for a specific amount of time (e.g.
 \texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
@@
 \texttt{TimeOut}, which is polled using the \texttt{await}
-statement. With the new delay constructs, these timers are not necessary
-and delays can be expressed directly in the code as shown in
+statement. With the new wait constructs, these timers are not necessary
+and waits can be expressed directly in the code as shown in
 Figure~\ref{delaycode:b}.
@@
       \scriptsize
-		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,delay,suspend}]
+		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,suspend}]
 emit TimerTriggered;
@@
   \hspace{1cm}%
-  \begin{SubFloat}{\label{delaycode:b}Program with delay constructs}
-		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,delay,suspend}]
+  \begin{SubFloat}{\label{delaycode:b}Program with wait constructs}
+		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,wait_atleast,suspend}]
 // No emit statement
@@
   abort(DoorOpened){
-    delay(10000 ms);
+    wait_atleast (10000 ms);
     exit(T);  
@@
 	\end{SubFloat}
-\caption{Replacing external timer with delay construct}
+\caption{Replacing external timer with wait\_atleast construct}
 \label{delaycode}
 \end{figure}
-
-Motor is a stepper motor controller from~\cite{Bourke2009a}, used for
-producing monochrome images on paper by correct application of current
-to a print head (a row of resistors). In this example, duration of the
-current applied is controlled by the feedback logic using real-time
-delays, which are expressed using reactive constructs. Printer is an
-example from~\cite{Schneider:1999:CRT:555233}, which models a printer
-and a spooler in timed CSP. Its operation can be described as follows:
-as soon as the spooler receives a print job it passes it to the printer
-through a channel where actual printing process take place. This example
-along with HRTCS demonstrates the use of non-deterministic delays.
 
@@
 
-HRTCS consists of 2 CDs, but only one with real-time delays. AECS also
-has 2 CDs; CD1 consist of a single delay statement, while CD2 has 2
-delay statements. In this case, as explained previously, delay
-statements replace the external timers and corresponding interface (I/O)
+HRTCS consists of 2 CDs, but only one with real-time waits. AECS also
+has 2 CDs; CD1 consist of a single wait statement, while CD2 has 2 wait
+statements. In this case, as explained previously, wait statements
+replace the external timers and corresponding interface (I/O)
 signals. The system triggers an alarm 10 seconds after detection of an
@@
 between 5.6 to 100.2 ms after the notification. In this example, bounded
-delay ensures that both; channel communication and print job should
-occur within the specified amount of time. Lastly, Motor is a
-synchronous program with 5 delay statements that control the amount of
-current applied to the stepper motor coils.
+wait ensures that both; channel communication and print job should occur
+within the specified amount of time. Lastly, Motor is a synchronous
+program with 5 exact wait statements that control the amount of current
+applied to the stepper motor coils.
 
@@
 each benchmark program. Note that we allowed the compiler to relax the
-upper real-time bounds even for bounded (\texttt{M..N}) or exact
-(\texttt{M..M}) delays when it failed to find a feasible solution.
+upper real-time bounds for \texttt{wait\_exactly} and
+\texttt{wait\_inbetween} when it failed to find a feasible solution.
 However, a user can prohibit such behavior using compiler switch as
@@
 As a first step, BCRT and WCRT were calculated for all examples. Given
-the real-time \mbox{\texttt{delay (M..N)}} provided by programmers,
-required logical delay \textit{d} was determined by applying
-Algorithm~\ref{alg:1}. In HRTCS, real-time delay of the range
-50.3..200.3 (ms) is only satisfied when \emph{d} is between 1272 and
-1739 (as shown previously in Section~\ref{sec:find-logic-delay}). Delay
+the real-time \mbox{\texttt{wait\_inbetween (M..N)}} provided by
+programmers, required logical delay \textit{d} was determined by
+applying Algorithm~\ref{alg:1}. In HRTCS, real-time delay of the range
+50.3..200.3 (ms) is only satisfied when \emph{d} is between 1506 and
+1788 (as shown previously in Section~\ref{sec:find-logic-delay}). Delay
 2.7..10.6 (ms) is satisfied in Printer/CD1 with \emph{d} of 143..163.
@@
 
-The previous results show that for HRTCS and Printer that the compiler
-was able to find \textit{d} without relaxing upper bound of
-\texttt{delay}. Figure~\ref{fig:graph1} shows the accuracy of our
-results by comparing differences between programmer specified upper
-bounds (Table~\ref{fig:benchlist}) and the actual bounds obtained using
-our techniques (Table~\ref{fig:comparison}) for the other examples
-requiring relaxation. AECS/CD1 delay is not included in
+The previous results show that for HRTCS and Printer the compiler was
+able to find \textit{d} without relaxing upper bound of wait
+statements. Figure~\ref{fig:graph1} shows the accuracy of our results by
+comparing differences between programmer specified upper bounds
+(Table~\ref{fig:benchlist}) and the actual bounds obtained using our
+techniques (Table~\ref{fig:comparison}) for the other examples requiring
+relaxation. AECS/CD1 wait specification is not included in
 Figure~\ref{fig:graph2}, because although it requires relaxation, the
-delay specification (Table~\ref{fig:benchlist}) does not have an upper
+wait specification (Table~\ref{fig:benchlist}) does not have an upper
 bound.
@@
 the architecture of the execution platform (i.e. JOP), the differece
-factor can can only be reduced by modifying the benchmark programs such
-that the critical paths are broken up using \texttt{pause} constructs,
+factor can only be reduced by modifying the benchmark programs such that
+the critical paths are broken up using \texttt{pause} constructs,
 thereby reducing the WCRT.  Figure~\ref{fig:improv} shows segments of
@@
 The result of the upper bound relaxation with reduced WCRT is shown in
-Figure~\ref{fig:graph2}.  The upper bounds for delay statements in
+Figure~\ref{fig:graph2}.  The upper bounds for wait statements in
 AECS/CD2 decreased by 11.81\%, while that for Motor example reduced from

### 2014-01-27 introduction.tex
Updated the paper

@@
 systems~\cite{amal121}, etc. Usually these systems also need to meet
-real-time constraints enforced by the environment. Yet, these languages
+real-time constraints imposed by the environment. Yet, these languages
 do not support describing real-time statements as first class language
@@
 
-The major motivations for introducing delay mechanism and at the same
-time contributions of the paper are: (1) delays allow real-time based
-synchronization between concurrent behaviors, (2) delay model relies on
-the use of relative instead of absolute real-time, i.e., a delay in
-selected time units is counted from the currently executing SystemJ
-statement (3) delays allow mixing behaviors with real-time features and
-others with only logical time and (4) real-time delay does not affect
+The major motivations for introducing these real-time \textit{wait}
+mechanisms and at the same time \textbf{contributions} of the paper are:
+(1) these waiting mechanisms allow real-time based synchronization
+between concurrent behaviors, (2) the waiting model relies on the use of
+relative instead of absolute real-time, i.e., a wait in selected time
+units is counted from the currently executing SystemJ statement, (3)
+these waiting mechanisms allow mixing behaviors with real-time features
+and others with only logical time and (4) real-time wait does not affect
 the model of logical time in SystemJ program as long as the amount of
@@
 analysis. All these as well as semantics of the delay construct are
-discussed and illustrated in the remaining part of the paper. Our
-contributions can be further refined as follows:
+discussed and illustrated in the rest of the paper. Our contributions
+can be further refined as follows:
 \begin{enumerate*}
@@
   introducing real-time delays.
-% \item Our solution does not change the functional or timing semantics of
-%   GALS and its subset synchronous programs, in fact, our solution does
-%   not require one to change the mid-end, the back-end or the related
-%   optimization phases of the compiler at all.
-\item We are the first, to our knowledge, to allow specification of both
-  deterministic and non-deterministic real-time in GALS and synchronous
-  languages. \textit{Deterministic delay:} is the ability to specify an
-  exact \textit{real-time} delay. \textit{Non-deterministic delay:} is
-  the ability to specify a range of real-time values for a delay
-  statement, e.g., delay statement in Figure~\ref{fig:1}.
+\item We are the first, to our knowledge, to allow specification of
+  both; exact and non-exact real-time in GALS and synchronous programs.
+\item We elucidate upon a key insight relating real-time and
+  logical-time -- \textit{the resolution of any real-time statement is
+    dependent on the worst case logical tick time} -- an aspect that
+  seems to have escaped in the current literature of real-time in
+  synchronous/GALS languages.
 \end{enumerate*}
@@
 techniques required for the rest of the
-paper. Section~\ref{sec:intr-real-time} introduces real-time delays in
-the GALS paradigm and describes their compilation into logical
+paper. Section~\ref{sec:intr-real-time} introduces real-time mechanisms
+in the GALS paradigm and describes their compilation into logical
 time. Section~\ref{sec:experimental-results} gives the experimental

### 2014-01-27 motivation.tex
Updated the paper

@@
-\subsection{Motivating Example}
+\subsection{Motivation via examples of real-time systems programmed in a
+reactive language -- SystemJ}
 \label{sec:motivating-example}
@@
 %  }
+
+Real-time operating systems~\cite{barry2009using} usually provide two
+types of mechanisms for introducing real-time in the developed program:
+(1) the ability to perform a timeout and (2) the ability to run tasks
+periodically with some real-time period $T$. We would like to provide
+similar mechanisms in reactive languages for real-time program
+development. In fact, we provide \textit{exact} and \textit{non exact}
+real-time control mechanisms -- a generalization of the two real-time
+mechanisms introduced in real-time operating systems.
+
+\subsubsection{Programming using the non exact real-time control
+  construct the \textrm{\texttt{wait\_inbetween}} statement}
+\label{sec:progr-using-non}
+
 \begin{figure}[t!]
-% \begin{center}
 	\vspace{-10pt}
-		\begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,delay,suspend}]
+        \begin{SubFloat}{\label{delaycode:a}A human responsiveness
+            system (HRTCS)}
+        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,suspend}]
 { // Clock-domain 1
@@
      //exit after any where from 50.3 to 200.3 ms.
-     delay (50.3 .. 200.3 ms); 
+     wait_inbetween (50.3 .. 200.3 ms); 
      exit(TRAP); 
@@
 \end{lstlisting}
-\caption{An example human responsiveness system (HRTCS)}
+\end{SubFloat}
+\begin{SubFloat}{\label{delaycode:b}Timing diagram -- the X-axis shows
+    the execution of the \texttt{wait\_inbetween} statement. The Y-axis
+    is the time elapsed.}
+  % \begin{figure}[h!]
+    \centering
+    \includegraphics[scale=0.7]{FF1}
+  % \end{figure}
+\end{SubFloat}
+\caption{Programming a human responsiveness system (HRTCS) with
+  \texttt{wait\_inbetween}}
 % }
@@
 
-The SystemJ pseudo-code~\cite{amal10} for such a system is shown in
-Figure~\ref{fig:1}. We use the SystemJ language in this paper, first of
-all, because being a \textit{Globally Asynchronous Locally Synchronous}
-(GALS) language it helps us easily describe a larger class of systems
-(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
-languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
-allows us to explore the marriage of real-time and logical time, which
-is equally applicable to the synchronous sub-set.
+The SystemJ pseudo-code~\cite{amal10} for such a system, along with the
+expected timing behavior of the \texttt{wait\_inbetween} statement, is
+shown in Figure~\ref{fig:1}. We use the SystemJ language in this paper,
+first of all, because being a \textit{Globally Asynchronous Locally
+  Synchronous} (GALS) language it helps us easily describe a larger
+class of systems (like HRTCS in Figure~\ref{fig:1}) compared to purely
+synchronous languages like Esterel. Moreover, being a super-set of
+Esterel, SystemJ allows us to explore the marriage of real-time and
+logical time, which is equally applicable to the synchronous sub-set.
 
@@
 clock-domain. All the SystemJ programming constructs used to implement
-this system are described in~\cite{amal10} and Table~\ref{tab:1} in
-Section~\ref{sec:background}. The delay specification, which models the
-non-deterministic delay between 50.3 ms and 200.3 ms is not supported in
-the current language and cannot be compiled with the current SystemJ
-compiler. We introduce real-time \texttt{delay} statements in SystemJ.
+this system are described in~\cite{amal10} and Table~\ref{tab:1}. The
+\texttt{wait\_inbetween} specification, which models the non exact delay
+between 50.3 ms and 200.3 ms is not supported in the current language
+and cannot be compiled with the current SystemJ compiler. We introduce
+real-time \texttt{wait\_inbetween} statements in SystemJ. Such non exact
+real-time mechanisms are useful in developing systems where the
+real-time delay is not know or should not be known a priori.
+
+\subsubsection{Programming a timeout using the non exact real-time
+  construct the \texttt{wait\_atleast} statement}
+\label{sec:progr-time-using}
+
+\begin{figure}[b!]
+	\centering
+	\vspace{-10pt}
+        \begin{SubFloat}{\label{dd:a}Timeout waiting for the input
+            signal \texttt{DoorOpened}}
+        \begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,wait_atleast,suspend}]
+trap(T){
+  abort(DoorOpened){
+    wait_atleast(10000 ms);
+    exit(T);  
+  }
+}
+\end{lstlisting}
+\end{SubFloat}
+\begin{SubFloat}{\label{dd:b}Timing diagram -- the X-axis shows the
+    execution of the \texttt{wait\_atleast} statement. The Y-axis is the
+    time elapsed.}
+\includegraphics[scale=0.7]{FF2}
+\end{SubFloat}
+\caption{Programming a timeout with \texttt{wait\_atleast}}
+\label{dd}
+\end{figure}
+
+There are many instances when one would like to wait on an input from
+environment for only a specified amount of time. If the signal is not
+received then a timeout is generated so that the program can make
+progress. Such a timeout programmed in the SystemJ pseudo-code along
+with its timing behavior is shown in Figure~\ref{dd}. The SystemJ
+program is waiting on an input signal, \texttt{DoorOpened}, from the
+environment for 10 seconds and makes progress irrespective of the
+reception of this signal after 10 seconds.
+
+The timing diagram, Figure~\ref{dd:b}, shows that the
+\texttt{wait\_atleast} statement waits for a \textit{at least} 10
+seconds, but unlike \texttt{wait\_inbetween} there is no upper
+bound. The statement following the \texttt{wait\_atleast} statement is
+executed after a finite countable, but unbounded delay. Such a statement
+is useful for programming low priority tasks.
+
+\subsubsection{Programming periodic tasks using the exact real-time
+  construct the \texttt{wait\_exactly} statement}
+\label{sec:progr-using-exact}
+
+The previous wait statements are non exact, i.e., the execution of the
+statement following the wait statement cannot be controlled exactly. The
+final variant we introduce allows doing exactly that. We call it
+\texttt{wait\_exactly} and this statement can be used to program exact
+timeouts, periodic tasks, etc.
+
+\begin{figure}[t!]
+  \centering
+	\vspace{-10pt}
+        \begin{SubFloat}{\label{pp:a}Periodically emitting signal S}
+        \begin{lstlisting}[style=sysj,morekeywords={emit,trap,pause,exit,wait_exactly}]
+while(true) { 
+ wait_exactly (1 ms); 
+ emit S; 
+ //do something 
+}
+\end{lstlisting}
+\end{SubFloat}
+\begin{SubFloat}{\label{pp:b}Timing diagram -- the X-axis shows the
+    execution of the \texttt{wait\_exactly} statement. The Y-axis is the
+    time elapsed.}
+  \includegraphics[scale=0.7]{FF3}
+\end{SubFloat}
+  \caption{Programming a periodic task with wait\_exactly}
+  \label{fig:periodic}
+  \vspace{-10pt}
+\end{figure}
+
+One such example; periodic emission of a signal to the environment is
+shown in Figure~\ref{fig:periodic}. As before upon execution of
+\texttt{wait\_exactly (1 ms)} statement, the time starts elapsing. But,
+unlike in previous cases the following statement needs to be executed
+once 1 ms have passed, as shown in Figure~\ref{pp:b}.
 

### 2014-01-27 related_work.tex
Updated the paper

@@
 which like SystemJ is an extension of synchronous language Esterel to
-asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
-external timers do not interact well with preemption constructs in these
-languages. Bourke et al. introduce real-time as first class constructs
-in the Esterel language. They like us provide real-time \texttt{delay}
-as first class Esterel programming constructs, and translate them into
-Esterel kernel constructs. But, unlike us they do not translate delays
+asynchorny. But, as mentioned in Section~\ref{sec:inter-timers-react}
+and~\ref{sec:resolution-real-time} external timers do not interact well
+with preemption constructs in these languages. Bourke et al. introduce
+real-time as first class constructs in the Esterel language. They like
+us provide real-time wait statements called \texttt{delay} as first
+class Esterel programming constructs, and translate them into Esterel
+kernel constructs. But, unlike us they do not translate wait statements
 into \texttt{pause} constructs directly. Instead, logical ticks are
@@
 need to be determined and present on the system for the solution to be
-realizable. Moreover, non-deterministic real-time delays and integration
-with non-maximal parallelism is not studied at all. Our solution does
-not require external timers (like~\cite{rsh94}) or logical notion of
-timers (like~\cite{Bourke2009a}). In our solution we calculate the delay
-in logical ticks \textit{after} allocation and scheduling, it gives the
-compiler developers the chance to optimize for many different criteria
-such as computation time, energy, power, etc, without worrying about or
-violating the real-time delay specification.
+realizable. Moreover, non exact real-time waits and integration with
+non-maximal parallelism is not studied at all. Our solution does not
+require external timers (like~\cite{rsh94}) or logical notion of timers
+(like~\cite{Bourke2009a}). In our solution we calculate the real-time
+waiting time in logical ticks \textit{after} allocation and scheduling,
+it gives the compiler developers the chance to optimize for many
+different criteria such as computation time, energy, power, etc, without
+worrying about or violating the real-time wait specification.
 
 Other works such as Quartz~\cite{glog02} and Bertin et
-al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate delay
-statements in Esterel using \texttt{pause} and special \textit{pragmas}
-(annotations), respectively.  The Quartz solution is targeted at
-studying timing properties using model checkers rather than
+al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate real-time
+wait statements in Esterel using \texttt{pause} and special
+\textit{pragmas} (annotations), respectively.  The Quartz solution is
+targeted at studying timing properties using model checkers rather than
 implementation, like us. The solution introduced by Bertin et

### 2014-01-27 conclusion.tex
Added proofs, can talk a bit more about completeness and soundness

@@
 ticks, which is invariant to the \textit{elasticity} of logical time. To
-our knowledge we are the first to introduce exact and inexact wait
+our knowledge we are the first to introduce exact and non-exact wait
 constructs in such languages and show their usefulness in designing

### 2014-01-27 correct.tex
Added proofs, can talk a bit more about completeness and soundness

\section{Proof of correctness}
\label{sec:proof-correctness}

Section~\ref{sec:intr-real-time} describes the techniques and algorithms
for introducing various types of wait statements providing real-time
control in the SystemJ language. In this section we give the formal
proof of correctness for the algorithms described in
Section~\ref{sec:intr-real-time}.

\begin{lemma}
  Algorithm~\ref{alg:1} gives a value for \texttt{d} such that: $\lceil
  \frac{M}{BCRT} \rceil \leq d \leq \lfloor \frac{N}{WCRT} \rfloor$
  provided $D$ is a non-empty set for construct \texttt{wait\_inbetween
    (M..N)}
\label{lemma:1}
\end{lemma}
\begin{proof}
  This lemma is trivially true from the definition of set $D$ in
  Algorithm~\ref{alg:1}.
\end{proof}

\begin{theorem}
 Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
 value \texttt{d}, such that for any given $BCRT \leq T \leq WCRT$,
 $M \leq T \times d \leq N$ holds.
\label{th:1}
\end{theorem}
\begin{proof}
  We use case analysis and contradiction to prove the above theorem.
  \begin{compactenum}[\hspace{0.25cm} 1.]
  \item Proof for case $T = WCRT$: Assume \mbox{$d \times WCRT < M$},
    then $d < \frac{M}{WCRT}$. $BCRT \leq WCRT$, by definition, hence,
    $\frac{M}{WCRT} \leq \frac{M}{BCRT}$. Thus for the assumption to
    hold, $d < \frac{M}{BCRT}$, which contradicts
    Lemma~\ref{lemma:1}. The assumption $d \times WCRT >N$ is trivially
    proven from Lemma~\ref{lemma:1}.
  \item Proof for case $T = BCRT$: Like above, the assumption $d \times
    BCRT < M$ is trivially proven from Lemma~\ref{lemma:1}. For $d
    \times BCRT > N$ to hold, $d > \frac{N}{BCRT}$ should hold. But,
    $BCRT \leq WCRT$, by definition, hence, $\frac{N}{BCRT} \geq
    \frac{N}{WCRT}$, which again contradicts Lemma~\ref{lemma:1}.
  \item Proof for case $BCRT < T < WCRT$: For $d \times T < M$ to hold,
    $d < \frac{M}{T}$ should hold. By definition, $T > BCRT$, thus,
    $\frac{M}{BCRT} < \frac{M}{T}$, hence, $d < {M}{BCRT}$ should hold,
    which contradicts Lemma~\ref{lemma:1}. We can prove the other case
    $d \times T \leq N$, using the upper bound from Lemma~\ref{lemma:1}
    similarly.
  \end{compactenum}
\end{proof}

The proof for correctness for Algorithm~\ref{alg:2} follows from
Theorem~\ref{th:1}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

### 2014-01-27 delays.tex
Added proofs, can talk a bit more about completeness and soundness

@@
   an exact delay, but rather the control is allowed to proceed to the
-  next statement anytime after waiting for \textit{at least} time $M$.
+  next statement anytime after waiting for \textit{at least} $M$ times
+  units.
 \item Given a SystemJ program \texttt{wait\_exactly (M); p}, where $M
@@
 
-Let us revisit our motivating example to illustrate the algorithm. From
-Figure~\ref{fig:1} we know that $M$ is 50.3 ms and $N$ is 200.3 ms,
-respectively. Let the WCRT and BCRT be: 0.112 ms and 0.0334 ms,
-respectively. Thus, the algorithm proceeds as follows:
+Let us revisit the HRTCS example (cf. Figure~\ref{fig:1}) to illustrate
+the algorithm. From Figure~\ref{fig:1} we know that $M$ is 50.3 ms and
+$N$ is 200.3 ms, respectively. Let the WCRT and BCRT be: 0.112 ms and
+0.0334 ms, respectively. Thus, the algorithm proceeds as follows:
 
@@
   calculate the logical ticks that are always running at the WCRT and
-  satisfy the required real-time wait.
+  satisfy the required real-time wait period.
 \item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
-  \lfloor 200.3/0.0334 \rfloor$. $l_1 = 1506$ and $u_1 = 5997$. We do
+  \lfloor 200.3/0.0334 \rfloor$. $l_2 = 1506$ and $u_2 = 5997$. We do
   the same for the BCRT case.
@@
 which can run at any physical clock-speed, bounded by the BCRT and the
-WCRT of the program and still result in the desired real-time
-wait. Figure~\ref{fig:hrtcssoln} gives the solution space for the
+WCRT of the program and still result in the desired real-time wait
+period. Figure~\ref{fig:hrtcssoln} gives the solution space for the
 aforementioned calculation. The patterned space is the only overlapping
@@
 
-\paragraph{Period execution of the clock domains}
+\paragraph{Periodic execution of the clock domains}
 \label{sec:peri-exec-clock}
@@
 semantics. The second option is to execute the CD periodically at the
-WCRT. WCRT determines the execution speed of any CD guaranteeing that no
-input events from the environment will be missed
+WCRT. WCRT determines the fastest execution speed of any CD with the
+guarantee that no input events from the environment will be missed
 (cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
@@
 
-We believe that the designer should be allowed to choose either of these
-techniques and hence, provide this option via compiler switches.
+We believe that the designer should be allowed to choose from either of
+these techniques and hence, provide this option via compiler switches.
 
@@
 are not met. If this compiler switch is not enabled, the compiler simply
-decides to execute the CD at the periodically, where the period is the
-WCRT. The system designer can then either set the compiler switch or
-change the program by shortening the critical path by introducing
-\texttt{pause} constructs. We show how introducing \texttt{pause}
-constructs can help reduce the WCRT in
+decides to execute the CD periodically, where the period is the
+WCRT. WCRT effects the over-approximation or the period of the CD,
+irrespective of the solution that the system designer chooses. A large
+WCRT results in reduced performance and hence it is essential that the
+WCRT be optimized. One can optimize the WCRT of any CD by breaking the
+critical paths using the \texttt{pause} construct as we show later in
 Section~\ref{sec:experimental-results}.
+
+% The system designer can then either set the compiler switch or change
+% the program by shortening the critical path by introducing
+% \texttt{pause} constructs. We show how introducing \texttt{pause}
+% constructs can help reduce the WCRT in
+% Section~\ref{sec:experimental-results}.
 

### 2014-01-27 discussion.tex
Added proofs, can talk a bit more about completeness and soundness

@@
 pauses introduces more logical ticks, consequently increasing the
-real-time wait constructs. Static analysis techniques from real-time
+real-time wait period. Static analysis techniques from real-time
 community can be used to automatically determine the wait specification,
@@
 
-As mentioned previously, the proposed solution might need
-over-approximation techniques (Algorithm~\ref{alg:2}) to guarantee
-timing requirements (cf. Section~\ref{sec:extend-tehcn-vari}). This
-might give the readers an impression that timer based solution proposed
-for standard real-time operating systems should be utilized for
-GALS/synchronous programs. This is a fallacy. Bourke et
-al~\cite{Bourke2009a} have investigated this technique
-unsuccessfully. We elaborate upon the interplay of timers and reactivity
-to make the problem of using timers explicit.
+% As mentioned previously, the proposed solution might need
+% over-approximation techniques (Algorithm~\ref{alg:2}) to guarantee
+% timing requirements (cf. Section~\ref{sec:extend-tehcn-vari}). This
+% might give the readers an impression that timer based solution proposed
+% for standard real-time operating systems should be utilized for
+% GALS/synchronous programs. This is a fallacy. Bourke et
+% al~\cite{Bourke2009a} have investigated this technique
+% unsuccessfully. We elaborate upon the interplay of timers and reactivity
+% to make the problem of using timers explicit.
 
@@
 preemption semantics of reactive languages. Previous attempts at
-incorporating delay (using external timers) have only been partially
+incorporating postponement using external timers has only been partially
 successful, because of the complex interplay between real-time and
@@
 As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
-with the external timer. The above program sends a signal to an external
-timer and waits for 10 ms to pass by. Consider what happens when signals
-\texttt{S} and \texttt{TIMER} occur in the same logical tick, the
-\texttt{await} statement is never executed (due to suspend) and hence,
-we enter a deadlock. Such problems are completely avoided in our
-technique, because we convert the real-time wait constructs into logical
-waits (\texttt{pause} constructs), which interact well with preemption.
-Furthermore, synchronous/GALS languages have always been targeted at
-real-time analyzable platforms and constrained
+with the external timer. The program in Figure~\ref{fig:preemp} sends a
+signal to an external timer and waits for 10 ms to pass by. Consider
+what happens when signals \texttt{S} and \texttt{TIMER} occur in the
+same logical tick, the \texttt{await} statement is never executed (due
+to suspend) and hence, we enter a deadlock. Such problems are completely
+avoided in our technique, because we convert the real-time wait
+constructs into logical waits (\texttt{pause} constructs), which
+interact well with preemption.  Furthermore, synchronous/GALS languages
+have always been targeted at real-time analyzable platforms and
+constrained
 environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}. One can use
@@
 }
-||
+|| //synchronous parallel
 {
@@
 example should work fine. An external timer is started in reaction R1,
-which counts down from 1 ms. Once this time has elapsed a signal
+which counts down from 1 ms. Once this time has elapsed, a signal
 \texttt{TIMER} is generated from this external timer, which should emit
@@
 the proposed technique, or external timer dependent, like above. This
-discovery is one of the key insights for proposing the solution
+discovery is one of the key insights for proposing the solutions
 described in this paper.

### 2014-01-27 exps.tex
Added proofs, can talk a bit more about completeness and soundness

@@
 with real-time constraints. The characteristics of the benchmark set
-including required real-time delays and program sizes are shown in
-Table~\ref{fig:benchlist}.
-{\renewcommand{\arraystretch}{0.9}
+including required real-time waits and program sizes are shown in
+Table~\ref{fig:benchlist}.  {\renewcommand{\arraystretch}{0.9}
 %\renewcommand{\texttt}[1] {\begin{lstlisting}[language=Java,morekeywords={delay}] #1 \end{lstlisting}}
@@
 timers as shown in Figure~\ref{delaycode}. In the original AECS, the
-system needed to trigger external timer by emitting the output signal in
-order to wait for a specific amount of time (e.g.
+system needed to trigger an external timer by emitting the output signal
+in order to wait for a specific amount of time (e.g.
 \texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer

### 2014-01-27 introduction.tex
Added proofs, can talk a bit more about completeness and soundness

@@
 languages in designing discrete systems, there is a need to introduce
-real-time since, often, implementation models are in the continuous or
+real-time, since often, implementation models are in the continuous or
 discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
@@
 delay is within boundaries that can be determined statically by program
-analysis. All these as well as semantics of the delay construct are
+analysis. All these as well as semantics of the wait constructs are
 discussed and illustrated in the rest of the paper. Our contributions
@@
   of any language compilation and needs to be considered when
-  introducing real-time delays.
+  introducing real-time waits.
 \item We are the first, to our knowledge, to allow specification of
@@
     dependent on the worst case logical tick time} -- an aspect that
-  seems to have escaped in the current literature of real-time in
-  synchronous/GALS languages.
+  seems to have escaped in the current literature on incorporation of
+  real-time in synchronous/GALS languages.
 \end{enumerate*}
@@
 in the GALS paradigm and describes their compilation into logical
-time. Section~\ref{sec:experimental-results} gives the experimental
-results for a set of real-time
-applications. Section~\ref{sec:related-work} positions this work in
-relation to other approaches to introducing real-time. Finally, we
-conclude in Section~\ref{sec:concl-future-work}.
+time. Section~\ref{sec:proof-correctness} sketches the correctness
+proofs for the introduced
+algorithms. Section~\ref{sec:experimental-results} gives the
+experimental results for a set of real-time applications. In
+Section~\ref{sec:disc-perc-limit} we discuss the advantages and
+limitations of the proposed approach. Section~\ref{sec:related-work}
+positions this work in relation to other approaches to introducing
+real-time. Finally, we conclude in Section~\ref{sec:concl-future-work}.
 

### 2014-01-27 motivation.tex
Added proofs, can talk a bit more about completeness and soundness

@@
 \subsubsection{Programming using the non exact real-time control
-  construct the \textrm{\texttt{wait\_inbetween}} statement}
+  construct; the \textrm{\texttt{wait\_inbetween}} statement}
 \label{sec:progr-using-non}
@@
 data for future research purposes. The machine switches on a green light
-on a touch screen for anywhere between 50.3 ms to 200.3 ms, if the user
-can touch this green light then she is successful and an integer
-\texttt{1} is sent to a database, if unsuccessful a \texttt{0} is sent
-instead. Different statistics and analyses on human responsiveness can
-be based on the collected data.
+on a touch screen for anywhere \textit{in between} 50.3 ms to 200.3 ms,
+if the user can touch this green light then she is successful and an
+integer \texttt{1} is sent to a database, if unsuccessful a \texttt{0}
+is sent instead. Different statistics and analyses on human
+responsiveness can be based on the collected data.
 
@@
 and cannot be compiled with the current SystemJ compiler. We introduce
-real-time \texttt{wait\_inbetween} statements in SystemJ. Such non exact
+such \texttt{wait\_inbetween} statements in SystemJ. Such non exact
 real-time mechanisms are useful in developing systems where the
-real-time delay is not know or should not be known a priori.
+real-time delay is not known or should not be known a priori.
 
 \subsubsection{Programming a timeout using the non exact real-time
-  construct the \texttt{wait\_atleast} statement}
+  construct; the \texttt{wait\_atleast} statement}
 \label{sec:progr-time-using}
@@
 \subsubsection{Programming periodic tasks using the exact real-time
-  construct the \texttt{wait\_exactly} statement}
+  construct; the \texttt{wait\_exactly} statement}
 \label{sec:progr-using-exact}
@@
 \end{SubFloat}
-\begin{SubFloat}{\label{pp:b}Timing diagram -- the X-axis shows the
+\begin{SubFloat}{\label{p1:b}Timing diagram -- the X-axis shows the
     execution of the \texttt{wait\_exactly} statement. The Y-axis is the
@@
   \caption{Programming a periodic task with wait\_exactly}
-  \label{fig:periodic}
+  \label{fig:p1}
   \vspace{-10pt}
@@
 One such example; periodic emission of a signal to the environment is
-shown in Figure~\ref{fig:periodic}. As before upon execution of
+shown in Figure~\ref{fig:p1}. As before upon execution of
 \texttt{wait\_exactly (1 ms)} statement, the time starts elapsing. But,
 unlike in previous cases the following statement needs to be executed
-once 1 ms have passed, as shown in Figure~\ref{pp:b}.
+exactly after 1 ms have passed, as shown in Figure~\ref{p1:b}.
 

### 2014-01-27 related_work.tex
Added proofs, can talk a bit more about completeness and soundness

@@
 require external timers (like~\cite{rsh94}) or logical notion of timers
-(like~\cite{Bourke2009a}). In our solution we calculate the real-time
-waiting time in logical ticks \textit{after} allocation and scheduling,
-it gives the compiler developers the chance to optimize for many
-different criteria such as computation time, energy, power, etc, without
-worrying about or violating the real-time wait specification.
+(like~\cite{Bourke2009a}). In our solution we calculate the real waiting
+time in logical ticks \textit{after} allocation and scheduling, it gives
+the compiler developers the chance to optimize for many different
+criteria such as computation time, energy, power, etc, without worrying
+about or violating the real-time wait specification.
 

### 2014-01-27 correct.tex
updated the proof section with more proof sketches

@@
-\section{Proof of correctness}
+\section{Proofs of correctness, completeness, and soundness}
 \label{sec:proof-correctness}
@@
   Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
- value \texttt{d}, such that for any given $BCRT \leq T \leq WCRT$,
- $M \leq T \times d \leq N$ holds.
+  value \texttt{d}, provided $D$ is a non empty set, such that for any
+  given $BCRT \leq T \leq WCRT$, $M \leq T \times d \leq N$ holds.
 \label{th:1}
@@
 
-The proof for correctness for Algorithm~\ref{alg:2} follows from
-Theorem~\ref{th:1}.
+Next, we prove the completeness property.
+\begin{theorem}
+  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
+  value \texttt{d}, provided $D$ is a non empty set, such that for
+  \textbf{all} given $BCRT \leq T \leq WCRT$, $M \leq T \times d \leq N$
+  holds.
+\end{theorem}
+\begin{proof}
+  From Theorem~\ref{th:1} case 3.
+\end{proof}
+
+Next, we sketch the proof for the opposite of completeness -- the
+soundness property of Algorithm~\ref{alg:1}.
+
+\begin{theorem}
+  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1}
+  \textbf{does not} give a value \texttt{d}, such that for \textbf{all}
+  given $BCRT \geq T \geq WCRT$, $M \leq T \times d \leq N$ holds.
+\end{theorem}
+
+\begin{proof}
+  The proof is the opposite of case 3 in Theorem~\ref{th:1}.
+\end{proof}
+
+The proofs for correctness, completeness, and soundness for
+Algorithm~\ref{alg:2} are similar to those of Algorithm~\ref{alg:1}.
 

### 2014-01-28 correct.tex
updated text

@@
   value \texttt{d}, provided $D$ is a non empty set, such that for any
-  given $BCRT \leq T \leq WCRT$, $M \leq T \times d \leq N$ holds.
+	given $\hj{T}$ \hj{in the range of} $BCRT \leq T \leq WCRT$, $M \leq T \times d \leq N$ holds.
 \label{th:1}

### 2014-01-28 delays.tex
updated text

@@
 the SystemJ language for real-time control. The other two constructs
-\mbox{\texttt{wait\_atleast (M)}} and \mbox{\texttt{wait\_exactly (M)}}
+\mbox{\texttt{wait\_atleast (M)}} and \mbox{\texttt{wait\_exact (M)}}
 are variants of this single derived construct.
@@
   units.
-\item Given a SystemJ program \texttt{wait\_exactly (M); p}, where $M
+\item Given a SystemJ program \texttt{wait\_exact (M); p}, where $M
   \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
@@
 Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
-$M$ and $N$, respectively of the \texttt{wait\_inbetween} construct. We
+$M$ and $N$, respectively, of the \texttt{wait\_inbetween} construct. We
 first divide $M$ and $N$ with the BCRT and WCRT, respectively. This
@@
 the algorithm. From Figure~\ref{fig:1} we know that $M$ is 50.3 ms and
-$N$ is 200.3 ms, respectively. Let the WCRT and BCRT be: 0.112 ms and
-0.0334 ms, respectively. Thus, the algorithm proceeds as follows:
+$N$ is 200.3 ms, respectively. \hj{Timing analysis of the program has
+found} WCRT and BCRT be: 0.112 ms and 0.0334 ms, respectively. Thus, the
+algorithm proceeds as follows:
 
@@
 
-\paragraph{The \texttt{wait\_exactly(M)} construct}
+\paragraph{The \texttt{wait\_exact(M)} construct}
 \label{sec:extend-techn-vari}
 
-The \texttt{wait\_exactly(M)} variant is more interesting. Like before,
+The \texttt{wait\_exact(M)} variant is more interesting. Like before,
 we find sets $S_1$ and $S_2$, and find the intersection of the two sets
@@
 the exact real-time wait of $M$ time units in case of
-\texttt{wait\_exactly(M)} statement, thereby violating its
+\texttt{wait\_exact(M)} statement, thereby violating its
 semantics. The second option is to execute the CD periodically at the
@@
 hypothesis, while at the same time guaranteeing the semantics of the
-\texttt{wait\_exactly(M)} statement. But, compared to the previous
+\texttt{wait\_exact(M)} statement. But, compared to the previous
 option, this technique is very inflexible and results in reduced

### 2014-01-28 exps.tex
updated text

@@
 \\ \midrule
-\multirow{5}{*}{Motor} & \texttt{wait\_exactly(2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
+\multirow{5}{*}{Motor} & \texttt{wait\_exact(2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
 											 &
-                                                                                         \texttt{wait\_exactly(1.667
+                                                                                         \texttt{wait\_exact(1.667
                                                                                            ms)}      &                    &                      \\ 
 											 &
-                                                                                    \texttt{wait\_exactly(0.05
+                                                                                    \texttt{wait\_exact(0.05
                                                                                            ms)}       &                    &                      \\ 
 											 &
-                                                                                    \texttt{wait\_exactly(0.3
+                                                                                    \texttt{wait\_exact(0.3
                                                                                            ms)}        &                    &                      \\ 
 											 &
-                                                                                         \texttt{wait\_exactly(0.733
+                                                                                         \texttt{wait\_exact(0.733
                                                                                            ms)}      &                    &
@@
 & \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
-& \texttt{wait\_exactly(10000 ms)} ($CD2$) & &
+& \texttt{wait\_exact(10000 ms)} ($CD2$) & &
 \\ \bottomrule
@@
 
+\hj{
 HRTCS is the human response system described in
-Section~\ref{sec:motivating-example}. Motor is a stepper motor
+Section~\ref{sec:motivating-example}. }
+
+\hj{
+Motor is a stepper motor
 controller from~\cite{Bourke2009a}, used for producing monochrome images
@@
 controlled by the feedback logic using real-time waits, which are
-expressed using reactive constructs. Printer is an example
+expressed using reactive constructs. }
+
+\hj{
+Printer is an example
 from~\cite{Schneider:1999:CRT:555233}, which models a printer and a
@@
 with HRTCS demonstrates the use of non exact wait statements.
+}
 
@@
 We use the \textit{Java Optimized Processor}
-(JOP)~\cite{jop:jnl:jsa2007} to carry out all our experiments. JOP
-implements the JVM specification in hardware directly and has been shown
-to be real-time amenable. We statically estimate the WCRT and BCRT of
-these benchmark applications using the real-time analysis tools provided
-by the JOP design framework. The \textit{Safety Critical Java} (SCJ)
-specification we use for real-time analysis avoids garbage collection
-(GC) and hence, GC overheads and unpredictability are completely
-avoided.
+(JOP)~\cite{jop:jnl:jsa2007}, \hj{which has been shown to be real-time
+amenable}, to carry out all our experiments. We statically estimate the
+WCRT and BCRT of these benchmark applications using the real-time
+analysis tools provided by the JOP design framework. The \textit{Safety
+Critical Java} (SCJ) specification we use for real-time analysis avoids
+garbage collection (GC) and hence, GC overheads and unpredictability are
+completely avoided.
 
@@
 each benchmark program. Note that we allowed the compiler to relax the
-upper real-time bounds for \texttt{wait\_exactly} and
+upper real-time bounds for \texttt{wait\_exact} and
 \texttt{wait\_inbetween} when it failed to find a feasible solution.
@@
 % than HRTCS to obtain a feasible \texttt{d}. Therefore,
-On the other hand, Algorithm \ref{alg:2} is applied to those examples,
-where relaxation was required, to determine the minimum upper real-time
+On the other hand, Algorithm \ref{alg:2} is applied to those examples
+where relaxation was required to determine the minimum upper real-time
 bound \emph{N} such that there exists a solution for \emph{D} in

### 2014-01-28 motivation.tex
updated text

@@
 \subsection{Motivation via examples of real-time systems programmed in a
-reactive language -- SystemJ}
+reactive language SystemJ}
 \label{sec:motivating-example}
@@
             system (HRTCS)}
-        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,suspend}]
+        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,wait_exact,suspend}]
 { // Clock-domain 1
 while(true) {
-  trap(TRAP){
+  trap(T){
    // Abort if user touches the screen
@@
      wait_inbetween (50.3 .. 200.3 ms); 
-     exit(TRAP); 
+     exit(T); 
     } // reaction R2
@@
   } do {send S(0);} // Else Send 0
-  pause; // only construct to consume logical time
+  wait_exact(1 s); // Wait for 1 second
  }
@@
 { // Clock-domain 2
-  while(true){ receive S; /* do something with S */}
+  while(true){ receive S; 
+      if(#S == 1) counter++; else counter--;}
 }
@@
 on a touch screen for anywhere \textit{in between} 50.3 ms to 200.3 ms,
-if the user can touch this green light then she is successful and an
-integer \texttt{1} is sent to a database, if unsuccessful a \texttt{0}
-is sent instead. Different statistics and analyses on human
+if the user can touch this green light then she is successful and \hj{an 
+integer \texttt{1} is sent to other process in order to increase an
+integer counter, if unsuccessful a \texttt{0} is sent instead to
+decrease the counter}. Different statistics and analyses on human
 responsiveness can be based on the collected data.
@@
 shown in Figure~\ref{fig:1}. We use the SystemJ language in this paper, 
-first of all, because being a \textit{Globally Asynchronous Locally
+because being a \textit{Globally Asynchronous Locally
   Synchronous} (GALS) language it helps us easily describe a larger
@@
 
-The SystemJ program is divided into two clock-domains, the first
-clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
-signal) on the touch screen sensor and waits for the user to respond. If
-the user is able to touch the screen within the specified time of 50.3
-ms to 200.3 ms a positive response is sent to the second
-clock-domain. All the SystemJ programming constructs used to implement
-this system are described in~\cite{amal10} and Table~\ref{tab:1}. The
-\texttt{wait\_inbetween} specification, which models the non exact delay
-between 50.3 ms and 200.3 ms is not supported in the current language
-and cannot be compiled with the current SystemJ compiler. We introduce
-such \texttt{wait\_inbetween} statements in SystemJ. Such non exact
-real-time mechanisms are useful in developing systems where the
-real-time delay is not known or should not be known a priori.
+The SystemJ program is divided into two \hj{mutually asynchronous}
+clock-domains, the first clock-domain continuously displays a green
+light (\texttt{GREEN\_LIGHT} signal) on the touch screen sensor and
+waits for the user to respond. If the user is able to touch the screen
+within the specified time of 50.3 ms to 200.3 ms a positive response is
+sent to the second clock-domain. All the SystemJ programming constructs
+used to implement this system are described in Table~\ref{tab:1} \hj{and
+more details can be found in~\cite{amal10}}. The
+\texttt{wait\_inbetween} specification, which models the non exact
+delay between 50.3 ms and 200.3 ms is not supported in the current
+language and cannot be compiled with the current SystemJ compiler. We
+introduce such \texttt{wait\_inbetween} statements in SystemJ. Such
+non exact real-time mechanisms are useful in developing systems where
+the real-time delay is not known or should not be known a priori.
 
@@
 \subsubsection{Programming periodic tasks using the exact real-time
-  construct; the \texttt{wait\_exactly} statement}
+  construct; the \texttt{wait\_exact} statement}
 \label{sec:progr-using-exact}
 
-The previous wait statements are non exact, i.e., the execution of the
-statement following the wait statement cannot be controlled exactly. The
-final variant we introduce allows doing exactly that. We call it
-\texttt{wait\_exactly} and this statement can be used to program exact
-timeouts, periodic tasks, etc.
+The previous wait statements are non exact, i.e., the \hj{time of}
+execution of the statement following the wait statement cannot be
+controlled exactly. The final variant we introduce allows doing exactly
+that. We call it \texttt{wait\_exact} and this statement can be used to
+program exact timeouts, periodic tasks, etc.
 
@@
         \begin{SubFloat}{\label{pp:a}Periodically emitting signal S}
-        \begin{lstlisting}[style=sysj,morekeywords={emit,trap,pause,exit,wait_exactly}]
+        \begin{lstlisting}[style=sysj,morekeywords={emit,trap,pause,exit,wait_exact}]
 while(true) { 
- wait_exactly (1 ms); 
+ wait_exact (1 ms); 
  emit S; 
@@
 \end{SubFloat}
+
 \begin{SubFloat}{\label{p1:b}Timing diagram -- the X-axis shows the
-    execution of the \texttt{wait\_exactly} statement. The Y-axis is the
+    execution of the \texttt{wait\_exact} statement. The Y-axis is the
     time elapsed.}
@@
 \end{SubFloat}
-  \caption{Programming a periodic task with wait\_exactly}
+  \caption{Programming a periodic task with wait\_exact}
   \label{fig:p1}
@@
 shown in Figure~\ref{fig:p1}. As before upon execution of
-\texttt{wait\_exactly (1 ms)} statement, the time starts elapsing. But,
+\texttt{wait\_exact (1 ms)} statement, the time starts elapsing. But,
 unlike in previous cases the following statement needs to be executed

### 2014-01-28 motivation.tex
Updated 3

@@
 if the user can touch this green light then she is successful and \hj{an
-integer \texttt{1} is sent to other process in order to increase an
+	integer \texttt{1} is sent to another process in order to increase an
 	integer counter, if unsuccessful a \texttt{0} is sent instead to

### 2014-01-28 exps.tex
updated the graph

@@
 		]
-			\addplot + [postaction={pattern=north east lines}] coordinates
+			\addplot + [fill=white,postaction={pattern=north east lines}] coordinates
 			{(HRTCS,1)(Printer/CD1,1)(Printer/CD2,1)(AECS/CD2,1.94)};
-			\addplot + [postaction={pattern=north west
+			\addplot + [fill=white,postaction={pattern=north west
                           lines}] coordinates 
@@
 		]
-			\addplot + [postaction={pattern=north east lines}] coordinates
+			\addplot + [fill=white,postaction={pattern=north east lines}] coordinates
                         {(Motor, 2.60)};
-			\addplot + [postaction={pattern=north west lines}] coordinates
+			\addplot + [fill=white,postaction={pattern=north west lines}] coordinates
                         {(Motor, 2.63)};
-			\addplot + [postaction={pattern=dots}] coordinates
+			\addplot + [fill=white,postaction={pattern=dots}] coordinates
                         {(Motor, 4.39)};
-			\addplot + [postaction={pattern=horizontal lines}] coordinates
+			\addplot + [fill=white,postaction={pattern=horizontal lines}] coordinates
                         {(Motor, 2.92)};
-			\addplot + [postaction={pattern=crosshatch}] coordinates
+			\addplot + [fill=white,postaction={pattern=crosshatch}] coordinates
                         {(Motor, 2.69)};

### 2014-01-28 delays.tex

@@
 (cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
-the CD at periodically at the WCRT, we can choose the value of
+the CD periodically at the WCRT, we can choose the value of
 \textit{d} from set $S1$. This technique allows meeting the synchronous
@@
 high-level technique of introducing real-time waits in the SystemJ
-language is shown in Figure~\ref{fig:4}. We use static-analysis
+language is shown in Figure~\ref{fig:4}. We use static analysis
 techniques to find out the WCRT/BCRT of the programs automatically after
@@
 and~\ref{alg:2} are performed after scheduling for different constraints
-such as faster execution time, or reduced power consumption, we,
+such as faster execution time, or reduced power consumption, we
 inherently satisfy the criteria of exploiting non-maximal

### 2014-01-28 discussion.tex

@@
 
-\subsection{Programming using the wait\_exactly construct -- the
+\subsection{Programming using the wait\_exact construct -- the
   periodic reactions}
@@
 	\vspace{-10pt}
-        \begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exactly}]
-          while(true){wait_exactly (1 ms); emit S; //do something
+        \begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
+          while(true){wait_exact (1 ms); emit S; //do something
           }
  \end{lstlisting}
-  \caption{Programming periodic tasks with wait\_exactly statement}
+  \caption{Programming periodic tasks with wait\_exact statement}
   \label{fig:periodic}
@@
     \centering
-			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exactly}]
+			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
 while(true) {
- wait_exactly(1 ms);
+ wait_exact(1 ms);
  //extra pause

### 2014-01-28 introduction.tex

@@
 (postponement) of an operation for 0.2 ms. These languages are based on
-formal semantics, essential for the formal reasoning about and
-verification of correctness of functional properties of the developed
-programs, but leave nonfunctional properties such as timing behavior as
-an implementation detail~\cite{boldt07}. Arguably, rightly so, because
-physical time cannot be incorporated without information about the
-underlying execution platform.  But, one can still reason in terms of
-logical time. These languages support discrete logical clock rather than
-a discrete physical (real-time) clock. The real-time period of the
-discrete logical clock (also referred to as a logical tick) is not fixed
-and it is determined by the responsiveness of the program to the input
-signals. Unlike a discrete physical clock, which has fixed real-time
-period, the period of the logical clock is elastic. Although the timing
-model with logical clock has worked well for reactive synchronous/GALS
-languages in designing discrete systems, there is a need to introduce
-real-time, since often, implementation models are in the continuous or
-discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
-Moreover, adding timing capabilities as first class language constructs
-create additional opportunity for formally verifying functional and
-non-functional requirements before system deployment.
+formal semantics, essential for the formal reasoning about \hj{code
+generation} and verification of correctness of functional properties of
+the developed programs, but leave nonfunctional properties such as
+timing behavior as an implementation detail~\cite{boldt07}. Arguably,
+rightly so, because physical time cannot be incorporated without
+information about the underlying execution platform.  But, one can still
+reason in terms of logical time. These languages support discrete
+logical clock rather than a discrete physical (real-time) clock. The
+real-time period of the discrete logical clock (also referred to as a
+logical tick) is not fixed and it is determined by the responsiveness of
+the program to the input signals. Unlike a discrete physical clock,
+which has fixed real-time period, the period of the logical clock is
+elastic. Although the timing model with logical clock has worked well
+for reactive synchronous/GALS languages in designing discrete systems,
+there is a need to introduce real-time since often, implementation
+models are in the continuous or discrete real-time
+domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
+capabilities as first class language constructs create additional
+opportunity for formally verifying functional and non-functional
+requirements before system deployment.
 
@@
 mechanisms and at the same time \textbf{contributions} of the paper are:
-(1) these waiting mechanisms allow real-time based synchronization
-between concurrent behaviors, (2) the waiting model relies on the use of
-relative instead of absolute real-time, i.e., a wait in selected time
-units is counted from the currently executing SystemJ statement, (3)
-these waiting mechanisms allow mixing behaviors with real-time features
-and others with only logical time and (4) real-time wait does not affect
-the model of logical time in SystemJ program as long as the amount of
-delay is within boundaries that can be determined statically by program
-analysis. All these as well as semantics of the wait constructs are
-discussed and illustrated in the rest of the paper. Our contributions
-can be further refined as follows:
+\hj{(1) real-time wait does not affect the model of logical time in
+	SystemJ program as long as the amount of delay is within boundaries
+	that can be determined statically by program analysis,} (2) the
+	waiting model relies on the use of relative instead of absolute
+	real-time, i.e., a wait in selected time units is counted from the
+	currently executing SystemJ statement and (3) these waiting mechanisms
+	allow mixing behaviors with real-time features and others with only
+	logical time. All these as well as semantics of the wait constructs
+	are discussed and illustrated in the rest of the paper. Our
+	contributions can be further refined as follows:
+
 \begin{enumerate*}
@@
   both; exact and non-exact real-time in GALS and synchronous programs.
-\item We elucidate upon a key insight relating real-time and
+\item We \hj{clarify} upon a key insight relating real-time and
   logical-time -- \textit{the resolution of any real-time statement is

### 2014-01-28 background.tex
Updated things

@@
 The perfect synchrony hypothesis is ideal for programming the
-synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
-requires real-time to compute. Formally, execution of every micro-step
-in the logical zero time requires $\delta$ physical time. Time for a
-logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
-where $\mathcal{F}$ is some function dependent upon the
-\underline{schedule}. We call $\Delta$ the reaction time. Depending upon
-the amount of computation required and schedule, $\Delta$ can vary,
-hence, in order to satisfy the implicit restriction imposed by the
-synchrony hypothesis -- no input event \textit{from the environment} can
-be missed -- one needs to calculate the \textit{Worst Case Reaction
-  Time} (WCRT) and the resultant WCRT needs to be smaller than the time
-between any two consecutive events on any input, else the synchrony
-hypothesis is violated. Techniques for calculating the WCRT of
-synchronous programs~\cite{boldt07} exist and this analysis is not the
-focus of this paper. The opposite of the WCRT, the shortest time for a
-reaction, is called the \textit{Best Case Reaction Time} (BCRT), we
-denote the WCRT and BCRT in Figure~\ref{fig:2c} using \texttt{W} and
+synchronous sub-set of SystemJ. Unfortunately, execution of every
+statement in the logical zero time requires $\delta$ physical time. Time
+for a logical tick can thus be summarized as $\Delta = \mathcal{F}
+(\delta)$, where $\mathcal{F}$ is the schedule. We call $\Delta$ the
+reaction time. Depending upon the amount of computation required and
+schedule, $\Delta$ can vary. In order to satisfy the implicit
+restriction imposed by the synchrony hypothesis -- no input event
+\textit{from the environment} can be missed -- one needs to calculate
+the \textit{Worst Case Reaction Time} (WCRT) and the resultant WCRT
+needs to be smaller than the time between any two consecutive events on
+any input, else the synchrony hypothesis is violated. Techniques for
+calculating the WCRT of synchronous programs~\cite{boldt07} exist and
+this analysis is not the focus of this paper. The opposite of the WCRT,
+the shortest time for a reaction, is called the \textit{Best Case
+  Reaction Time} (BCRT). Formally, let $\{\Delta_1, \Delta_2,\ldots,
+\Delta_i\}$, be the set of all possible reaction times for some
+CD. Then, $WCRT= max (\Delta_i)$ and $BCRT = min (\Delta_i)$.
+
+We denote the WCRT and BCRT in Figure~\ref{fig:2c} using \texttt{W} and
 \texttt{B}, respectively. Figure~\ref{fig:2c} shows a continuously

### 2014-01-28 correct.tex
Updated correct.tex
Updated things

--- 2014-01-27 correct.tex
@@
   value \texttt{d}, provided $D$ is a non empty set, such that for any
-  given $BCRT \leq T \leq WCRT$, $M \leq T \times d \leq N$ holds.
+  given reaction time $\hj{\Delta}$ \hj{in the range of} $BCRT \leq
+  \Delta \leq WCRT$, $M \leq \Delta \times d \leq N$ holds.
 \label{th:1}
@@
   \begin{compactenum}[\hspace{0.25cm} 1.]
-  \item Proof for case $T = WCRT$: Assume \mbox{$d \times WCRT < M$},
-    then $d < \frac{M}{WCRT}$. $BCRT \leq WCRT$, by definition, hence,
-    $\frac{M}{WCRT} \leq \frac{M}{BCRT}$. Thus for the assumption to
-    hold, $d < \frac{M}{BCRT}$, which contradicts
+  \item Proof for case $\Delta = WCRT$: Assume \mbox{$d \times WCRT <
+      M$}, then $d < \frac{M}{WCRT}$. $BCRT \leq WCRT$, by definition,
+    hence, $\frac{M}{WCRT} \leq \frac{M}{BCRT}$. Thus for the assumption
+    to hold, $d < \frac{M}{BCRT}$, which contradicts
     Lemma~\ref{lemma:1}. The assumption $d \times WCRT >N$ is trivially
     proven from Lemma~\ref{lemma:1}.
-  \item Proof for case $T = BCRT$: Like above, the assumption $d \times
-    BCRT < M$ is trivially proven from Lemma~\ref{lemma:1}. For $d
-    \times BCRT > N$ to hold, $d > \frac{N}{BCRT}$ should hold. But,
+  \item Proof for case $\Delta = BCRT$: Like above, the assumption $d
+    \times BCRT < M$ is trivially proven from Lemma~\ref{lemma:1}. For
+    $d \times BCRT > N$ to hold, $d > \frac{N}{BCRT}$ should hold. But,
     $BCRT \leq WCRT$, by definition, hence, $\frac{N}{BCRT} \geq
     \frac{N}{WCRT}$, which again contradicts Lemma~\ref{lemma:1}.
-  \item Proof for case $BCRT < T < WCRT$: For $d \times T < M$ to hold,
-    $d < \frac{M}{T}$ should hold. By definition, $T > BCRT$, thus,
-    $\frac{M}{BCRT} < \frac{M}{T}$, hence, $d < {M}{BCRT}$ should hold,
-    which contradicts Lemma~\ref{lemma:1}. We can prove the other case
-    $d \times T \leq N$, using the upper bound from Lemma~\ref{lemma:1}
-    similarly.
+  \item Proof for case $BCRT < \Delta < WCRT$: For $d \times \Delta < M$
+    to hold, $d < \frac{M}{\Delta}$ should hold. By definition, $\Delta
+    > BCRT$, thus, $\frac{M}{BCRT} < \frac{M}{\Delta}$, hence, $d <
+    {M}{BCRT}$ should hold, which contradicts Lemma~\ref{lemma:1}. We
+    can prove the other case $d \times \Delta \leq N$, using the upper
+    bound from Lemma~\ref{lemma:1} similarly.
   \end{compactenum}
@@
   value \texttt{d}, provided $D$ is a non empty set, such that for
-  \textbf{all} given $BCRT \leq T \leq WCRT$, $M \leq T \times d \leq N$
-  holds.
+  \textbf{all} given reaction times $\Delta$ in the range of $BCRT \leq
+  \Delta \leq WCRT$, $M \leq \Delta \times d \leq N$ holds.
 \end{theorem}
@@
   \textbf{does not} give a value \texttt{d}, such that for \textbf{all}
-  given $BCRT \geq T \geq WCRT$, $M \leq T \times d \leq N$ holds.
+  given reaction times $T$ in the range of $BCRT > T > WCRT$, $M \leq T
+  \times d \leq N$ holds.
 \end{theorem}

### 2014-01-28 introduction.tex
Updated things

@@
 (postponement) of an operation for 0.2 ms. These languages are based on
-formal semantics, essential for the formal reasoning about \hj{code
-generation} and verification of correctness of functional properties of
-the developed programs, but leave nonfunctional properties such as
-timing behavior as an implementation detail~\cite{boldt07}. Arguably,
-rightly so, because physical time cannot be incorporated without
-information about the underlying execution platform.  But, one can still
-reason in terms of logical time. These languages support discrete
-logical clock rather than a discrete physical (real-time) clock. The
-real-time period of the discrete logical clock (also referred to as a
-logical tick) is not fixed and it is determined by the responsiveness of
-the program to the input signals. Unlike a discrete physical clock,
-which has fixed real-time period, the period of the logical clock is
-elastic. Although the timing model with logical clock has worked well
-for reactive synchronous/GALS languages in designing discrete systems,
-there is a need to introduce real-time since often, implementation
-models are in the continuous or discrete real-time
+formal semantics, essential for correct by construction code generation
+and verification of correctness of functional properties of the
+developed programs, but leave nonfunctional properties such as timing
+behavior as an implementation detail~\cite{boldt07}. Arguably, rightly
+so, because physical time cannot be incorporated without information
+about the underlying execution platform.  But, one can still reason in
+terms of logical time. These languages support discrete logical clock
+rather than a discrete physical (real-time) clock. The real-time period
+of the discrete logical clock (also referred to as a logical tick) is
+not fixed and it is determined by the responsiveness of the program to
+the input signals. Unlike a discrete physical clock, which has fixed
+real-time period, the period of the logical clock is elastic. Although
+the timing model with logical clock has worked well for reactive
+synchronous/GALS languages in designing discrete systems, there is a
+need to introduce real-time since often, implementation models are in
+the continuous or discrete real-time
 domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
@@
   SystemJ program as long as the amount of delay is within boundaries
-	that can be determined statically by program analysis,} (2) the
-	waiting model relies on the use of relative instead of absolute
-	real-time, i.e., a wait in selected time units is counted from the
-	currently executing SystemJ statement and (3) these waiting mechanisms
-	allow mixing behaviors with real-time features and others with only
-	logical time. All these as well as semantics of the wait constructs
-	are discussed and illustrated in the rest of the paper. Our
-	contributions can be further refined as follows:
+  that can be determined statically by program analysis, assuming that
+  real-time analyzable platforms are available,} (2) the waiting model
+relies on the use of relative instead of absolute real-time, i.e., a
+wait in selected time units is counted from the currently executing
+SystemJ statement and (3) these waiting mechanisms allow mixing
+behaviors with real-time features and others with only logical time. All
+these as well as semantics of the wait constructs are discussed and
+illustrated in the rest of the paper. Our contributions can be further
+refined as follows:
 
@@
  i.e., non-negative rational numbers.
-\item We consider the property of \textit{non
-    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
-  not always sufficient resources for processes to execute, so
-  effectively processes are allocated and scheduled with different
-  optimization criteria -- reducing computation time, reducing power
-  consumption, etc. This allocation and scheduling is an integral part
-  of any language compilation and needs to be considered when
-  introducing real-time waits.
+% \item We consider the property of \textit{non
+%     maximal-parallelism}. \textit{Non-maximal parallelism:} There are
+%   not always sufficient resources for processes to execute, so
+%   effectively processes are allocated and scheduled with different
+%   optimization criteria -- reducing computation time, reducing power
+%   consumption, etc. This allocation and scheduling is an integral part
+%   of any language compilation and needs to be considered when
+%   introducing real-time waits.
 \item We are the first, to our knowledge, to allow specification of
   both; exact and non-exact real-time in GALS and synchronous programs.
-\item We \hj{clarify} upon a key insight relating real-time and
+\item We shed light upon a key insight relating real-time and
   logical-time -- \textit{the resolution of any real-time statement is

### 2014-01-28 motivation.tex
Updated things

@@
   } do {send S(0);} // Else Send 0
-  wait_exact(1 s); // Wait for 1 second
+  pause; pause; pause; // pass some time before restarting
  }
@@
 \end{SubFloat}
-\begin{SubFloat}{\label{delaycode:b}Timing diagram -- the X-axis shows
+\begin{SubFloat}{\label{d:b}Timing diagram -- the X-axis shows
     the execution of the \texttt{wait\_inbetween} statement. The Y-axis
@@
 \end{figure}
+
+The \texttt{wait\_inbetween (M..N)} statement postpones the program
+control flow to move to the proceeding statement for a minimum of $M$
+time units and a maximum of $N$ time units, as shown in
+Figure~\ref{d:b}.
 
@@
 \texttt{wait\_inbetween} specification, which models the non exact
-delay between 50.3 ms and 200.3 ms is not supported in the current
-language and cannot be compiled with the current SystemJ compiler. We
-introduce such \texttt{wait\_inbetween} statements in SystemJ. Such
-non exact real-time mechanisms are useful in developing systems where
-the real-time delay is not known or should not be known a priori.
+postponement between 50.3 ms and 200.3 ms is not supported in the
+current language and cannot be compiled with the current SystemJ
+compiler. We introduce such \texttt{wait\_inbetween} statements in
+SystemJ. Such non exact real-time mechanisms are useful in developing
+systems where the real-time postponement is not known or should not be
+known a priori.
 
@@
 
+The \texttt{wait\_atleast (M)} statement is used to postpone the program
+control flow from proceeding to the next statment for a minimum of $M$
+time units. The maximum postponement is unbounded, but countable
+(Figure~\ref{dd:b}).
+
 There are many instances when one would like to wait on an input from
@@
 bound. The statement following the \texttt{wait\_atleast} statement is
-executed after a finite countable, but unbounded delay. Such a statement
-is useful for programming low priority tasks.
+executed after a finite countable, but unbounded postponement. Such a
+statement is useful for programming low priority tasks.
 
@@
 %%% End:
+
+%  LocalWords:  statment

### 2014-01-28 discussion.tex
Updated

@@
 \begin{figure}[h!]
-  \centering
 	\vspace{-10pt}
-        \begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
-          while(true){wait_exact (1 ms); emit S; //do something
+        \begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
+while(true){wait_exact (1 ms); 
+	emit S; //do something
 }
@@
     \centering
-			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
+			\begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
 while(true) {
@@
   \begin{SubFloat}{\label{pp:b}Rewritten program}
-			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay}]
+			\begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,delay}]
 while(true) {
@@
 	\vspace{-10pt}
-		\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay,suspend}]
+		\begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,delay,suspend}]
   suspend(S) {
@@
 	\vspace{-10pt}
-		\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay,suspend}]
+		\begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,delay,suspend}]
 {

### 2014-01-28 exps.tex
Updated

--- 2014-01-28 exps.tex
@@
 
-\hj{
-HRTCS is the human response system described in
-Section~\ref{sec:motivating-example}. }
+\hj{ HRTCS is the human response system described in
+	Section~\ref{sec:motivating-example}. HRTCS consists of 2 CDs, but
+only one with real-time waits.}
 
-\hj{
-Motor is a stepper motor
-controller from~\cite{Bourke2009a}, used for producing monochrome images
-on paper by correct application of current to a print head (a row of
-resistors). In this example, duration of the current applied is
-controlled by the feedback logic using real-time waits, which are
-expressed using reactive constructs. }
+\hj{ Motor is a stepper motor controller from~\cite{Bourke2009a}, used
+for producing monochrome images on paper by correct application of
+current to a print head (a row of resistors). In this example, duration
+of the current applied is controlled by the feedback logic using
+real-time waits, which are expressed using reactive constructs.} It is a
+synchronous program with 5 exact wait statements.
 
-\hj{
-Printer is an example
-from~\cite{Schneider:1999:CRT:555233}, which models a printer and a
-spooler in timed CSP. Its operation can be described as follows: as soon
-as the spooler receives a print job it passes it to the printer through
-a channel where actual printing process take place. This example along
-with HRTCS demonstrates the use of non exact wait statements.
-}
+\hj{ Printer is an example from~\cite{Schneider:1999:CRT:555233}, which
+models a printer and a spooler in timed CSP. Its operation can be
+described as follows: as soon as the spooler receives a print job it
+passes it to the printer through a channel where actual printing process
+take place.} Each spooler and printer process is mapped to a separate
+CD. Spooler CD sends a job to the Printer CD between 2.7 to 10.6 ms
+after receiving it from the environment (e.g. a user). On the other
+hand, Printer CD notifies the environment that it starts printing
+between 1.2 to 30.3 ms once it receives the job from the Spooler CD.
+Actual print job is started between 5.6 to 100.2 ms after the
+notification. In this example, bounded wait ensures that both; channel
+communication and print job should occur within the specified amount of
+time. This example along with HRTCS demonstrates the use of non exact
+wait statements.
 
 AECS (Access and Environmental Control System) is an application that
-controls an intelligent room environment described
-in~\cite{aecs_ispa}. The system interacts with two external timers
-through input/output signals in order to measure how long the entrance
-door to the room is being opened to detect intruders.  The program has
-been modified such that the wait constructs are used instead of external
-timers as shown in Figure~\ref{delaycode}. In the original AECS, the
-system needed to trigger an external timer by emitting the output signal
-in order to wait for a specific amount of time (e.g.
+controls an intelligent room environment described in~\cite{aecs_ispa}.
+AECS has 2 CDs; CD1 consist of a single wait statement, while CD2 has 2
+wait statements. The system is designed to measure how long the entrance
+door to the room is being opened or to detect unknown entries (i.e.
+intruders). For instance, the system triggers an alarm 10 seconds after
+detection of an intruder or generates a beep sound when an entrance door
+is being opened for the same amount of time. In the original AECS, the
+system needed to interacts with two external timers through input/output
+signals in order to wait for a specific amount of time (e.g.
 \texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
 expires, it notifies the system through the input signal
-\texttt{TimeOut}, which is polled using the \texttt{await}
-statement. With the new wait constructs, these timers are not necessary
-and waits can be expressed directly in the code as shown in
-Figure~\ref{delaycode:b}.
+\texttt{TimeOut}, which is polled using the \texttt{await} statement.
+The program has been modified such that the wait constructs replace
+external timers and corresponding interface (I/O) signals, which is
+shown in Figure~\ref{delaycode:b}.  With the new wait constructs, these
+timers are not necessary and waits can be expressed directly in the
+code.
 
@@
 %their volumes.
-
-HRTCS consists of 2 CDs, but only one with real-time waits. AECS also
-has 2 CDs; CD1 consist of a single wait statement, while CD2 has 2 wait
-statements. In this case, as explained previously, wait statements
-replace the external timers and corresponding interface (I/O)
-signals. The system triggers an alarm 10 seconds after detection of an
-intruder or generates a beep sound when an entrance door is being opened
-for the same amount of time. In the Printer example, spooler and printer
-processes are mapped to separate CDs. Spooler CD sends a job to the
-Printer CD between 2.7 to 10.6 ms after receiving it from the
-environment (e.g. a user). On the other hand, Printer CD notifies the
-environment that it starts printing between 1.2 to 30.3 ms once it
-receives the job from the Spooler CD. Actual print job is started
-between 5.6 to 100.2 ms after the notification. In this example, bounded
-wait ensures that both; channel communication and print job should occur
-within the specified amount of time. Lastly, Motor is a synchronous
-program with 5 exact wait statements that control the amount of current
-applied to the stepper motor coils.
 
@@
 		]
-			\addplot + [postaction={pattern=north east lines}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=north east lines}] coordinates
 			{(HRTCS,1)(Printer/CD1,1)(Printer/CD2,1)(AECS/CD2,1.94)};
-			\addplot + [postaction={pattern=north west
+			\addplot + [black,fill=white,postaction={pattern=north west
                           lines}] coordinates 
@@
 		]
-			\addplot + [postaction={pattern=north east lines}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=north east lines}] coordinates
                         {(Motor, 2.60)};
-			\addplot + [postaction={pattern=north west lines}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=north west lines}] coordinates
                         {(Motor, 2.63)};
-			\addplot + [postaction={pattern=dots}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=dots}] coordinates
                         {(Motor, 4.39)};
-			\addplot + [postaction={pattern=horizontal lines}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=horizontal lines}] coordinates
                         {(Motor, 2.92)};
-			\addplot + [postaction={pattern=crosshatch}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=crosshatch}] coordinates
                         {(Motor, 2.69)};
@@
 		]
-			\addplot + [fill=white,postaction={pattern=north east lines}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=north east lines}] coordinates
                         {(Motor, 8.05) (AECS/CD2, 11.82)};
-			\addplot + [fill=white,postaction={pattern=north west lines}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=north west lines}] coordinates
                         {(Motor, 8.00) (AECS/CD2, 11.82)};
-			\addplot + [fill=white,postaction={pattern=dots}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=dots}] coordinates
                         {(Motor, 6.42)};
-			\addplot + [fill=white,postaction={pattern=horizontal lines}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=horizontal lines}] coordinates
                         {(Motor, 7.53)};
-			\addplot + [fill=white,postaction={pattern=crosshatch}] coordinates
+			\addplot + [black,fill=white,postaction={pattern=crosshatch}] coordinates
                         {(Motor, 7.89)};

### 2014-01-29 correct.tex
Updated

@@
     > BCRT$, thus, $\frac{M}{BCRT} < \frac{M}{\Delta}$, hence, $d <
-    {M}{BCRT}$ should hold, which contradicts Lemma~\ref{lemma:1}. We
+    \frac{M}{BCRT}$ should hold, which contradicts Lemma~\ref{lemma:1}. We
     can prove the other case $d \times \Delta \leq N$, using the upper
@@
   \textbf{does not} give a value \texttt{d}, such that for \textbf{all}
-  given reaction times $T$ in the range of $BCRT > T > WCRT$, $M \leq T
-  \times d \leq N$ holds.
+  given reaction times $\Delta$ in the range of $BCRT > \Delta > WCRT$,
+  $M \leq \Delta \times d \leq N$ holds.
 \end{theorem}
@@
 The proofs for correctness, completeness, and soundness for
-Algorithm~\ref{alg:2} are similar to those of Algorithm~\ref{alg:1}.
+Algorithm~\ref{alg:2} are similar to those of
+Algorithm~\ref{alg:1}. Furthermore, the proofs for \texttt{wait\_atleast
+  (M)} and \texttt{wait\_exact (M)} can be proven using the same
+techniques as shown in Theorem~\ref{th:1}.
 

### 2014-01-29 delays.tex
Updated things
Updated

@@
 aforementioned calculation. The patterned space is the only overlapping
-region and hence gives the solution space for the above example.
+region and hence gives the solution space for the above example. It is
+possible that $D$ is an empty set, thereby, no solution can be found. We
+provide solutions for such cases in the next section.
 
@@
 analyzable platforms, in our case \textit{Java Optimized Processor}
-(JOP)~\cite{jop:jnl:jsa2007}. Moreover, since algorithms~\ref{alg:1}
-and~\ref{alg:2} are performed after scheduling for different constraints
-such as faster execution time, or reduced power consumption, we
-inherently satisfy the criteria of exploiting non-maximal
-parallelism. The low-level analysis guarantees a precise WCRT/BCRT for
-the logical tick, which is then used for calculating $d$ in
-Figure~\ref{fig:3}.
+(JOP)~\cite{jop:jnl:jsa2007}. It is worth pointing out that since all
+rewrites, including the rewrite for the wait constructs happen before
+WCRT/BCRT analysis, this analysis incorporates the affects of the
+introduction of the wait statements. Moreover, since
+algorithms~\ref{alg:1} and~\ref{alg:2} are performed after scheduling
+for different constraints such as faster execution time, or reduced
+power consumption, we inherently satisfy the criteria of exploiting
+non-maximal parallelism. The low-level analysis guarantees a precise
+WCRT/BCRT for the logical tick, which is then used for calculating $d$
+in Figure~\ref{fig:3}.
 
@@
 decides to execute the CD periodically, where the period is the
-WCRT. WCRT effects the over-approximation or the period of the CD,
+WCRT. WCRT effects the over-approximation and the period of the CD,
 irrespective of the solution that the system designer chooses. A large

### 2014-01-29 motivation.tex
Updated

@@
 shown in Figure~\ref{fig:1}. We use the SystemJ language in this paper,
-because being a \textit{Globally Asynchronous Locally
-  Synchronous} (GALS) language it helps us easily describe a larger
-class of systems (like HRTCS in Figure~\ref{fig:1}) compared to purely
-synchronous languages like Esterel. Moreover, being a super-set of
-Esterel, SystemJ allows us to explore the marriage of real-time and
-logical time, which is equally applicable to the synchronous sub-set.
+because being a \textit{Globally Asynchronous Locally Synchronous}
+(GALS) language it helps us easily describe a larger class of systems
+(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
+languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
+allows us to explore the marriage of real-time and logical time not only
+in the GALS, but also in the synchronous setting.
 

### 2014-01-31 background.tex
Updated
Updated the paper

@@
 signals and with each other via channels and CSP~\cite{choa85} style
-rendezvous. Each CD has its own notion of logical time. CDs adhere to
+rendezvous. Each CD has its own notion of logical time. CDs are composed
+of one or more units called reactions. These reactions themselves can be
+composed of children reaction forming a hierarchy. Each CD adheres to
 the \textit{perfect synchrony} hypothesis, i.e., all statements execute
@@
 The perfect synchrony hypothesis is ideal for programming the
-synchronous sub-set of SystemJ. Unfortunately, execution of every
+synchronous subset of SystemJ. Unfortunately, execution of every
 statement in the logical zero time requires $\delta$ physical time. Time
@@
 calculating the WCRT of synchronous programs~\cite{boldt07} exist and
-this analysis is not the focus of this paper. The opposite of the WCRT,
+WCRT analysis is not the focus of this paper. The opposite of the WCRT,
 the shortest time for a reaction, is called the \textit{Best Case
   Reaction Time} (BCRT). Formally, let $\{\Delta_1, \Delta_2,\ldots,
-\Delta_i\}$, be the set of all possible reaction times for some
-CD. Then, $WCRT= max (\Delta_i)$ and $BCRT = min (\Delta_i)$.
+\Delta_N\}$, be the set of all possible reaction times for some
+CD. Then, $\exists i \in N, \mathrm{where\ } WCRT= max (\Delta_i)$,
+similarly, $\exists j \in N$, where $BCRT = min (\Delta_j)$.
 

### 2014-01-31 correct.tex
Updated the paper

@@
 \begin{proof}
-  From Theorem~\ref{th:1} case 3.
+  Follows from Theorem~\ref{th:1}.
 \end{proof}
@@
 Algorithm~\ref{alg:2} are similar to those of
-Algorithm~\ref{alg:1}. Furthermore, the proofs for \texttt{wait\_atleast
-  (M)} and \texttt{wait\_exact (M)} can be proven using the same
-techniques as shown in Theorem~\ref{th:1}.
+Algorithm~\ref{alg:1}. Furthermore, even correctness, completeness, and
+soundness for \texttt{wait\_atleast (M)} and \texttt{wait\_exact (M)}
+can be proven using the same techniques as shown in Theorem~\ref{th:1}.
 

### 2014-01-31 delays.tex
Updated the paper

@@
 \in  \mathbb{Q}^{>0}$  and $N  \in  \mathbb{Q}^{>0}$,  statement $p$  is
-executed after real-time delay of $\tau$ units, such that, $M \leq \tau
-\leq N$.
+executed after  real-time postponement  of $\tau$  units, such  that, $M
+\leq \tau \leq N$.
 
@@
 \item Given a SystemJ program \texttt{wait\_atleast (M); p}, where $M
-  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
-  of $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
-  note that the lower bound of the wait construct; $M$ is \textit{not}
-  an exact delay, but rather the control is allowed to proceed to the
-  next statement anytime after waiting for \textit{at least} $M$ times
-  units.
-\item Given a SystemJ program \texttt{wait\_exact (M); p}, where $M
-  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
-  of $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
-  waiting time is exact.
+  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time
+  postponement of $\tau$ units, such that, $M \leq \tau < \infty$. It is
+  important to note that the lower bound of the wait construct; $M$ is
+  \textit{not} an exact postponement, but rather the control is allowed
+  to proceed to the next statement anytime after waiting for \textit{at
+    least} $M$ times units.
+\item Given a SystemJ program \texttt{wait\_exact (M); p}, where $M \in
+  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time
+  postponement of $\tau$ units, such that, $M \leq \tau \leq M$. In this
+  variant, the waiting time is exact.
 \end{enumerate*}
@@
 The fundamental observation is that real-time is converted into logical
-time via the \texttt{pause} construct. The rewrite basically maps the
-physical notion of time back to the elegant logical notion of time. The
-rewrite \textit{waits} a certain number of logical ticks, before
-proceeding to the next statement. The number of logical ticks \texttt{d}
-to \textit{wait} (Figure~\ref{fig:3}) is determined by the compiler
+time via the \texttt{pause} construct. The rewrite basically maps
+real-time back to the elegant logical notion of time. The rewrite
+\textit{waits} a certain number of logical ticks, before proceeding to
+the next statement. The number of logical ticks \texttt{d} to
+\textit{wait} (Figure~\ref{fig:3}) is determined by the compiler
 statically at compile time. The value of \texttt{d} is intricately tied
@@
 region and hence gives the solution space for the above example. It is
-possible that $D$ is an empty set, thereby, no solution can be found. We
-provide solutions for such cases in the next section.
+possible that $D$ is an empty set, that means no overlapping region
+could be found. We provide solutions for such cases in the next
+subsection.
 
@@
 \texttt{wait\_atleast(M)}). Note that $S_1$ and $S_2$ can never be empty
-sets and will always contain at minimum a single element. For such cases
-we provide two solutions listed below.
+sets and will always contain at minimum a single element. For such
+cases, we provide two solutions listed below.
 
@@
 the exact real-time wait of $M$ time units in case of
-\texttt{wait\_exact(M)} statement, thereby violating its
+\texttt{wait\_exact(M)} statement, or is unable to provide a
+postponement in the region of $[M, N]$ time units in the case of
+\texttt{wait\_inbetween (M..N)} statement, thereby violating their
 semantics. The second option is to execute the CD periodically at the
@@
 (cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
-the CD periodically at the WCRT, we can choose the value of
-\textit{d} from set $S1$. This technique allows meeting the synchronous
-hypothesis, while at the same time guaranteeing the semantics of the
-\texttt{wait\_exact(M)} statement. But, compared to the previous
-option, this technique is very inflexible and results in reduced
-response time and increased power consumption.
+the CD periodically at the WCRT, we can choose the value of \textit{d}
+from set $S1$. This technique allows meeting the synchronous hypothesis,
+while at the same time guaranteeing the semantics of the
+\texttt{wait\_exact(M)} statement. But, compared to the previous option,
+this technique is very inflexible and results in reduced response time
+and increased power consumption.
 

### 2014-01-31 discussion.tex
Updated the paper

@@
 
-We dedicate this section to discuss the details of the wait construct
-semantics, their usage and the advantages and limitations of our
-proposed technique to introducing these real-time constructs in a
-reactive setting.
+We dedicate this section to discuss the details of the real-time wait
+construct semantics, its advantages and limitations in a reactive
+setting.
 
@@
 
-\begin{figure}[h!]
-	\vspace{-10pt}
-        \begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
-while(true){wait_exact (1 ms); 
-	emit S; //do something
-}
- \end{lstlisting}
-  \caption{Programming periodic tasks with wait\_exact statement}
-  \label{fig:periodic}
-	\vspace{-10pt}
-\end{figure}
+% \begin{figure}[h!]
+% 	\vspace{-10pt}
+%         \begin{lstlisting}[style=sysj,basicstyle=\footnotesize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
+% while(true){wait_exact (1 ms); 
+% 	emit S; //do something
+% }
+%  \end{lstlisting}
+%   \caption{Programming periodic tasks with wait\_exact statement}
+%   \label{fig:periodic}
+% 	\vspace{-10pt}
+% \end{figure}
 
-The program in Figure~\ref{fig:preemp}, reproduced from
-Section~\ref{sec:progr-using-exact}, is supposed to emit signal
-\texttt{S} every \texttt{1 ms}. Such periodic reactions (or CDs) require
-special consideration, like any other real-time periodic task. One
-should be aware of the interaction of wait and other \texttt{pause}
+% The program in Figure~\ref{fig:preemp}, reproduced from
+% Section~\ref{sec:progr-using-exact}, is supposed to emit signal
+% \texttt{S} every \texttt{1 ms}.
+
+One should be aware of the interaction of wait and other \texttt{pause}
 constructs in the body of the reaction (or CD). For example, consider
-the (perceived) periodic program and its rewrite below.
+the (perceived) periodic code snippet in Figure~\ref{pp:a} and its
+rewrite in Figure~\ref{pp:b}.
 
@@
 The system designer needs to be aware that just introducing a wait
-construct does not make the reaction (or CD) periodic. The program on
-the left does \textit{not} emit signal \texttt{S} every \texttt{1 ms},
-because of the extra \texttt{pause} statement. Introducing additional
-pauses introduces more logical ticks, consequently increasing the
-real-time wait period. Static analysis techniques from real-time
-community can be used to automatically determine the wait specification,
-in such cases, but describing these techniques is out of the scope of
-this work.
+construct does not make the reaction (or CD) periodic. The code snippet
+in Figure~\ref{pp:a} does \textit{not} emit signal \texttt{S} every
+\texttt{1 ms}, because of the extra \texttt{pause}
+statement. Introducing additional pauses introduces more logical ticks,
+consequently increasing the real-time wait period. Static analysis
+techniques from real-time community can be used to automatically
+determine the specification for the \texttt{wait\_exact} statement in
+Figure~\ref{pp:a}, in such cases, but describing these techniques is out
+of the scope of this work.
 
@@
 
-The proposed solution relies on time analyzable platforms, i.e.,
-platforms where WCRT and BCRT can be calculated using static analysis
-tools. Many might consider this to be too much of a restriction. In this
-section we debunk this \textit{perceived} disadvantage.
+The solution for introducing real-time as first class language
+constructs introduced in this paper rely on platforms that allow
+computing the WCRT and the BCRT of the program. Many might consider this
+to be too much of a restriction. In this section we debunk this
+\textit{perceived} disadvantage.
 
@@
 constructs into logical waits (\texttt{pause} constructs), which
-interact well with preemption.  Furthermore, synchronous/GALS languages
-have always been targeted at real-time analyzable platforms and
-constrained
+interact well with preemption.  Furthermore, synchronous languages have
+always been targeted at real-time analyzable platforms and constrained
 environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}. One can use
@@
  //reaction R2
- // do something
+ // do something with heavy computation
 }
@@
 There is yet another problem with timer based systems. Let us consider
-the program in Figure~\ref{fig:resolution}. On a cursory look this
+the code snippet in Figure~\ref{fig:resolution}. On a cursory look this
 example should work fine. An external timer is started in reaction R1,
@@
 (cf. Section~\ref{sec:mapping-logical-time}). In
-Figure~\ref{fig:resolution}, reaction R2 determines the length of the
-tick, because it is performing the heavier computation. Consider what
-happens if the external timer generates the \texttt{TIMER} signal, but
-the system is still performing computation and the logical tick has not
-yet finished -- we \textit{miss} this \texttt{TIMER} signal from the
+Figure~\ref{fig:resolution}, assuming reaction R2 performs some large
+data computation the length of the tick is determined by R2. Consider
+what happens if the external timer generates the \texttt{TIMER} signal,
+but the system is still performing computation and the logical tick has
+not yet finished -- we \textit{miss} this \texttt{TIMER} signal from the
 environment. Thus, even in the absence of the \texttt{suspend}
@@
 external timer signal. In fact, the WCRT needs to be smaller than 1 ms
-for the above program to perform as expected. But, this again implies
-need for a time analyzable platform even when using external timers.
+for the above code snippet to perform as expected. But, this again
+implies need for a time analyzable platform even when using external
+timers.
 
@@
 the proposed technique, or external timer dependent, like above. This
-discovery is one of the key insights for proposing the solutions
+analysis is one of the key insights for proposing the solutions
 described in this paper.

### 2014-01-31 exps.tex
Updated the paper

@@
 	\midrule
-	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0.0396 ms	& 0.0425 ms & 0.0190 ms &
+	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0334 ms	& 0.0425 ms & 0.0190 ms &
 	0.0263 ms & 0.1784 ms & 0.1284 ms\\ 
 	\cmidrule(r){1-8}
-	\multicolumn{2}{c}{\emph{WCRT}} &0.1152 ms & 0.1096 ms & 0.0647 ms
+	\multicolumn{2}{c}{\emph{WCRT}} &0.112 ms & 0.1096 ms & 0.0647 ms
 	& 0.0491 ms & 0.3476 ms & 0.2991 ms\\ 
@@
   Section~\ref{sec:motivating-example}. HRTCS consists of 2 CDs, but
-only one with real-time waits.}
+  only one with real-time wait.}
 
@@
 is being opened for the same amount of time. In the original AECS, the
-system needed to interacts with two external timers through input/output
+system needed to interact with two external timers through input/output
 signals in order to wait for a specific amount of time (e.g.
@@
 shown in Figure~\ref{delaycode:b}.  With the new wait constructs, these
-timers are not necessary and waits can be expressed directly in the
-code.
+timers are not necessary and real-time waits can be expressed directly
+in the SystemJ code.
 
@@
 statements. Figure~\ref{fig:graph1} shows the accuracy of our results by
-comparing differences between programmer specified upper bounds
+comparing between the programmer specified upper bounds
 (Table~\ref{fig:benchlist}) and the actual bounds obtained using our
-techniques (Table~\ref{fig:comparison}) for the other examples requiring
+techniques (Table~\ref{fig:comparison}) for the examples requiring
 relaxation. AECS/CD1 wait specification is not included in
@@
 logical tick increase WCRT of the program.  Assuming we do not change
-the architecture of the execution platform (i.e. JOP), the differece
+the architecture of the execution platform (i.e. JOP), the difference
 factor can only be reduced by modifying the benchmark programs such that

### 2014-01-31 introduction.tex
Updated
Updated the paper

@@
 of the discrete logical clock (also referred to as a logical tick) is
-not fixed and it is determined by the responsiveness of the program to
-the input signals. Unlike a discrete physical clock, which has fixed
-real-time period, the period of the logical clock is elastic. Although
-the timing model with logical clock has worked well for reactive
-synchronous/GALS languages in designing discrete systems, there is a
-need to introduce real-time since often, implementation models are in
-the continuous or discrete real-time
+not fixed and it is determined by the speed of computation of the
+program for different input signals. Unlike a discrete physical clock,
+which has fixed real-time period, the period of the logical clock is
+elastic. Although, the timing model with logical clock has worked well
+for reactive synchronous/GALS languages in designing discrete systems,
+there is a need to introduce real-time since often, implementation
+models are in the continuous or discrete real-time
 domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
@@
 \begin{enumerate*}
-\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
+\item Real-time is specified in the $\mathbb{Q}^{>0}$ domain of numbers,
   i.e., non-negative rational numbers.

### 2014-01-31 motivation.tex
Updated the paper

@@
 \end{SubFloat}
-\begin{SubFloat}{\label{d:b}Timing diagram -- the X-axis shows
-    the execution of the \texttt{wait\_inbetween} statement. The Y-axis
-    is the time elapsed.}
+\begin{SubFloat}{\label{d:b}Timing diagram -- the X-axis shows the
+    execution of the \texttt{wait\_inbetween} statement. The Y-axis is
+    the time elapsed. The arrow shows one example of the expiration of
+    the \mbox{\texttt{wait\_inbetween (50.3..200.3 ms)}} statement.}
   % \begin{figure}[h!]
@@
 The \texttt{wait\_inbetween (M..N)} statement postpones the program
-control flow to move to the proceeding statement for a minimum of $M$
-time units and a maximum of $N$ time units, as shown in
+control flow from moving to the proceeding statement for a minimum of
+$M$ time units and a maximum of $N$ time units, as shown in
 Figure~\ref{d:b}.
@@
 waits for the user to respond. If the user is able to touch the screen
-within the specified time of 50.3 ms to 200.3 ms a positive response is
-sent to the second clock-domain. All the SystemJ programming constructs
-used to implement this system are described in Table~\ref{tab:1} \hj{and
-  more details can be found in~\cite{amal10}}. The
-\texttt{wait\_inbetween} specification, which models the non exact
-postponement between 50.3 ms and 200.3 ms is not supported in the
-current language and cannot be compiled with the current SystemJ
-compiler. We introduce such \texttt{wait\_inbetween} statements in
-SystemJ. Such non exact real-time mechanisms are useful in developing
-systems where the real-time postponement is not known or should not be
-known a priori.
+within the specified time of at least 50.3 ms and at most 200.3 ms, a
+positive response is sent to the second clock-domain. All the SystemJ
+programming constructs used to implement this system are described in
+Table~\ref{tab:1} \hj{and more details can be found
+  in~\cite{amal10}}. The \texttt{wait\_inbetween} specification, which
+models the non exact postponement between 50.3 ms and 200.3 ms is not
+supported in the current language and cannot be compiled with the
+current SystemJ compiler. We introduce such \texttt{wait\_inbetween}
+statements in SystemJ. Such non exact real-time mechanisms are useful
+for developing systems where the real-time postponement is not known or
+should not be known a priori.
 
@@
     execution of the \texttt{wait\_atleast} statement. The Y-axis is the
-    time elapsed.}
+    time elapsed. The arrow shows one possible expiration of
+    \texttt{wait\_atleast (10000 ms)} statement.}
 \includegraphics[scale=0.7]{FF2}
@@
 environment for only a specified amount of time. If the signal is not
-received then a timeout is generated so that the program can make
+received, then a timeout is generated so that the program can make
 progress. Such a timeout programmed in the SystemJ pseudo-code along
@@
     execution of the \texttt{wait\_exact} statement. The Y-axis is the
-    time elapsed.}
+    time elapsed. The arrow shows the \textit{only} possible expiration
+    of the \texttt{wait\_exact (1 ms)} statement.}
   \includegraphics[scale=0.7]{FF3}
@@
 One such example; periodic emission of a signal to the environment is
-shown in Figure~\ref{fig:p1}. As before upon execution of
+shown in Figure~\ref{fig:p1}. As before, upon execution of
 \texttt{wait\_exact (1 ms)} statement, the time starts elapsing. But,
 unlike in previous cases the following statement needs to be executed
-exactly after 1 ms have passed, as shown in Figure~\ref{p1:b}.
+exactly after 1 ms has passed, as shown in Figure~\ref{p1:b}.
 

### 2014-02-02 correct.tex
Paper updates

@@
 \begin{proof}
-  We use case analysis and contradiction to prove the above theorem.
+  We use contradiction to prove the theorem.
   \begin{compactenum}[\hspace{0.25cm} 1.]
@@
     to hold, $d < \frac{M}{\Delta}$ should hold. By definition, $\Delta
-    > BCRT$, thus, $\frac{M}{BCRT} < \frac{M}{\Delta}$, hence, $d <
-    \frac{M}{BCRT}$ should hold, which contradicts Lemma~\ref{lemma:1}. We
-    can prove the other case $d \times \Delta \leq N$, using the upper
-    bound from Lemma~\ref{lemma:1} similarly.
+    > BCRT$, thus, $\frac{M}{\Delta} < \frac{M}{BCRT}$, hence, $d <
+    \frac{M}{BCRT}$ should hold, which contradicts
+    Lemma~\ref{lemma:1}. We can prove the other case $d \times \Delta >
+    N$, using the upper bound from Lemma~\ref{lemma:1} similarly.
   \end{compactenum}
@@
 \begin{proof}
-  The proof is the opposite of case 3 in Theorem~\ref{th:1}.
+  Soundness is trivially proven knowing that there exists no $\Delta <
+  BCRT$ or $\Delta > WCRT$ by definition
+  (cf. Section~\ref{sec:mapping-logical-time}).
 \end{proof}
@@
 Algorithm~\ref{alg:2} are similar to those of
-Algorithm~\ref{alg:1}. Furthermore, even correctness, completeness, and
+Algorithm~\ref{alg:1}. Furthermore, correctness, completeness, and
 soundness for \texttt{wait\_atleast (M)} and \texttt{wait\_exact (M)}
-can be proven using the same techniques as shown in Theorem~\ref{th:1}.
+are dependent on Theorem~\ref{th:1}, because these are variants of the
+\texttt{wait\_inbetween (M..N)} statement.
 

### 2014-02-02 delays.tex
Paper updates

@@
 first divide $M$ and $N$ with the BCRT and WCRT, respectively. This
-division gives us the number of individual logical ticks required to
-postpone the CD (or synchronous program) by the real-time
-specification. We always $ceil$ when dividing $M$ and $floor$ when
-dividing $N$ to make sure that the resultant values are integers (in
-domain $\mathbb{N}^{>0}$) and these functions guarantee that the
-resultant logical ticks result in real-time postponements between the
-required range $[M,N]$. Next, a function $F$ maps these calculated
-values to a set of equidistant integer points (values) separated by a
-unit value -- these points represent all the logical ticks running at
-the WCRT and the BCRT, respectively that satisfy the real-time wait
-requirements. The intersection of these two sets gives all the logical
-ticks that satisfy the real-time requirements invariant of the logical
-time and its elasticity.
+division gives us the number of individual logical ticks, at the fastest
+(BCRT) and slowest (WCRT) program execution speeds, required to postpone
+the CD (or synchronous program) by the real-time specification. We
+always $ceil$ when dividing $M$ and $floor$ when dividing $N$ to make
+sure that the resultant values are integers (in domain
+$\mathbb{N}^{>0}$) and these functions guarantee that the resultant
+logical ticks result in real-time postponements between the required
+range $[M,N]$. Next, a function $F$ maps these calculated values to a
+set of equidistant integer points (values) separated by a unit value --
+these points represent all the logical ticks running at the WCRT and the
+BCRT, respectively that satisfy the real-time wait requirements. The
+intersection of these two sets gives all the logical ticks that satisfy
+the real-time requirements invariant of the logical time and its
+elasticity.
 
@@
 $N$ is 200.3 ms, respectively. \hj{Timing analysis of the program has
-found} WCRT and BCRT be: 0.112 ms and 0.0334 ms, respectively. Thus, the
-algorithm proceeds as follows:
+  found} WCRT and BCRT to be: 0.112 ms and 0.0334 ms,
+respectively. Thus, the algorithm proceeds as follows:
 
@@
 while at the same time guaranteeing the semantics of the
-\texttt{wait\_exact(M)} statement. But, compared to the previous option,
-this technique is very inflexible and results in reduced response time
-and increased power consumption.
+\texttt{wait\_exact(M)} statement. This technique trades-off reduced
+average reaction time and increased power consumption while providing
+exact postponement.
 

### 2014-02-02 exps.tex
Paper updates

@@
 communication and print job should occur within the specified amount of
-time. This example along with HRTCS demonstrates the use of non exact
-wait statements.
+time. % This example along with HRTCS demonstrates the use of non exact
+% wait statements.
 
@@
 controls an intelligent room environment described in~\cite{aecs_ispa}.
-AECS has 2 CDs; CD1 consist of a single wait statement, while CD2 has 2
-wait statements. The system is designed to measure how long the entrance
-door to the room is being opened or to detect unknown entries (i.e.
-intruders). For instance, the system triggers an alarm 10 seconds after
-detection of an intruder or generates a beep sound when an entrance door
-is being opened for the same amount of time. In the original AECS, the
-system needed to interact with two external timers through input/output
-signals in order to wait for a specific amount of time (e.g.
-\texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
-expires, it notifies the system through the input signal
-\texttt{TimeOut}, which is polled using the \texttt{await} statement.
-The program has been modified such that the wait constructs replace
-external timers and corresponding interface (I/O) signals, which is
-shown in Figure~\ref{delaycode:b}.  With the new wait constructs, these
-timers are not necessary and real-time waits can be expressed directly
-in the SystemJ code.
+AECS has 2 CDs; each with one wait statement. The system is designed to
+measure how long the entrance door to the room is being opened or to
+detect unknown entries (i.e.  intruders). For instance, the system
+triggers an alarm 10 seconds after detection of an intruder or generates
+a beep sound when an entrance door is being opened for the same amount
+of time. In the original AECS, the system needed to interact with two
+external timers through input/output signals in order to wait for a
+specific amount of time (e.g.  \texttt{TimerTriggered} in
+Figure~\ref{delaycode:a}). When the timer expires, it notifies the
+system through the input signal \texttt{TimeOut}, which is polled using
+the \texttt{await} statement.  The program has been modified such that
+the wait constructs replace external timers and corresponding interface
+(I/O) signals, which is shown in Figure~\ref{delaycode:b}.  With the new
+wait constructs, these timers are not necessary and real-time waits can
+be expressed directly in the SystemJ code.
 
@@
 (Table~\ref{fig:benchlist}) and the actual bounds obtained using our
-techniques (Table~\ref{fig:comparison}) for the examples requiring
-relaxation. AECS/CD1 wait specification is not included in
-Figure~\ref{fig:graph2}, because although it requires relaxation, the
-wait specification (Table~\ref{fig:benchlist}) does not have an upper
-bound.
+techniques (Table~\ref{fig:comparison}).
+% relaxation. % AECS/CD1 wait specification is not included in
+% Figure~\ref{fig:graph1}, because although it requires relaxation, the
+% wait specification (Table~\ref{fig:benchlist}) does not have an upper
+% bound.
 
@@
 			symbolic x
-			coords={HRTCS,Printer/CD1,Printer/CD2,AECS/CD2},
+			coords={HRTCS,Printer/CD1,Printer/CD2,AECS/CD1,AECS/CD2},
 			height=5cm,
@@
 			\addplot + [black,fill=white,postaction={pattern=north east lines}] coordinates
-			{(HRTCS,1)(Printer/CD1,1)(Printer/CD2,1)(AECS/CD2,1.94)};
+			{(HRTCS,1)(Printer/CD1,1)(Printer/CD2,0)(AECS/CD1,1)(AECS/CD2,0)};
 			\addplot + [black,fill=white,postaction={pattern=north west
@@
 factor can only be reduced by modifying the benchmark programs such that
-the critical paths are broken up using \texttt{pause} constructs,
-thereby reducing the WCRT.  Figure~\ref{fig:improv} shows segments of
-code from AECS whose WCRT is reduced by such technique.
+the critical paths are broken up using additional \texttt{pause}
+constructs, thereby reducing the WCRT.  Figure~\ref{fig:improv} shows
+segments of code from AECS whose WCRT is reduced by such technique.
 
@@
 present(EntryRequest)
+// obtaining signal value
   id = #EntryRequest;
@@
 			\addplot + [black,fill=white,postaction={pattern=north west lines}] coordinates
-                        {(Motor, 8.00) (AECS/CD2, 11.82)};
+                        {(Motor, 8.00)};
 			\addplot + [black,fill=white,postaction={pattern=dots}] coordinates

### 2014-02-02 introduction.tex
Paper updates

--- 2014-01-28 introduction.tex
@@
 formal semantics, essential for correct by construction code generation
-and verification of correctness of functional properties of the
-developed programs, but leave nonfunctional properties such as timing
-behavior as an implementation detail~\cite{boldt07}. Arguably, rightly
-so, because physical time cannot be incorporated without information
-about the underlying execution platform.  But, one can still reason in
-terms of logical time. These languages support discrete logical clock
-rather than a discrete physical (real-time) clock. The real-time period
-of the discrete logical clock (also referred to as a logical tick) is
-not fixed and it is determined by the responsiveness of the program to
-the input signals. Unlike a discrete physical clock, which has fixed
-real-time period, the period of the logical clock is elastic. Although
-the timing model with logical clock has worked well for reactive
-synchronous/GALS languages in designing discrete systems, there is a
-need to introduce real-time since often, implementation models are in
-the continuous or discrete real-time
+and formally verifying functional properties of developed programs, but
+leave nonfunctional properties such as timing behavior as an
+implementation detail~\cite{boldt07}. Arguably, rightly so, because
+physical time cannot be incorporated without information about the
+underlying execution platform.  But, one can still reason in terms of
+logical time. These languages support discrete logical clock rather than
+a discrete physical (real-time) clock. The real-time period of the
+discrete logical clock (also referred to as a logical tick) is not fixed
+and it is determined by the speed of computation of the program for
+different input signals. Unlike a discrete physical clock, which has
+fixed real-time period, the period of the logical clock is
+elastic. Although the timing model with logical clock has worked well
+for reactive synchronous/GALS languages in designing discrete systems,
+there is a need to introduce real-time since often, implementation
+models are in the continuous or discrete real-time
 domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
@@
 relies on the use of relative instead of absolute real-time, i.e., a
-wait in selected time units is counted from the currently executing
-SystemJ statement and (3) these waiting mechanisms allow mixing
-behaviors with real-time features and others with only logical time. All
-these as well as semantics of the wait constructs are discussed and
-illustrated in the rest of the paper. Our contributions can be further
-refined as follows:
+wait in selected time units is counted from the beginning of the wait
+statement and (3) these waiting mechanisms allow mixing behaviors with
+real-time features and others with only logical time. All these as well
+as semantics of the wait constructs are discussed and illustrated in the
+rest of the paper. Our contributions can be further refined as follows:
 
 \begin{enumerate*}
-\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
+\item Real-time is specified in the $\mathbb{Q}^{>0}$ domain of numbers,
   i.e., non-negative rational numbers.
@@
 positions this work in relation to other approaches to introducing
-real-time. Finally, we conclude in Section~\ref{sec:concl-future-work}.
+real-time in other languages. Finally, we conclude in
+Section~\ref{sec:concl-future-work}.
 

### 2014-02-02 motivation.tex
Paper updates

@@
 	\vspace{-10pt}
-        \begin{SubFloat}{\label{delaycode:a}A human responsiveness
+        \begin{SubFloat}{\label{delay:a}A human responsiveness
             system (HRTCS)}
@@
 The \texttt{wait\_inbetween (M..N)} statement postpones the program
-control flow from moving to the proceeding statement for a minimum of
+control flow from proceeding to the following statement for a minimum of
 $M$ time units and a maximum of $N$ time units, as shown in
@@
 (GALS) language it helps us easily describe a larger class of systems
-(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
+(like HRTCS in Figure~\ref{delay:a}) compared to purely synchronous
 languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
@@
 
-The SystemJ program is divided into two \hj{mutually asynchronous}
-clock-domains, the first clock-domain continuously displays a green
-light (\texttt{GREEN\_LIGHT} signal) on the touch screen sensor and
-waits for the user to respond. If the user is able to touch the screen
-within the specified time of at least 50.3 ms and at most 200.3 ms, a
-positive response is sent to the second clock-domain. All the SystemJ
-programming constructs used to implement this system are described in
-Table~\ref{tab:1} \hj{and more details can be found
-  in~\cite{amal10}}. The \texttt{wait\_inbetween} specification, which
-models the non exact postponement between 50.3 ms and 200.3 ms is not
-supported in the current language and cannot be compiled with the
-current SystemJ compiler. We introduce such \texttt{wait\_inbetween}
-statements in SystemJ. Such non exact real-time mechanisms are useful
-for developing systems where the real-time postponement is not known or
-should not be known a priori.
+The SystemJ program in Figure~\ref{delay:a} is divided into two
+\hj{mutually asynchronous} clock-domains, the first clock-domain
+continuously displays a green light (\texttt{GREEN\_LIGHT} signal) on
+the touch screen sensor and waits for the user to respond. If the user
+is able to touch the screen within the specified time of at least 50.3
+ms and at most 200.3 ms, a positive response is sent to the second
+clock-domain. All the SystemJ programming constructs used to implement
+this system are described in Table~\ref{tab:1} \hj{and more details can
+  be found in~\cite{amal10}}. The \texttt{wait\_inbetween}
+specification, which models the non exact postponement between 50.3 ms
+and 200.3 ms is not supported in the current language and cannot be
+compiled with the current SystemJ compiler. We introduce such
+\texttt{wait\_inbetween} statements in SystemJ. Such non exact real-time
+mechanisms are useful for developing systems where the real-time
+postponement is not known or should not be known a priori.
 

### 2014-03-31 exps.tex
ACM format -- not completed

@@
 \\ \midrule
-\multirow{2}{*}{AECS} & \texttt{wait\_atleast(10000 ms)} ($CD1$) 
-& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
+\multirow{3}{*}{AECS} & \texttt{wait\_atleast(10000 ms)} ($CD1$) 
+& \multirow{3}{*}{2} & \multirow{3}{*}{267} \\
+& \texttt{wait\_exact(10000 ms)} ($CD2$) & & \\
 & \texttt{wait\_exact(10000 ms)} ($CD2$) & &
@@
 	\cmidrule(r){2-8}
-	& d &1506-1788 & 57 & 143-163 & 46-617 & 56062 &77844\\ 
+	& d &1272-1739 & 57 & 143-163 & 46-617 & 56062 &77844\\ 
 	\cmidrule(r){1-8}
@@
 	\cmidrule(r){2-8}
-	& d & N/A & 40 & N/A & 213-2041 & N/A &53160\\ 
+	& d & N/A & 40 & N/A & 213-2041 & N/A &77844\\ 
 	\cmidrule(r){1-8}
@@
 applying Algorithm~\ref{alg:1}. In HRTCS, real-time delay of the range
-50.3..200.3 (ms) is only satisfied when \emph{d} is between 1506 and
-1788 (as shown previously in Section~\ref{sec:find-logic-delay}). Delay
+50.3..200.3 (ms) is only satisfied when \emph{d} is between 1272 and
+1739 (as shown previously in Section~\ref{sec:find-logic-delay}). Delay
 2.7..10.6 (ms) is satisfied in Printer/CD1 with \emph{d} of 143..163.
@@
 			\addplot + [black,fill=white,postaction={pattern=north east lines}] coordinates
-			{(HRTCS,1)(Printer/CD1,1)(Printer/CD2,0)(AECS/CD1,1)(AECS/CD2,0)};
+			{(HRTCS,1)(Printer/CD1,1)(Printer/CD2,1)(AECS/CD1,1)(AECS/CD2,1.94)};
 			\addplot + [black,fill=white,postaction={pattern=north west
@@
 			\addplot + [black,fill=white,postaction={pattern=north west lines}] coordinates
-                        {(Motor, 8.00)};
+                        {(Motor, 8.00) (AECS/CD2, 11.82)};
 			\addplot + [black,fill=white,postaction={pattern=dots}] coordinates

### 2014-03-31 introduction.tex
ACM format -- not completed

@@
 different input signals. Unlike a discrete physical clock, which has
-fixed real-time period, the period of the logical clock is
-elastic. Although the timing model with logical clock has worked well
-for reactive synchronous/GALS languages in designing discrete systems,
-there is a need to introduce real-time since often, implementation
-models are in the continuous or discrete real-time
-domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
-capabilities as first class language constructs create additional
-opportunity for formally verifying functional and non-functional
-requirements before system deployment.
+fixed real-time period, the period of the logical clock is elastic.
+Although the timing model with logical clock has worked well for
+reactive synchronous/GALS languages in designing discrete systems, there
+is a need to introduce real-time since often, implementation models are
+in the continuous or discrete real-time
+domain~\cite{DBLP:journals/pieee/SifakisTY03}. This paper targets
+enhancing formal GALS and synchronous languages, by adding real-time
+constructs. The real-time constructs speed up a design process by
+allowing system designers to explicitly program real-time postponements
+within the language itself without need of external resources such as
+timers.
+
+%Moreover, adding timing
+%capabilities as first class language constructs create additional
+%opportunity for formally verifying functional and non-functional
+%requirements before system deployment.
 
@@
 	SystemJ program as long as the amount of delay is within boundaries
-  that can be determined statically by program analysis, assuming that
-  real-time analyzable platforms are available,} (2) the waiting model
-relies on the use of relative instead of absolute real-time, i.e., a
-wait in selected time units is counted from the beginning of the wait
+that can be determined statically by program analysis.} (2) the waiting
+model relies on the use of relative instead of absolute real-time, i.e.,
+a wait in selected time units is counted from the beginning of the wait
 statement and (3) these waiting mechanisms allow mixing behaviors with
@@
   real-time in synchronous/GALS languages.
+\item We \textbf{assume} that real-time analyzable platforms are
+	available on which SystemJ programs are executed.
 \end{enumerate*}

### 2014-03-31 motivation.tex
ACM format -- not completed

@@
 \texttt{wait\_atleast} statement waits for a \textit{at least} 10
-seconds, but unlike \texttt{wait\_inbetween} there is no upper
-bound. The statement following the \texttt{wait\_atleast} statement is
+seconds, but unlike \texttt{wait\_inbetween} there is no upper bound
+i.e. program can proceed to the next statement at any time after 10
+seconds. The statement following the \texttt{wait\_atleast} statement is
 executed after a finite countable, but unbounded postponement. Such a

### 2014-03-31 introduction.tex
Updated the paper

@@
 domain~\cite{DBLP:journals/pieee/SifakisTY03}. This paper targets
-enhancing formal GALS and synchronous languages, by adding real-time
-constructs. The real-time constructs speed up a design process by
-allowing system designers to explicitly program real-time postponements
-within the language itself without need of external resources such as
-timers.
+enhancing the expressiveness of formal GALS and synchronous languages by
+adding real-time constructs. The real-time constructs speed up the
+design process by allowing system designers to explicitly program
+real-time control logic within the language itself without relying on
+external resources such as timers.
 
@@
   real-time in synchronous/GALS languages.
-\item We \textbf{assume} that real-time analyzable platforms are
-	available on which SystemJ programs are executed.
 \end{enumerate*}
 
-The rest of the paper is arranged as follows:
-Section~\ref{sec:background} gives the background on SystemJ and related
-techniques required for the rest of the
-paper. Section~\ref{sec:intr-real-time} introduces real-time mechanisms
-in the GALS paradigm and describes their compilation into logical
-time. Section~\ref{sec:proof-correctness} sketches the correctness
-proofs for the introduced
-algorithms. Section~\ref{sec:experimental-results} gives the
-experimental results for a set of real-time applications. In
+All aforementioned contributions hold under a single assumption that
+real-time analyzable platforms are available. The rest of the paper is
+arranged as follows: Section~\ref{sec:background} gives the background
+on SystemJ and related techniques required for the rest of the paper.
+Section~\ref{sec:intr-real-time} introduces real-time mechanisms in the
+GALS paradigm and describes their compilation into logical time.
+Section~\ref{sec:proof-correctness} sketches the correctness proofs for
+the introduced algorithms. Section~\ref{sec:experimental-results} gives
+the experimental results for a set of real-time applications. In
 Section~\ref{sec:disc-perc-limit} we discuss the advantages and

### 2014-03-31 motivation.tex
Updated the paper

--- 2014-02-02 motivation.tex
@@
 \texttt{wait\_atleast} statement waits for a \textit{at least} 10
-seconds, but unlike \texttt{wait\_inbetween} there is no upper
-bound. The statement following the \texttt{wait\_atleast} statement is
-executed after a finite countable, but unbounded postponement. Such a
-statement is useful for programming low priority tasks.
+seconds, but unlike \texttt{wait\_inbetween} there is no upper bound.
+For example in Figure~\ref{dd:b} program can proceed to the next
+statement at 11 seconds. The statement following the
+\texttt{wait\_atleast} statement is executed after a finite countable,
+but unbounded postponement. Such a statement is useful for programming
+low priority tasks.
 

### 2014-04-01 introduction.tex
Updated paper

@@
 enhancing the expressiveness of formal GALS and synchronous languages by
-adding real-time constructs. The real-time constructs speed up the
+adding real-time constructs. These real-time constructs speed up the
 design process by allowing system designers to explicitly program
 real-time control logic within the language itself without relying on
-external resources such as timers.
+external resources such as timers. Introduction of real-time as a
+language construct allows verification of temporal properties of
+programs before deployment.
 
@@
 
-All aforementioned contributions hold under a single assumption that
-real-time analyzable platforms are available. The rest of the paper is
-arranged as follows: Section~\ref{sec:background} gives the background
-on SystemJ and related techniques required for the rest of the paper.
+All aforementioned contributions require a single \textbf{assumption}
+that real-time analyzable platforms are available that allow computing
+the timing bounds of SystemJ programs. This assumption is necessary for
+all hard real-time system implementations~\cite{wilhelm08}.
+
+The rest of the paper is arranged as follows:
+Section~\ref{sec:background} gives the background on SystemJ and related
+techniques required for the rest of the paper.
 Section~\ref{sec:intr-real-time} introduces real-time mechanisms in the
